### 2024-12-04 Paper/main.tex
add manuscript

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Certified Symbolic Transducer with the Application of String Solver} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Shuanglong Kan}{Barkhausen Institut, Germany \and \url{https://github.com/ShlKan} }{shuanglongkan@gmail.com}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\authorrunning{J. Open Access and J.\,R. Public} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. 
\end{abstract}

\section{Introduction}
\label{sec:introduction}

\section{Lorem ipsum dolor sit amet}

Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. Donec eget odio et magna ullamcorper vehicula ut vitae libero. Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis. Pellentesque id ante erat. In hac habitasse platea dictumst. Curabitur a scelerisque odio. Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam. Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.

\begin{lemma}[Lorem ipsum]
\label{lemma:lorem}
Vestibulum sodales dolor et dui cursus iaculis. Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper. Proin facilisis gravida rutrum. Etiam sed sollicitudin lorem. Proin pellentesque risus at elit hendrerit pharetra. Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
\end{lemma}

\begin{proof}
Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.

\proofsubparagraph*{Just some paragraph within the proof.}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\begin{claim}
content...
\end{claim}
\begin{claimproof}
content...
    \begin{enumerate}
        \item abc abc abc \claimqedhere{}
    \end{enumerate}
\end{claimproof}

\end{proof}

\begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
\label{lemma:curabitur}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\end{corollary}

\begin{proposition}\label{prop1}
This is a proposition
\end{proposition}

\autoref{prop1} and \cref{prop1} \ldots

\subsection{Curabitur dictum felis id sapien}

Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat. Curabitur sed velit diam. Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu. Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet. Duis rutrum dapibus ornare. Nulla vehicula vulputate iaculis. Proin a consequat neque. Donec ut rutrum urna. Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nec faucibus leo. Cras ut nisl odio, non tincidunt lorem. Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat. Fusce risus libero, convallis at fermentum in, dignissim sed sem. Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}. Donec non suscipit lorem. Nam sit amet enim vitae nisl accumsan pretium.

\begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
for i:=maxint to 0 do 
begin 
    j:=square(root(i));
end;
\end{lstlisting}

\subsection{Proin ac fermentum augue}

Proin ac fermentum augue. Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis. Nulla facilisi. Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac. Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna. Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui. Praesent a pretium quam. Mauris non mauris sit amet eros rutrum aliquam id ut sapien. Nulla aliquet fringilla sagittis. Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor. Nulla cursus aliquet enim. Cras sapien risus, accumsan eu cursus ut, commodo vel velit. Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel. Integer faucibus faucibus felis.

\begin{itemize}
\item Ut vitae diam augue. 
\item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
\item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
\end{itemize}

\begin{remark}
content...
\end{remark}

\section{Pellentesque quis tortor}

Nec urna malesuada sollicitudin. Nulla facilisi. Vivamus aliquam tempus ligula eget ornare. Praesent eget magna ut turpis mattis cursus. Aliquam vel condimentum orci. Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi. Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.

\begin{lemma} [Quisque blandit tempus nunc]
Sed interdum nisl pretium non. Mauris sodales consequat risus vel consectetur. Aliquam erat volutpat. Nunc sed sapien ligula. Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus. Nunc vestibulum nunc ac massa pretium pharetra. Nulla facilisis turpis id augue venenatis blandit. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
\end{lemma}

Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.

\section{Morbi eros magna}

Morbi eros magna, vestibulum non posuere non, porta eu quam. Maecenas vitae orci risus, eget imperdiet mauris. Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis. Sed condimentum convallis dolor, a dignissim est ultrices eu. Donec consectetur volutpat eros, et ornare dui ultricies id. Vivamus eu augue eget dolor euismod ultrices et sit amet nisi. Vivamus malesuada leo ac leo ullamcorper tempor. Donec justo mi, tempor vitae aliquet non, faucibus eu lacus. Donec dictum gravida neque, non porta turpis imperdiet eget. Curabitur quis euismod ligula.

%%
%% Bibliography
%%

%% Please use bibtex,

\bibliography{lipics-v2021-sample-article}

\appendix

\section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}

List of different predefined enumeration styles:

\begin{itemize}
\item \verb|\begin{itemize}...\end{itemize}|
\item \dots
\item \dots
%\item \dots
\end{itemize}

\begin{enumerate}
\item \verb|\begin{enumerate}...\end{enumerate}|
\item \dots
\item \dots
%\item \dots
\end{enumerate}

\begin{alphaenumerate}
\item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
\item \dots
\item \dots
%\item \dots
\end{alphaenumerate}

\begin{romanenumerate}
\item \verb|\begin{romanenumerate}...\end{romanenumerate}|
\item \dots
\item \dots
%\item \dots
\end{romanenumerate}

\begin{bracketenumerate}
\item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
\item \dots
\item \dots
%\item \dots
\end{bracketenumerate}

\begin{description}
\item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
\item[Description 2] Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\item[Description 3]  \dots
%\item \dots
\end{description}

\cref{testenv-proposition} and \autoref{testenv-proposition} ...

\section{Theorem-like environments}\label{sec:theorem-environments}

List of different predefined enumeration styles:

\begin{theorem}\label{testenv-theorem}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{theorem}

\begin{lemma}\label{testenv-lemma}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{lemma}

\begin{corollary}\label{testenv-corollary}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{corollary}

\begin{proposition}\label{testenv-proposition}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{proposition}

\begin{conjecture}\label{testenv-conjecture}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{conjecture}

\begin{observation}\label{testenv-observation}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{observation}

\begin{exercise}\label{testenv-exercise}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{exercise}

\begin{definition}\label{testenv-definition}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{definition}

\begin{example}\label{testenv-example}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{example}

\begin{note}\label{testenv-note}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{note}

\begin{note*}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{note*}

\begin{remark}\label{testenv-remark}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{remark}

\begin{remark*}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{remark*}

\begin{claim}\label{testenv-claim}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claim}

\begin{claim*}\label{testenv-claim2}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claim*}

\begin{proof}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{proof}

\begin{claimproof}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claimproof}

\end{document}

### 2024-12-09 Paper/main.tex
add introduction

@@
 \begin{abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Praesent convallis orci arcu, eu mollis dolor.
-Aliquam eleifend suscipit lacinia.
-Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse potenti.
+Finite-State Automata (FA) and Finite-State Transducers (FT) are extensively utilized in programming languages and software engineering applications.
+For instance, regular expressions and their variations play a pivotal role in programming languages like JavaScript, Python, and others.
+Formalizing FAs and FTs in Coq, Isabelle/HOL, and other proof assistants are a popular topics.
+However, all these formalization are not practical in real-world applications.
+One of the reasons the transition labels are only single characters in the alphabet, for instance $q\xrightarrow{a} q'$ is a transition and $a$ is a single character.
+For real-world applications, the alphabet an FA or FT may be enormous or even \emph{infinite}.
+This classic way of transition definition can yield transitions explosion.
+
+A more practical way is to formalize symoblic FAs [1] and FTs [2], in which transition labels are symbolic and can be infinite. The work of [3] has done the work for FAs, but FTs 
+have not yet been formalized, which pose more challenges on proving the correctness of the formalization.
+
+In this paper, we aim to filling this gap.
+We gave a symbolic transducer formalization in Isabelle/HOL.
+The formalization is refinement-based and extensible with different symbolic representations of transition labels.
+In order to evaluate its effect and efficiency, we applied it to a  SMT string solver for modeling replacement operations in modern programming languages.
+The experimental results show the formalized symbolic transducer can efficient solver string constraints.
+
 \end{abstract}
@@
 \label{sec:introduction}
+
+Automata and Transducers are crucial concepts in formal languages and have widely applications in programming languages and software engineering.
+For instance, [1] has shown the correspondence between  regular expressions in modern regular expressions and variants of FAs and FTs.
+Other industrial usage of FAs and FTs is AWS access control polices checking.
+
+Even though there various  formalization of FAs and FTs in Coq, Isabelle, and other proof assistants.
+They are mainly based on classic definitions of FAs and FTs.
+There some drawbacks when come to practical applications.
+(1) transition labels are
+non-symbolic and usually finite.
+A classic and normal definition of a transition is $q\xrightarrow{a}q'$, where $a$ is a character in an finite alphabet.  This simple way will yield transition explosions. 
+For instance, if the alphabet $\Sigma$ is of the size $10,000$, then a transition from $q$ to $q'$ that accept any characters in $\Sigma$ need to split into $10,000$ transitions.
+Automata product in this way will be very inefficient.
+Moreover,
+The alphabet are usually finite. For practical applications, infinite alphabet are often necessary. For instance, if the alphabet is all integers.
+
+Symblic FAs and FTs [2,3,4] are extensions of classic FAs and FTs that make their application more practical.
+The transition labels are represented by algebra.
+For instances, intervals ($'a'-'z'$), boolean algebras ($x \% 2 == 0$), or others.
+This symbolic way is more succinct and its support for infinite alphabet extend the expressive power of FAs and FTs.
+
+Formalizing FAs and FTs in Isabelle/HOL, Coq, and other proof assists are more challenging compare with formalizing classic ones.
+Moreover, how to make the formalization extensible is also an important point to think, because, symbolic FAs and FTs support different algebras symbolic representations.
+For new algebra representations, we do not want repeat some proof works.
+
+Fortunately, symblic FAs has been formalized in Isabelle/HOL [certistr] and experiments in [certistr] illustrates the efficiency and effective of symbolic FAs.
+Unfortunately FTs are not formalized yet.
+FTs are more powerful and expressive than symbolic FAs.
+For instance, when FTs are support, replacement operation in modern programming languages, such as Javascript, python and others can be modelled as FTs.
+And CertiStr can be extended to support replacement operations.
+
+In this paper, we formalize symblic FTs based on symblic FAs.
+In order to solve the extensive problem, the formalization is refinement-based, in which at the abstraction level, transition labels are modeled as a general concept: sets.
 

### 2024-12-17 Paper/main.tex
add FT definitions

@@
 
+\lstset{
+	basicstyle=\ttfamily,
+	keywordstyle=\color{blue}\bfseries,
+	keywords={definition,if,then,else, where, record, fun, lemma},
+	escapeinside={|}{|}
+}
+
 \begin{document}
@@
 In this paper, we formalize symblic FTs based on symblic FAs.
-In order to solve the extensive problem, the formalization is refinement-based, in which at the abstraction level, transition labels are modeled as a general concept: sets.
+In order to solve the potential extension problem of FTs to support additional transition label theories, the formalization is refinement-based, in which at the abstraction level, transition labels are modeled as a general concept: sets.
+We can project any theory to sets.
+\emph{extend this}.
+The most important operation that we formalized and proved is the application operation on an input language.
+More precisely, given an FA $\mathcal{A}$ (to represent a regular language) as the input language and a FT $\mathcal{T}$, the application operation $\mathcal{F}~\mathcal{A}$ is the  of output languages of FT after receiving the input language
 
-\section{Lorem ipsum dolor sit amet}
+In the next refinement level, the states and transitions are stored in efficient data structure based on Isabelle/HOL refinement framework [2].
+For transition labels, we implemented an interval algebra, which are efficient for creating, checking, interaction.
+%
+In the future the refinement of transition label from sets to other algebras instead of intervals can be done easily by implementing the common interfaces with interval algebras.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}.
-Praesent convallis orci arcu, eu mollis dolor.
-Aliquam eleifend suscipit lacinia.
-Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse potenti.
-Donec eget odio et magna ullamcorper vehicula ut vitae libero.
-Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis.
-Pellentesque id ante erat.
-In hac habitasse platea dictumst.
-Curabitur a scelerisque odio.
-Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam.
-Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.
+The paper is organized as follows. 
+Section 2 presents the formalization of symbolic FTs.
+Section 3 presents the Core operations of symbolic FTs.
+
+\section{Formalization of Symbolic FTs}
+
+We firstly present a mathematical form of FTs.
+
+\begin{definition}[Label Theory]
+	A label theory
+\end{definition}
+
+\begin{definition}[Symbolic FTs]
+	A symbolic FT is defined a quadruple $(Q, \Delta , I, F)$, where
+	\begin{itemize}
+		\item $Q$ is a finite set of states,
+		\item $\Delta$ is a finite set of transitions of the form $(q, \phi, f, q')$ or in a more succinct form $q\xrightarrow{(\phi, f)}q'$.
+$\phi$ is the input of the transition, which is a predicate, a symbolic representation of input characters.
+$f$ is a label theory that generates the output of this transition,
+		\item $I$ is the set of initial states,
+		\item $F$ is the set of accepting states.
+	\end{itemize} 
+\end{definition}
+
+\begin{figure}[t]
+	\begin{lstlisting}
+record (|$'q,$||$~'a,$| |$'b$|) |$\mathbf{NFT}$| =
+	|$\mathcal{Q}$| :: "|$'q$| set"
+	|$\Delta$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
+	|$\mathcal{I}$| :: "|$'q$| set"
+	|$\mathcal{F}$| :: "|$'q$| set"
+	\end{lstlisting}
+\caption{The definitions of $FTs$}
+\label{fig-def-wf-acyclic}
+\end{figure}
 

### 2025-01-11 Paper/main.tex
clean files with gitignore

@@
 %\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-
+\usepackage{stmaryrd}
 %\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
@@
 
-\authorrunning{J. Open Access and J.\,R. Public} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
+\authorrunning{S. Kan} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
 
@@
 \EventNoEds{2}
-\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
-\EventShortTitle{CVIT 2016}
-\EventAcronym{CVIT}
-\EventYear{2016}
-\EventDate{December 24--27, 2016}
+\EventLongTitle{15th Conference on
+Interactive Theorem Proving (ITP 2025)}
+\EventShortTitle{ITP 2025}
+\EventAcronym{ITP}
+\EventYear{2025}
+\EventDate{July 24--27, 2025}
 \EventLocation{Little Whinging, United Kingdom}
@@
 	keywordstyle=\color{blue}\bfseries,
-	keywords={definition,if,then,else, where, record, fun, lemma},
+	keywords={definition,if,then,else, where, record, fun, lemma, type_synonym, fixes, assumes, shows},
 	escapeinside={|}{|}
@@
 
-Symblic FAs and FTs [2,3,4] are extensions of classic FAs and FTs that make their application more practical.
+Symbolic FAs and FTs [2,3,4] are extensions of classic FAs and FTs that make their application more practical.
 The transition labels are represented by algebra.
@@
 
-Fortunately, symblic FAs has been formalized in Isabelle/HOL [certistr] and experiments in [certistr] illustrates the efficiency and effective of symbolic FAs.
+Fortunately, symbolic FAs has been formalized in Isabelle/HOL [certistr] and experiments in [certistr] illustrates the efficiency and effective of symbolic FAs.
 Unfortunately FTs are not formalized yet.
@@
 
-In this paper, we formalize symblic FTs based on symblic FAs.
+In this paper, we formalize symbolic FTs based on symbolic FAs.
 In order to solve the potential extension problem of FTs to support additional transition label theories, the formalization is refinement-based, in which at the abstraction level, transition labels are modeled as a general concept: sets.
@@
 
-We firstly present a mathematical form of FTs.
+We first present a mathematical form of FTs, which is independent of Isabelle's formalization. 
+The key is the label theory, which defines how do output labels generated from input labels.
 
-\begin{definition}[Label Theory]
-	A label theory
+Let $\mathcal{U}$ be a multi-sorted carrier set or background universe, which is equipped with functions and relations over the elements.
+We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
+We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type. Assume $\mathbb{N}$ is the set of elements of the sort natural numbers.
+
+A lambda term is defined as $\lambda x. t$ of the type $\tau_1 \rightarrow \tau_2$.
+When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate.
+Assume $\psi$ is a predicate.
+We write $a\in \llbracket\psi \rrbracket$ if $\psi~a=\top$.
+For the non-predicate lambda term, we view it as a function, which generates an output element of type $\tau_2$ with the input term of type $\tau_1$.
+
+\begin{definition}[Symbolic Finite Transducer]
+\label{def-sft}
+   A Symbolic Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $(\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
+   \begin{itemize}
+   \item $\mathcal{Q}$ is a finite set of states,
+   \item $\mathcal{I}\subseteq \mathcal{Q}$ is the set of initial states,
+   \item $\mathcal{F} \subseteq\mathcal{Q}$ is the set of accepting states,
+   \item $\Delta$ is the set of transition relations.
+Each element in $\Delta$ is of the form $(q, \phi, f, q') (\text{or written in } q\xrightarrow{\phi, f} q')$, where $q$ and $q'$ are the states in $\mathcal{Q}$.
+   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
+   $f$ is a lambda term of the type $\tau_1\rightarrow \tau_2$. We call it output function in the following.
+   \end{itemize}
+
+For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called input,  then the application $f~a$ is the output.
+
 \end{definition}
 
-\begin{definition}[Symbolic FTs]
-	A symbolic FT is defined a quadruple $(Q, \Delta , I, F)$, where
-	\begin{itemize}
-		\item $Q$ is a finite set of states,
-		\item $\Delta$ is a finite set of transitions of the form $(q, \phi, f, q')$ or in a more succinct form $q\xrightarrow{(\phi, f)}q'$.
-$\phi$ is the input of the transition, which is a predicate, a symbolic representation of input characters.
-$f$ is a label theory that generates the output of this transition,
-		\item $I$ is the set of initial states,
-		\item $F$ is the set of accepting states.
-	\end{itemize} 
-\end{definition}
+SFTs accept an input word and generate an output. This can be defined by runs of SFTs.
+An SFT run $\sigma$ is a sequence of $(q_0, \psi_0, f_0, q_1),(q_1, \psi_1, f_1, q_2),\ldots, (q_{n-1}, \psi_{n-1}, q_n)$ such that $q_0\in \mathcal{I}_T$ and $(q_i, 
+\psi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta_T$.
+$\sigma$ is an accepting run when $q_n$ is an accepting state.
+For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ iff $a_i \in \llbracket\psi_i \rrbracket$ for $0 \leq i \leq n - 1$.
+Run $\sigma$ will generate an output sequence if $w$ is accepted by it, which is $w'= (f_0~a_0), \ldots (f_{n_1}~a_{n_1})$.
+We denote this operation as $\mathcal{N}~w = w'$.
+
+\subsection{The Isabelle/HOL Formalization of SFT}
 
@@
 record (|$'q,$||$~'a,$| |$'b$|) |$\mathbf{NFT}$| =
+	|$\mathcal{Q}_T$| :: "|$'q$| set"
+	|$\Delta_T$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
+	|$\mathcal{I}_T$| :: "|$'q$| set"
+	|$\mathcal{F}_T$| :: "|$'q$| set"
+        |$\mathcal{M}$| :: "|$'b\Rightarrow 'a$| Tlabel"
+
+type_synonym |$('q, 'a, 'b)$| LTTS = "|$('q \times ('a \text{ set option }\times b) \times 'q)$| set"
+
+type_synonym |$'a$| Tlabel = "|$'a \text{ option} \Rightarrow 'a$ set"
+	\end{lstlisting}
+\caption{The formalization of $FTs$ in Isabelle/HOL}
+\label{fig-def-FT}
+\end{figure}
+
+Figure \ref{fig-def-FT} shows our formalization of SFTs in Isabelle / HOL.
+The elements $\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$ are intuitive when referring to Definition \ref{def-sft}.
+However, the transitions are not exactly the same.
+%
+The set of transitions is defined by \texttt{LTTS} ( which is short for Labeled Transducer Transition System).
+Each transition is defined as $'q \times ('a \text{ set option }\times b) \times 'q$.
+There are some design decisions, which are made to make the FT formalization more flexible.
+
+Firstly, $'a \text{ set option }$ is the type input of the transition, it accepts a set of elements of type $'a$ or \texttt{None} regarding the semantics of the type \texttt{option}. 
+Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
+A predicate's semantics as introduced before represents a set of elements that make the predicate true.
+But predicates have various different forms.
+For instance, the interval $[1, 9]$ is a predicate that contains all elements $e$ between $1$ and $9$, i.e., $1 \leq e \leq 9$.
+The predicate $\lambda x : \mathbb{N}.~ x \text{ mod } 2 = 0$ denotes the set of even natural numbers.
+All these different forms of labels are abstracted as sets.
+
+%
+The value $\texttt{None}$ is used to denote $\varepsilon$-move, that is, it accepts nothing but may still output some elements.
+This design makes it easier model real world applications in FTs.
+We will show an example after introducing the output functions.
+
+The second element of type $'b$ is the \emph{indexes} of the output functions.
+$\mathcal{M}$ maps each of these indexes to an output function.
+The output function, defined by \texttt{Tlabel} maps a single input element to a set of output elements instead of single one.
+The reason is to allow FTs output non-deterministically, which means that the output may randomly or according to some criteria select one of the element in the set to output.
+
+Now let us look at an example.
+In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
+This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}).
+We can easily model this operation using our formalization of NFTs with the following steps:
+\begin{itemize}
+\item Step 1, construct a corresponding NFA from \texttt{pattern}).
+\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
+\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
+\item Step 4, unlabel previous accepting states as non-accepting states. 
+\end{itemize}
+
+We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
+
+\section{The Key Operation of NFTs}
+
+Given an NFT $\mathcal{N}$ and a word $w$, it is crucial and interesting that the possible output words can be calculated.
+This is the operation, we want to formalize, but we will generalize this operation.
+
+Instead of calculating the outputs for a single word, the operation calculates the outputs for a set of input words. Single word input is a special case with the set contains exactly only one element.
+
+The set of input words is modeled as a regular language.
+To be more precise, let $\mathcal{N}$ be an NFT, and $\mathcal{A}$ be an NFA.
+We define the operation $\texttt{PRODUCT} ~\mathcal{N}~\mathcal{A} = \{w'\mid w\in \mathcal{L}(\mathcal{A})\text{ and } w' = \mathcal{N}~\mathcal{A}\}$, where $\mathcal{L}(\mathcal{A})$ denotes the language of NFA $\mathcal{A}$.
+
+But there is one thing we need to consider, as our NFT formalization accepts $\varepsilon$-transitions, the output NFAs may contain $\varepsilon$-transitions as well.
+Unfortunately, [CertiStr] does not formalize $\varepsilon-$NFA.
+We formalize symbolic $\varepsilon-$NFA and the conversion from $\varepsilon-$NFA to NFA.
+We only present the definitions of $\varepsilon-$NFA and NFA in this paper.
+The more details including correctness proofs can be found in the Isabelle code.
+
+\begin{figure}[t]
+	\begin{lstlisting}
+record (|$'q,$||$~'a$|) |$\mathbf{NFA}$| =
 	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
+	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
 	|$\mathcal{I}$| :: "|$'q$| set"
 	|$\mathcal{F}$| :: "|$'q$| set"
+
+record (|$'q,$||$~'a$|) |$\mathbf{eNFA}$| =
+	|$\mathcal{Q}_e$| :: "|$'q$| set"
+	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
+        |$\Delta_e'$| :: "|$('q * 'q)$| set"
+	|$\mathcal{I}_e$| :: "|$'q$| set"
+	|$\mathcal{F}_e$| :: "|$'q$| set"
+
+type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
 	\end{lstlisting}
-\caption{The definitions of $FTs$}
-\label{fig-def-wf-acyclic}
+\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\label{fig-def-FAs}
 \end{figure}
 
-\begin{lemma}[Lorem ipsum]
-\label{lemma:lorem}
-Vestibulum sodales dolor et dui cursus iaculis.
-Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper.
-Proin facilisis gravida rutrum.
-Etiam sed sollicitudin lorem.
-Proin pellentesque risus at elit hendrerit pharetra.
-Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
-\end{lemma}
+Figure \ref{fig-def-FAs} shows the definitions of NFA and $\varepsilon-$NFA.
+The record type \texttt{eNFA} corresponds to $\varepsilon-$NFA.
+$\Delta_e$ is same as $\Delta$ in \texttt{NFA}.
+$\varepsilon$-transitions are defined by $\Delta_e'$.
+Now we are ready to present the operation \emph{APPLICATION}.
+Let $\mathcal{T}$ be an SFT and  $\mathcal{A}$ be an NFA.
+The output language generated by $\mathcal{T}$ with the input language $\mathcal{A}$ is an an \texttt{eNFA} $\mathcal{A}'$.
 
-\begin{proof}
-Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.
+\begin{figure}[ht]
+	\begin{lstlisting}
+definition productT :: |$"('q,'a,'b,'c)"$| NFT |$\Rightarrow$| ('q,'a) NFA |$\Rightarrow$| 
+           |$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}\Rightarrow 'b\text{ set option})\Rightarrow$|
+           |$('q\times 'q, ~'b)$| eNFA where
+  "productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
+    |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
+    |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
+          |$(p, (\texttt{None}, f), q)\in \Delta_T \mathcal{T} \land \exists S. (\mathcal{M}~\mathcal{T})~f~\text{None} = \text{Some } S\} ~\cup$|
+        |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
+          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_T~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 = \emptyset \land$|
+            |$ \exists S.~F~((\mathcal{M}~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
+    |$\Delta_e' = \{((p,p'),~\text{the }(((\mathcal{M}~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
+          |$(p, (None, f), q)\in \Delta_T~\mathcal{T} \land (\mathcal{M}~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
+        |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
+          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_T~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 = \emptyset \land$|
+            |$ \exists x\in(\sigma_1\cap\sigma_2).~((\mathcal{M}~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
+    |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
+    |$\mathcal{F}_e = \mathcal{F}_t~\mathcal{T} \times \mathcal{F}~\mathcal{A}$| |$\rrparenthesis$|"
+	\end{lstlisting}
+\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\label{fig-def-FTProd}
+\end{figure}
 
-\proofsubparagraph*{Just some paragraph within the proof.}
-Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-\begin{claim}
-content...
-\end{claim}
-\begin{claimproof}
-content...
-    \begin{enumerate}
-        \item abc abc abc \claimqedhere{}
-    \end{enumerate}
-\end{claimproof}
+Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an NFA.
+The parameters $\mathcal{T}$ and $\mathcal{A}$ are intuitive.
+But we need to explain the role of \texttt{F}.
+The output function is of type $'a \Rightarrow 'b \text{set option}$, which applies to a single element of the type $'a$.
+$\texttt{F}$ extends this function to apply it to a set of elements.
+More precisely, let $f$ be an output function, the semantics of $\texttt{F}$ is defined as $\texttt{F}~f~A=\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset)$.
+The transition relations $\Delta_e$ and $\Delta_e'$ are defined by splitting into two cases, whether the transitions in SFTs are $\varepsilon$-transitions or not.
 
-\end{proof}
+In order to prove this abstract definition of product operation's correctness, we firstly formalize the outputs of an NFT regarding the input of an NFA.
+Figure \ref{fig-def-output} has two definitions: (1) \texttt{outputE} extracts the output (second element) from a path.
+(2) \texttt{outputL} defines the set of all outputs of an NFT $\mathcal{T}$, regarding the input language $\mathcal{A}$.
 
-\begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
-\label{lemma:curabitur}
-Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-\end{corollary}
+\begin{figure}[ht]
+	\begin{lstlisting}
+fun outputE :: "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" where
+  "outputE [] = []" |$\mid$|
+  "outputE ((_, Some a) # l) = a # (outputE l)" |$\mid$|
+  "outputE ((_, None) # l) = (outputE l)"
 
-\begin{proposition}\label{prop1}
-This is a proposition
-\end{proposition}
+definition outputL :: "|$('q, 'a, 'b, 'c)\text{  NFT} \Rightarrow $|
+            |$ ('c \Rightarrow ('a, 'b)\text{  Tlabel}) \Rightarrow 
+                               ('q, 'a) \text{ NFA\_rec} \Rightarrow 'b\text{ list set}$|" where
+  "outputL |$\mathcal{T}$| M |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. |$q\in \mathcal{I}_T~\mathcal{T} \land q' \in \mathcal{F}_T~\mathcal{T} \land$|
+                |$ \text{LTTS\_reachable }(\Delta_T~\mathcal{T})~\text{M}~q~\pi~q'\land \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}" 
+	\end{lstlisting}
+\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\label{fig-def-output}
+\end{figure}
 
-\autoref{prop1} and \cref{prop1} \ldots
+\begin{figure}[ht]
+	\begin{lstlisting}
+lemma productT_correct:
+  fixes |$\mathcal{T}$| |$\mathcal{A}$| F
+  assumes F_ok1: "|$\forall~f~s.~(\forall e \in s. f (\text{ Some } e) = \text{None}) \mapsto \text{ F}~f~s = \text{None}"$|
+      and F_ok2: "|$\forall~f~s.~\text{F }f~s = $|
+            |$\text{Some }(\cup \{\text{the }(f (\text{ Some }e))\mid e. e \in s \land f~(\text{Some }e) \neq \text{ None}\})$|"
+      and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
+    shows "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL |$\mathcal{T}~(\mathcal{M}~\mathcal{T})~\mathcal{A}$|"
 
-\subsection{Curabitur dictum felis id sapien}
+	\end{lstlisting}
+\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\label{fig-def-product-correct}
+\end{figure}
 
-Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat.
-Curabitur sed velit diam.
-Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu.
-Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet.
-Duis rutrum dapibus ornare.
-Nulla vehicula vulputate iaculis.
-Proin a consequat neque.
-Donec ut rutrum urna.
-Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Aenean nec faucibus leo.
-Cras ut nisl odio, non tincidunt lorem.
-Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat.
-Fusce risus libero, convallis at fermentum in, dignissim sed sem.
-Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}.
-Donec non suscipit lorem.
-Nam sit amet enim vitae nisl accumsan pretium.
+Figure \ref{fig-def-product-correct} shows the correctness of the product operation and Figure \ref{fig-def-output} contains the definitions of \texttt{outputE} and \texttt{outputL} that are used in Lemma \texttt{productT\_correct}.
+The two assumptions define the constraints of the function \texttt{F}.
 
-\begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
-for i:=maxint to 0 do 
-begin 
-    j:=square(root(i));
-end;
-\end{lstlisting}
-
-\subsection{Proin ac fermentum augue}
-
-Proin ac fermentum augue.
-Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis.
-Nulla facilisi.
-Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac.
-Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna.
-Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui.
-Praesent a pretium quam.
-Mauris non mauris sit amet eros rutrum aliquam id ut sapien.
-Nulla aliquet fringilla sagittis.
-Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor.
-Nulla cursus aliquet enim.
-Cras sapien risus, accumsan eu cursus ut, commodo vel velit.
-Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel.
-Integer faucibus faucibus felis.
-
-\begin{itemize}
-\item Ut vitae diam augue. 
-\item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
-\item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
-\end{itemize}
-
-\begin{remark}
-content...
-\end{remark}
-
-\section{Pellentesque quis tortor}
-
-Nec urna malesuada sollicitudin.
-Nulla facilisi.
-Vivamus aliquam tempus ligula eget ornare.
-Praesent eget magna ut turpis mattis cursus.
-Aliquam vel condimentum orci.
-Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi.
-Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.
-
-\begin{lemma} [Quisque blandit tempus nunc]
-Sed interdum nisl pretium non.
-Mauris sodales consequat risus vel consectetur.
-Aliquam erat volutpat.
-Nunc sed sapien ligula.
-Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus.
-Nunc vestibulum nunc ac massa pretium pharetra.
-Nulla facilisis turpis id augue venenatis blandit.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-\end{lemma}
-
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-
-\section{Morbi eros magna}
-
-Morbi eros magna, vestibulum non posuere non, porta eu quam.
-Maecenas vitae orci risus, eget imperdiet mauris.
-Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis.
-Sed condimentum convallis dolor, a dignissim est ultrices eu.
-Donec consectetur volutpat eros, et ornare dui ultricies id.
-Vivamus eu augue eget dolor euismod ultrices et sit amet nisi.
-Vivamus malesuada leo ac leo ullamcorper tempor.
-Donec justo mi, tempor vitae aliquet non, faucibus eu lacus.
-Donec dictum gravida neque, non porta turpis imperdiet eget.
-Curabitur quis euismod ligula.
-
-%%
-%% Bibliography
-%%
-
-%% Please use bibtex,
-
-\bibliography{lipics-v2021-sample-article}
-
-\appendix
-
-\section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}
-
-List of different predefined enumeration styles:
-
-\begin{itemize}
-\item \verb|\begin{itemize}...\end{itemize}|
-\item \dots
-\item \dots
-%\item \dots
-\end{itemize}
-
-\begin{enumerate}
-\item \verb|\begin{enumerate}...\end{enumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{enumerate}
-
-\begin{alphaenumerate}
-\item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{alphaenumerate}
-
-\begin{romanenumerate}
-\item \verb|\begin{romanenumerate}...\end{romanenumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{romanenumerate}
-
-\begin{bracketenumerate}
-\item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{bracketenumerate}
-
-\begin{description}
-\item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
-\item[Description 2] Fusce eu leo nisi.
-Cras eget orci neque, eleifend dapibus felis.
-Duis et leo dui.
-Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\item[Description 3]  \dots
-%\item \dots
-\end{description}
-
-\cref{testenv-proposition} and \autoref{testenv-proposition} ...
-
-\section{Theorem-like environments}\label{sec:theorem-environments}
-
-List of different predefined enumeration styles:
-
-\begin{theorem}\label{testenv-theorem}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{theorem}
-
-\begin{lemma}\label{testenv-lemma}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{lemma}
-
-\begin{corollary}\label{testenv-corollary}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{corollary}
-
-\begin{proposition}\label{testenv-proposition}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{proposition}
-
-\begin{conjecture}\label{testenv-conjecture}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{conjecture}
-
-\begin{observation}\label{testenv-observation}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{observation}
-
-\begin{exercise}\label{testenv-exercise}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{exercise}
-
-\begin{definition}\label{testenv-definition}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{definition}
-
-\begin{example}\label{testenv-example}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{example}
-
-\begin{note}\label{testenv-note}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{note}
-
-\begin{note*}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{note*}
-
-\begin{remark}\label{testenv-remark}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{remark}
-
-\begin{remark*}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{remark*}
-
-\begin{claim}\label{testenv-claim}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{claim}
-
-\begin{claim*}\label{testenv-claim2}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{claim*}
-
-\begin{proof}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{proof}
-
-\begin{claimproof}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{claimproof}
+\section{Algorithm Level Refinement}
+We have
 

### 2025-01-21 Paper/main.tex
add Automata

@@
 In the future the refinement of transition label from sets to other algebras instead of intervals can be done easily by implementing the common interfaces with interval algebras.
+
+We highlight our formulization with the following points.
+(1) Efficiency, the algebraic representation
+support compressed storage of transition labels and efficiency checking
+of membership.
+(2) Extensible, the refinement based formulization allows the extension
+of the SNFA to support different theories of transition labels.
 
@@
 The two assumptions define the constraints of the function \texttt{F}.
+The conclusion in \textcolor{blue}{\texttt{show}} illustrates 
+that the language equivalence between the production definition and mathematical semantics.
 

### 2025-02-05 Paper/main.tex
stage saving

@@
 \usepackage{stmaryrd}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+\usetikzlibrary{intersections}
+\usetikzlibrary{automata}
+\usetikzlibrary{positioning}
 %\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
@@
 \begin{abstract}
-Finite-State Automata (FA) and Finite-State Transducers (FT) are extensively utilized in programming languages and software engineering applications.
-For instance, regular expressions and their variations play a pivotal role in programming languages like JavaScript, Python, and others.
-Formalizing FAs and FTs in Coq, Isabelle/HOL, and other proof assistants are a popular topics.
-However, all these formalization are not practical in real-world applications.
+Finite-State Automata (FAs) and Finite-State Transducers (FTs) are extensively used in programming languages and software engineering applications.
+Regular expressions and their variations, for example, play a crucial role in languages like JavaScript, Python, and others.
+While formalizing FAs and FTs in proof assistants such as Coq and Isabelle/HOL is a popular topic, these formalizations often lack practicality for real-world applications.
+
 One of the reasons the transition labels are only single characters in the alphabet, for instance $q\xrightarrow{a} q'$ is a transition and $a$ is a single character.
@@
 
-A more practical way is to formalize symoblic FAs [1] and FTs [2], in which transition labels are symbolic and can be infinite. The work of [3] has done the work for FAs, but FTs 
-have not yet been formalized, which pose more challenges on proving the correctness of the formalization.
+A more pragmatic approach involves the formalization of symbolic Finite Automata (FAs) [1] and Finite Transducers (FTs) [2], where transition labels are symbolic and potentially infinite.
+Although the formalization of symbolic FAs has been comprehensively addressed in the work of [3], the formalization of symbolic FTs remains uncharted.
+This presents additional challenges in establishing the correctness of such formalizations.
 
@@
 \section{The Key Operation of NFTs}
+\label{sec-form-fts}
 
@@
 $\varepsilon$-transitions are defined by $\Delta_e'$.
-Now we are ready to present the operation \emph{APPLICATION}.
+Now we are ready to present the operation \emph{Product}.
 Let $\mathcal{T}$ be an SFT and  $\mathcal{A}$ be an NFA.
@@
 But we need to explain the role of \texttt{F}.
-The output function is of type $'a \Rightarrow 'b \text{set option}$, which applies to a single element of the type $'a$.
+The output function is of type "$'a \Rightarrow 'b \text{set option}$", which applies to a single element of the type $'a$.
 $\texttt{F}$ extends this function to apply it to a set of elements.
@@
 
+$\Delta_e$ has two cases:(1) when $(p, (\text{None}, f),  q)$ is a transition in the NFT.
+Because the input can be empty ($\varepsilon$), therefore the NFA does not need to move.
+Therefore the product transtion just output of $f~\text{None}$.
+(2) when $(p, (Some~\sigma_1, f), q)$ is a transition in the NFT.
+This transition can only synchronize with the transitions in the NFA that have shared characters with $\sigma_1$.
+And the output is $F~f~(\sigma_1\cap\sigma_2)$, where $\sigma_2$ is the input label of the NFA.
+Similarly, $\Delta_e'$ has two cases too.
+
 In order to prove this abstract definition of product operation's correctness, we firstly formalize the outputs of an NFT regarding the input of an NFA.
 Figure \ref{fig-def-output} has two definitions: (1) \texttt{outputE} extracts the output (second element) from a path.
-(2) \texttt{outputL} defines the set of all outputs of an NFT $\mathcal{T}$, regarding the input language $\mathcal{A}$.
+(2) \texttt{outputL} defines the set of all outputs of an NFT $\mathcal{T}$, when the input is in the  language $\mathcal{A}$.
 
@@
 \section{Algorithm Level Refinement}
-We have
+\label{sec_alg_refinement}
+In Section \ref{sec-form-fts}, we give an abstract definition of NFT's product operation. In this section, we present the possible way to compute the Product from programming language's perspective.
+In this level, we will also refine the transition labels from \emph{sets} to interval algebra.
+
+\subsection{Intervals}
+An interval is defined as a pair $[i, j]$, which corresponds to the set $\{e \mid e \geq i \land e \leq j\}$.
+In our formalization, we extends the interval definition to a list, that is $[(i_1, j_1), \ldots, (i_n, j_n)]$, which corresponds to the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq i_n\}$.
+This definition enables us to merge more transitions in NFAs.
+In the sequent, intervals always mean interval lists.
+There are a collection interfaces for intervals. 
+\begin{enumerate}
+  \item $\texttt{semI}~i$, formalize the semantics of an interval as a set.
+  \item $\texttt{empty\_intervals}~i$, checks the emptiness of an interval.
+  \item $\texttt{intersectIs}~i_1~i_2$, computes the intersection of two intervals.
+  \item $\texttt{diffIs}~i_1~i_2$, computes the intervals difference, that is the interval corresponds to the set $\texttt{semI}(i_1)-\texttt{semI}(i_2)$.
+\end{enumerate}
+
+\subsection{Algorithm of Product Computation.}
+
+Fig.
+\ref{fig-compute-nft-product} illustrates the computation of NFT products.
+\texttt{prods\_imp} computes the product of two sets.
+It is illustrated in \ref{fig-def-prods_imp}.
+We use the constructs from \emph{Refine\_Monadic}.
+The algorithm itself is intuitive: the nested two loops traverse all possible combination of states in the states.
+
+\begin{figure}[ht]
+	\begin{lstlisting}
+definition productT_impl where
+  "productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
+    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) (nfa_states |$\mathcal{A}$|);
+    (D1, D2) |$\leftarrow$| trans_comp_imp (nft_tranfun |$\mathcal{T}$|) F fe 
+        (nft_trans |$\mathcal{T}$|) (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
+    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|) (nfa_initial |$\mathcal{A}$|);
+    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) (nfa_accepting |$\mathcal{A}$|);
+    RETURN (Q,D1,D2,I, F)
+  }"
+\end{lstlisting}
+\caption{The computation of NFT product}
+\label{fig-compute-nft-product}
+\end{figure}
+
+\begin{figure}[ht]
+	\begin{lstlisting}
+definition prods_imp where
+  "prods_imp Q1 Q2 =
+   FOREACH {q. q |$\in$| s.|$\alpha$| Q1} (|$\lambda$| q Q. do {
+    S |$\leftarrow$| FOREACH {q. q |$\in$| s.|$\alpha$| Q2}
+         (|$\lambda$| q' Q'. RETURN (ss.ins (q,q') Q')) (ss.empty ());
+    RETURN (ss.union Q S)
+   }) (ss.empty ())"
+\end{lstlisting}
+\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\label{fig-def-prods_imp}
+\end{figure}
+
+The challenging part is how to compute the transitions sets.
+It is illustrated in Fig.
+\ref{fig-def-prods-imp}.
+The function \texttt{trans\_comp\_imp} traverse the transtions in the NFT and split each transtion into two cases.
+This function uses another two functions:
+
+1. \texttt{copy\_tran\_impl}
+
+2. \texttt{copy\_tran'\_impl}
+
+\begin{figure}[ht]
+	\begin{lstlisting}
+definition trans_comp_imp where
+  "trans_comp_imp M F fe T1 T2 Q =
+    FOREACH {t. t |$\in$| dt.|$\alpha$| T1}
+     (|$\lambda$| (q, (|$\alpha$|, f), q') (D1, D2). 
+       (if (|$\alpha$| = None) then 
+          (if ((M f) None = None) then 
+            do {
+              D2' |$\leftarrow$| copy_tran_imp q q' Q;
+              RETURN (D1, ssd.union D2 D2')
+            }
+            else do {
+              D1' |$\leftarrow$| copy_tran'_imp q (the (M f None)) q' Q;
+              RETURN (ddt.union D1 D1', D2)
+            })
+          else 
+            (subtrans_comp_imp M q (the |$\alpha$|) f q' F fe T2 D1 D2)
+        )) (ddt.empty, ssd.empty ())"
+
+\end{lstlisting}
+\caption{The computation of transition sets}
+\label{fig-def-prods-imp}
+\end{figure}
+
+\section{Refinement to Efficient Data Structure}
+
+\section{An Application to String Solver}
+\label{sec-app-str-solver}
+In this section, we will show how to use the formalized symbolic transducer to solve the string constraints with replacement operations.
+
+A replacement operation has the following form: $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
+This operation replaces the occurrence of the substring in \texttt{str} that matches \texttt{pattern} with \texttt{replacement}.
+\texttt{str} and \texttt{replacement} are strings and \texttt{pattern} is a regular expression.
+
+The semantics of the replacement operation can be described in two cases: (1) if there are strings in \texttt{str} that match \texttt{pattern}, then replace them with \texttt{replacement}.
+(2) If there are no matches, then the result is \texttt{str}.
+
+The first case can be modeled by the SFTs.
+We illustrate this case by an example.
+Given a replacement operation $\texttt{replace}(s, $"$[0-9]+$"$, $"$\text{ITP}$"$)$, we can model this operation by an SFT.
+
+There are three steps to model this operation:
+
+\begin{enumerate}
+  \item Generate an SNFA from the regular expression "$[0-9]+$". Adding output functions to the transitions of the SNFA that outputs empty characters to make it an SFT.
+  \item Generate an SNFA from the replacement string "$"$\text{ITP}$"$".
+Note that a single string is also a regular expression.
+Adding input labels and output functions to the transitions of the SNFA to make it an SFT.
+  \item Concatenate the two SFTs and adding preceding and succeeding transitions to make it accept the whole input string.
+\end{enumerate}
+
+We explain it step by step.
+
+\noindent\emph{Step 1.}
+The SNFA corresponding to the regular expression "$[0-9]+$" is shown in Figure \ref{fig-snfa-pattern}.
+On the left side is the SNFA and on the right side is the corresponding SFT.
+The function \texttt{f} is the output function that outputs empty characters.
+It can be defined as $\texttt{f} = \lambda x.~\texttt{None}$.
+The function \texttt{g} is the output function that outputs the characters in the replacement string.
+It can be defined as (the ascii code of $'I'$, $'T'$, $'P'$ are 73, 84, 80 respectively):
+
+\[
+\texttt{g} = \lambda~i~x.~\texttt{match }~i~ with~ 1 \rightarrow [(73, 73)] ~|~ 2 \rightarrow [(84, 84)] ~|~ 3 \rightarrow [(80, 80)] ~|~ \_ ~\rightarrow \texttt{None}
+\]
+
+\begin{figure}[h] \centering
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+  \node[state, initial] (q0)   {$q_0$}; 
+  \node[state, accepting] (q1) [right=of q0] {$q_1$};
+
+   \path[->] 
+   (q0) edge [bend left] node {[0-9]} (q1)
+   (q1) edge [loop above] node {[0-9]} (q1);
+
+  \node[state, initial] (q0') [right=4cm of q1] {$q_0'$}; 
+  \node[state, accepting] (q1') [right=of q0'] {$q_1'$};
+
+   \path[->] 
+   (q0') edge [bend left] node {[0-9]/\texttt{f}} (q1')
+   (q1') edge [loop above] node {[0-9]/\texttt{f}} (q1');
+\end{tikzpicture}
+\caption{The SFA for "$[0-9]+$" duplicated}
+\label{fig-snfa-pattern}
+\end{figure}
+
+\noindent\emph{Step 2.}
+The SNFA corresponding to the replacement string "\text{ITP}" is shown in Figure \ref{fig-snfa-replacement}.
+On the left side is the SNFA and on the right side is the corresponding SFT.
+The function \texttt{f} is the output function that outputs the characters in the replacement string.
+It can be defined as $\texttt{f} = \lambda x.~\texttt{Some } x$.
+
+\begin{figure}[h] \centering
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+  % Original NFA for "ITP"
+  \node[state, initial] (q0)   {$q_0$}; 
+  \node[state] (q1) [right=of q0] {$q_1$}; 
+  \node[state] (q2) [below=of q0] {$q_2$}; 
+  \node[state, accepting] (q3) [right=of q2] {$q_3$};
+
+  \path[->] 
+  (q0) edge node {I} (q1)
+  (q1) edge node {T} (q2)
+  (q2) edge node {P} (q3);
+
+  % Duplicate NFA for "ITP" on the right
+  \node[state, initial] (q0') [right=4cm of q1] {$q_0'$}; 
+  \node[state] (q1') [right=of q0'] {$q_1'$}; 
+  \node[state] (q2') [below=of q0'] {$q_2'$}; 
+  \node[state, accepting] (q3') [right=of q2'] {$q_3'$};
+
+  \path[->] 
+  (q0') edge node {I/\texttt{g 1}} (q1')
+  (q1') edge node {T/\texttt{g 2}} (q2')
+  (q2') edge node {P/\texttt{g 3}} (q3');
+\end{tikzpicture}
+\caption{The SFA for "$"\text{ITP}"$" duplicated}
+\label{fig-snfa-replacement}
+\end{figure}
+
+\noindent\emph{Step 3.}
+The concatenation of the two SFTs is shown in Figure \ref{fig-rearranged-automata}.
+It contains two steps: (1) rearrange the transitions of the two SFTs to make them accept the whole input string.
+(2) add the preceding and succeeding transitions to make it accept the whole input string.
+
+\begin{figure}[h] \centering
+  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+    % Second automaton from the first figure
+    \node[state, initial] (q0') {$q_0'$}; 
+    \node[state] (q1') [right=of q0'] {$q_1'$};
+
+     \path[->] 
+     (q0') edge [bend left] node {[0-9]/\texttt{f}} (q1')
+     (q1') edge [loop above] node {[0-9]/\texttt{f}} (q1')
+     (q0') edge [loop below] node {[\texttt{None}/\texttt{f}]} (q0');
+
+    % Second automaton from the second figure
+    \node[state] (q0'') [right=4cm of q1'] {$q_0''$}; 
+    \node[state] (q1'') [right=of q0''] {$q_1''$}; 
+    \node[state] (q2'') [below=of q0''] {$q_2''$}; 
+    \node[state, accepting] (q3'') [right=of q2''] {$q_3''$};
+
+    \path[->] 
+    (q0'') edge node {I/\texttt{g 1}} (q1'')
+    (q1'') edge node {T/\texttt{g 2}} (q2'')
+    (q2'') edge node {P/\texttt{g 3}} (q3'')
+    (q3'') edge [loop below] node {[\texttt{None}/\texttt{f}]} (q3'')
+    (q1') edge node {[\texttt{None}/\texttt{f}]} (q0'');
+  \end{tikzpicture}
+  \caption{The SFT for the replacement operation $\texttt{replace}(s, $"$[0-9]+$"$, $"$\text{ITP}$"$)$}
+  \label{fig-rearranged-automata}
+  \end{figure}
 

### 2025-02-10 Paper/main.tex
polish and proceed the paper

@@
 
-\begin{figure}[t]
+\begin{figure}[hbt!]
 	\begin{lstlisting}
@@
 
-\begin{figure}[ht]
+\begin{figure}[hbt!]
 	\begin{lstlisting}
@@
 \subsection{Intervals}
-An interval is defined as a pair $[i, j]$, which corresponds to the set $\{e \mid e \geq i \land e \leq j\}$.
+An interval is defined as a pair $(i, j)$, which corresponds to the set $\{e \mid e \geq i \land e \leq j\}$.
 In our formalization, we extends the interval definition to a list, that is $[(i_1, j_1), \ldots, (i_n, j_n)]$, which corresponds to the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq i_n\}$.
 This definition enables us to merge more transitions in NFAs.
+Moreover, when an interval operation produces more than two intervals, we do not need to split the intervals.
+For instance, the interval result of $(1, 5) - (3, 4)$ (the difference between the interval $[1, 5]$ and $[3, 4]$) is equivalent to $[(1, 2), (5, 5)]$.
+
 In the sequent, intervals always mean interval lists.
@@
 \begin{enumerate}
-  \item $\texttt{semI}~i$, formalize the semantics of an interval as a set.
-  \item $\texttt{empty\_intervals}~i$, checks the emptiness of an interval.
+  \item $\texttt{semI}~i$, formalize the semantics of an interval as a set, that is $\texttt{semI }(i,j) = \{e \mid e \geq i \land e \leq j\}$.
+  \item $\texttt{emptyI}~i$, checks the emptiness of an interval.
+   \item $\texttt{nemptyI}~i$, checks the non-emptiness of an interval.
   \item $\texttt{intersectIs}~i_1~i_2$, computes the intersection of two intervals.
@@
 
+We define a canonical form for intervals to facilitate the proof. An interval $[i_1, j_1), \ldots, (i_n, j_n)]$ is in canonical form if 
+(1) $i_k \leq j_k, 1 \leq k \leq n$,
+(2) $j_k\leq i_{k+1}, 1 \leq k < n$.
+All the interval operations are proved to keep the canonical form after the operation if the input intervals are in canonical form.
+
 \subsection{Algorithm of Product Computation.}
@@
 Fig.
-\ref{fig-compute-nft-product} illustrates the computation of NFT products.
+\ref{fig-compute-nft-product} illustrates the product computation of a SFT and an SFA.
+\footnote{To make it easier to understand, the Isabelle code presented in this section is not exactly the same as the real one.}
 \texttt{prods\_imp} computes the product of two sets.
@@
 We use the constructs from \emph{Refine\_Monadic}.
-The algorithm itself is intuitive: the nested two loops traverse all possible combination of states in the states.
+\texttt{prods\_imp} itself is intuitive: the nested two loops traverse all possible combination of states in the states.
+The \texttt{FOREACH} construct in \emph{Refine\_Monadic} is similar to OCaml's \texttt{Set.fold}.
+$\texttt{FOREACH}~S~f~I$ traverses all elements in $S$ and applies $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$ to each element.
+$I$ is the initial value of type $'b$.
 
-\begin{figure}[ht]
+\begin{figure}[hbt!]
 	\begin{lstlisting}
@@
   "prods_imp Q1 Q2 =
-   FOREACH {q. q |$\in$| s.|$\alpha$| Q1} (|$\lambda$| q Q. do {
-    S |$\leftarrow$| FOREACH {q. q |$\in$| s.|$\alpha$| Q2}
-         (|$\lambda$| q' Q'. RETURN (ss.ins (q,q') Q')) (ss.empty ());
-    RETURN (ss.union Q S)
-   }) (ss.empty ())"
+   FOREACH {q. q |$\in$| Q1} (|$\lambda$| q Q. do {
+    S |$\leftarrow$| FOREACH {q. q |$\in$| Q2}
+         (|$\lambda$| q|$'$| Q|$'$|. RETURN ({(q,q|$'$|)} |$\cup$| Q|$'$|)) |$\emptyset$|;
+    RETURN (Q |$\cup$| S)
+   }) |$\emptyset$|"
 \end{lstlisting}
@@
 
-The challenging part is how to compute the transitions sets.
-It is illustrated in Fig.
-\ref{fig-def-prods-imp}.
-The function \texttt{trans\_comp\_imp} traverse the transtions in the NFT and split each transtion into two cases.
-This function uses another two functions:
-
-1. \texttt{copy\_tran\_impl}
-
-2. \texttt{copy\_tran'\_impl}
-
-\begin{figure}[ht]
+\begin{figure}[hbt!]
 	\begin{lstlisting}
 definition trans_comp_imp where
-  "trans_comp_imp M F fe T1 T2 Q =
-    FOREACH {t. t |$\in$| dt.|$\alpha$| T1}
-     (|$\lambda$| (q, (|$\alpha$|, f), q') (D1, D2). 
+  "trans_comp_imp M F T1 T2 Q =
+    FOREACH {t. t |$\in$| T1}
+      (|$\lambda$| (q, (|$\alpha$|, f), q|$'$|) (D1, D2). 
         (if (|$\alpha$| = None) then 
-          (if ((M f) None = None) then 
-            do {
-              D2' |$\leftarrow$| copy_tran_imp q q' Q;
-              RETURN (D1, ssd.union D2 D2')
-            }
-            else do {
-              D1' |$\leftarrow$| copy_tran'_imp q (the (M f None)) q' Q;
-              RETURN (ddt.union D1 D1', D2)
-            })
+          (subtrans_comp_|$\varepsilon$| M q  f q|$'$| F T2 D1 D2)
         else 
-            (subtrans_comp_imp M q (the |$\alpha$|) f q' F fe T2 D1 D2)
-        )) (ddt.empty, ssd.empty ())"
-
+          (subtrans_comp M q (the |$\alpha$|) f q|$'$| F T2 D1 D2)
+      )) (|$\emptyset$|, |$\emptyset$|)"
 \end{lstlisting}
@@
 
-\section{Refinement to Efficient Data Structure}
+The challenging part is how to compute the transition sets.
+It is illustrated in Fig.
+\ref{fig-def-prods-imp}.
+The function \texttt{trans\_comp\_imp} traverses the transitions in the SFT and it calls two the following two functions to compute the synchronized transitions:
+(1) \texttt{subtrans\_comp\_$\varepsilon$}: handles the case that the transition in the SFT is $\varepsilon$-transition.
+(2) \texttt{subtrans\_comp}: handles the case that the transition in the SFT is not $\varepsilon$-transition.
+
+We elaborate only on the function \texttt{subtrans\_comp} in Fig. \ref{fig-def-subtrans_comp}. The other function is similar. 
+This function traverses the transitions SFA's transitions $\texttt{T}$. If it find the a transtion $(q1, \alpha', q1')\in \texttt{T}$ satisfies the following condition:
+$\alpha \cap \alpha' \neq \emptyset$ (Checked by \texttt{nemptyIs}),
+it started to compute the non-$\varepsilon$-transitions set $\texttt{D1}$ and the $\varepsilon$-transitions set $\texttt{D2}$ by checking whether the output set of $\alpha \cap \alpha'$ is empty or not.
+
+\begin{figure}[hbt!]
+	\begin{lstlisting}
+definition subtrans_comp_imp where
+  "subtrans_comp_imp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
+    FOREACH
+      {t. t |$\in$| T} (|$\lambda$| (q1, |$\alpha'$|, q1|$'$|) (D1, D2).
+        (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then do {
+          D1 |$\leftarrow$| (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) |$\neq$| None) 
+            then RETURN 
+            let |$\alpha_i$| = (the (F (M f) (intersectIs |$\alpha$| |$\alpha'$|))) in
+            ({(q,q1) |$\alpha_i$| (q|$'$|,q1|$'$|)} |$\cup$| D1)
+            else RETURN D1);
+          D2 |$\leftarrow$| (if (M f) (intersectIs |$\alpha$| |$\alpha'$|) |$\neq$| |$\emptyset$| then 
+            RETURN (((q,q1), (q|$'$|,q1|$'$|)) |$\cup$| D2) else RETURN D2);
+          RETURN (D1, D2)
+        }
+        else (RETURN (D1, D2)))) (D1, D2)"
+    \end{lstlisting}
+    \caption{The computation of transition sets}
+    \label{fig-def-subtrans_comp}
+    \end{figure}
+
+  \begin{figure}[hbt!]
+    \begin{lstlisting}
+  lemma productT_imp_correct:
+  assumes finite_TT: "|$\text{finite } (\Delta_T~\mathcal{T})$|"    
+      and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
+      and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
+      and finite_TQ: "|$\text{finite } (\mathcal{Q}_T~\mathcal{T})$|"
+      and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
+      and finite_TI: "|$\text{finite } (\mathcal{I}_T~\mathcal{T})$|"
+      and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
+      and finite_TF: "|$\text{finite } (\mathcal{F}_T~\mathcal{T})$|"
+  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe} \leq \text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
+  \end{lstlisting}
+  \caption{The refinement of the product computation}
+  \label{fig-def-productT_imp_correct}
+  \end{figure}
+
+  To prove the correctness of the product computation introduced above, we just need to show that the product result of \texttt{productT\_imp} in Fig.
+\ref{fig-def-prods_imp} is a refinement of the product result of \texttt{productT} in Figure \ref{fig-compute-nft-product}.
+This refinement specification is illustrated in Figure \ref{fig-def-productT_imp_correct}.
+The proof is under the framework of \emph{Refine\_Monadic}.
+The specification $\mathbf{C} \leq\texttt{SPEC}~\mathbf{A}$ means that $\mathbf{C}$ is an element in the set $\mathbf{A}$.
+In Fig.
+\ref{fig-def-productT_imp_correct}, "$\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F$" specifies a set of elements $A$, such that $A$ equals to $\texttt{productT}~\mathcal{T}~\mathcal{A}~F$.
+  The specification will force us to prove the result $\varepsilon$-NFAs of \texttt{productT\_imp}  and \texttt{productT} have the same sets of initial states, transition relations, $\varepsilon$-transitions, and accepting states.
+
+\subsection{Refinement to Efficient Data Structure}
+
+The last subsection shows the way to compute the product of two NFAs and the transition labels have been refined to interval algebra. But it still uses the data structure of sets in Isabelle/HOL.
+
+The \emph{Refine\_Monadic} framework provides a way to refine the data structure of sets to more efficient data structure, such as red-black trees.
+The basic idea is to provide a set of \emph{set} operations on the refined data structure and prove the refinement relation between the operations on the refined data structure and the operations on the abstract data structure.
+%
+In our formalization, we refine the data structure of sets to red-black trees.
 
@@
 
+CertiStr is an existing string solver based on SFAs.
+However, due to the limitation of SFAs, it cannot solve the string constraints with replacement operations.
+CertiStr is based on forward-propagations.
+We will not elaborate and repeat the idea of forward-propagation in this paper.
+In this section, we only focus on how to model the replacement operation in SFTs.
+
+\subsection{Modeling the Replacement Operation}
+
 A replacement operation has the following form: $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
@@
 
-The semantics of the replacement operation can be described in two cases: (1) if there are strings in \texttt{str} that match \texttt{pattern}, then replace them with \texttt{replacement}.
-(2) If there are no matches, then the result is \texttt{str}.
+The semantics of the replacement operation can be described in two cases: (1) if there are a substring in \texttt{str} that match \texttt{pattern}, then replace it with \texttt{replacement}.
+(2) If there are no matches, then the result is \texttt{str} itself.
 
-The first case can be modeled by the SFTs.
+The first case can be modeled by SFTs.
 We illustrate this case by an example.
@@
 \begin{enumerate}
-  \item Generate an SNFA from the regular expression "$[0-9]+$". Adding output functions to the transitions of the SNFA that outputs empty characters to make it an SFT.
+  \item Generate an SFA from the regular expression "$[0-9]+$". Adding output functions to the transitions of the SFA that output empty symbol (\texttt{None}) to make it an SFT.
   \item Generate an SNFA from the replacement string "$"$\text{ITP}$"$".
 Note that a single string is also a regular expression.
-Adding input labels and output functions to the transitions of the SNFA to make it an SFT.
-  \item Concatenate the two SFTs and adding preceding and succeeding transitions to make it accept the whole input string.
+Add input labels and output functions to the transitions of the SNFA to make it an SFT.
+  \item Concatenate the two SFTs and add preceding and succeeding self-loop transitions to make it accept the whole input string.
 \end{enumerate}
 
-We explain it step by step.
+We explain it step by step by the example.
 
@@
 On the left side is the SNFA and on the right side is the corresponding SFT.
-The function \texttt{f} is the output function that outputs empty characters.
+The function \texttt{f} is the output function that outputs empty symbol.
 It can be defined as $\texttt{f} = \lambda x.~\texttt{None}$.
-The function \texttt{g} is the output function that outputs the characters in the replacement string.
-It can be defined as (the ascii code of $'I'$, $'T'$, $'P'$ are 73, 84, 80 respectively):
-
-\[
-\texttt{g} = \lambda~i~x.~\texttt{match }~i~ with~ 1 \rightarrow [(73, 73)] ~|~ 2 \rightarrow [(84, 84)] ~|~ 3 \rightarrow [(80, 80)] ~|~ \_ ~\rightarrow \texttt{None}
-\]
 
 \begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+\begin{tikzpicture}[shorten >=1pt, node distance=1.5cm, auto]
   \node[state, initial] (q0)   {$q_0$}; 
@@
 The SNFA corresponding to the replacement string "\text{ITP}" is shown in Figure \ref{fig-snfa-replacement}.
-On the left side is the SNFA and on the right side is the corresponding SFT.
-The function \texttt{f} is the output function that outputs the characters in the replacement string.
-It can be defined as $\texttt{f} = \lambda x.~\texttt{Some } x$.
+On the left side is the SFA and on the right side is the corresponding SFT.
+The function \texttt{g} is the output function that outputs the characters in the replacement string.
+It can be defined as (the ascii code of $'I'$, $'T'$, $'P'$ are 73, 84, 80 respectively):
+\[
+\texttt{g} = \lambda~i~x.~\texttt{match }~i~ with~ 1 \rightarrow [(73, 73)] ~|~ 2 \rightarrow [(84, 84)] ~|~ 3 \rightarrow [(80, 80)] ~|~ \_ ~\rightarrow \texttt{None}
+\]
 
 \begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+\begin{tikzpicture}[shorten >=1pt, node distance=1.5cm, auto]
   % Original NFA for "ITP"
@@
   \path[->] 
-  (q0') edge node {I/\texttt{g 1}} (q1')
-  (q1') edge node {T/\texttt{g 2}} (q2')
-  (q2') edge node {P/\texttt{g 3}} (q3');
+  (q0') edge node {\texttt{None}/\texttt{g 1}} (q1')
+  (q1') edge node {\texttt{None}/\texttt{g 2}} (q2')
+  (q2') edge node {\texttt{None}/\texttt{g 3}} (q3');
 \end{tikzpicture}
@@
 The concatenation of the two SFTs is shown in Figure \ref{fig-rearranged-automata}.
-It contains two steps: (1) rearrange the transitions of the two SFTs to make them accept the whole input string.
+It contains two minor steps: (1) add a transtion (\texttt{None}/\texttt{f}) from each accepting state of the first SFT to the initial state of the second SFT.
 (2) add the preceding and succeeding transitions to make it accept the whole input string.
@@
 \begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+  \begin{tikzpicture}[shorten >=1pt, node distance=1.5cm, auto]
     % Second automaton from the first figure
@@
      (q1') edge [loop above] node {[0-9]/\texttt{f}} (q1')
-     (q0') edge [loop below] node {[\texttt{None}/\texttt{f}]} (q0');
+     (q0') edge [loop below] node {\texttt{None}/\texttt{f}} (q0');
 
@@
     \path[->] 
-    (q0'') edge node {I/\texttt{g 1}} (q1'')
-    (q1'') edge node {T/\texttt{g 2}} (q2'')
-    (q2'') edge node {P/\texttt{g 3}} (q3'')
-    (q3'') edge [loop below] node {[\texttt{None}/\texttt{f}]} (q3'')
-    (q1') edge node {[\texttt{None}/\texttt{f}]} (q0'');
+    (q0'') edge node {\texttt{None}/\texttt{g 1}} (q1'')
+    (q1'') edge node {\texttt{None}/\texttt{g 2}} (q2'')
+    (q2'') edge node {\texttt{None}/\texttt{g 3}} (q3'')
+    (q3'') edge [loop below] node {\texttt{None}/\texttt{f}} (q3'')
+    (q1') edge node {\texttt{None}/\texttt{f}} (q0'');
   \end{tikzpicture}
@@
 
+  With this SFT, we can compute the forward image of the replacement operation.
+  Let $s' = \texttt{replace}(s, $"$[0-9]+$"$, $"$\text{ITP}$"$)$ be a string constraint and we want to compute the domain of $s'$ regarding this replacement operation.
+  Assume the SFT generated from this replacement operation is $\mathcal{T}$ and the string $s$'s domain ranges over an SFA $\mathcal{A}$.
+The foward image of this replacement operation is $\mathcal{T}~\mathcal{A}$.
+Therefore, the domain of $s'$ is $\mathcal{T}~\mathcal{A}$.
+
+\subsection{Experiments}
+
+We have implemented a prototype of the string solver to support the replacement operation based on CertiStr[1], which is called CertiStr-R in this paper.
+
+CertiStr-R supports two kinds of replacement operations: (1) \texttt{str.replace s p r}, where the pattern \texttt{p} is a string and (2) \texttt{str.replace\_re s p r}, where the pattern \texttt{p} is a regular expression.
+
+The limitation of CertiStr-R is that it does not support \texttt{replace\_all} and \texttt{replace\_re\_all}.
+CertiStr-R is possible to extend in the future.
+Another limitation is that the semantics of \texttt{replace} and \texttt{replace\_re} are different from the definition in SMTLIB.
+
+\section{Related Work}
+
+\section{Conclusion}
+
 \end{document}

### 2025-02-14 Paper/main.tex
add benchmark and support for str.allchar

@@
 \begin{abstract}
-Finite-State Automata (FAs) and Finite-State Transducers (FTs) are extensively used in programming languages and software engineering applications.
-Regular expressions and their variations, for example, play a crucial role in languages like JavaScript, Python, and others.
-While formalizing FAs and FTs in proof assistants such as Coq and Isabelle/HOL is a popular topic, these formalizations often lack practicality for real-world applications.
+Finite-State Automata (FAs) and Finite-State Transducers (FTs) are integral components in the domains of programming languages and software engineering.
+Regular expressions, which are often implemented using FAs, play a pivotal role in languages such as JavaScript and Python.
+FTs extend the capabilities of FAs by allowing transformations of input strings into output strings, thus providing a more expressive framework for operations that involve both recognition and transformation.
+Despite the popularity of formalizing FAs and FTs within proof assistants like Coq and Isabelle/HOL, these formalizations frequently fall short in terms of applicability to real-world scenarios.
 
-One of the reasons the transition labels are only single characters in the alphabet, for instance $q\xrightarrow{a} q'$ is a transition and $a$ is a single character.
-For real-world applications, the alphabet an FA or FT may be enormous or even \emph{infinite}.
-This classic way of transition definition can yield transitions explosion.
+One limitation of traditional finite-state models is that transition labels are typically restricted to single characters from a finite alphabet.
+For example, a transition is often represented as $q \xrightarrow{a} q'$, where $a$ is a single character.
+However, in real-world applications, the alphabet of a FA or FT can be vast or even \emph{infinite}.
+This conventional approach to defining transitions can lead to a phenomenon known as transition explosion, where the number of transitions becomes unmanageably large.
 
-A more pragmatic approach involves the formalization of symbolic Finite Automata (FAs) [1] and Finite Transducers (FTs) [2], where transition labels are symbolic and potentially infinite.
-Although the formalization of symbolic FAs has been comprehensively addressed in the work of [3], the formalization of symbolic FTs remains uncharted.
-This presents additional challenges in establishing the correctness of such formalizations.
+A more pragmatic approach involves the formalization of symbolic FAs [1] and FTs [2], wherein transition labels are symbolic and potentially infinite.
+While the formalization of symbolic FAs has been comprehensively explored in the work of [3], the formalization of symbolic FTs remains largely unexplored.
+This gap presents additional challenges in establishing the correctness and applicability of such formalizations.
 
-In this paper, we aim to filling this gap.
-We gave a symbolic transducer formalization in Isabelle/HOL.
-The formalization is refinement-based and extensible with different symbolic representations of transition labels.
-In order to evaluate its effect and efficiency, we applied it to a  SMT string solver for modeling replacement operations in modern programming languages.
-The experimental results show the formalized symbolic transducer can efficient solver string constraints.
+In this paper, we endeavor to address this gap by presenting a formalization of symbolic transducers within the Isabelle/HOL framework.
+This formalization is refinement-based and is designed to be extensible with various symbolic representations of transition labels.
+To evaluate its effectiveness and efficiency, we applied this formalization to an SMT string solver for modeling replacement operations.
+The experimental results demonstrate that the formalized symbolic transducer can efficiently solve string constraints.
 

### 2025-02-16 CertiStr/test/benchmark/benchmark.sh
add tests

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    result=$(CertiStr $test)
+    echo "Test ($test): $result"
+done

### 2025-02-16 CertiStr/test/sat/sat.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    result=$(CertiStr $test)
+    if ( [ $result != 'SAT' ])
+    then
+        echo "Test ($test) Fail:"
+        echo $result
+        exit
+    fi
+done
+
+echo 'Success'

### 2025-02-16 CertiStr/test/unsat/unsat.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    result=$(CertiStr $test)
+    if ( [ $result != 'UNSAT' ])
+    then
+        echo "Test ($test) Fail:"
+        echo $result
+        exit
+    fi
+done
+
+echo 'Success'

### 2025-02-17 Paper/main.tex
remove unnecessary file

@@
 A more pragmatic approach involves the formalization of symbolic FAs [1] and FTs [2], wherein transition labels are symbolic and potentially infinite.
-While the formalization of symbolic FAs has been comprehensively explored in the work of [3], the formalization of symbolic FTs remains largely unexplored.
+While the formalization of symbolic FAs has been comprehensively explored in the work of \cite{cpp/KanLRS22}, the formalization of symbolic FTs remains largely unexplored.
 This gap presents additional challenges in establishing the correctness and applicability of such formalizations.
@@
 
-We first present a mathematical form of FTs, which is independent of Isabelle's formalization. 
-The key is the label theory, which defines how do output labels generated from input labels.
+We first present a pure mathematical form of FTs \cite{VeanesHLMB12Transducer}, which is independent of Isabelle's formalization. 
+The central concept is the label theory, which characterizes the process by which output labels are derived from input labels.
 
@@
 In order to prove this abstract definition of product operation's correctness, we firstly formalize the outputs of an NFT regarding the input of an NFA.
-Figure \ref{fig-def-output} has two definitions: (1) \texttt{outputE} extracts the output (second element) from a path.
+Figure \ref{fig-def-output} has two definitions: (1) \texttt{outputE} extracts the output (second element) from a path in the NFT.
 (2) \texttt{outputL} defines the set of all outputs of an NFT $\mathcal{T}$, when the input is in the  language $\mathcal{A}$.
@@
 
+\bibliography{literature}
 \end{document}

### 2025-02-17 Paper/main.tex
polish: papers

@@
 
-We first present a pure mathematical form of FTs \cite{VeanesHLMB12Transducer}, which is independent of Isabelle's formalization. 
+We first present a pure mathematical form of SFTs \cite{VeanesHLMB12Transducer}, which is independent of Isabelle's formalization. 
 The central concept is the label theory, which characterizes the process by which output labels are derived from input labels.
@@
 For the non-predicate lambda term, we view it as a function, which generates an output element of type $\tau_2$ with the input term of type $\tau_1$.
+With these notations and the above definitions, we can define SFTs as follows.
 
@@
 \label{def-sft}
-   A Symbolic Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $(\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
+   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
    \begin{itemize}
@@
    \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \phi, f, q') (\text{or written in } q\xrightarrow{\phi, f} q')$, where $q$ and $q'$ are the states in $\mathcal{Q}$.
-   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of the type $\tau_1\rightarrow \tau_2$. We call it output function in the following.
+Each element in $\Delta$ is of the form $(q, \psi, f, q') \text{ or}$ $\text{written in }q\xrightarrow{\psi, f} q'$, where $q$ and $q'$ are the states in $\mathcal{Q}$.
+   $\psi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
+   $f$ is a lambda term of the type $\tau_1\rightarrow \tau_2$. $f$ is called "output function".
    \end{itemize}
 
-For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called input,  then the application $f~a$ is the output.
+For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called an input,  then the application $(f~a)$ is the output.
 
@@
 
-SFTs accept an input word and generate an output. This can be defined by runs of SFTs.
-An SFT run $\sigma$ is a sequence of $(q_0, \psi_0, f_0, q_1),(q_1, \psi_1, f_1, q_2),\ldots, (q_{n-1}, \psi_{n-1}, q_n)$ such that $q_0\in \mathcal{I}_T$ and $(q_i, 
-\psi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta_T$.
+SFTs accept an input word and generate an output. This can be defined by \emph{runs} of SFTs.
+An SFT run $\sigma$ is a sequence $(q_0, \psi_0, f_0, q_1),(q_1, \psi_1, f_1, q_2),\ldots, (q_{n-1}, \psi_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i, 
+\psi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
 $\sigma$ is an accepting run when $q_n$ is an accepting state.
+
 For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ iff $a_i \in \llbracket\psi_i \rrbracket$ for $0 \leq i \leq n - 1$.
 Run $\sigma$ will generate an output sequence if $w$ is accepted by it, which is $w'= (f_0~a_0), \ldots (f_{n_1}~a_{n_1})$.
-We denote this operation as $\mathcal{N}~w = w'$.
+We call in the sequent:
+\begin{itemize}
+\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{path},
+\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the path and $(f_0~a_0), \ldots, (f_{n-1}~a_{n-1})$ the \emph{output} of the path.
+\end{itemize}
 
-\subsection{The Isabelle/HOL Formalization of SFT}
+If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the path is an accepting path of $\mathcal{T}$.
+For a path $\pi$, we denote the input of $\pi$ as $in(\pi)$ and the output of $\pi$ as $out(\pi)$.
+%
+Given an SFT $\mathcal{T}$ and a word $w$, we define the operation product of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with the input $w$.
+More precisely,
+\[
+\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting path }\land in(\pi) = w \land out(\pi) = w'\}.
+\]
+
+To make the operation \emph{product} more general, we extend the operation to an SFT and a set of intput words represented by a regular language, which can be denoted as an NFA. More precisely,
+\[
+\mathcal{T}\times\{\mathcal{A}\} = \{w'\mid w.~w\in \mathcal{L}(\mathcal{A})\land w' = \mathcal{T}\times\{w\} \}, \text{where $\mathcal{L}(\mathcal{A})$ denotes the language of $\mathcal{A}$}.
+\]
+
+\begin{figure}[hbt!]
+  \centering
+  \begin{tikzpicture}[shorten >=1pt, node distance=3.5cm, on grid, auto]
+    \node[state, initial] (q_0)   {$q_0$}; 
+    \node[state] (q_1) [right=of q_0] {$q_1$}; 
+    \node[state, accepting] (q_2) [right=of q_1] {$q_2$};
+
+    \path[->] 
+    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
+    (q_1) edge node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
+    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
+    (q_2) edge[loop above] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
+ \end{tikzpicture}
+    \caption{The formalization of $FTs$ in Isabelle/HOL}
+    \label{fig-example-ft}
+    \end{figure}   
+We now illustrate SFTs with an example in Figure \ref{fig-example-ft}. It accepts strings of the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$. 
+It uses intervals $[i-j]$ as the input label, which is a predicate that returns true if the input letter is between $i$ and $j$.
+The function $f_1=\lambda x.~ \texttt{toUpper}(x)$ maps each lowercase letter to uppercase letter, and $f_2=\lambda x.~\texttt{toLower}(x)$ maps each uppercase letter to lowercase letter.
+Give an string "$\texttt{bigSMALL}$", the SFT will generate an output "$\texttt{BIGsmall}$".
+
+\subsection{The Isabelle/HOL Formalization of SFTs}
 
@@
 
-\section{The Key Operation of NFTs}
+\section{The Product Operation of NFTs}
 \label{sec-form-fts}
@@
 
-In order to prove this abstract definition of product operation's correctness, we firstly formalize the outputs of an NFT regarding the input of an NFA.
-Figure \ref{fig-def-output} has two definitions: (1) \texttt{outputE} extracts the output (second element) from a path in the NFT.
-(2) \texttt{outputL} defines the set of all outputs of an NFT $\mathcal{T}$, when the input is in the  language $\mathcal{A}$.
+To establish the correctness of the abstract definition of the product operation, we first formalize the outputs of a Non-deterministic Finite Transducer (NFT) with respect to the input of a Non-deterministic Finite Automaton (NFA).
+Figure \ref{fig-def-output} presents two key definitions: (1) \texttt{outputE}, which extracts the sequence of outputs $o_1, \ldots, o_n$ from a given NFT path $(i_1, o_1), \ldots, (i_n, o_n)$; and (2) \texttt{outputL}, which represents the set of all possible outputs generated by an NFT $\mathcal{T}$ for the input language defined by an NFA $\mathcal{A}$.
 

### 2025-02-17 Paper/main.tex
add support for variable equal

@@
 Each transition is defined as $'q \times ('a \text{ set option }\times b) \times 'q$.
-There are some design decisions, which are made to make the FT formalization more flexible.
+There are some design decisions, which are made to make the FT formalization more abstract and flexible.
 
-Firstly, $'a \text{ set option }$ is the type input of the transition, it accepts a set of elements of type $'a$ or \texttt{None} regarding the semantics of the type \texttt{option}. 
+Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or \texttt{None} regarding the semantics of the type \texttt{option}. 
 Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
@@
 The predicate $\lambda x : \mathbb{N}.~ x \text{ mod } 2 = 0$ denotes the set of even natural numbers.
-All these different forms of labels are abstracted as sets.
+All these different forms of labels are abstracted as sets in the formalization.
 

### 2025-02-17 CertiStr/test/benchmark/benchmark.sh

@@ -1,6 +1,7 @@
 #!/bin/bash
 
 for test in ./*.smt2; do
+    echo "Test ($test):"
     result=$(CertiStr $test)
-    echo "Test ($test): $result"
+    echo "$result"
 done

### 2025-02-20 Paper/main.tex
add support regex pattern

@@
 
+We extend CertiStr \cite{cpp/KanLRS22} not only to support SFTs but also $\varepsilon$SFA, including the conversion from $\varepsilon$SFA to SFA and the the correctness of the conversion.
+
 \section{An Application to String Solver}
@@
 We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}(s, $"$[0-9]+$"$, $"$\text{ITP}$"$)$, we can model this operation by an SFT.
-
-There are three steps to model this operation:
+Given a replacement operation $\texttt{replace}(s, $\texttt{/[0-9]+/}$, $"$\texttt{NUM}$"$)$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string"\texttt{NUM}".
+We model this replacement operation by an SFT with the following 3 steps:
 
 \begin{enumerate}
-  \item Generate an SFA from the regular expression "$[0-9]+$". Adding output functions to the transitions of the SFA that output empty symbol (\texttt{None}) to make it an SFT.
-  \item Generate an SNFA from the replacement string "$"$\text{ITP}$"$".
+  \item Generate an SFA from the regular expression $\texttt{/[0-9]+/}$. Adding output functions to the transitions of the SFA that output empty string (\texttt{None}) to make it an SFT.
+  \item Generate an SNFA from the replacement string "\texttt{NUM}".
 Note that a single string is also a regular expression.
@@
 
-We explain it step by step by the example.
-
 \noindent\emph{Step 1.}
-The SNFA corresponding to the regular expression "$[0-9]+$" is shown in Figure \ref{fig-snfa-pattern}.
-On the left side is the SNFA and on the right side is the corresponding SFT.
-The function \texttt{f} is the output function that outputs empty symbol.
-It can be defined as $\texttt{f} = \lambda x.~\texttt{None}$.
+The SFA corresponding to the regular expression "$\texttt{/[0-9]+/}$" is shown in Figure \ref{fig-snfa-pattern}.
+On the left side is the SFA and on the right side is the corresponding SFT.
+The function $\texttt{f} = \lambda x.~\texttt{None}$ is the output function that outputs empty string.
 
 \begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.5cm, auto]
+\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
   \node[state, initial] (q0)   {$q_0$}; 
@@
    \path[->] 
-   (q0) edge [bend left] node {[0-9]} (q1)
-   (q1) edge [loop above] node {[0-9]} (q1);
+   (q0) edge [bend left] node {\texttt{[0-9]}} (q1)
+   (q1) edge [loop above] node {\texttt{[0-9]}} (q1);
 
@@
    \path[->] 
-   (q0') edge [bend left] node {[0-9]/\texttt{f}} (q1')
-   (q1') edge [loop above] node {[0-9]/\texttt{f}} (q1');
+   (q0') edge [bend left] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
+   (q1') edge [loop above] node {\texttt{[0-9]}/$\texttt{f}$} (q1');
 \end{tikzpicture}
-\caption{The SFA for "$[0-9]+$" duplicated}
+\caption{Corresponding SFA and SFT for \texttt{/[0-9]+/}}
 \label{fig-snfa-pattern}
@@
 \noindent\emph{Step 2.}
-The SNFA corresponding to the replacement string "\text{ITP}" is shown in Figure \ref{fig-snfa-replacement}.
+The SNFA corresponding to the replacement string "\texttt{NUM}" is shown in Figure \ref{fig-snfa-replacement}.
 On the left side is the SFA and on the right side is the corresponding SFT.
 The function \texttt{g} is the output function that outputs the characters in the replacement string.
-It can be defined as (the ascii code of $'I'$, $'T'$, $'P'$ are 73, 84, 80 respectively):
+\texttt{g} is defined as:
 \[
-\texttt{g} = \lambda~i~x.~\texttt{match }~i~ with~ 1 \rightarrow [(73, 73)] ~|~ 2 \rightarrow [(84, 84)] ~|~ 3 \rightarrow [(80, 80)] ~|~ \_ ~\rightarrow \texttt{None}
+\texttt{g} = \lambda~i~x.~\texttt{match }~i~ with~ 1 \rightarrow [(78, 78)] ~|~ 2 \rightarrow [(85, 85)] ~|~ 3 \rightarrow [(77, 77)] ~|~ \_ ~\rightarrow \texttt{None}
 \]
+Note that the ASCII code of  $\texttt{'N'}$, $\texttt{'U'}$, $\texttt{'M'}$ are 78, 85, 77 respectively.
 
 \begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.5cm, auto]
+\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
   % Original NFA for "ITP"
@@
   \path[->] 
-  (q0) edge node {I} (q1)
-  (q1) edge node {T} (q2)
-  (q2) edge node {P} (q3);
+  (q0) edge node {\texttt{N}} (q1)
+  (q1) edge node {\texttt{U}} (q2)
+  (q2) edge node {\texttt{M}} (q3);
 
@@
   (q1') edge node {\texttt{None}/\texttt{g 2}} (q2')
-  (q2') edge node {\texttt{None}/\texttt{g 3}} (q3');
+  (q2') edge node[below] {\texttt{None}/\texttt{g 3}} (q3');
 \end{tikzpicture}
-\caption{The SFA for "$"\text{ITP}"$" duplicated}
+\caption{Corresponding SFA and SFT for "\texttt{NUM}"}
 \label{fig-snfa-replacement}
@@
 \noindent\emph{Step 3.}
-The concatenation of the two SFTs is shown in Figure \ref{fig-rearranged-automata}.
+The concatenation of the two SFTs in Fig.
+\ref{fig-snfa-pattern} and Fig.
+\ref{fig-snfa-replacement} is shown in Figure \ref{fig-rearranged-automata}.
 It contains two minor steps: (1) add a transtion (\texttt{None}/\texttt{f}) from each accepting state of the first SFT to the initial state of the second SFT.
@@
 \begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=1.5cm, auto]
+  \begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
     % Second automaton from the first figure
@@
      \path[->] 
-     (q0') edge [bend left] node {[0-9]/\texttt{f}} (q1')
-     (q1') edge [loop above] node {[0-9]/\texttt{f}} (q1')
+     (q0') edge [bend left] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
+     (q1') edge [loop above] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
      (q0') edge [loop below] node {\texttt{None}/\texttt{f}} (q0');
@@
     % Second automaton from the second figure
-    \node[state] (q0'') [right=4cm of q1'] {$q_0''$}; 
+    \node[state] (q0'') [right=3cm of q1'] {$q_0''$}; 
     \node[state] (q1'') [right=of q0''] {$q_1''$}; 
@@
   \end{tikzpicture}
-  \caption{The SFT for the replacement operation $\texttt{replace}(s, $"$[0-9]+$"$, $"$\text{ITP}$"$)$}
+  \caption{The SFT for the replacement operation $\texttt{replace}(s, $\texttt{/[0-9]+/}$, $\text{NUM}$)$}
   \label{fig-rearranged-automata}
@@
   With this SFT, we can compute the forward image of the replacement operation.
-  Let $s' = \texttt{replace}(s, $"$[0-9]+$"$, $"$\text{ITP}$"$)$ be a string constraint and we want to compute the domain of $s'$ regarding this replacement operation.
-  Assume the SFT generated from this replacement operation is $\mathcal{T}$ and the string $s$'s domain ranges over an SFA $\mathcal{A}$.
-The foward image of this replacement operation is $\mathcal{T}~\mathcal{A}$.
-Therefore, the domain of $s'$ is $\mathcal{T}~\mathcal{A}$.
+  Let $s' = \texttt{replace}(s,~$\texttt{/[0-9]+/}$,~\text{"}$\texttt{NUM}$\text{"})$ be a string constraint and we want to compute the value range of $s'$ regarding this replacement operation.
+  Assume the SFT generated from this replacement operation is $\mathcal{T}$ and the string variable $s$'s values ranges over an SFA $\mathcal{A}$.
+The foward image of this replacement operation is $\mathcal{T}\times \mathcal{A}$.
+It also means that the value of $s$' must ranges over this image.
+
+  \emph{Compare with the semantics of SMTLIB.} SMTLIB's replacement's semantics is not exactly the same as the semantics  in this paper.
+In SMTLIB, the single replacement operation is defined as replace the \emph{first} occurrence of the substring that matches the pattern with the replacement string.
+Our semantics is not limited to the \emph{first} occurrence.
+
+  From replacement operation modeling above, we can see that our SFT's formalization with the support of empty string for input labels and output functions make it easy to model the replacement operation.
+  However,
+  The power of our SFT's formalization is not fully explored in the string solver as the transition labels and output functions are simple.
+For other applications, the transition labels and output functions can indeed support more complex modelings.
 
@@
 
-We have implemented a prototype of the string solver to support the replacement operation based on CertiStr[1], which is called CertiStr-R in this paper.
+We have implemented a prototype of the string solver to support the replacement operation based on CertiStr \cite{cpp/KanLRS22} \footnote{The code is available at \url{https://github.com/ShlKan/certified-str-solver}}, which is called CertiStrR in this paper.
+We select a benchmark from the SMT-LIB \cite{smtlib_benchmarks} to test the effective and efficiency of CertiStrR. The benchmark contains 100 string constraints.
 
-CertiStr-R supports two kinds of replacement operations: (1) \texttt{str.replace s p r}, where the pattern \texttt{p} is a string and (2) \texttt{str.replace\_re s p r}, where the pattern \texttt{p} is a regular expression.
-
-The limitation of CertiStr-R is that it does not support \texttt{replace\_all} and \texttt{replace\_re\_all}.
-CertiStr-R is possible to extend in the future.
-Another limitation is that the semantics of \texttt{replace} and \texttt{replace\_re} are different from the definition in SMTLIB.
+CertiStr-R supports two kinds of SMT-LIB's replacement operations: (1) \texttt{str.replace s p r}, where the pattern \texttt{p} is a string and (2) \texttt{str.replace\_re s p r}, where the pattern \texttt{p} is a regular expression.
 

### 2025-02-24 Paper/main.tex
papers

@@
 \usetikzlibrary{positioning}
+\usepackage{booktabs}
 %\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
+\usepackage{listings}
+\usepackage{xcolor}
+
+\lstdefinelanguage{SMTLIB}{
+  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
+  sensitive=true,
+  morecomment=[l];,
+  morestring=[b]"
+}
+\lstset{
+  language=SMTLIB,
+  basicstyle=\ttfamily\small,
+  keywordstyle=\color{blue}\bfseries,
+  commentstyle=\color{green!50!black},
+  numbers=left,
+  numberstyle=\tiny\color{gray},
+  stepnumber=1,
+  numbersep=5pt,
+  backgroundcolor=\color{gray!10},
+  frame=single,
+  breaklines=true,
+  captionpos=b,
+  showstringspaces=false
+}
 
@@
 
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of intput words represented by a regular language, which can be denoted as an NFA. More precisely,
+To make the operation \emph{product} more general, we extend the operation to an SFT and a set of intput words represented by a regular language, which can be denoted as an NFA $\mathcal{A}$.
+More precisely,
 \[
-\mathcal{T}\times\{\mathcal{A}\} = \{w'\mid w.~w\in \mathcal{L}(\mathcal{A})\land w' = \mathcal{T}\times\{w\} \}, \text{where $\mathcal{L}(\mathcal{A})$ denotes the language of $\mathcal{A}$}.
+\mathcal{T}\times \mathcal{A} = \{w'\mid w.~w\in \mathcal{L}(\mathcal{A})\land w' = \mathcal{T}\times\{w\} \}, \text{where $\mathcal{L}(\mathcal{A})$ denotes the language of $\mathcal{A}$}.
 \]
@@
 We now illustrate SFTs with an example in Figure \ref{fig-example-ft}. It accepts strings of the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$. 
-It uses intervals $[i-j]$ as the input label, which is a predicate that returns true if the input letter is between $i$ and $j$.
+It uses intervals $[i-j]$ as the input label, which is a predicate as $[i-j] = \lambda x.~ i \leq x \leq j$.
 The function $f_1=\lambda x.~ \texttt{toUpper}(x)$ maps each lowercase letter to uppercase letter, and $f_2=\lambda x.~\texttt{toLower}(x)$ maps each uppercase letter to lowercase letter.
@@
 
+The transition labels can be more complex.
+For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$.
+This symbolic label corresponds to the set of even natural numbers.
+
 \subsection{The Isabelle/HOL Formalization of SFTs}
+
+We have seen the diversity of the transition labels in the examples.
+To make the transition labels more flexible, our formalization of SFTs should be extensible.
+We will illustrate how can we leverage the refinement framework in Isabelle/HOL to achieve this goal.
 
@@
 We define a canonical form for intervals to facilitate the proof. An interval $[i_1, j_1), \ldots, (i_n, j_n)]$ is in canonical form if 
-(1) $i_k \leq j_k, 1 \leq k \leq n$,
-(2) $j_k\leq i_{k+1}, 1 \leq k < n$.
+(1) $i_k \leq j_k, \text{ for } 1 \leq k \leq n$,
+(2) $j_k\leq i_{k+1}, \text{ for } 1 \leq k < n$.
 All the interval operations are proved to keep the canonical form after the operation if the input intervals are in canonical form.
+Canonical form not only simplifies the proof, but also improves the performance of interval operations.
 
@@
 We elaborate only on the function \texttt{subtrans\_comp} in Fig. \ref{fig-def-subtrans_comp}. The other function is similar. 
-This function traverses the transitions SFA's transitions $\texttt{T}$. If it find the a transtion $(q1, \alpha', q1')\in \texttt{T}$ satisfies the following condition:
+This function traverses the transitions SFA's transitions $\texttt{T}$. If it find a transtion $(q_1, \alpha', q_1')\in \texttt{T}$ satisfies the following condition:
 $\alpha \cap \alpha' \neq \emptyset$ (Checked by \texttt{nemptyIs}),
@@
 We have implemented a prototype of the string solver to support the replacement operation based on CertiStr \cite{cpp/KanLRS22} \footnote{The code is available at \url{https://github.com/ShlKan/certified-str-solver}}, which is called CertiStrR in this paper.
-We select a benchmark from the SMT-LIB \cite{smtlib_benchmarks} to test the effective and efficiency of CertiStrR. The benchmark contains 100 string constraints.
+%
+CertiStr-R supports two kinds of SMT-LIB's replacement operations: (1) \texttt{str.replace s p r}, where the pattern \texttt{p} is a string and (2) \texttt{str.replace\_re s p r}, where the pattern \texttt{p} is a regular expression.
+Listing \ref{lst-smtlib-code} shows an example of the SMT-LIB code corresponding to the replacement operation $\texttt{replace}(s,~$\texttt{/[0-9]+/}$,~\text{"}$\texttt{NUM}$\text{"})$ that we illustrated in Section \ref{sec-app-str-solver}.
+The constants are satisfiable because the
+string \texttt{a} can be fully matched by the regular expression \texttt{/[0-9]+/} and then replaced by "\texttt{NUM}".
+\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}]
+  (set-logic QF_S)
+  (declare-fun a () String)
+  (declare-fun b () String)
+  (assert (= a "2025"))
+  (assert (= b "NUM"))
+  (assert (= b (str.replace_re a (re.+ (re.range "0" "9")) "NUM")))
+  (check-sat)
+  \end{lstlisting}
 
-CertiStr-R supports two kinds of SMT-LIB's replacement operations: (1) \texttt{str.replace s p r}, where the pattern \texttt{p} is a string and (2) \texttt{str.replace\_re s p r}, where the pattern \texttt{p} is a regular expression.
+We select benchmarks from the SMT-LIB 2024 \cite{smtlib_benchmarks} to test the effective and efficiency of CertiStrR.
+The benchmarks contain mainly from QF\_S and QF\_SLIA.
+The benchmarks are devided into 2 groups: (1) tests with \texttt{str.replace} and (2) tests with \texttt{str.replace\_re}.
+Because CertiStrR currently has a limited front-end to support all SMT-LIB operations.
+We remove the operations that are not supported by CertiStrR, for example, the operation \texttt{or}.
+Totally the benchmark contains 300 tests.
+
+\begin{table}[h]
+  \centering
+  \begin{tabular}{lcccc}
+      \toprule
+      & \textbf{Result} & \textbf{Time} & \textbf{Memory} & \textbf{Number of Tests} \\
+      \midrule
+      \texttt{str.replace} & & & & \\
+      \texttt{str.replace\_re} & & & & \\
+      \bottomrule
+  \end{tabular}
+  \caption{Performance metrics for string operations}
+  \label{tab:string_operations}
+\end{table}
+
+Among all these tests, the test that consumes the most time is 2s.
+
+TODO: 
+(1) The complicated model.
+(2) which test consume the most time.
+(3) which test consume the least time.
+
+\subsection{Effort of Certified Development}
 

### 2025-02-25 Paper/main.tex
add more benchmarks and modify paper

@@
 The constants are satisfiable because the
-string \texttt{a} can be fully matched by the regular expression \texttt{/[0-9]+/} and then replaced by "\texttt{NUM}".
+substring "\texttt{2025}" in \texttt{a} can be matched by the regular expression \texttt{/[0-9]+/} and then replaced by "\texttt{NUM}".
 \begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}]
@@
   (declare-fun b () String)
-  (assert (= a "2025"))
-  (assert (= b "NUM"))
+  (assert (= a "2024,2025"))
+  (assert (= b "2024,NUM"))
   (assert (= b (str.replace_re a (re.+ (re.range "0" "9")) "NUM")))
@@
 Because CertiStrR currently has a limited front-end to support all SMT-LIB operations.
-We remove the operations that are not supported by CertiStrR, for example, the operation \texttt{or}.
-Totally the benchmark contains 300 tests.
+We rewrite the operations that are not supported by CertiStrR, for example, \texttt{(assert (and c1 c2))} is rewritten as two separate \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints.
 

### 2025-02-27 Paper/main.tex
rewrite benchmarks and modify papers

@@
 We have implemented a prototype of the string solver to support the replacement operation based on CertiStr \cite{cpp/KanLRS22} \footnote{The code is available at \url{https://github.com/ShlKan/certified-str-solver}}, which is called CertiStrR in this paper.
+The frontend of CertiStrR is not certified and implemented with the dolmen \cite{dolmen} library (An OCaml library of SMTLIB parser) and ocaml-re-nfa \cite{ocaml-re-nfa} library (An OCaml library to construct an NFA from a regular expression).
+
 %
@@
 Listing \ref{lst-smtlib-code} shows an example of the SMT-LIB code corresponding to the replacement operation $\texttt{replace}(s,~$\texttt{/[0-9]+/}$,~\text{"}$\texttt{NUM}$\text{"})$ that we illustrated in Section \ref{sec-app-str-solver}.
-The constants are satisfiable because the
+The constraints are satisfiable because the
 substring "\texttt{2025}" in \texttt{a} can be matched by the regular expression \texttt{/[0-9]+/} and then replaced by "\texttt{NUM}".
+
 \begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}]
@@
 
+  As we discussed in Section \ref{sec-app-str-solver}, our replacement semantics is not exactly the same as SMTLIB's standard semantics.
+If we run the SMT solver CVC5 \cite{cvc5} with the example code in Listing \ref{lst-smtlib-code}, the result is UNSAT. Because CVC5 only matches the first occurrence of the pattern.
+
 We select benchmarks from the SMT-LIB 2024 \cite{smtlib_benchmarks} to test the effective and efficiency of CertiStrR.
@@
 Because CertiStrR currently has a limited front-end to support all SMT-LIB operations.
-We rewrite the operations that are not supported by CertiStrR, for example, \texttt{(assert (and c1 c2))} is rewritten as two separate \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints.
+We rewrite the operations that are not supported by CertiStrR, for example, \texttt{(assert (and c1 c2))} is rewritten as two separate \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints that CertiStrR supports.
+
+% replace_re
+% Average real time: 0.37 seconds
+% Total number of real entries: 98
+% Average user time: 0.35 seconds
+% Total number of user entries: 98
+% Average sys time: 0.01 seconds
+% Total number of sys entries: 98
+% Total number of SAT entries: 61
+% Total number of UNSAT entries: 4
+% Total number of inconclusive entries: 30
+
+% replace_str   
+% Average real time: 0.27 seconds
+% Total number of real entries: 323
+% Average user time: 0.25 seconds
+% Total number of user entries: 323
+% Average sys time: 0.01 seconds
+% Total number of sys entries: 323
+% Total number of SAT entries: 315
+% Total number of UNSAT entries: 173
+% Total number of inconclusive entries: 8
 
@@
   \centering
-  \begin{tabular}{lcccc}
+  \begin{tabular}{lccccc}
       \toprule
-      & \textbf{Result} & \textbf{Time} & \textbf{Memory} & \textbf{Number of Tests} \\
+      & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inconclusive} & \textbf{Time} & \textbf{Number of Tests} \\
       \midrule
-      \texttt{str.replace} & & & & \\
-      \texttt{str.replace\_re} & & & & \\
+      \texttt{replace\_str} & 142 & 173 & 8 & 0.27 & 323\\
+      \texttt{replace\_re} & 57 & 4 & 30 & 0.37 & 98\\
       \bottomrule
@@
 
-Among all these tests, the test that consumes the most time is 2s.
+The experiments were conducted on a laptop equipped with an Apple M4 processor, featuring 24 GB of memory.
+The timelimit for each test is $1$ minutes.
+We split the tests into 2 groups: (1) tests with \texttt{str.replace} (\texttt{replace\_str}, 323 tests) and (2) tests with \texttt{str.replace\_re} (\texttt{replace\_re}, 98 tests).
+The average time of \texttt{replace\_str} is 0.27 seconds and the average time of \texttt{replace\_re} is 0.37 seconds.
+The test results are classified into 3 categories: (1) SAT, (2) UNSAT, and (3) Inconclusive.
+The result inconclusive means that the solver cannot prove the result.
+This does not mean the the test runs out of time.
+It means the solver cannot prove the result with the current forward-propagation algorithm used by CertiStr \cite{cpp/KanLRS22}.
+Forward-propagation algorithm cannot decide the result if the constraints do not satisfy tree-property, even after the forward-propagation all variables's domain are not empty as illustrated in CertiStr \cite{cpp/KanLRS22}.
 
-TODO: 
-(1) The complicated model.
-(2) which test consume the most time.
-(3) which test consume the least time.
+After analyzing the tests that consume more time than others.
+We found the replacement operation modeling itself is efficient.
+The time consumption of is mainly from the automata accumulation during the forward-propagation algorithm.
+For instance, given the constraints as follows:
+\[
+x=x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r);~x = x_4;~x = \texttt{replace}(x, p_1, r_1);
+\]
+where \texttt{++} is the concatenation operation.
+As the varible $x$ appears many times on the left-hand side of the constraints, the forward-propagation algorithm will accumulate the automata of $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace}($x, p_1, r_1$) to $x$ during the forward-propagation.
+Each accumulation will perform the product operation with the previous automata.
+The complexity of the product operation is $O(n^2)$ in the worst case, where $n$ is the size of the automata.
+This leads to state explosion.
 
 \subsection{Effort of Certified Development}
+
+We discuss the effort of certified development in this subsection.
+The effort of certified development is mainly from the following two aspects: (1) the abstract level Formalization of SFTs and (2) the low-level refinement of the SFTs to efficient data structure.
 

### 2025-02-27 CertiStr/test/benchmark/benchmark.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-for test in ./*.smt2; do
-    echo "Test ($test):"
-    result=$(CertiStr $test)
-    echo "$result"
-done

### 2025-02-27 CertiStr/test/benchmark/replace_re/benchmark.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    echo "Test ($test):"
+    result=$(dune exec CertiStr -- $test);
+    echo "$result"
+done

### 2025-02-28 Paper/main.tex
remove old tests

@@
 \begin{itemize}
-\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{path},
-\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the path and $(f_0~a_0), \ldots, (f_{n-1}~a_{n-1})$ the \emph{output} of the path.
+\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
+\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, (f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
 \end{itemize}
 
-If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the path is an accepting path of $\mathcal{T}$.
-For a path $\pi$, we denote the input of $\pi$ as $in(\pi)$ and the output of $\pi$ as $out(\pi)$.
+If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
+For a trace $\pi$, we denote the input of $\pi$ as $in(\pi)$ and the output of $\pi$ as $out(\pi)$.
 %
@@
 \[
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting path }\land in(\pi) = w \land out(\pi) = w'\}.
+\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace }\land in(\pi) = w \land out(\pi) = w'\}.
 \]

### 2025-03-05 Paper/main.tex
polish the paper to product operation

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+\documentclass[a4paper,UKenglish,cleveref, autoref, anonymous, thm-restate]{lipics-v2021}
 %This is a template for producing LIPIcs articles. 
@@
 \begin{abstract}
-Finite-State Automata (FAs) and Finite-State Transducers (FTs) are integral components in the domains of programming languages and software engineering.
-Regular expressions, which are often implemented using FAs, play a pivotal role in languages such as JavaScript and Python.
-FTs extend the capabilities of FAs by allowing transformations of input strings into output strings, thus providing a more expressive framework for operations that involve both recognition and transformation.
-Despite the popularity of formalizing FAs and FTs within proof assistants like Coq and Isabelle/HOL, these formalizations frequently fall short in terms of applicability to real-world scenarios.
+Finite-state Automata (FAs) are fundamental components in the domains of programming languages.
+For instance, regular expressions, which are pivotal in languages such as JavaScript and Python, are frequently implemented using FAs.
+%
+Finite-state Transducers (FTs) extend the capabilities of FAs by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
+Despite the widespread adoption of formalizing FAs and FTs within proof assistants such as Coq and Isabelle/HOL, these formalizations often fall short in terms of applicability to real-world scenarios.
+A significant limitation of classical finite-state models is that transition labels are typically confined to a single character from a finite alphabet.
+However, in practical applications, the alphabet of an FA or FT can be extensive or even \emph{infinite}.
+This classical approach to formalizing transitions can result in transition explosion, leading to critical performance bottlenecks of FA operations.
 
-One limitation of traditional finite-state models is that transition labels are typically restricted to single characters from a finite alphabet.
-For example, a transition is often represented as $q \xrightarrow{a} q'$, where $a$ is a single character.
-However, in real-world applications, the alphabet of a FA or FT can be vast or even \emph{infinite}.
-This conventional approach to defining transitions can lead to a phenomenon known as transition explosion, where the number of transitions becomes unmanageably large.
-
-A more pragmatic approach involves the formalization of symbolic FAs [1] and FTs [2], wherein transition labels are symbolic and potentially infinite.
-While the formalization of symbolic FAs has been comprehensively explored in the work of \cite{cpp/KanLRS22}, the formalization of symbolic FTs remains largely unexplored.
-This gap presents additional challenges in establishing the correctness and applicability of such formalizations.
-
-In this paper, we endeavor to address this gap by presenting a formalization of symbolic transducers within the Isabelle/HOL framework.
+A more pragmatic approach involves the formalization of symbolic FAs \cite{cav/DAntoniV17} and FTs \cite{VeanesHLMB12Transducer}, where transition labels are symbolic and potentially infinite.
+While the formalization of symbolic FAs has been explored in the work of CertiStr \cite{cpp/KanLRS22}, the formalization of symbolic FTs in interactive proof assistants remains largely unexplored due to the increased complexity challenges.
+%
+In this paper, we aim to formalize symbolic FTs within the Isabelle/HOL framework.
 This formalization is refinement-based and is designed to be extensible with various symbolic representations of transition labels.
-To evaluate its effectiveness and efficiency, we applied this formalization to an SMT string solver for modeling replacement operations.
-The experimental results demonstrate that the formalized symbolic transducer can efficiently solve string constraints.
+To assess its performance, we applied the formalized symbolic FTs to an SMT string solver for modeling replacement operations.
+The experimental results indicate that the formalized symbolic transducer can efficiently solve string constraints.
 
@@
 
-Automata and Transducers are crucial concepts in formal languages and have widely applications in programming languages and software engineering.
-For instance, [1] has shown the correspondence between  regular expressions in modern regular expressions and variants of FAs and FTs.
-Other industrial usage of FAs and FTs is AWS access control polices checking.
+Finite-state Automata (FA) and Finite-state Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
+For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
+Additionally, FAs and FTs find significant industrial applications, such as in the verification of AWS access control policies \cite{DBLP:conf/fmcad/BackesBCDGLRTV18}.
 
-Even though there various  formalization of FAs and FTs in Coq, Isabelle, and other proof assistants.
-They are mainly based on classic definitions of FAs and FTs.
-There some drawbacks when come to practical applications.
-(1) transition labels are
-non-symbolic and usually finite.
-A classic and normal definition of a transition is $q\xrightarrow{a}q'$, where $a$ is a character in an finite alphabet.  This simple way will yield transition explosions. 
-For instance, if the alphabet $\Sigma$ is of the size $10,000$, then a transition from $q$ to $q'$ that accept any characters in $\Sigma$ need to split into $10,000$ transitions.
-Automata product in this way will be very inefficient.
-Moreover,
-The alphabet are usually finite. For practical applications, infinite alphabet are often necessary. For instance, if the alphabet is all integers.
+Even though there are various formalizations of FAs and FTs in interactive proof assistants such as Isabelle \cite{isabelle-homepage} and Coq \cite{coq-homepage}, these are predominantly based on classical definitions.
+However, these traditional approaches present certain limitations when applied to practical scenarios.
+One significant drawback is that transition labels are typically non-symbolic and finite.
+A conventional transition is represented as $q\xrightarrow{a}q'$, where $a$ is a character from a finite alphabet.
+This simplistic representation can lead to a phenomenon known as transition explosion.
+For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x10FFFF}$ distinct characters.
+Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x10FFFF}$ individual transitions, rendering the product of two FAs highly inefficient.
+Furthermore, in practical applications, it is often necessary to consider infinite alphabets, such as the set of all integers.
 
-Symbolic FAs and FTs [2,3,4] are extensions of classic FAs and FTs that make their application more practical.
-The transition labels are represented by algebra.
-For instances, intervals ($'a'-'z'$), boolean algebras ($x \% 2 == 0$), or others.
-This symbolic way is more succinct and its support for infinite alphabet extend the expressive power of FAs and FTs.
+Symbolic Finite Automata (SFA) and Symbolic Finite-state Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic Finite Automata (FAs) and Finite-state Transducers (FTs), enhancing their applicability in practical scenarios.
+These symbolic models utilize transition labels defined by boolean algebras, allowing for more expressive representations.
+For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \% 2 = 0$), denoting all even numbers.
+This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
 
-Formalizing FAs and FTs in Isabelle/HOL, Coq, and other proof assists are more challenging compare with formalizing classic ones.
-Moreover, how to make the formalization extensible is also an important point to think, because, symbolic FAs and FTs support different algebras symbolic representations.
-For new algebra representations, we do not want repeat some proof works.
+The formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
+Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
 
-Fortunately, symbolic FAs has been formalized in Isabelle/HOL [certistr] and experiments in [certistr] illustrates the efficiency and effective of symbolic FAs.
-Unfortunately FTs are not formalized yet.
-FTs are more powerful and expressive than symbolic FAs.
-For instance, when FTs are support, replacement operation in modern programming languages, such as Javascript, python and others can be modelled as FTs.
-And CertiStr can be extended to support replacement operations.
+Prior work has successfully formalized symbolic FAs within Isabelle/HOL through CertiStr \cite{cpp/KanLRS22}, demonstrating both the efficiency and effectiveness of symbolic FAs in practice.
+However, the formalization of FTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
+FTs constitute a significantly more expressive and powerful theoretical framework compared to symbolic FAs, as evidenced by their capability to model complex string transformations such as replacement operations.
+Furthermore, the seminal work of \cite{VeanesHLMB12Transducer} demonstrates the broad applicability of SFTs across diverse domains, including security-critical applications such as cross-site scripting (XSS) prevention, image transformation operations, and privacy-preserving location data processing.
 
-In this paper, we formalize symbolic FTs based on symbolic FAs.
-In order to solve the potential extension problem of FTs to support additional transition label theories, the formalization is refinement-based, in which at the abstraction level, transition labels are modeled as a general concept: sets.
-We can project any theory to sets.
-\emph{extend this}.
-The most important operation that we formalized and proved is the application operation on an input language.
-More precisely, given an FA $\mathcal{A}$ (to represent a regular language) as the input language and a FT $\mathcal{T}$, the application operation $\mathcal{F}~\mathcal{A}$ is the  of output languages of FT after receiving the input language
+In this work, we present a comprehensive formalization of symbolic FTs that builds upon the formalization of symbolic FAs.
+To address the extensibility challenges inherent in supporting diverse transition label theories, we adopt a refinement-based approach.
+At the abstraction level, transition labels are formalized through the fundamental mathematical concept of \emph{sets}.
+This abstraction facilitates subsequent refinement to various boolean algebras, including intervals and arithmetic conditions, while maintaining theoretical consistency.
 
-In the next refinement level, the states and transitions are stored in efficient data structure based on Isabelle/HOL refinement framework [2].
-For transition labels, we implemented an interval algebra, which are efficient for creating, checking, interaction.
-%
-In the future the refinement of transition label from sets to other algebras instead of intervals can be done easily by implementing the common interfaces with interval algebras.
+The key operation of our formalization is the product operation between an SFT and an input regular language.
+Specifically, given an SFA $\mathcal{A}$ representing a regular language and an SFT $\mathcal{T}$, we define the product operation $\mathcal{T} \times\mathcal{A}$ that characterizes the set of output languages generated by $\mathcal{T}$ when processing inputs from the language recognized by $\mathcal{A}$.
 
-We highlight our formulization with the following points.
-(1) Efficiency, the algebraic representation
-support compressed storage of transition labels and efficiency checking
-of membership.
-(2) Extensible, the refinement based formulization allows the extension
-of the SNFA to support different theories of transition labels.
+Leveraging the data refinement framework \cite{DBLP:conf/itp/Lammich13} formalized in Isabelle/HOL, we implement an efficient representation of states and transitions using sophisticated data structures such as hashmaps and red-black trees.
+Furthermore, we develop a comprehensive interval algebra that provides efficient operations for creation, verification, and manipulation of intervals, thereby facilitating the implementation of replacement operations within our string solver framework.
+
+In summary, Our formalization makes two primary contributions.
+First, we achieve computational efficiency through an algebraic representation that enables compressed storage of transition labels.
+Second, our refinement-based formalization framework demonstrates significant extensibility, facilitating the incorporation of diverse theories of transition labels.
 
 The paper is organized as follows.
-Section 2 presents the formalization of symbolic FTs.
-Section 3 presents the Core operations of symbolic FTs.
+Section \ref{sec:formalization} presents the formalization of symbolic FTs.
+Section \ref{sec:product-operation} presents the product operation of symbolic FTs in the abstraction level.
+Section \ref{sec_alg_refinement} presents the refinement of the product operation in Isabelle/HOL.
+Section \ref{sec-app-str-solver} presents the application of SFTs in the modeling of replacement operations in string solvers.
+Section \ref{sec:related-work} presents the related work.
+Section \ref{sec:conclusion} concludes the paper.
 
 \section{Formalization of Symbolic FTs}
+\label{sec:formalization}
 
-We first present a pure mathematical form of SFTs \cite{VeanesHLMB12Transducer}, which is independent of Isabelle's formalization. 
-The central concept is the label theory, which characterizes the process by which output labels are derived from input labels.
+We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL.
+The foundational concept underlying this formalization is the label theory, which characterizes the process by which output labels are derived from input labels.
 
@@
 We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
-We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type. Assume $\mathbb{N}$ is the set of elements of the sort natural numbers.
+We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type.
 
-A lambda term is defined as $\lambda x. t$ of the type $\tau_1 \rightarrow \tau_2$.
+A lambda term is defined as $\lambda x.~t$ of type $\tau_1 \rightarrow \tau_2$.
 When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate.
-Assume $\psi$ is a predicate.
-We write $a\in \llbracket\psi \rrbracket$ if $\psi~a=\top$.
-For the non-predicate lambda term, we view it as a function, which generates an output element of type $\tau_2$ with the input term of type $\tau_1$.
+Let $\phi$ be a predicate.
+We write $a\in \llbracket\phi \rrbracket$ if $\phi~a=\top$.
+For non-predicate lambda terms, we view them as functions that generate output elements of type $\tau_2$ given input terms of type $\tau_1$.
 With these notations and the above definitions, we can define SFTs as follows.
@@
    \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \psi, f, q') \text{ or}$ $\text{written in }q\xrightarrow{\psi, f} q'$, where $q$ and $q'$ are the states in $\mathcal{Q}$.
-   $\psi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of the type $\tau_1\rightarrow \tau_2$. $f$ is called "output function".
+Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
+   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
+   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
    \end{itemize}
@@
 SFTs accept an input word and generate an output. This can be defined by \emph{runs} of SFTs.
-An SFT run $\sigma$ is a sequence $(q_0, \psi_0, f_0, q_1),(q_1, \psi_1, f_1, q_2),\ldots, (q_{n-1}, \psi_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i, 
-\psi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
+An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i, 
+\phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
 $\sigma$ is an accepting run when $q_n$ is an accepting state.
 
-For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ iff $a_i \in \llbracket\psi_i \rrbracket$ for $0 \leq i \leq n - 1$.
-Run $\sigma$ will generate an output sequence if $w$ is accepted by it, which is $w'= (f_0~a_0), \ldots (f_{n_1}~a_{n_1})$.
-We call in the sequent:
+For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ if and only if $a_i \in \llbracket\phi_i \rrbracket$ for $0 \leq i \leq n - 1$.
+When $w$ is accepted, run $\sigma$ generates an output sequence $w'= f_0~a_0, \ldots, f_{n-1}~a_{n-1}$.
+We define:
 \begin{itemize}
@@
 If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
-For a trace $\pi$, we denote the input of $\pi$ as $in(\pi)$ and the output of $\pi$ as $out(\pi)$.
+For a trace $\pi$, we denote its input as $\mathit{in}(\pi)$ and its output as $\mathit{out}(\pi)$.
 %
-Given an SFT $\mathcal{T}$ and a word $w$, we define the operation product of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with the input $w$.
+Given an SFT $\mathcal{T}$ and a word $w$, we define the \emph{product} operation of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with input $w$.
 More precisely,
 \[
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace }\land in(\pi) = w \land out(\pi) = w'\}.
+\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land in(\pi) = w \land out(\pi) = w'\}.
 \]
 
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of intput words represented by a regular language, which can be denoted as an NFA $\mathcal{A}$.
+To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an NFA $\mathcal{A}$.
 More precisely,
 \[
-\mathcal{T}\times \mathcal{A} = \{w'\mid w.~w\in \mathcal{L}(\mathcal{A})\land w' = \mathcal{T}\times\{w\} \}, \text{where $\mathcal{L}(\mathcal{A})$ denotes the language of $\mathcal{A}$}.
+\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
 \]
@@
  \end{tikzpicture}
-    \caption{The formalization of $FTs$ in Isabelle/HOL}
+    \caption{An example of SFT}
     \label{fig-example-ft}
     \end{figure}
-We now illustrate SFTs with an example in Figure \ref{fig-example-ft}. It accepts strings of the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$. 
-It uses intervals $[i-j]$ as the input label, which is a predicate as $[i-j] = \lambda x.~ i \leq x \leq j$.
-The function $f_1=\lambda x.~ \texttt{toUpper}(x)$ maps each lowercase letter to uppercase letter, and $f_2=\lambda x.~\texttt{toLower}(x)$ maps each uppercase letter to lowercase letter.
-Give an string "$\texttt{bigSMALL}$", the SFT will generate an output "$\texttt{BIGsmall}$".
 
-The transition labels can be more complex.
+    Figure \ref{fig-example-ft} illustrates an SFT that accepts strings matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
+The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
+The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation.
+For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.
+
+%The transition labels can be more complex.
 For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$.
@@
 
-We have seen the diversity of the transition labels in the examples.
-To make the transition labels more flexible, our formalization of SFTs should be extensible.
-We will illustrate how can we leverage the refinement framework in Isabelle/HOL to achieve this goal.
+As we have discussed the diversity of transition labels in Section \ref{sec:introduction}, we now present an extensible formalization of SFTs that accommodates this variety.
+Our approach leverages the refinement framework (Refine\_Monadic \cite{Refine_Monadic-AFP}) in Isabelle/HOL to achieve the flexibility and extensibility of transition labels modeling.
 
@@
 record (|$'q,$||$~'a,$| |$'b$|) |$\mathbf{NFT}$| =
-	|$\mathcal{Q}_T$| :: "|$'q$| set"
-	|$\Delta_T$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
-	|$\mathcal{I}_T$| :: "|$'q$| set"
-	|$\mathcal{F}_T$| :: "|$'q$| set"
+	|$\mathcal{Q}$| :: "|$'q$| set"
+	|$\Delta$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
+	|$\mathcal{I}$| :: "|$'q$| set"
+	|$\mathcal{F}$| :: "|$'q$| set"
         |$\mathcal{M}$| :: "|$'b\Rightarrow 'a$| Tlabel"
@@
 
-type_synonym |$'a$| Tlabel = "|$'a \text{ option} \Rightarrow 'a$ set"
+type_synonym |$'a$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
 	\end{lstlisting}
@@
 
-Figure \ref{fig-def-FT} shows our formalization of SFTs in Isabelle / HOL.
-The elements $\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$ are intuitive when referring to Definition \ref{def-sft}.
-However, the transitions are not exactly the same.
+Figure \ref{fig-def-FT} presents our formalization of SFTs in Isabelle/HOL.
+While the elements $\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$ directly correspond to their counterparts in Definition \ref{def-sft}, the transition relations are not exactly the same.
 %
-The set of transitions is defined by \texttt{LTTS} ( which is short for Labeled Transducer Transition System).
-Each transition is defined as $'q \times ('a \text{ set option }\times b) \times 'q$.
-There are some design decisions, which are made to make the FT formalization more abstract and flexible.
+The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \text{ set option }\times b) \times 'q$.
+This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
 
-Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or \texttt{None} regarding the semantics of the type \texttt{option}. 
+Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
 Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
@@
 But predicates have various different forms.
-For instance, the interval $[1, 9]$ is a predicate that contains all elements $e$ between $1$ and $9$, i.e., $1 \leq e \leq 9$.
-The predicate $\lambda x : \mathbb{N}.~ x \text{ mod } 2 = 0$ denotes the set of even natural numbers.
-All these different forms of labels are abstracted as sets in the formalization.
+For instance, the interval $[1, 9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
+The predicate $\lambda x \colon \mathbb{N}.~ x \text{ mod } 2 = 0$ denotes the set of even natural numbers.
+All these different forms of labels are abstracted as sets in our formalization.
+%
+The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements.
+This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
 
-%
-The value $\texttt{None}$ is used to denote $\varepsilon$-move, that is, it accepts nothing but may still output some elements.
-This design makes it easier model real world applications in FTs.
-We will show an example after introducing the output functions.
+The second element of type $'b$ serves as an index into the output function space.
+The mapping $\mathcal{M}$ associates each index with a specific output function.
+These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
+This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria.
+Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
 
-The second element of type $'b$ is the \emph{indexes} of the output functions.
-$\mathcal{M}$ maps each of these indexes to an output function.
-The output function, defined by \texttt{Tlabel} maps a single input element to a set of output elements instead of single one.
-The reason is to allow FTs output non-deterministically, which means that the output may randomly or according to some criteria select one of the element in the set to output.
-
-Now let us look at an example.
+%Now let us look at an example.
 In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
 This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}).
-We can easily model this operation using our formalization of NFTs with the following steps:
-\begin{itemize}
-\item Step 1, construct a corresponding NFA from \texttt{pattern}).
-\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
-\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
-\item Step 4, unlabel previous accepting states as non-accepting states. 
-\end{itemize}
+%We can easily model this operation using our formalization of NFTs with the following steps:
+%\begin{itemize}
+%\item Step 1, construct a corresponding NFA from \texttt{pattern}).
+%\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
+%\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
+%\item Step 4, unlabel previous accepting states as non-accepting states. 
+%\end{itemize}
 
-We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
+%We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
 
 \section{The Product Operation of NFTs}
-\label{sec-form-fts}
+\label{sec:product-operation}
 
-Given an NFT $\mathcal{N}$ and a word $w$, it is crucial and interesting that the possible output words can be calculated.
-This is the operation, we want to formalize, but we will generalize this operation.
-
-Instead of calculating the outputs for a single word, the operation calculates the outputs for a set of input words. Single word input is a special case with the set contains exactly only one element.
-
-The set of input words is modeled as a regular language.
-To be more precise, let $\mathcal{N}$ be an NFT, and $\mathcal{A}$ be an NFA.
-We define the operation $\texttt{PRODUCT} ~\mathcal{N}~\mathcal{A} = \{w'\mid w\in \mathcal{L}(\mathcal{A})\text{ and } w' = \mathcal{N}~\mathcal{A}\}$, where $\mathcal{L}(\mathcal{A})$ denotes the language of NFA $\mathcal{A}$.
-
-But there is one thing we need to consider, as our NFT formalization accepts $\varepsilon$-transitions, the output NFAs may contain $\varepsilon$-transitions as well.
-Unfortunately, [CertiStr] does not formalize $\varepsilon-$NFA.
-We formalize symbolic $\varepsilon-$NFA and the conversion from $\varepsilon-$NFA to NFA.
-We only present the definitions of $\varepsilon-$NFA and NFA in this paper.
-The more details including correctness proofs can be found in the Isabelle code.
+In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an NFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
+An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
+Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of automata with $\varepsilon$-transitions ($\varepsilon$NFAs), we extend the framework with a formalization of symbolic $\varepsilon$NFAs (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
+In this paper, we present the core definitions of both $\varepsilon$NFAs and NFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
 
@@
 	\begin{lstlisting}
-record (|$'q,$||$~'a$|) |$\mathbf{NFA}$| =
+record (|$'q,$||$~'a$|) |$\texttt{NFA}$| =
 	|$\mathcal{Q}$| :: "|$'q$| set"
@@
 
-record (|$'q,$||$~'a$|) |$\mathbf{eNFA}$| =
-	|$\mathcal{Q}_e$| :: "|$'q$| set"
-	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
-        |$\Delta_e'$| :: "|$('q * 'q)$| set"
-	|$\mathcal{I}_e$| :: "|$'q$| set"
-	|$\mathcal{F}_e$| :: "|$'q$| set"
+record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
+	|$\mathcal{Q}$| :: "|$'q$| set"
+	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
+	|$\Delta'$| :: "|$('q * 'q)$| set"
+	|$\mathcal{I}$| :: "|$'q$| set"
+	|$\mathcal{F}$| :: "|$'q$| set"
 
@@
 	\end{lstlisting}
-\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\caption{The formalization of $\varepsilon$NFA and SFA in Isabelle/HOL}
 \label{fig-def-FAs}
@@
 
-Figure \ref{fig-def-FAs} shows the definitions of NFA and $\varepsilon-$NFA.
-The record type \texttt{eNFA} corresponds to $\varepsilon-$NFA.
-$\Delta_e$ is same as $\Delta$ in \texttt{NFA}.
-$\varepsilon$-transitions are defined by $\Delta_e'$.
-Now we are ready to present the operation \emph{Product}.
-Let $\mathcal{T}$ be an SFT and  $\mathcal{A}$ be an NFA.
-The output language generated by $\mathcal{T}$ with the input language $\mathcal{A}$ is an an \texttt{eNFA} $\mathcal{A}'$.
+Figure \ref{fig-def-FAs} presents the formalization of both SFAs and $\varepsilon$SFAs using Isabelle/HOL record types \texttt{NFA} and \texttt{eNFA}, respectively.
+The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
+
+Having established these foundational definitions, we can now formalize the product operation.
 
@@
 
+\begin{table}[h]
+  \centering
+  \begin{tabular}{lccc}
+      \toprule
+      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs} \\
+      \midrule
+      Abstract & & & \\
+      Implementation & & & \\
+      \bottomrule
+  \end{tabular}
+  \caption{Overview of the effort of certified development}
+  \label{tab:abstract_impl}
+\end{table}
+
 \section{Related Work}
+\label{sec:related-work}
 
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2025-03-06 Paper/main.tex
paper polish to page 9

@@
 	keywordstyle=\color{blue}\bfseries,
-	keywords={definition,if,then,else, where, record, fun, lemma, type_synonym, fixes, assumes, shows},
+	keywords={definition,if,then,else, where, record, fun, lemma, type_synonym, fixes, shows, assumes, shows},
 	escapeinside={|}{|}
@@
 	\begin{lstlisting}
-record (|$'q,$||$~'a,$| |$'b$|) |$\mathbf{NFT}$| =
-	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
-	|$\mathcal{I}$| :: "|$'q$| set"
-	|$\mathcal{F}$| :: "|$'q$| set"
-        |$\mathcal{M}$| :: "|$'b\Rightarrow 'a$| Tlabel"
+record (|$'q,$||$~'a,$| |$'b$|) |$\texttt{NFT}$| =
+	|$\mathcal{Q}_t$| :: "|$'q$| set"
+	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
+	|$\mathcal{I}_t$| :: "|$'q$| set"
+	|$\mathcal{F}_t$| :: "|$'q$| set"
+        |$\mathcal{M}_t$| :: "|$'b\Rightarrow 'a$| Tlabel"
 
@@
 Figure \ref{fig-def-FT} presents our formalization of SFTs in Isabelle/HOL.
-While the elements $\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$ directly correspond to their counterparts in Definition \ref{def-sft}, the transition relations are not exactly the same.
+While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
 %
@@
 
-In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an NFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
+In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
 An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
 Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of automata with $\varepsilon$-transitions ($\varepsilon$NFAs), we extend the framework with a formalization of symbolic $\varepsilon$NFAs (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
-In this paper, we present the core definitions of both $\varepsilon$NFAs and NFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
+In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
 
@@
 record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
-	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\Delta'$| :: "|$('q * 'q)$| set"
-	|$\mathcal{I}$| :: "|$'q$| set"
-	|$\mathcal{F}$| :: "|$'q$| set"
+	|$\mathcal{Q}_e$| :: "|$'q$| set"
+	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
+	|$\Delta_e'$| :: "|$('q * 'q)$| set"
+	|$\mathcal{I}_e$| :: "|$'q$| set"
+	|$\mathcal{F}_e$| :: "|$'q$| set"
 
@@
 Figure \ref{fig-def-FAs} presents the formalization of both SFAs and $\varepsilon$SFAs using Isabelle/HOL record types \texttt{NFA} and \texttt{eNFA}, respectively.
-The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
+The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
 
@@
 
-Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an NFA.
-The parameters $\mathcal{T}$ and $\mathcal{A}$ are intuitive.
-But we need to explain the role of \texttt{F}.
-The output function is of type "$'a \Rightarrow 'b \text{set option}$", which applies to a single element of the type $'a$.
-$\texttt{F}$ extends this function to apply it to a set of elements.
-More precisely, let $f$ be an output function, the semantics of $\texttt{F}$ is defined as $\texttt{F}~f~A=\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset)$.
-The transition relations $\Delta_e$ and $\Delta_e'$ are defined by splitting into two cases, whether the transitions in SFTs are $\varepsilon$-transitions or not.
+Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA.
+The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively.
+But we need to explain the role of $\text{\texttt{F}}$.
+The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$.
+$\text{\texttt{F}}$ extends $f$ to apply it to a set of elements.
+More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as
 
-$\Delta_e$ has two cases:(1) when $(p, (\text{None}, f),  q)$ is a transition in the NFT.
-Because the input can be empty ($\varepsilon$), therefore the NFA does not need to move.
-Therefore the product transtion just output of $f~\text{None}$.
-(2) when $(p, (Some~\sigma_1, f), q)$ is a transition in the NFT.
-This transition can only synchronize with the transitions in the NFA that have shared characters with $\sigma_1$.
-And the output is $F~f~(\sigma_1\cap\sigma_2)$, where $\sigma_2$ is the input label of the NFA.
-Similarly, $\Delta_e'$ has two cases too.
+\[\texttt{F}~f~A=\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset)\].
 
-To establish the correctness of the abstract definition of the product operation, we first formalize the outputs of a Non-deterministic Finite Transducer (NFT) with respect to the input of a Non-deterministic Finite Automaton (NFA).
-Figure \ref{fig-def-output} presents two key definitions: (1) \texttt{outputE}, which extracts the sequence of outputs $o_1, \ldots, o_n$ from a given NFT path $(i_1, o_1), \ldots, (i_n, o_n)$; and (2) \texttt{outputL}, which represents the set of all possible outputs generated by an NFT $\mathcal{T}$ for the input language defined by an NFA $\mathcal{A}$.
+The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions.
+In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels:
+For transitions in $\Delta_e$,
+
+\begin{enumerate}
+\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, the input character is $\varepsilon$.
+Consequently, the SFA remains in its current state, and the product transition produces the output $(f~\text{None})$.
+
+\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$.
+The resulting output is $F~f~(\sigma_1\cap\sigma_2)$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
+\end{enumerate}
+
+The transitions in $\Delta_e'$ follow a similar pattern with analogous cases for $\varepsilon$ and non-$\varepsilon$ transitions.
+
+To establish the correctness specification of the product operation, we begin by formalizing the concept of SFT \emph{traces} as introduced in Definition \ref{def-sft}.
+In our formalization, traces are represented by the type $('a\;\text{option} \times 'b\;\text{option})\;\texttt{list}$, as shown in Figure \ref{fig-def-output}.
+%
+Given a trace $\pi$, we define two key projection functions:
+(1) \texttt{inputE}, which corresponds to $\text{in}(\pi)$ and extracts the input sequence
+(2) \texttt{outputE}, which corresponds to $\text{out}(\pi)$ and extracts the output sequence.
+
+The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
+The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\Delta_t~M~q~\pi~q'$.
 
@@
 	\begin{lstlisting}
+fun inputE :: |$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$| where
+  "inputE [] = []" |$\mid$|
+  "inputE ((Some a, |$\_$|) |$\#$| l) = a |$\#$| (inputE l)" |$\mid$|
+  "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"
+
 fun outputE :: "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" where
@@
                                ('q, 'a) \text{ NFA\_rec} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| M |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. |$q\in \mathcal{I}_T~\mathcal{T} \land q' \in \mathcal{F}_T~\mathcal{T} \land$|
-                |$ \text{LTTS\_reachable }(\Delta_T~\mathcal{T})~\text{M}~q~\pi~q'\land \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}" 
+  "outputL |$\mathcal{T}$| M |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T} \land$|
+                |$ \text{LTTS\_reachable }(\Delta_t~\mathcal{T})~\text{M}~q~\pi~q'\land \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}" 
 	\end{lstlisting}
@@
       and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-    shows "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL |$\mathcal{T}~(\mathcal{M}~\mathcal{T})~\mathcal{A}$|"
+    |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL |$\mathcal{T}~(\mathcal{M}~\mathcal{T})~\mathcal{A}$|"
 
@@
 
-Figure \ref{fig-def-product-correct} shows the correctness of the product operation and Figure \ref{fig-def-output} contains the definitions of \texttt{outputE} and \texttt{outputL} that are used in Lemma \texttt{productT\_correct}.
-The two assumptions define the constraints of the function \texttt{F}.
-The conclusion in \textcolor{blue}{\texttt{show}} illustrates 
-that the language equivalence between the production definition and mathematical semantics.
+Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation.
+The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
+The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.
 
@@
 \label{sec_alg_refinement}
-In Section \ref{sec-form-fts}, we give an abstract definition of NFT's product operation. In this section, we present the possible way to compute the Product from programming language's perspective.
-In this level, we will also refine the transition labels from \emph{sets} to interval algebra.
+Having established the abstract definition of the SFT product operation in Section \ref{sec:product-operation}, we now present its algorithmic refinement.
+This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
 
 \subsection{Intervals}
-An interval is defined as a pair $(i, j)$, which corresponds to the set $\{e \mid e \geq i \land e \leq j\}$.
-In our formalization, we extends the interval definition to a list, that is $[(i_1, j_1), \ldots, (i_n, j_n)]$, which corresponds to the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq i_n\}$.
-This definition enables us to merge more transitions in NFAs.
-Moreover, when an interval operation produces more than two intervals, we do not need to split the intervals.
-For instance, the interval result of $(1, 5) - (3, 4)$ (the difference between the interval $[1, 5]$ and $[3, 4]$) is equivalent to $[(1, 2), (5, 5)]$.
 
-In the sequent, intervals always mean interval lists.
-There are a collection interfaces for intervals. 
+An interval is defined as a pair $(i, j)$ representing the set $\{e \mid i \leq e \leq j\}$.
+To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
+This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
+For example, the set difference between intervals $[1, 5]$ and $[3, 4]$ can be directly represented as the interval list $[(1, 2), (5, 5)]$.
+
+Throughout the following discussion, we use the term "interval" to refer to interval lists. Our formalization provides a collection of interval operations through the following interface:
+
 \begin{enumerate}
-  \item $\texttt{semI}~i$, formalize the semantics of an interval as a set, that is $\texttt{semI }(i,j) = \{e \mid e \geq i \land e \leq j\}$.
-  \item $\texttt{emptyI}~i$, checks the emptiness of an interval.
-   \item $\texttt{nemptyI}~i$, checks the non-emptiness of an interval.
-  \item $\texttt{intersectIs}~i_1~i_2$, computes the intersection of two intervals.
-  \item $\texttt{diffIs}~i_1~i_2$, computes the intervals difference, that is the interval corresponds to the set $\texttt{semI}(i_1)-\texttt{semI}(i_2)$.
+  \item $\texttt{semI}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semI}~(i,j) = \{e \mid i \leq e \leq j\}$.
+  \item $\texttt{emptyI}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semI}~i = \emptyset$.
+  \item $\texttt{nemptyI}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semI}~i \neq \emptyset$.
+  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semI}~i = \texttt{semI}~i_1 \cap \texttt{semI}~i_2$.
+  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semI}~i = \texttt{semI}~i_1 \setminus \texttt{semI}~i_2$.
 \end{enumerate}
 
-We define a canonical form for intervals to facilitate the proof. An interval $[i_1, j_1), \ldots, (i_n, j_n)]$ is in canonical form if 
-(1) $i_k \leq j_k, \text{ for } 1 \leq k \leq n$,
-(2) $j_k\leq i_{k+1}, \text{ for } 1 \leq k < n$.
-All the interval operations are proved to keep the canonical form after the operation if the input intervals are in canonical form.
-Canonical form not only simplifies the proof, but also improves the performance of interval operations.
+To facilitate formal reasoning and optimize performance, we introduce a canonical form for interval.
+An interval $[(i_1, j_1), \ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
+\begin{enumerate}
+  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
+  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
+\end{enumerate}
 
-\subsection{Algorithm of Product Computation.}
+We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
+This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
 
-Fig.
-\ref{fig-compute-nft-product} illustrates the product computation of a SFT and an SFA.
-\footnote{To make it easier to understand, the Isabelle code presented in this section is not exactly the same as the real one.}
-\texttt{prods\_imp} computes the product of two sets.
-It is illustrated in \ref{fig-def-prods_imp}.
-We use the constructs from \emph{Refine\_Monadic}.
-\texttt{prods\_imp} itself is intuitive: the nested two loops traverse all possible combination of states in the states.
-The \texttt{FOREACH} construct in \emph{Refine\_Monadic} is similar to OCaml's \texttt{Set.fold}.
-$\texttt{FOREACH}~S~f~I$ traverses all elements in $S$ and applies $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$ to each element.
-$I$ is the initial value of type $'b$.
+\subsection{Algorithmic Implementation of the Product Operation}
+
+In this subsection, we present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.}.
+Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
+The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets.
+This function employs the \texttt{FOREACH} construct, a higher-order iteration operator analogous to OCaml's \texttt{Set.fold}.
+Specifically, given a set $S$, a function $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$, and an initial accumulator $I$ of type $'b$, the expression $\texttt{FOREACH}~S~f~I$ systematically applies $f$ to each element in $S$, accumulating results in a principled manner.
 
@@
 
-The challenging part is how to compute the transition sets.
-It is illustrated in Fig.
-\ref{fig-def-prods-imp}.
-The function \texttt{trans\_comp\_imp} traverses the transitions in the SFT and it calls two the following two functions to compute the synchronized transitions:
-(1) \texttt{subtrans\_comp\_$\varepsilon$}: handles the case that the transition in the SFT is $\varepsilon$-transition.
-(2) \texttt{subtrans\_comp}: handles the case that the transition in the SFT is not $\varepsilon$-transition.
+A central algorithmic challenge in our implementation lies in the computation of transition sets.
+As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
+This function decomposes the synchronization process into two distinct cases, each handled by a specialized function:
+
+\begin{enumerate}
+  \item \texttt{subtrans\_comp\_$\varepsilon$}: Processes $\varepsilon$-transitions in the SFT, where transitions consume no input but may produce output
+  \item \texttt{subtrans\_comp}: Processes standard transitions in the SFT, where both input consumption and output generation may occur
+\end{enumerate}
 

### 2025-03-08 Paper/main.tex
finish the first round revision

@@
 
-\title{Certified Symbolic Transducer with the Application of String Solver} %TODO Please add
+\title{Certified Symbolic Transducer with Applications in String Solving} %TODO Please add
 
@@
 The second element of type $'b$ serves as an index into the output function space.
-The mapping $\mathcal{M}$ associates each index with a specific output function.
+The mapping $\mathcal{M}_t$ associates each index with a specific output function.
 These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
@@
 In this subsection, we present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.}.
+%
 Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
+Note that in the implementation, there are some new operations: \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nfa\_accepting}, and \texttt{nft\_tranfun}.
+There are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
+
 The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets.
@@
 definition trans_comp_imp where
-  "trans_comp_imp M F T1 T2 Q =
+  "trans_comp_imp M F fe T1 T2 Q =
     FOREACH {t. t |$\in$| T1}
@@
         (if (|$\alpha$| = None) then 
-          (subtrans_comp_|$\varepsilon$| M q  f q|$'$| F T2 D1 D2)
+          (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe T2 D1 D2)
         else 
@@
 
-We elaborate only on the function \texttt{subtrans\_comp} in Fig. \ref{fig-def-subtrans_comp}. The other function is similar. 
-This function traverses the transitions SFA's transitions $\texttt{T}$. If it find a transtion $(q_1, \alpha', q_1')\in \texttt{T}$ satisfies the following condition:
-$\alpha \cap \alpha' \neq \emptyset$ (Checked by \texttt{nemptyIs}),
-it started to compute the non-$\varepsilon$-transitions set $\texttt{D1}$ and the $\varepsilon$-transitions set $\texttt{D2}$ by checking whether the output set of $\alpha \cap \alpha'$ is empty or not.
-
 \begin{figure}[hbt!]
 	\begin{lstlisting}
-definition subtrans_comp_imp where
-  "subtrans_comp_imp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
+definition subtrans_comp where
+  "subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
     FOREACH
@@
             else RETURN D1);
-          D2 |$\leftarrow$| (if (M f) (intersectIs |$\alpha$| |$\alpha'$|) |$\neq$| |$\emptyset$| then 
+          D2 |$\leftarrow$| (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) |$\neq$| |$\emptyset$| then 
             RETURN (((q,q1), (q|$'$|,q1|$'$|)) |$\cup$| D2) else RETURN D2);
@@
 
+    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
+This function systematically processes transitions in the SFA, represented by the set \texttt{T}.
+For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
+
+    \begin{enumerate}
+      \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
+      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists an input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
+It is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
+    \end{enumerate}
+
   \begin{figure}[hbt!]
@@
   lemma productT_imp_correct:
-  assumes finite_TT: "|$\text{finite } (\Delta_T~\mathcal{T})$|"    
+  assumes finite_TT: "|$\text{finite } (\Delta_t~\mathcal{T})$|"    
       and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
       and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
-      and finite_TQ: "|$\text{finite } (\mathcal{Q}_T~\mathcal{T})$|"
+      and finite_TQ: "|$\text{finite } (\mathcal{Q}_t~\mathcal{T})$|"
       and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
-      and finite_TI: "|$\text{finite } (\mathcal{I}_T~\mathcal{T})$|"
+      and finite_TI: "|$\text{finite } (\mathcal{I}_t~\mathcal{T})$|"
       and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
-      and finite_TF: "|$\text{finite } (\mathcal{F}_T~\mathcal{T})$|"
+      and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
   shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe} \leq \text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
@@
 
-  To prove the correctness of the product computation introduced above, we just need to show that the product result of \texttt{productT\_imp} in Fig.
-\ref{fig-def-prods_imp} is a refinement of the product result of \texttt{productT} in Figure \ref{fig-compute-nft-product}.
-This refinement specification is illustrated in Figure \ref{fig-def-productT_imp_correct}.
-The proof is under the framework of \emph{Refine\_Monadic}.
-The specification $\mathbf{C} \leq\texttt{SPEC}~\mathbf{A}$ means that $\mathbf{C}$ is an element in the set $\mathbf{A}$.
-In Fig.
-\ref{fig-def-productT_imp_correct}, "$\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F$" specifies a set of elements $A$, such that $A$ equals to $\texttt{productT}~\mathcal{T}~\mathcal{A}~F$.
-  The specification will force us to prove the result $\varepsilon$-NFAs of \texttt{productT\_imp}  and \texttt{productT} have the same sets of initial states, transition relations, $\varepsilon$-transitions, and accepting states.
+  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
+This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.
 
-\subsection{Refinement to Efficient Data Structure}
+  The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$.
+More precisely,  the concrete implementation must produce an SFA that is structurally equivalent to the one produced by the abstract algorithm \texttt{productT}.
 
-The last subsection shows the way to compute the product of two NFAs and the transition labels have been refined to interval algebra. But it still uses the data structure of sets in Isabelle/HOL.
+  To establish this equivalence, we must prove that the SFAs produced by \texttt{productT\_imp} and \texttt{productT} are identical in all essential components:
+  initial state sets, transition relations, $\varepsilon$-transition sets, accepting state sets.
 
-The \emph{Refine\_Monadic} framework provides a way to refine the data structure of sets to more efficient data structure, such as red-black trees.
-The basic idea is to provide a set of \emph{set} operations on the refined data structure and prove the refinement relation between the operations on the refined data structure and the operations on the abstract data structure.
-%
-In our formalization, we refine the data structure of sets to red-black trees.
-
-We extend CertiStr \cite{cpp/KanLRS22} not only to support SFTs but also $\varepsilon$SFA, including the conversion from $\varepsilon$SFA to SFA and the the correctness of the conversion.
+The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
+The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 
+In our formalization, we refine the sets of storing states and transitions to red-black trees.
 
@@
 \label{sec-app-str-solver}
-In this section, we will show how to use the formalized symbolic transducer to solve the string constraints with replacement operations.
 
-CertiStr is an existing string solver based on SFAs.
-However, due to the limitation of SFAs, it cannot solve the string constraints with replacement operations.
-CertiStr is based on forward-propagations.
-We will not elaborate and repeat the idea of forward-propagation in this paper.
-In this section, we only focus on how to model the replacement operation in SFTs.
+In this section, we demonstrate the practical application of our formalized SFTs to string constraint solving, specifically focusing on replacement operations.
+While CertiStr \cite{cpp/KanLRS22} provides a verified framework for string constraint solving using SFAs, its capabilities are inherently limited by the expressiveness of SFAs.
+In particular, SFAs cannot directly model string transformation operations such as replacement.
+
+Our string solver is based on CertiStr \cite{cpp/KanLRS22} by extending it with the modeling of replacement operations.
+CertiStr exploits forward-propagation to solve the string constraints.
+Our SFT modeling for replacement operations can seamlessly be integrated into CertiStr's forward-propagation framework.
 
@@
 
-A replacement operation has the following form: $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
-This operation replaces the occurrence of the substring in \texttt{str} that matches \texttt{pattern} with \texttt{replacement}.
-\texttt{str} and \texttt{replacement} are strings and \texttt{pattern} is a regular expression.
+The string replacement operation, denoted as $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
+\begin{itemize}
+  \item $\texttt{str}$: The input string to be transformed
+  \item $\texttt{pattern}$: A regular expression defining the matching criteria
+  \item $\texttt{replacement}$: The string to be substituted for matched substrings
+\end{itemize}
 
-The semantics of the replacement operation can be described in two cases: (1) if there are a substring in \texttt{str} that match \texttt{pattern}, then replace it with \texttt{replacement}.
-(2) If there are no matches, then the result is \texttt{str} itself.
+The semantics of the replacement operation can be formally characterized by two distinct cases:
+\begin{enumerate}
+  \item When there exists at least one substring $s'$ in \texttt{str} such that $s'$ matches $\texttt{pattern}$, then $s'$ is replaced with \texttt{replacement}.
+  \item When no substring of \texttt{str} matches $\texttt{pattern}$, the operation returns \texttt{str} unchanged.
+\end{enumerate}
 
@@
 We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}(s, $\texttt{/[0-9]+/}$, $"$\texttt{NUM}$"$)$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string"\texttt{NUM}".
+Given a replacement operation $\texttt{replace}(s, $\texttt{/[0-9]+/}$, $"$\texttt{NUM}$"$)$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
 We model this replacement operation by an SFT with the following 3 steps:
@@
 \begin{enumerate}
-  \item Generate an SFA from the regular expression $\texttt{/[0-9]+/}$. Adding output functions to the transitions of the SFA that output empty string (\texttt{None}) to make it an SFT.
-  \item Generate an SNFA from the replacement string "\texttt{NUM}".
-Note that a single string is also a regular expression.
-Add input labels and output functions to the transitions of the SNFA to make it an SFT.
-  \item Concatenate the two SFTs and add preceding and succeeding self-loop transitions to make it accept the whole input string.
+  \item First, construct an SFA that recognizes the regular expression pattern $\texttt{/[0-9]+/}$.
+Transform this SFA into an SFT by augmenting each transition with an output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
+  \item Second, construct an SFA that accepts the replacement string "\texttt{NUM}".
+Since a constant string can be viewed as a specialized regular expression, we can construct its SFA representation.
+Convert this SNFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
+  \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
+These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and empty output function, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
 \end{enumerate}
@@
 \noindent\emph{Step 1.}
-The SFA corresponding to the regular expression "$\texttt{/[0-9]+/}$" is shown in Figure \ref{fig-snfa-pattern}.
-On the left side is the SFA and on the right side is the corresponding SFT.
-The function $\texttt{f} = \lambda x.~\texttt{None}$ is the output function that outputs empty string.
+Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
+The left side shows the SFA that recognizes the regular expression $\texttt{/[0-9]+/}$, while the right side presents its transformation into an SFT.
+This transformation is achieved by augmenting each transition with the output function $\texttt{f} = \lambda x.~\texttt{None}$, which consistently produces the empty string, effectively "consuming" the matched digits without generating output.
 
@@
 \noindent\emph{Step 2.}
-The SNFA corresponding to the replacement string "\texttt{NUM}" is shown in Figure \ref{fig-snfa-replacement}.
-On the left side is the SFA and on the right side is the corresponding SFT.
-The function \texttt{g} is the output function that outputs the characters in the replacement string.
-\texttt{g} is defined as:
+Figure \ref{fig-snfa-replacement} depicts the automata for the replacement string "\texttt{NUM}".
+The left side shows the SFA that accepts this constant string, while the right side presents its SFT transformation.
+The transformation employs an indexed output function \texttt{g} that emits characters of the replacement string sequentially:
 \[
-\texttt{g} = \lambda~i~x.~\texttt{match }~i~ with~ 1 \rightarrow [(78, 78)] ~|~ 2 \rightarrow [(85, 85)] ~|~ 3 \rightarrow [(77, 77)] ~|~ \_ ~\rightarrow \texttt{None}
+\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
+\begin{cases}
+1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
+2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
+3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
+\_ \mapsto \texttt{None}
+\end{cases}
 \]
-Note that the ASCII code of  $\texttt{'N'}$, $\texttt{'U'}$, $\texttt{'M'}$ are 78, 85, 77 respectively.
+This function maps transition indices to their corresponding character outputs, using ASCII codes to represent the string "\texttt{NUM}" character by character.
 
@@
 \noindent\emph{Step 3.}
-The concatenation of the two SFTs in Fig.
-\ref{fig-snfa-pattern} and Fig.
-\ref{fig-snfa-replacement} is shown in Figure \ref{fig-rearranged-automata}.
-It contains two minor steps: (1) add a transtion (\texttt{None}/\texttt{f}) from each accepting state of the first SFT to the initial state of the second SFT.
-(2) add the preceding and succeeding transitions to make it accept the whole input string.
+Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT (Fig.
+\ref{fig-snfa-pattern}) with the replacement-generating SFT (Fig.
+\ref{fig-snfa-replacement}).
+The composition process involves two key modifications:
+\begin{enumerate}
+  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with \texttt{None}/\texttt{f}) from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
+  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with $\Sigma/\texttt{f}$, where $\Sigma$ represents the full alphabet.
+These transitions enable the SFT to process arbitrary input prefixes and suffixes while preserving the matched substring for replacement
+\end{enumerate}
 
@@
      (q1') edge [loop above] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
-     (q0') edge [loop below] node {\texttt{None}/\texttt{f}} (q0');
+     (q0') edge [loop below] node {$\Sigma$/$\texttt{f}$} (q0');
 
@@
     (q2'') edge node {\texttt{None}/\texttt{g 3}} (q3'')
-    (q3'') edge [loop below] node {\texttt{None}/\texttt{f}} (q3'')
+    (q3'') edge [loop below] node {$\Sigma$/$\texttt{f}$} (q3'')
     (q1') edge node {\texttt{None}/\texttt{f}} (q0'');
@@
 
-  With this SFT, we can compute the forward image of the replacement operation.
-  Let $s' = \texttt{replace}(s,~$\texttt{/[0-9]+/}$,~\text{"}$\texttt{NUM}$\text{"})$ be a string constraint and we want to compute the value range of $s'$ regarding this replacement operation.
-  Assume the SFT generated from this replacement operation is $\mathcal{T}$ and the string variable $s$'s values ranges over an SFA $\mathcal{A}$.
-The foward image of this replacement operation is $\mathcal{T}\times \mathcal{A}$.
-It also means that the value of $s$' must ranges over this image.
+  Having constructed the SFT, we can now compute the forward image of the replacement operation.
+Consider the string constraint $s' = \texttt{replace}(s,~\texttt{/[0-9]+/},~\text{"}\texttt{NUM}\text{"})$, where we aim to characterize the possible values of $s'$.
+Let $\mathcal{T}$ denote the constructed SFT modeling the replacement operation, and let $\mathcal{A}$ be the SFA representing the domain of possible values for the input string $s$.
+The forward image of this replacement operation is given by the product $\mathcal{T} \times \mathcal{A}$, which precisely captures the set of all possible output strings that can be produced by applying the replacement operation to any input string accepted by $\mathcal{A}$.
+  Our string solver uses this forward image to do forward propagation for further string solving.
 
-  \emph{Compare with the semantics of SMTLIB.} SMTLIB's replacement's semantics is not exactly the same as the semantics  in this paper.
-In SMTLIB, the single replacement operation is defined as replace the \emph{first} occurrence of the substring that matches the pattern with the replacement string.
-Our semantics is not limited to the \emph{first} occurrence.
-
-  From replacement operation modeling above, we can see that our SFT's formalization with the support of empty string for input labels and output functions make it easy to model the replacement operation.
-  However,
-  The power of our SFT's formalization is not fully explored in the string solver as the transition labels and output functions are simple.
-For other applications, the transition labels and output functions can indeed support more complex modelings.
+  \emph{Comparison with SMT-LIB Semantics.} It is important to note that our formalization differs from the standard SMT-LIB semantics for the replacement operation.
+In SMT-LIB, the \texttt{str.replace} operation is defined to replace only the \emph{first} occurrence of a substring matching the given pattern.
+In contrast, our semantics allows for a more general interpretation where any matching substring may be replaced.
 
@@
 
-We have implemented a prototype of the string solver to support the replacement operation based on CertiStr \cite{cpp/KanLRS22} \footnote{The code is available at \url{https://github.com/ShlKan/certified-str-solver}}, which is called CertiStrR in this paper.
-The frontend of CertiStrR is not certified and implemented with the dolmen \cite{dolmen} library (An OCaml library of SMTLIB parser) and ocaml-re-nfa \cite{ocaml-re-nfa} library (An OCaml library to construct an NFA from a regular expression).
+We have implemented CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}, to support string replacement operations\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}.
+While the core solving algorithm maintains the certification guarantees of CertiStr, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
 
+CertiStrR implements two kinds of SMT-LIB's replacement operations:
+\texttt{str.replace s p r}: a string-based replacement where pattern \texttt{p} is a constant string.
+\texttt{str.replace\_re s p r}: a regular expression-based replacement where pattern \texttt{p} is a regular expression
+
+Listing \ref{lst-smtlib-code} demonstrates the regular expression variant through an example that replaces numeric sequences with the string "\texttt{NUM}".
+The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
 %
-CertiStr-R supports two kinds of SMT-LIB's replacement operations: (1) \texttt{str.replace s p r}, where the pattern \texttt{p} is a string and (2) \texttt{str.replace\_re s p r}, where the pattern \texttt{p} is a regular expression.
-Listing \ref{lst-smtlib-code} shows an example of the SMT-LIB code corresponding to the replacement operation $\texttt{replace}(s,~$\texttt{/[0-9]+/}$,~\text{"}$\texttt{NUM}$\text{"})$ that we illustrated in Section \ref{sec-app-str-solver}.
-The constraints are satisfiable because the
-substring "\texttt{2025}" in \texttt{a} can be matched by the regular expression \texttt{/[0-9]+/} and then replaced by "\texttt{NUM}".
+
+As discussed in the previous subsection, our replacement semantics differs from the SMT-LIB standard semantics.
+When we run the SMT solver CVC5 \cite{cvc5} with the example code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
 
@@
 
-  As we discussed in Section \ref{sec-app-str-solver}, our replacement semantics is not exactly the same as SMTLIB's standard semantics.
-If we run the SMT solver CVC5 \cite{cvc5} with the example code in Listing \ref{lst-smtlib-code}, the result is UNSAT. Because CVC5 only matches the first occurrence of the pattern.
-
-We select benchmarks from the SMT-LIB 2024 \cite{smtlib_benchmarks} to test the effective and efficiency of CertiStrR.
-The benchmarks contain mainly from QF\_S and QF\_SLIA.
-The benchmarks are devided into 2 groups: (1) tests with \texttt{str.replace} and (2) tests with \texttt{str.replace\_re}.
-Because CertiStrR currently has a limited front-end to support all SMT-LIB operations.
-We rewrite the operations that are not supported by CertiStrR, for example, \texttt{(assert (and c1 c2))} is rewritten as two separate \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints that CertiStrR supports.
+  We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
+The benchmarks are divided into two categories based on the replacement operations: (1) string-based replacement (\texttt{str.replace}) and (2) regular expression-based replacement (\texttt{str.replace\_re}).
+Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB specification, we preprocess certain operations.
+For instance, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
 
@@
 
-The experiments were conducted on a laptop equipped with an Apple M4 processor, featuring 24 GB of memory.
-The timelimit for each test is $1$ minutes.
-We split the tests into 2 groups: (1) tests with \texttt{str.replace} (\texttt{replace\_str}, 323 tests) and (2) tests with \texttt{str.replace\_re} (\texttt{replace\_re}, 98 tests).
-The average time of \texttt{replace\_str} is 0.27 seconds and the average time of \texttt{replace\_re} is 0.37 seconds.
-The test results are classified into 3 categories: (1) SAT, (2) UNSAT, and (3) Inconclusive.
-The result inconclusive means that the solver cannot prove the result.
-This does not mean the the test runs out of time.
-It means the solver cannot prove the result with the current forward-propagation algorithm used by CertiStr \cite{cpp/KanLRS22}.
-Forward-propagation algorithm cannot decide the result if the constraints do not satisfy tree-property, even after the forward-propagation all variables's domain are not empty as illustrated in CertiStr \cite{cpp/KanLRS22}.
+The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a time limit of one minute per test. 
+The results show average execution times of 0.27 seconds for \texttt{str.replace} and 0.37 seconds for \texttt{str.replace\_re} operations.
+Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
+%
+An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
+Specifically, when the string constraints do not satisfy the tree property, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
 
-After analyzing the tests that consume more time than others.
-We found the replacement operation modeling itself is efficient.
-The time consumption of is mainly from the automata accumulation during the forward-propagation algorithm.
-For instance, given the constraints as follows:
+Our performance analysis revealed that while the SFT-based replacement operation modeling is efficient, the primary computational bottleneck stems from automata accumulation during forward-propagation.
+Consider the following constraint system:
 \[
-x=x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r);~x = x_4;~x = \texttt{replace}(x, p_1, r_1);
+x = x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r);~x = x_4;~x = \texttt{replace}(x, p_1, r_1);
 \]
-where \texttt{++} is the concatenation operation.
-As the varible $x$ appears many times on the left-hand side of the constraints, the forward-propagation algorithm will accumulate the automata of $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace}($x, p_1, r_1$) to $x$ during the forward-propagation.
-Each accumulation will perform the product operation with the previous automata.
-The complexity of the product operation is $O(n^2)$ in the worst case, where $n$ is the size of the automata.
-This leads to state explosion.
+where \texttt{++} denotes string concatenation.
+The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for each constraint: $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace}($x, p_1, r_1$).
+Each accumulation step requires computing the product of the current automaton with the previous result.
+Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state and transition explosion.
 
@@
 
-We discuss the effort of certified development in this subsection.
-The effort of certified development is mainly from the following two aspects: (1) the abstract level Formalization of SFTs and (2) the low-level refinement of the SFTs to efficient data structure.
+We discuss the effort of certified SFT development in this subsection.
+Table \ref{tab:abstract_impl} provides an overview.
+The abstract-level development means all formalizations in Section \ref{sec:product-operation}.
+The implementation-level development means all formalizations in Section \ref{sec_alg_refinement}.
 
@@
       \toprule
-      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs} \\
+      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs} (lines of code) \\
       \midrule

### 2025-03-09 Paper/main.tex
first version of the paper

@@
 
-We have implemented CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}, to support string replacement operations\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}.
+We have implemented CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
+%\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
 While the core solving algorithm maintains the certification guarantees of CertiStr, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
@@
 Table \ref{tab:abstract_impl} provides an overview.
-The abstract-level development means all formalizations in Section \ref{sec:product-operation}.
-The implementation-level development means all formalizations in Section \ref{sec_alg_refinement}.
+The abstract-level development means all formalizations in Section \ref{sec:formalization} and  \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs .
+The implementation-level development means all formalizations in Section \ref{sec_alg_refinement} including the refinement of the product operation.
+The last row corresponds to the effort of the interval formalization.
+The most difficult part is the correctness proof of the product operation at the abstract level.
+Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
 
@@
       \midrule
-      Abstract & & & \\
-      Implementation & & & \\
+      Abstract-level & 17 & 21 & 3274 \\
+      Implementation-level & 51 & 43 & 2700 \\
+      Interval & 15 & 29 & 1500 \\
       \bottomrule
@@
 
+Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer,popl/VeanesHLMB12, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
+This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
+Recent developments have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
+
+The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
+These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
+Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.
+
+While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
+To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
+Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.
+
+A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
+While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
+We believe that our work will benefit the community by providing a formal foundation for string solvers.
+
+Beyond string theories, certification efforts in SMT solving have extended to other domains.
+Notable among these is the work by Shi et al.
+\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of formal verification in automated reasoning.
+
 \section{Conclusion}
@@
 
+In this paper, we present the first formalization of SFTs in the proof assistant Isabelle/HOL.
+The formalization features flexible interfaces for easily modeling different applications by supporting $\varepsilon$-transitions for both input and output, and extensive to different boolean algebras such as intervals, arithmetic operations, among others by leveraging the refinement framework.
+In order to assess the effciency and effectiveness of our formalization, we have implemented CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}, to support string replacement operations and evaluated it on a set of benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
+The experimental results show that our formalization is efficient and effective.
+Even though we only illustrate the application of our formalization of SFTs in string solvers, out extensible framework on transition labels is applicable to other domains such as program verification.
+
+In the future, we plan to extend our symbolic formalization to different variant SFTs.
+For example, attaching priority to transtions so that we can model the standard semantics of replacement operations in SMT-LIB.
+
 \bibliography{literature}

### 2025-03-09 Paper/main.tex
polish the conclusion

@@
 
-In this paper, we present the first formalization of SFTs in the proof assistant Isabelle/HOL.
-The formalization features flexible interfaces for easily modeling different applications by supporting $\varepsilon$-transitions for both input and output, and extensive to different boolean algebras such as intervals, arithmetic operations, among others by leveraging the refinement framework.
-In order to assess the effciency and effectiveness of our formalization, we have implemented CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}, to support string replacement operations and evaluated it on a set of benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
-The experimental results show that our formalization is efficient and effective.
-Even though we only illustrate the application of our formalization of SFTs in string solvers, out extensible framework on transition labels is applicable to other domains such as program verification.
+In this paper, we have presented the first formalization of Symbolic Finite Transducers (SFTs) in the proof assistant Isabelle/HOL.
+Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both input and output, and extensibility to various boolean algebras (including intervals and arithmetic operations) via the refinement framework.
 
-In the future, we plan to extend our symbolic formalization to different variant SFTs.
-For example, attaching priority to transtions so that we can model the standard semantics of replacement operations in SMT-LIB.
+To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}.
+This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
+The experimental results confirm both the efficiency and effectiveness of our approach.
+While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.
+
+Future work includes extending our symbolic formalization to handle variants of SFTs.
+One particularly promising direction is the addition of transition priorities, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations.
 

### 2025-03-10 Paper/main.tex
polish the paper

@@
 This simplistic representation can lead to a phenomenon known as transition explosion.
-For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x10FFFF}$ distinct characters.
-Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x10FFFF}$ individual transitions, rendering the product of two FAs highly inefficient.
+For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
+Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the product of two FAs highly inefficient.
 Furthermore, in practical applications, it is often necessary to consider infinite alphabets, such as the set of all integers.
 
-Symbolic Finite Automata (SFA) and Symbolic Finite-state Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic Finite Automata (FAs) and Finite-state Transducers (FTs), enhancing their applicability in practical scenarios.
+Symbolic Finite Automata (SFA) and Symbolic Finite-state Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
+%
 These symbolic models utilize transition labels defined by boolean algebras, allowing for more expressive representations.
@@
 FTs constitute a significantly more expressive and powerful theoretical framework compared to symbolic FAs, as evidenced by their capability to model complex string transformations such as replacement operations.
-Furthermore, the seminal work of \cite{VeanesHLMB12Transducer} demonstrates the broad applicability of SFTs across diverse domains, including security-critical applications such as cross-site scripting (XSS) prevention, image transformation operations, and privacy-preserving location data processing.
+%
+Furthermore, numerous studies have demonstrated the broad applicability of SFTs across diverse domains.
+The work of \cite{VeanesHLMB12Transducer} showcases SFTs in security-critical applications, particularly for cross-site scripting (XSS) prevention.
+\cite{uss/HooimeijerLMSV11} extends this security focus by applying SFTs to web application sanitizer analysis.
+In system verification, \cite{osdi/YaseenABCL20} employs SFTs for runtime behavior monitoring, while \cite{pldi/HuD17} demonstrates their utility in automated program transformation through systematic inversion techniques.
 
@@
 
-Leveraging the data refinement framework \cite{DBLP:conf/itp/Lammich13} formalized in Isabelle/HOL, we implement an efficient representation of states and transitions using sophisticated data structures such as hashmaps and red-black trees.
-Furthermore, we develop a comprehensive interval algebra that provides efficient operations for creation, verification, and manipulation of intervals, thereby facilitating the implementation of replacement operations within our string solver framework.
+In the refinement level, we implemented transition labels using an interval-based representation.
+The formalization of interval algebra provides efficient set-theoretic operations—including membership checking, intersection, and difference computations—facilitating the refinement of transition labels from abstract sets to concrete intervals.
+Furthermore, leveraging the data refinement framework \cite{DBLP:conf/itp/Lammich13} in Isabelle/HOL, we store states and transitions using sophisticated data structures such as hashmaps and red-black trees, ensuring efficient automata manipulation.
 
-In summary, Our formalization makes two primary contributions.
-First, we achieve computational efficiency through an algebraic representation that enables compressed storage of transition labels.
-Second, our refinement-based formalization framework demonstrates significant extensibility, facilitating the incorporation of diverse theories of transition labels.
+To evaluate the effectiveness of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
+We evaluated our implementation using a comprehensive set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
+The experimental results demonstrate that our formalization achieves computational efficiency in practical constraint-solving scenarios.
 
-The paper is organized as follows.
-Section \ref{sec:formalization} presents the formalization of symbolic FTs.
-Section \ref{sec:product-operation} presents the product operation of symbolic FTs in the abstraction level.
-Section \ref{sec_alg_refinement} presents the refinement of the product operation in Isabelle/HOL.
-Section \ref{sec-app-str-solver} presents the application of SFTs in the modeling of replacement operations in string solvers.
-Section \ref{sec:related-work} presents the related work.
-Section \ref{sec:conclusion} concludes the paper.
+In summary, our formalization makes the following contributions:
+\begin{enumerate}
+\item We present the first formalization of symbolic finite-state transducers in the Isabelle/HOL proof assistant.
+\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodates various boolean algebras, ensuring adaptability to future transition label developments.
+\item We develop a comprehensive interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
+\item We demonstrate the practical utility of our formalization through its application to string constraint solving, specifically in modeling replacement operations with verified correctness guarantees.
+\end{enumerate}
+
+The remainder of this paper is organized as follows:
+Section \ref{sec:formalization} introduces the formal foundations of symbolic finite-state transducers.
+Section \ref{sec:product-operation} presents the product operation at the abstract level.
+Section \ref{sec_alg_refinement} details the algorithmic refinement of the product operation.
+Section \ref{sec-app-str-solver} demonstrates the application to string constraint solving.
+Section \ref{sec:related-work} discusses related work.
+Section \ref{sec:conclusion} concludes with future directions.
 
@@
 
-Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer,popl/VeanesHLMB12, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
+\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer,popl/VeanesHLMB12, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
 This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
@@
 
-The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
+\emph{Applications of Symbolic Automata and Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
 These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
@@
 
-While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
+\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
 To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
@@
 
-A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
+\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
 While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
-We believe that our work will benefit the community by providing a formal foundation for string solvers.
+As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
+We believe that our work will benefit the community by providing a formal foundation for string solvers development.
 
-Beyond string theories, certification efforts in SMT solving have extended to other domains.
-Notable among these is the work by Shi et al.
-\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of formal verification in automated reasoning.
+\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains.
+For example, the work by Shi et al.
+\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.
 
@@
 
-Future work includes extending our symbolic formalization to handle variants of SFTs.
-One particularly promising direction is the addition of transition priorities, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations.
+Future work includes extending our symbolic formalization in two key directions:
+%
+(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
+(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
+These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.
 

### 2025-03-14 Paper/main.tex
finish second round revision

@@
 \EventNoEds{2}
-\EventLongTitle{15th Conference on
+\EventLongTitle{16th Conference on
 Interactive Theorem Proving (ITP 2025)}
@@
 Despite the widespread adoption of formalizing FAs and FTs within proof assistants such as Coq and Isabelle/HOL, these formalizations often fall short in terms of applicability to real-world scenarios.
-A significant limitation of classical finite-state models is that transition labels are typically confined to a single character from a finite alphabet.
+A significant limitation of classical finite-state models is that transition labels are typically defined as a single character from a finite alphabet.
 However, in practical applications, the alphabet of an FA or FT can be extensive or even \emph{infinite}.
-This classical approach to formalizing transitions can result in transition explosion, leading to critical performance bottlenecks of FA operations.
+This classical approach to formalizing transitions can result in transition explosion, leading to critical performance bottlenecks of FA and FT operations.
 
@@
 To assess its performance, we applied the formalized symbolic FTs to an SMT string solver for modeling replacement operations.
-The experimental results indicate that the formalized symbolic transducer can efficiently solve string constraints.
+The experimental results indicate that the formalized symbolic transducer can efficiently and effectively solve string constraints with replacement operations.
 
@@
 For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
-Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the product of two FAs highly inefficient.
+Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the FA and FT operations highly inefficient.
 Furthermore, in practical applications, it is often necessary to consider infinite alphabets, such as the set of all integers.
 
-Symbolic Finite Automata (SFA) and Symbolic Finite-state Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
+Symbolic Finite-state Automata (SFA) and Symbolic Finite-state Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
 %
 These symbolic models utilize transition labels defined by boolean algebras, allowing for more expressive representations.
-For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \% 2 = 0$), denoting all even numbers.
+For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \mod 2 = 0$), denoting all even numbers.
 This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
@@
 
-Prior work has successfully formalized symbolic FAs within Isabelle/HOL through CertiStr \cite{cpp/KanLRS22}, demonstrating both the efficiency and effectiveness of symbolic FAs in practice.
+Prior work of CertiStr \cite{cpp/KanLRS22} has successfully formalized SFAs within Isabelle/HOL, demonstrating both the efficiency and effectiveness of SFAs in practice.
 However, the formalization of FTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
@@
 
-In this work, we present a comprehensive formalization of symbolic FTs that builds upon the formalization of symbolic FAs.
+In this work, we present a comprehensive formalization of SFTs.
 To address the extensibility challenges inherent in supporting diverse transition label theories, we adopt a refinement-based approach.
@@
 
-To evaluate the effectiveness of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
-We evaluated our implementation using a comprehensive set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
-The experimental results demonstrate that our formalization achieves computational efficiency in practical constraint-solving scenarios.
+To evaluate the effectiveness and efficiency of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
+We evaluated our implementation using a set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
+The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
 
@@
 \item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodates various boolean algebras, ensuring adaptability to future transition label developments.
-\item We develop a comprehensive interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
+\item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
 \item We demonstrate the practical utility of our formalization through its application to string constraint solving, specifically in modeling replacement operations with verified correctness guarantees.
@@
 
-\begin{definition}[Symbolic Finite Transducer]
+\begin{definition}[Symbolic Finite-state Transducer]
 \label{def-sft}
-   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
+   A Symbolic Finite-state Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
    \begin{itemize}
@@
 
-SFTs accept an input word and generate an output. This can be defined by \emph{runs} of SFTs.
+SFTs accept an input word and generate an output word. This can be defined by \emph{runs} of SFTs.
 An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i, 
@@
 
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an NFA $\mathcal{A}$.
+To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an SFA $\mathcal{A}$.
 More precisely,
@@
 For instance, the interval $[1, 9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
-The predicate $\lambda x \colon \mathbb{N}.~ x \text{ mod } 2 = 0$ denotes the set of even natural numbers.
+The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
 All these different forms of labels are abstracted as sets in our formalization.
@@
 
-Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA.
-The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively.
+Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA. The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively. The 
+result of the product operation is an $\varepsilon$SFA.
 But we need to explain the role of $\text{\texttt{F}}$.
@@
 In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels:
-For transitions in $\Delta_e$,
 
 \begin{enumerate}
-\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, the input character is $\varepsilon$.
-Consequently, the SFA remains in its current state, and the product transition produces the output $(f~\text{None})$.
+\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e.
+the input character is $\varepsilon$.
+Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output $(f~\text{None})$.
 
-\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$.
+\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$.
 The resulting output is $F~f~(\sigma_1\cap\sigma_2)$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
@@
 The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\Delta_t~M~q~\pi~q'$.
+The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\Delta_t~\texttt{M}~q~\pi~q'$.
 
@@
 This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $[1, 5]$ and $[3, 4]$ can be directly represented as the interval list $[(1, 2), (5, 5)]$.
+For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$.
 
@@
 Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some new operations: \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nfa\_accepting}, and \texttt{nft\_tranfun}.
-There are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
+Note that in the implementation, there are some refined operations: \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nfa\_accepting}, and \texttt {nft\_tranfun}.
+These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
 
@@
 
-A central algorithmic challenge in our implementation lies in the computation of transition sets.
+A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ and $\texttt{D2}$ in Figure \ref{fig-compute-nft-product}.
 As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
@@
     We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
-This function systematically processes transitions in the SFA, represented by the set \texttt{T}.
+For a transtion $(q, \alpha, f, q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
 For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
@@
       \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
-      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists an input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
+      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
 It is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
@@
 \begin{itemize}
-  \item $\texttt{str}$: The input string to be transformed
-  \item $\texttt{pattern}$: A regular expression defining the matching criteria
-  \item $\texttt{replacement}$: The string to be substituted for matched substrings
+  \item $\texttt{str}$: The input string to be transformed.
+  \item $\texttt{pattern}$: A regular expression defining the matching criteria.  
+  \item $\texttt{replacement}$: The string to be substituted for the matched substring.
 \end{itemize}
@@
 Since a constant string can be viewed as a specialized regular expression, we can construct its SFA representation.
-Convert this SNFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
+Convert this SFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
   \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
@@
    \path[->] 
-   (q0') edge [bend left] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
-   (q1') edge [loop above] node {\texttt{[0-9]}/$\texttt{f}$} (q1');
+   (q0') edge [bend left] node {\texttt{[0-9]},$\texttt{f}$} (q1')
+   (q1') edge [loop above] node {\texttt{[0-9]},$\texttt{f}$} (q1');
 \end{tikzpicture}
@@
 \begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
+\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
   % Original NFA for "ITP"
@@
   \path[->] 
-  (q0') edge node {\texttt{None}/\texttt{g 1}} (q1')
-  (q1') edge node {\texttt{None}/\texttt{g 2}} (q2')
-  (q2') edge node[below] {\texttt{None}/\texttt{g 3}} (q3');
+  (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
+  (q1') edge node {\texttt{None}, \texttt{(g 2)}} (q2')
+  (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
 \end{tikzpicture}
@@
 \begin{enumerate}
-  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with \texttt{None}/\texttt{f}) from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
-  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with $\Sigma/\texttt{f}$, where $\Sigma$ represents the full alphabet.
+  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
+  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{f}$", where $\Sigma$ represents the full alphabet.
+For the string solver, $\Sigma$ is the set of all unicode characters.
 These transitions enable the SFT to process arbitrary input prefixes and suffixes while preserving the matched substring for replacement
@@
 \begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
+  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
     % Second automaton from the first figure
@@
      \path[->] 
-     (q0') edge [bend left] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
-     (q1') edge [loop above] node {\texttt{[0-9]}/$\texttt{f}$} (q1')
-     (q0') edge [loop below] node {$\Sigma$/$\texttt{f}$} (q0');
+     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
+     (q1') edge [loop above] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
+     (q0') edge [loop below] node {$\Sigma$, $\texttt{f}$} (q0');
 
@@
     \path[->] 
-    (q0'') edge node {\texttt{None}/\texttt{g 1}} (q1'')
-    (q1'') edge node {\texttt{None}/\texttt{g 2}} (q2'')
-    (q2'') edge node {\texttt{None}/\texttt{g 3}} (q3'')
-    (q3'') edge [loop below] node {$\Sigma$/$\texttt{f}$} (q3'')
-    (q1') edge node {\texttt{None}/\texttt{f}} (q0'');
+    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
+    (q1'') edge node {\texttt{None}, \texttt{(g 2)}} (q2'')
+    (q2'') edge [below] node {\texttt{None}, \texttt{(g 3)}} (q3'')
+    (q3'') edge [loop below] node {$\Sigma$, $\texttt{f}$} (q3'')
+    (q1') edge node {\texttt{None}, \texttt{f}} (q0'');
   \end{tikzpicture}
@@
 In contrast, our semantics allows for a more general interpretation where any matching substring may be replaced.
+But it still replaces only one occurrence of them.
 
@@
 %\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
-While the core solving algorithm maintains the certification guarantees of CertiStr, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
+While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
 
@@
 As discussed in the previous subsection, our replacement semantics differs from the SMT-LIB standard semantics.
-When we run the SMT solver CVC5 \cite{cvc5} with the example code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
+Run the SMT solver CVC5 \cite{cvc5} with the code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
 
@@
 %
-An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
+An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests are finished in one minute) but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
 Specifically, when the string constraints do not satisfy the tree property, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
@@
 where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for each constraint: $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace}($x, p_1, r_1$).
+The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace}($x, p_1, r_1$).
 Each accumulation step requires computing the product of the current automaton with the previous result.
@@
 The last row corresponds to the effort of the interval formalization.
-The most difficult part is the correctness proof of the product operation at the abstract level.
+The most difficult part is the correctness proof of the product operation at the abstract level (Theorem \ref{fig-def-product-correct}).
 Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
@@
 This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
-Recent developments have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
+Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
 
@@
 
-To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}.
+To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}.
 This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.

### 2025-03-15 CertiStr/test/benchmark/replace_str/benchmark.sh
remove dune test in benchmark

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-for test in ./*.smt2; do
-    echo "Test ($test):"
-    result=$(dune exec CertiStr -- $test);
-    echo "$result"
-done

### 2025-03-29 Paper/main.tex
rerun test
update code

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, anonymous, thm-restate]{lipics-v2021}
+\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
 %This is a template for producing LIPIcs articles. 
@@
 \begin{abstract}
-Finite-state Automata (FAs) are fundamental components in the domains of programming languages.
+Finite Automata (FAs) are fundamental components in the domains of programming languages.
 For instance, regular expressions, which are pivotal in languages such as JavaScript and Python, are frequently implemented using FAs.
 %
-Finite-state Transducers (FTs) extend the capabilities of FAs by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
-Despite the widespread adoption of formalizing FAs and FTs within proof assistants such as Coq and Isabelle/HOL, these formalizations often fall short in terms of applicability to real-world scenarios.
-A significant limitation of classical finite-state models is that transition labels are typically defined as a single character from a finite alphabet.
-However, in practical applications, the alphabet of an FA or FT can be extensive or even \emph{infinite}.
-This classical approach to formalizing transitions can result in transition explosion, leading to critical performance bottlenecks of FA and FT operations.
+Finite Transducers (FTs) extend the capabilities of FAs by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
+Despite the various formalizations of FAs in proof assistants such as Coq and Isabelle/HOL, these formalizations often fall short in terms of applicability to real-world scenarios.
+A significant limitation of classical FAs and FTs is that transition labels are typically defined as a single character from a finite alphabet.
+However, in practical applications, the alphabet of an FA or FT can be enormously large or even \emph{infinite}.
+The classical approach to formalizing transitions can result in transition explosion, leading to critical performance bottlenecks of FA and FT operations.
 
@@
 
-Finite-state Automata (FA) and Finite-state Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
+Finite Automata (FA) and Finite Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
 For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
@@
 
-Symbolic Finite-state Automata (SFA) and Symbolic Finite-state Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
+Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
 %
-These symbolic models utilize transition labels defined by boolean algebras, allowing for more expressive representations.
+These symbolic models utilize transition labels defined by first-order predicates over boolean algebras, allowing for more expressive representations.
 For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \mod 2 = 0$), denoting all even numbers.
@@
 
-The formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
-Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
+However, the formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
+Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse predicate representations of boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
 
 Prior work of CertiStr \cite{cpp/KanLRS22} has successfully formalized SFAs within Isabelle/HOL, demonstrating both the efficiency and effectiveness of SFAs in practice.
-However, the formalization of FTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
-FTs constitute a significantly more expressive and powerful theoretical framework compared to symbolic FAs, as evidenced by their capability to model complex string transformations such as replacement operations.
+However, the formalization of SFTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
+SFTs constitute a significantly more expressive and powerful theoretical framework compared to  SFAs, as evidenced by their capability to model complex string transformations such as replacement operations.
 %
@@
 At the abstraction level, transition labels are formalized through the fundamental mathematical concept of \emph{sets}.
-This abstraction facilitates subsequent refinement to various boolean algebras, including intervals and arithmetic conditions, while maintaining theoretical consistency.
+This abstraction facilitates subsequent refinement to various representations of Boolean algebras, such as intervals and arithmetic predicates, while maintaining theoretical consistency.
 
 The key operation of our formalization is the product operation between an SFT and an input regular language.
-Specifically, given an SFA $\mathcal{A}$ representing a regular language and an SFT $\mathcal{T}$, we define the product operation $\mathcal{T} \times\mathcal{A}$ that characterizes the set of output languages generated by $\mathcal{T}$ when processing inputs from the language recognized by $\mathcal{A}$.
+Specifically, given an SFA $\mathcal{A}$ representing a regular language and an SFT $\mathcal{T}$, we define the product operation $\mathcal{T} \times\mathcal{A}$ that characterizes the output language generated by $\mathcal{T}$ when processing inputs from the language recognized by $\mathcal{A}$.
 
-In the refinement level, we implemented transition labels using an interval-based representation.
+In the refinement level, we implemented transition labels using an interval-based representation to examplify the refinement process.
 The formalization of interval algebra provides efficient set-theoretic operations—including membership checking, intersection, and difference computations—facilitating the refinement of transition labels from abstract sets to concrete intervals.
@@
 \begin{enumerate}
-\item We present the first formalization of symbolic finite-state transducers in the Isabelle/HOL proof assistant.
+\item We present the first formalization of symbolic finite transducers in Isabelle/HOL proof assistant.
 \item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodates various boolean algebras, ensuring adaptability to future transition label developments.
@@
 The remainder of this paper is organized as follows:
-Section \ref{sec:formalization} introduces the formal foundations of symbolic finite-state transducers.
+Section \ref{sec:formalization} introduces the formalization of symbolic finite transducers.
 Section \ref{sec:product-operation} presents the product operation at the abstract level.
@@
 
-\section{Formalization of Symbolic FTs}
+\section{Formalization of SFTs}
 \label{sec:formalization}
@@
 We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL. 
-The foundational concept underlying this formalization is the label theory, which characterizes the process by which output labels are derived from input labels.
+\subsection{The Mathematical Definition of SFTs}
 
@@
 
-\begin{definition}[Symbolic Finite-state Transducer]
+\begin{definition}[Symbolic Finite Transducer]
 \label{def-sft}
-   A Symbolic Finite-state Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
+   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
    \begin{itemize}
@@
 
-    Figure \ref{fig-example-ft} illustrates an SFT that accepts strings matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
+    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
 The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
@@
 	\begin{lstlisting}
-record (|$'q,$||$~'a,$| |$'b$|) |$\texttt{NFT}$| =
+record (|$'q,$||$~'a,$| |$'i,$| |$'b$|) |$\texttt{NFT}$| =
 	|$\mathcal{Q}_t$| :: "|$'q$| set"
-	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'b$|) LTTS"
+	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'i$|) LTTS"
 	|$\mathcal{I}_t$| :: "|$'q$| set"
 	|$\mathcal{F}_t$| :: "|$'q$| set"
-        |$\mathcal{M}_t$| :: "|$'b\Rightarrow 'a$| Tlabel"
-
-type_synonym |$('q, 'a, 'b)$| LTTS = "|$('q \times ('a \text{ set option }\times b) \times 'q)$| set"
-
-type_synonym |$'a$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
+        |$\mathcal{M}_t$| :: "|$'i\Rightarrow ('a,~'b)$| Tlabel"
+type_synonym |$('q,~'a,~'i)$| LTTS = "|$('q \times ('a \text{ set option }\times 'i) \times 'q)$| set"
+type_synonym |$('a,~'b)$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
 	\end{lstlisting}
-\caption{The formalization of $FTs$ in Isabelle/HOL}
+\caption{The formalization of $SFTs$ in Isabelle/HOL}
 \label{fig-def-FT}
@@
 %
-The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \text{ set option }\times b) \times 'q$.
+The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
 This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
@@
 But predicates have various different forms.
-For instance, the interval $[1, 9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
+For instance, the interval $[1-9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
 The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
@@
 
-The second element of type $'b$ serves as an index into the output function space.
+The second element of type $'i$ serves as an index into the output function space. 
+%
+We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
+%
 The mapping $\mathcal{M}_t$ associates each index with a specific output function.
@@
 
-\section{The Product Operation of NFTs}
+\section{The Product Operation of SFTs}
 \label{sec:product-operation}
@@
 An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
-Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of automata with $\varepsilon$-transitions ($\varepsilon$NFAs), we extend the framework with a formalization of symbolic $\varepsilon$NFAs (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
+Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of SFAs with $\varepsilon$-transitions, we extend the framework in CertiStr with a formalization of symbolic SFAs with $\varepsilon$-transitions (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
 In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
@@
 	\end{lstlisting}
-\caption{The formalization of $\varepsilon$NFA and SFA in Isabelle/HOL}
+\caption{The formalization of $\varepsilon$SFAs and SFAs}
 \label{fig-def-FAs}
@@
 	\begin{lstlisting}
-definition productT :: |$"('q,'a,'b,'c)"$| NFT |$\Rightarrow$| ('q,'a) NFA |$\Rightarrow$| 
+definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| |$('q,'a)$| NFA |$\Rightarrow$| 
            |$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}\Rightarrow 'b\text{ set option})\Rightarrow$|
@@
     |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-    |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-          |$(p, (\texttt{None}, f), q)\in \Delta_T \mathcal{T} \land \exists S. (\mathcal{M}~\mathcal{T})~f~\text{None} = \text{Some } S\} ~\cup$|
-        |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_T~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 = \emptyset \land$|
-            |$ \exists S.~F~((\mathcal{M}~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
-    |$\Delta_e' = \{((p,p'),~\text{the }(((\mathcal{M}~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-          |$(p, (None, f), q)\in \Delta_T~\mathcal{T} \land (\mathcal{M}~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
-        |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_T~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 = \emptyset \land$|
-            |$ \exists x\in(\sigma_1\cap\sigma_2).~((\mathcal{M}~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
+    |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$|
+          |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T} \land \exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
+        |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
+          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
+            |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
+    |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
+          |$(p, (None, f), q)\in \Delta_t~\mathcal{T} \land (\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
+        |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
+          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
+            |$ \exists x\in(\sigma_1\cap\sigma_2).~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
     |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
@@
 	\end{lstlisting}
-\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\caption{The formalization of product operation}
 \label{fig-def-FTProd}
@@
 result of the product operation is an $\varepsilon$SFA.
-But we need to explain the role of $\text{\texttt{F}}$.
+But we need to explain the role of parameter $\text{\texttt{F}}$.
 The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$.
-$\text{\texttt{F}}$ extends $f$ to apply it to a set of elements.
-More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as
+$\text{\texttt{F}}$ extends $f$ to apply $f$ to a set of elements.
+More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as follows:
 
-\[\texttt{F}~f~A=\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset)\].
+\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].
 
 The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions.
-In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels:
+In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels.
+Let us consider $\Delta_e$ first.
 
@@
 \item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e.
-the input character is $\varepsilon$.
-Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output $(f~\text{None})$.
+the input character is $\varepsilon$, and $f~\text{None} \neq \text{None}$.
+Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output "$((\mathcal{M}_t~\mathcal{T})~f)~\text{None}$".
+Remember that $f$ is just an index, $(\mathcal{M}_t~\mathcal{T})~f$ is the output function.
 
-\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$.
-The resulting output is $F~f~(\sigma_1\cap\sigma_2)$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
+\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
+The resulting output is $\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)$.
 \end{enumerate}
@@
 Given a trace $\pi$, we define two key projection functions:
-(1) \texttt{inputE}, which corresponds to $\text{in}(\pi)$ and extracts the input sequence
-(2) \texttt{outputE}, which corresponds to $\text{out}(\pi)$ and extracts the output sequence.
+(1) \texttt{inputE}, which corresponds to ${in}(\pi)$ and extracts the input sequence
+(2) \texttt{outputE}, which corresponds to ${out}(\pi)$ and extracts the output sequence.
 
 The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\Delta_t~\texttt{M}~q~\pi~q'$.
+The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
 
@@
 
-definition outputL :: "|$('q, 'a, 'b, 'c)\text{  NFT} \Rightarrow $|
-            |$ ('c \Rightarrow ('a, 'b)\text{  Tlabel}) \Rightarrow 
-                               ('q, 'a) \text{ NFA\_rec} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| M |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T} \land$|
-                |$ \text{LTTS\_reachable }(\Delta_t~\mathcal{T})~\text{M}~q~\pi~q'\land \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}" 
+definition outputL :: "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
+                               ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
+  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T} \land$|
+                |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}" 
 	\end{lstlisting}
-\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\caption{The formalization of traces in SFTs}
 \label{fig-def-output}
@@
   fixes |$\mathcal{T}$| |$\mathcal{A}$| F
-  assumes F_ok1: "|$\forall~f~s.~(\forall e \in s. f (\text{ Some } e) = \text{None}) \mapsto \text{ F}~f~s = \text{None}"$|
-      and F_ok2: "|$\forall~f~s.~\text{F }f~s = $|
-            |$\text{Some }(\cup \{\text{the }(f (\text{ Some }e))\mid e. e \in s \land f~(\text{Some }e) \neq \text{ None}\})$|"
+  assumes F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None}) \longleftrightarrow\text{ F}~f~s = \text{None}"$|
+      and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s \land f~(\text{Some }e) = \text{ Some S}\})$|"
       and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-    |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL |$\mathcal{T}~(\mathcal{M}~\mathcal{T})~\mathcal{A}$|"
+    |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL |$\mathcal{T}~\mathcal{A}$|"
 
 	\end{lstlisting}
-\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\caption{The correctness lemma of the product operation}
 \label{fig-def-product-correct}
@@
 The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
+The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.
 The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.
@@
 
-An interval is defined as a pair $(i, j)$ representing the set $\{e \mid i \leq e \leq j\}$.
+An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
 To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
 This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$.
+For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
 
@@
 \begin{enumerate}
-  \item $\texttt{semI}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semI}~(i,j) = \{e \mid i \leq e \leq j\}$.
-  \item $\texttt{emptyI}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semI}~i = \emptyset$.
-  \item $\texttt{nemptyI}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semI}~i \neq \emptyset$.
-  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semI}~i = \texttt{semI}~i_1 \cap \texttt{semI}~i_2$.
-  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semI}~i = \texttt{semI}~i_1 \setminus \texttt{semI}~i_2$.
+  \item $\texttt{semIs}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semIs}~(i,j) = \{e \mid i \leq e \leq j\}$.
+  \item $\texttt{emptyIs}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semIs}~i = \emptyset$.
+  \item $\texttt{nemptyIs}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semIs}~i \neq \emptyset$.
+  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \cap \texttt{semIs}~i_2$.
+  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \setminus \texttt{semIs}~i_2$.
 \end{enumerate}
@@
     F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) (nfa_accepting |$\mathcal{A}$|);
-    RETURN (Q,D1,D2,I, F)
+    RETURN (Q, D1, D2, I, F)
   }"
 \end{lstlisting}
-\caption{The computation of NFT product}
+\caption{The computation of SFT product}
 \label{fig-compute-nft-product}
@@
 \end{lstlisting}
-\caption{The formalization of $\varepsilon-$NFA and NFA in Isabelle/HOL}
+\caption{The computation of Cartesian product of two state sets}
 \label{fig-def-prods_imp}
@@
         else 
-          (subtrans_comp M q (the |$\alpha$|) f q|$'$| F T2 D1 D2)
+          (subtrans_comp M q (the |$\alpha$|) f q|$'$| F fe T2 D1 D2)
       )) (|$\emptyset$|, |$\emptyset$|)"
 \end{lstlisting}
-\caption{The computation of transition sets}
+\caption{The computation of \texttt{trans\_comp\_imp}}
 \label{fig-def-prods-imp}
@@
 
-A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ and $\texttt{D2}$ in Figure \ref{fig-compute-nft-product}.
+A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ (corresponding to $\Delta_e$) and $\texttt{D2}$ (corresponding to $\Delta_e'$) in Figure \ref{fig-compute-nft-product}.
 As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
@@
           D1 |$\leftarrow$| (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) |$\neq$| None) 
-            then RETURN 
-            let |$\alpha_i$| = (the (F (M f) (intersectIs |$\alpha$| |$\alpha'$|))) in
-            ({(q,q1) |$\alpha_i$| (q|$'$|,q1|$'$|)} |$\cup$| D1)
+            then  
+            let |$\alpha_i$| = the (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) in
+            RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} |$\cup$| D1
             else RETURN D1);
-          D2 |$\leftarrow$| (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) |$\neq$| |$\emptyset$| then 
-            RETURN (((q,q1), (q|$'$|,q1|$'$|)) |$\cup$| D2) else RETURN D2);
+          D2 |$\leftarrow$| (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) then 
+            RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 else RETURN D2);
           RETURN (D1, D2)
@@
     \end{lstlisting}
-    \caption{The computation of transition sets}
+    \caption{The computation of \texttt{subtrans\_comp}}
     \label{fig-def-subtrans_comp}
@@
     We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
-For a transtion $(q, \alpha, f, q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
+For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
 For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
@@
       \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
-It is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
+The checking is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
     \end{enumerate}
@@
   \end{lstlisting}
-  \caption{The refinement of the product computation}
+  \caption{The refinement relation between \texttt{productT\_imp} and \texttt{productT}}
   \label{fig-def-productT_imp_correct}
@@
   The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$.
-More precisely,  the concrete implementation must produce an SFA that is structurally equivalent to the one produced by the abstract algorithm \texttt{productT}.
+More precisely,  the concrete implementation must produce an $\varepsilon$SFA that is structurally equivalent or isomorphic to the one produced by the abstract algorithm \texttt{productT}.
 
-  To establish this equivalence, we must prove that the SFAs produced by \texttt{productT\_imp} and \texttt{productT} are identical in all essential components:
-  initial state sets, transition relations, $\varepsilon$-transition sets, accepting state sets.
+  To establish this equivalence, we must prove that the $\varepsilon$SFAs produced by \texttt{productT\_imp} and \texttt{productT} are isomorphic in all essential components: the set of states, transition relations, $\varepsilon$-transition relations, initial and accepting state sets.
 
@@
 
-\section{An Application to String Solver}
+\section{An Application to String Solving}
 \label{sec-app-str-solver}
@@
   \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
-These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and empty output function, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
+These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and the identity function $\texttt{id}=\lambda x.~x$, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
 \end{enumerate}
@@
   \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
-  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{f}$", where $\Sigma$ represents the full alphabet.
+  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{id}$", where $\Sigma$ represents the full alphabet.
 For the string solver, $\Sigma$ is the set of all unicode characters.
@@
      (q1') edge [loop above] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q0') edge [loop below] node {$\Sigma$, $\texttt{f}$} (q0');
+     (q0') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q0');
 
@@
     (q2'') edge [below] node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q3'') edge [loop below] node {$\Sigma$, $\texttt{f}$} (q3'')
+    (q3'') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q3'')
     (q1') edge node {\texttt{None}, \texttt{f}} (q0'');
@@
 
-We have implemented CertiStrR, an extension of the Certified String Solver CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
+We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
 %\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
@@
 The benchmarks are divided into two categories based on the replacement operations: (1) string-based replacement (\texttt{str.replace}) and (2) regular expression-based replacement (\texttt{str.replace\_re}).
-Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB specification, we preprocess certain operations.
+Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB language, we preprocess certain operations.
 For instance, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
@@
       \texttt{replace\_str} & 142 & 173 & 8 & 0.27 & 323\\
-      \texttt{replace\_re} & 57 & 4 & 30 & 0.37 & 98\\
+      \texttt{replace\_re} & 84 & 4 & 10 & 0.36 & 98\\
       \bottomrule
   \end{tabular}
-  \caption{Performance metrics for string operations}
+  \caption{Experimental results}
   \label{tab:string_operations}
@@
 The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a time limit of one minute per test. 
-The results show average execution times of 0.27 seconds for \texttt{str.replace} and 0.37 seconds for \texttt{str.replace\_re} operations.
+The results show average execution times of 0.27 seconds for \texttt{str.replace\_str} and 0.36 seconds for \texttt{str.replace\_re} operations.
 Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
@@
 An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests are finished in one minute) but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
-Specifically, when the string constraints do not satisfy the tree property, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
+Specifically, when the string constraints do not satisfy the tree property defined in \cite{cpp/KanLRS22}, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
 
 Our performance analysis revealed that while the SFT-based replacement operation modeling is efficient, the primary computational bottleneck stems from automata accumulation during  forward-propagation.
-Consider the following constraint system:
+Consider the following string constraints:
 \[
-x = x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r);~x = x_4;~x = \texttt{replace}(x, p_1, r_1);
+x = x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r);~x = x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1);
 \]
 where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace}($x, p_1, r_1$).
+The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
 Each accumulation step requires computing the product of the current automaton with the previous result.
-Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state and transition explosion.
+Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state explosion.
 
@@
 The last row corresponds to the effort of the interval formalization.
-The most difficult part is the correctness proof of the product operation at the abstract level (Theorem \ref{fig-def-product-correct}).
+The most difficult part is the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
 Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
@@
 
-\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer,popl/VeanesHLMB12, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
+\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
 This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
@@
 
-\emph{Applications of Symbolic Automata and Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
+\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
 These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
@@
 
-In this paper, we have presented the first formalization of Symbolic Finite Transducers (SFTs) in the proof assistant Isabelle/HOL.
-Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both input and output, and extensibility to various boolean algebras (including intervals and arithmetic operations) via the refinement framework.
+In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL.
+Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.
 

### 2025-07-20 Paper/main.tex
minor
update the paper

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
+\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %This is a template for producing LIPIcs articles. 
@@
 
-\title{Certified Symbolic Transducer with Applications in String Solving} %TODO Please add
-
 %\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
 
-\author{Shuanglong Kan}{Barkhausen Institut, Germany \and \url{https://github.com/ShlKan} }{shuanglongkan@gmail.com}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
-Please provide at least the name of the affiliation and the country.
-The full address is optional.
-Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.
-
-\authorrunning{S. Kan} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
-
-\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
-
-\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
-
-\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords
-
-\category{} %optional, e.g. invited paper
-
-\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
-%\relatedversiondetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g.
-Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional
-
-%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
-%\supplementdetails[linktext={opt.
-text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g.
-Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional
-
-%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors.
-Please enter author specific funding statements as fifth argument of the \author macro.
-
-\acknowledgements{I want to thank \dots}%optional
-
-%\nolinenumbers %uncomment to disable line numbering
-
-%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\EventEditors{John Q. Open and Joan R. Access}
-\EventNoEds{2}
-\EventLongTitle{16th Conference on
-Interactive Theorem Proving (ITP 2025)}
-\EventShortTitle{ITP 2025}
-\EventAcronym{ITP}
-\EventYear{2025}
-\EventDate{July 24--27, 2025}
-\EventLocation{Little Whinging, United Kingdom}
-\EventLogo{}
-\SeriesVolume{42}
-\ArticleNo{23}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\special{papersize=8.5in,11in}
+\setlength{\pdfpageheight}{\paperheight}
+\setlength{\pdfpagewidth}{\paperwidth}
+
+\title{Certified Symbolic Finite Transducers: Formalization and Applications to String Manipulation}
+
+\begin{abstract}
+Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
+FTs have wide applications, including program analysis, string constraint solving, and security-critical sanitizer analysis.
+However, there is currently no scalable formalization of FTs.
+This is primarily because transition labels in FTs are not formalized symbolically, resulting in transition explosion and significant performance bottlenecks in practical applications.
+
+In this paper, we present a formalization of FTs in the symbolic setting using Isabelle/HOL, where transition labels can be instantiated with various Boolean algebras, such as intervals and arithmetic predicates.
+This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
+%
+To evaluate the effectiveness of our formalization, we apply the formalized SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for preventing XSS attacks, and (2) a string solver that models replacement operations with various semantics, including left-most and replace-all matching.
+Experimental results demonstrate that our approach achieves computational efficiency in constraint-solving scenarios.
+\end{abstract}
+
 \maketitle
-
-%TODO mandatory: add short abstract of the document
-\begin{abstract}
-Finite Automata (FAs) are fundamental components in the domains of programming languages.
-For instance, regular expressions, which are pivotal in languages such as JavaScript and Python, are frequently implemented using FAs.
-%
-Finite Transducers (FTs) extend the capabilities of FAs by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
-Despite the various formalizations of FAs in proof assistants such as Coq and Isabelle/HOL, these formalizations often fall short in terms of applicability to real-world scenarios.
-A significant limitation of classical FAs and FTs is that transition labels are typically defined as a single character from a finite alphabet.
-However, in practical applications, the alphabet of an FA or FT can be enormously large or even \emph{infinite}.
-The classical approach to formalizing transitions can result in transition explosion, leading to critical performance bottlenecks of FA and FT operations.
-
-A more pragmatic approach involves the formalization of symbolic FAs \cite{cav/DAntoniV17} and FTs \cite{VeanesHLMB12Transducer}, where transition labels are symbolic and potentially infinite.
-While the formalization of symbolic FAs has been explored in the work of CertiStr \cite{cpp/KanLRS22}, the formalization of symbolic FTs in interactive proof assistants remains largely unexplored due to the increased complexity challenges.
-%
-In this paper, we aim to formalize symbolic FTs within the Isabelle/HOL framework.
-This formalization is refinement-based and is designed to be extensible with various symbolic representations of transition labels.
-To assess its performance, we applied the formalized symbolic FTs to an SMT string solver for modeling replacement operations.
-The experimental results indicate that the formalized symbolic transducer can efficiently and effectively solve string constraints with replacement operations.
-
-\end{abstract}
 
@@
 \item We present the first formalization of symbolic finite transducers in Isabelle/HOL proof assistant.
-\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodates various boolean algebras, ensuring adaptability to future transition label developments.
+\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
 \item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
@@
 \item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
-\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, (f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
+\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, $ $(f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
 \end{itemize}
@@
 \[
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land in(\pi) = w \land out(\pi) = w'\}.
+\begin{split}
+\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
+\end{split}
 \]
@@
 \[
-\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
+\begin{split}
+\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \\\text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
+\end{split}
 \]
@@
   \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=3.5cm, on grid, auto]
+  \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm and 2cm, on grid, auto]
+    % Make q0 and q1 horizontally wide apart
     \node[state, initial] (q_0)   {$q_0$}; 
-    \node[state] (q_1) [right=of q_0] {$q_1$}; 
-    \node[state, accepting] (q_2) [right=of q_1] {$q_2$};
+    \node[state] (q_1) [right=4cm of q_0] {$q_1$}; 
+    \node[state, accepting] (q_2) [below=2.2cm of q_1] {$q_2$};
 
@@
     (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_1) edge node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
+    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
     (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_2) edge[loop above] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
+    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
   \end{tikzpicture}
@@
 \subsection{The Isabelle/HOL Formalization of SFTs}
+
+Firstly, we present the basic idea of our formalization of SFTs in Isabelle/HOL, which consists of three layers.
+
+\begin{figure}[hbt!]
+\centering
+\begin{tikzpicture}[
+  node distance=1cm,
+  every node/.style={font=\small},
+  layer/.style={rectangle, draw, rounded corners, fill=blue!10, minimum width=5.5cm, minimum height=1.2cm, align=center},
+  arrow/.style={->, thick}
+]
+
+% Abstract Layer
+\node[layer] (abstract) {%
+  \textbf{Abstract Layer}\\
+  FT with transition labels as \textbf{sets}
+};
+
+% Locale Layer
+\node[layer, below=of abstract] (locale) {%
+  \textbf{Implementation Layer}\\
+  FT with transition labels as a \\boolean algebra locale
+};
+
+% Implementation Layer
+\node[layer, below=of locale] (impl) {%
+  \textbf{Boolean Algebra Layer}\\
+  Concrete Boolean algebra implementations\\
+  (e.g., intervals, bitvectors, predicates)
+};
+
+% Draw arrows with refinement labels
+\draw[arrow] (abstract) -- node[right] {refinement} (locale);
+\draw[arrow] (locale) -- node[right] {refinement} (impl);
+
+\end{tikzpicture}
+\caption{Three-layered formalization of SFTs in Isabelle/HOL. 
+\textbf{Abstract Layer:} Transition labels are modeled as sets, representing the most general case.
+\textbf{Locale Layer:} Transition labels are abstracted as a type variable \texttt{'a} within a locale, supporting any Boolean algebra.
+\textbf{Implementation Layer:} Concrete Boolean algebra instances, such as intervals, bitvectors, or predicates, are provided.}
+\label{fig:layered-formalization}
+\end{figure}
 
@@
 	\begin{lstlisting}
-definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| |$('q,'a)$| NFA |$\Rightarrow$| 
-           |$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}\Rightarrow 'b\text{ set option})\Rightarrow$|
-           |$('q\times 'q, ~'b)$| eNFA where
+definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| 
+   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$| 
+   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
 "productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
   |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-    |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$|
-          |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T} \land \exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
-        |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
+  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
+       |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
+        |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
+      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
+       |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
+       |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
           |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
   |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-          |$(p, (None, f), q)\in \Delta_t~\mathcal{T} \land (\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
+       |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
+       |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
       |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-          |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
-            |$ \exists x\in(\sigma_1\cap\sigma_2).~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
+        |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}$|
+        |$\land~\sigma_1\cap \sigma_2 \neq \emptyset~\land\exists x\in(\sigma_1\cap\sigma_2).$|
+        |$~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
   |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
@@
 	\begin{lstlisting}
-fun inputE :: |$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$| where
+fun inputE :: 
+   "|$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$|"
+where
   "inputE [] = []" |$\mid$|
-  "inputE ((Some a, |$\_$|) |$\#$| l) = a |$\#$| (inputE l)" |$\mid$|
+  "inputE ((Some a, |$\_$|) |$\#$| l) = 
+                      a |$\#$| (inputE l)" |$\mid$|
   "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"
 
-fun outputE :: "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" where
+fun outputE :: 
+    "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" 
+where
   "outputE [] = []" |$\mid$|
-  "outputE ((_, Some a) # l) = a # (outputE l)" |$\mid$|
-  "outputE ((_, None) # l) = (outputE l)"
+  "outputE ((_,Some a) # l) = 
+                    a # (outputE l)" |$\mid$|
+  "outputE ((_,None) # l) = (outputE l)"
 
-definition outputL :: "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
+definition outputL :: 
+   "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
                                ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T} \land$|
-                |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}" 
+  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. 
+    |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T}~\land$| 
+    |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land$||$ \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}"
 	\end{lstlisting}
@@
   fixes |$\mathcal{T}$| |$\mathcal{A}$| F
-  assumes F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None}) \longleftrightarrow\text{ F}~f~s = \text{None}"$|
-      and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s \land f~(\text{Some }e) = \text{ Some S}\})$|"
+  assumes 
+   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
+       |$\text{ F}~f~s = \text{None}"$|
+   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
+        |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
+        |$f~(\text{Some }e) = \text{ Some S}\})$|"
       and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-    |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL |$\mathcal{T}~\mathcal{A}$|"
+  |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL 
+          |$\mathcal{T}~\mathcal{A}$|"
 
@@
 To facilitate formal reasoning and optimize performance, we introduce a canonical form for interval.
-An interval $[(i_1, j_1), \ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
+An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
 \begin{enumerate}
@@
 Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations: \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nfa\_accepting}, and \texttt {nft\_tranfun}.
+Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}.
 These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
@@
   "productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
-    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) (nfa_states |$\mathcal{A}$|);
-    (D1, D2) |$\leftarrow$| trans_comp_imp (nft_tranfun |$\mathcal{T}$|) F fe 
-        (nft_trans |$\mathcal{T}$|) (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
-    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|) (nfa_initial |$\mathcal{A}$|);
-    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) (nfa_accepting |$\mathcal{A}$|);
+    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) 
+    (nfa_states |$\mathcal{A}$|);
+    (D1, D2) |$\leftarrow$| trans_comp_imp 
+     (nft_tranfun |$\mathcal{T}$|) F fe (nft_trans |$\mathcal{T}$|) 
+      (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
+    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|)
+      (nfa_initial |$\mathcal{A}$|);
+    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) 
+      (nfa_accepting |$\mathcal{A}$|);
     RETURN (Q, D1, D2, I, F)
@@
   FOREACH {t. t |$\in$| T1}
-      (|$\lambda$| (q, (|$\alpha$|, f), q|$'$|) (D1, D2). 
-        (if (|$\alpha$| = None) then 
-          (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe T2 D1 D2)
+   (|$\lambda$|(q,(|$\alpha$|,f),q|$'$|) (D1,D2). 
+    (if (|$\alpha$|=None) then 
+     (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe 
+                              T2 D1 D2)
      else
-          (subtrans_comp M q (the |$\alpha$|) f q|$'$| F fe T2 D1 D2)
-      )) (|$\emptyset$|, |$\emptyset$|)"
+     (subtrans_comp M q (the |$\alpha$|) f 
+            q|$'$| F fe T2 D1 D2))) (|$\emptyset$|, |$\emptyset$|)"
 \end{lstlisting}
@@
 "subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
-    FOREACH
-      {t. t |$\in$| T} (|$\lambda$| (q1, |$\alpha'$|, q1|$'$|) (D1, D2).
-        (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then do {
-          D1 |$\leftarrow$| (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) |$\neq$| None) 
+ FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
+  (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then
+    do {
+      D1 |$\leftarrow$| 
+       (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) 
+          |$\neq$| None) then
+         let |$\alpha_i$| = the (F (M f) 
+              (intersectIs |$\alpha$| |$\alpha'$|)) in
+         RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
+                |$\cup$| D1
+        else RETURN D1);
+      D2 |$\leftarrow$| 
+       (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) 
         then 
-            let |$\alpha_i$| = the (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) in
-            RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} |$\cup$| D1
-            else RETURN D1);
-          D2 |$\leftarrow$| (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) then 
-            RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 else RETURN D2);
-          RETURN (D1, D2)
-        }
+         RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 
+        else RETURN D2);
+      RETURN (D1, D2) }
    else (RETURN (D1, D2)))) (D1, D2)"
@@
       and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
-  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe} \leq \text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
+  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe}\leq$|
+            |$\text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
   \end{lstlisting}
@@
 
-The string replacement operation, denoted as $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
+The string replacement operation, denoted as 
+$\texttt{replace}(\texttt{str}, \newline\texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
 \begin{itemize}
@@
 We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}(s, $\texttt{/[0-9]+/}$, $"$\texttt{NUM}$"$)$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
+Given a replacement operation $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
 We model this replacement operation by an SFT with the following 3 steps:
@@
 \begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
+  % Left: SFA for [0-9]+ (vertical)
   \node[state, initial] (q0)   {$q_0$}; 
-  \node[state, accepting] (q1) [right=of q0] {$q_1$};
+  \node[state, accepting, below=of q0] (q1) {$q_1$};
 
    \path[->] 
-   (q0) edge [bend left] node {\texttt{[0-9]}} (q1)
-   (q1) edge [loop above] node {\texttt{[0-9]}} (q1);
+   (q0) edge [bend left=20] node[right] {\texttt{[0-9]}} (q1)
+   (q1) edge [loop below] node {\texttt{[0-9]}} (q1);
 
-  \node[state, initial] (q0') [right=4cm of q1] {$q_0'$}; 
-  \node[state, accepting] (q1') [right=of q0'] {$q_1'$};
+  % Right: SFT for [0-9]+ (vertical)
+  \node[state, initial, right=4cm of q0] (q0') {$q_0'$}; 
+  \node[state, accepting, below=of q0'] (q1') {$q_1'$};
 
    \path[->] 
-   (q0') edge [bend left] node {\texttt{[0-9]},$\texttt{f}$} (q1')
-   (q1') edge [loop above] node {\texttt{[0-9]},$\texttt{f}$} (q1');
+   (q0') edge [bend left=20] node[right] {\texttt{[0-9]},$\texttt{f}$} (q1')
+   (q1') edge [loop below] node {\texttt{[0-9]},$\texttt{f}$} (q1');
 \end{tikzpicture}
@@
 \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-  % Original NFA for "ITP"
-  \node[state, initial] (q0)   {$q_0$}; 
-  \node[state] (q1) [right=of q0] {$q_1$}; 
-  \node[state] (q2) [below=of q0] {$q_2$}; 
-  \node[state, accepting] (q3) [right=of q2] {$q_3$};
+  % SFA for "NUM" (top)
+  \node[state, initial above] (q0)   {$q_0$}; 
+  \node[state] (q1) [right=1cm of q0] {$q_1$}; 
+  \node[state] (q2) [below=1cm of q1] {$q_2$}; 
+  \node[state, accepting] (q3) [below=1cm of q0] {$q_3$};
 
@@
 
-  % Duplicate NFA for "ITP" on the right
-  \node[state, initial] (q0') [right=4cm of q1] {$q_0'$}; 
+  % SFT for "NUM" (right)
+  \node[state, initial above] (q0') [right=1.5cm of q1] {$q_0'$}; 
   \node[state] (q1') [right=of q0'] {$q_1'$}; 
-  \node[state] (q2') [below=of q0'] {$q_2'$}; 
-  \node[state, accepting] (q3') [right=of q2'] {$q_3'$};
+  \node[state] (q2') [below=1cm of q1'] {$q_2'$}; 
+  \node[state, accepting] (q3') [below=1cm of q0'] {$q_3'$};
 
@@
   (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
-  (q1') edge node {\texttt{None}, \texttt{(g 2)}} (q2')
+  (q1') edge node[left] {\texttt{None}, \texttt{(g 2)}} (q2')
   (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
@@
     % Second automaton from the second figure
-    \node[state] (q0'') [right=3cm of q1'] {$q_0''$}; 
+    \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
     \node[state] (q1'') [right=of q0''] {$q_1''$}; 
-    \node[state] (q2'') [below=of q0''] {$q_2''$}; 
-    \node[state, accepting] (q3'') [right=of q2''] {$q_3''$};
+    \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
+    \node[state] (q2'') [right=of q3''] {$q_2''$};
 
@@
     (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
-    (q1'') edge node {\texttt{None}, \texttt{(g 2)}} (q2'')
-    (q2'') edge [below] node {\texttt{None}, \texttt{(g 3)}} (q3'')
+    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
+    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
     (q3'') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q3'')
-    (q1') edge node {\texttt{None}, \texttt{f}} (q0'');
+    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
   \end{tikzpicture}
-  \caption{The SFT for the replacement operation $\texttt{replace}(s, $\texttt{/[0-9]+/}$, $\text{NUM}$)$}
+  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
   \label{fig-rearranged-automata}
@@
   (assert (= b "2024,NUM"))
-  (assert (= b (str.replace_re a (re.+ (re.range "0" "9")) "NUM")))
+  (assert (= b (str.replace_re a 
+      (re.+ (re.range "0" "9")) "NUM")))
   (check-sat)
@@
   \centering
-  \begin{tabular}{lccccc}
+  \small
+  \begin{tabular}{lcccc}
       \toprule
-      & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inconclusive} & \textbf{Time} & \textbf{Number of Tests} \\
+      & \textbf{SAT/UNSAT/Inc} & \textbf{Time (s)} & \textbf{Tests} \\
       \midrule
-      \texttt{replace\_str} & 142 & 173 & 8 & 0.27 & 323\\
-      \texttt{replace\_re} & 84 & 4 & 10 & 0.36 & 98\\
+      \texttt{replace\_str} & 142/173/8 & 0.27 & 323\\
+      \texttt{replace\_re} & 84/4/10 & 0.36 & 98\\
       \bottomrule
@@
 Consider the following string constraints:
-\[
-x = x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r);~x = x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1);
-\]
+\begin{align}
+x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber \\
+x &= x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
+\end{align}
 where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
+The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}\newline($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
 Each accumulation step requires computing the product of the current automaton with the previous result.
@@
   \centering
+  \small
   \begin{tabular}{lccc}
       \toprule
-      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs} (lines of code) \\
+      & \textbf{Defs} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
       \midrule
-      Abstract-level & 17 & 21 & 3274 \\
-      Implementation-level & 51 & 43 & 2700 \\
+      Abstract & 17 & 21 & 3274 \\
+      Implementation & 51 & 43 & 2700 \\
       Interval & 15 & 29 & 1500 \\

### 2025-07-20 Paper/main.tex
fix the figure

@@
 We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
+Given a replacement operation $\texttt{replace}\newline(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
 We model this replacement operation by an SFT with the following 3 steps:
@@
      (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop above] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
+     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
      (q0') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q0');
@@
     (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q3'') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q3'')
+    (q3'') edge [loop left] node {$\Sigma$, $\texttt{id}$} (q3'')
     (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');

### 2025-07-23 Paper/main.tex
add paper

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\lstdefinelanguage{Isabelle}{
+    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
+              where,fixes,assumes,shows,if,then,else,let,in,case,of,
+              record,datatype,type_synonym,fun,function,primrec},
+    sensitive=true,
+    morecomment=[s]{(*}{*)},
+    morestring=[b]",
+  }
 
 \lstset{
+  language=Isabelle,
 	basicstyle=\ttfamily,
 	keywordstyle=\color{blue}\bfseries,
-	keywords={definition,if,then,else, where, record, fun, lemma, type_synonym, fixes, shows, assumes, shows},
-	escapeinside={|}{|}
+	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and},
+	escapeinside={|}{|},
+  sensitive=true
 }
@@
 \begin{figure}[hbt!]
-	\begin{lstlisting}
+\begin{lstlisting}[language=SMTLIB]
 definition subtrans_comp where

### 2025-07-29 Paper/main.tex
add experiment result for sanitizer
add figures

@@
 
-\section{An Application to String Solving}
-\label{sec-app-str-solver}
-
-In this section, we demonstrate the practical application of our formalized SFTs to string constraint solving, specifically focusing on replacement operations.
-While CertiStr \cite{cpp/KanLRS22} provides a verified framework for string constraint solving using SFAs, its capabilities are inherently limited by the expressiveness of SFAs.
-In particular, SFAs cannot directly model string transformation operations such as replacement.
-
-Our string solver is based on CertiStr \cite{cpp/KanLRS22} by extending it with the modeling of replacement operations.
-CertiStr exploits forward-propagation to solve the string constraints.
-Our SFT modeling for replacement operations can seamlessly be integrated into CertiStr's forward-propagation framework.
+\section{Applications and Evaluation}
+
+We evaluate our formalization and implementation of the SFT product operation in two key application domains: string sanitization and string replacement operations in string solving.
+These applications demonstrate the practical utility of our certified SFT product operation in real-world scenarios.
+\subsection{Sanitizer}
+
+One Important application of our formalization is in the context of string sanitization, specifically for web applications.
+Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before being rendered in a web browser.
+Listing \ref{lst-html-encode} presents a C\# code snippet that implements an HTML encoding function, which is a sanitization operation.
+The function iterates through each character in the input string and encodes characters that are not safe for HTML rendering.
+
+\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
+static string EncodeHtml(string t)
+{
+  if (t == null) { return null; }
+  if (t.Length == 0) {
+    return string.Empty;
+  }
+  StringBuilder builder = new
+    StringBuilder("", t.Length * 2);
+  foreach (char c in t) {
+    if ((((c > ''') && (c < '{')) |$\mid\mid$|
+         ((c > '@') && (c < '['))) |$\mid\mid$|
+         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
+         (((c == '.') || (c == ',')) |$\mid\mid$|
+         ((c == '-') || (c == '_'))))) {
+          builder.Append(c);
+      }
+      else {
+          builder.Append("&#" +
+          ((int) c).ToString() + ";");
+      }
+  }
+  return builder.ToString();
+}
+\end{lstlisting}
+
+Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
+The transducer accepts any input string and produces an output where safe characters are preserved using the identity function $\texttt{id}$, while unsafe characters are encoded as HTML entities.
+The set of safe characters is defined as $\texttt{safe\_chars} = [\texttt{' '}, \texttt{','}, \texttt{'.'}, \texttt{'-'}, \texttt{'\_'}] \cup [\texttt{'0'}-\texttt{'9'}] \cup [\texttt{'A'}-\texttt{'Z'}] \cup [\texttt{'a'}-\texttt{'z'}]$, which can be efficiently represented using the interval algebra we formalized.
+Unsafe characters are transformed using the encoding function $\texttt{encode}(c) = \texttt{``\&\#''} + \texttt{toString}(\texttt{ord}(c)) + \texttt{``\;''}$.
+When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
+  % States
+  \node[state, initial, accepting] (q0) {$q_0$};
+  \node[state] (q1) [right=3.5cm of q0] {$q_1$};
+  \node[state] (q2) [below=of q0] {$q_2$};
+
+  \path[->] 
+    % Transition for safe characters
+    (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
+    % Transition for unsafe characters  
+    (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
+    % Epsilon transitions back to q0
+    (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
+    (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
+\end{tikzpicture}
+\caption{SFT representation of the HTML encoding sanitizer function.}
+\label{fig:html-sanitizer-sft}
+\end{figure}
+
+We evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT with SFAs constructed from randomly generated Unicode user inputs.
+This experimental setup reflects realistic usage scenarios where diverse character sets and input patterns are encountered.
+
+Figure \ref{fig:benchmark-results} demonstrates the scalability characteristics of our SFT product implementation.
+The computation requires approximately 1 second for 100,000-character inputs and 10 seconds for 1,000,000-character inputs.
+These performance metrics confirm that our certified implementation maintains computational efficiency suitable for practical string processing applications at scale.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.45]{benchmark_loglog.pdf}
+\caption{Experimental results for the sanitizer product}
+\label{fig:benchmark-results}
+\end{figure}
+
+Beyond performance evaluation, our SFT formalization enables formal verification of string sanitizer correctness.
+For example, this verification framework models user input as an SFA, the sanitizer as an SFT, and potential attacks as an SFA accepting unsafe strings.
+A sanitizer is verified correct if the product of the sanitizer SFT and user input SFA produces no outputs that intersect with the attack SFA's language.
+This correctness property can be checked by computing the intersection of the product SFT's output language with the attack SFA, where emptiness of this intersection guarantees sanitizer safety.
 
@@
 
+  The modeling of the aformentioned replacement operation will replace one occurrence of the substring that matches the regular expression pattern.
+We may also want to model left-most replacement and replace all occurrences of the substring that matches the regular expression pattern.
+
+  One way to model left-most replacement is illustrated in the following.
+
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
+  % First SFT rectangle
+  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
+  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
+
+  % States inside first SFT
+  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
+  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
+  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
+
+  % Second SFT rectangle
+  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
+  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
+
+  % States inside second SFT
+  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
+  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
+  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
+
+  % Epsilon transition from accepting state of first SFT to initial state of second SFT
+  \path[->] (p1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
+
+\end{tikzpicture}
+\caption{The modeling of left-most replacement operation.}
+\label{fig:replace-left-most}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
+  % First SFT rectangle
+  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
+  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
+
+  % States inside first SFT
+  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
+  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
+  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
+
+  % Second SFT rectangle
+  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
+  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
+
+  % States inside second SFT
+  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
+  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
+  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
+
+  % Epsilon transition from accepting state of first SFT to initial state of second SFT
+  \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
+
+  % Epsilon transition from accepting state of second SFT to initial state of first SFT
+  \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
+
+\end{tikzpicture}
+\caption{The modeling of replace all operation.}
+\label{fig:replace-all}
+\end{figure}
+
 \subsection{Experiments}
@@
 
-\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}]
+\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}, float=htbp]
   (set-logic QF_S)

### 2025-07-30 Paper/main.tex
paper polish

@@
 	keywordstyle=\color{blue}\bfseries,
-	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and},
+	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and, locale},
 	escapeinside={|}{|},
@@
 
-\section{Formalization of SFTs}
+\section{Symbolic Finite Transducers}
 \label{sec:formalization}
@@
 We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL.
-\subsection{The Mathematical Definition of SFTs}
 
@@
 \begin{split}
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
+\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\ \mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
 \end{split}
@@
 
-\subsection{The Isabelle/HOL Formalization of SFTs}
-
-Firstly, we present the basic idea of our formalization of SFTs in Isabelle/HOL, which consists of three layers.
-
-\begin{figure}[hbt!]
-\centering
-\begin{tikzpicture}[
-  node distance=1cm,
-  every node/.style={font=\small},
-  layer/.style={rectangle, draw, rounded corners, fill=blue!10, minimum width=5.5cm, minimum height=1.2cm, align=center},
-  arrow/.style={->, thick}
-]
-
-% Abstract Layer
-\node[layer] (abstract) {%
-  \textbf{Abstract Layer}\\
-  FT with transition labels as \textbf{sets}
-};
-
-% Locale Layer
-\node[layer, below=of abstract] (locale) {%
-  \textbf{Implementation Layer}\\
-  FT with transition labels as a \\boolean algebra locale
-};
-
-% Implementation Layer
-\node[layer, below=of locale] (impl) {%
-  \textbf{Boolean Algebra Layer}\\
-  Concrete Boolean algebra implementations\\
-  (e.g., intervals, bitvectors, predicates)
-};
-
-% Draw arrows with refinement labels
-\draw[arrow] (abstract) -- node[right] {refinement} (locale);
-\draw[arrow] (locale) -- node[right] {refinement} (impl);
-
-\end{tikzpicture}
-\caption{Three-layered formalization of SFTs in Isabelle/HOL. 
-\textbf{Abstract Layer:} Transition labels are modeled as sets, representing the most general case.
-\textbf{Locale Layer:} Transition labels are abstracted as a type variable \texttt{'a} within a locale, supporting any Boolean algebra.
-\textbf{Implementation Layer:} Concrete Boolean algebra instances, such as intervals, bitvectors, or predicates, are provided.}
-\label{fig:layered-formalization}
-\end{figure}
-
-As we have discussed the diversity of transition labels in Section \ref{sec:introduction}, we now present an extensible formalization of SFTs that accommodates this variety.
-Our approach leverages the refinement framework (Refine\_Monadic \cite{Refine_Monadic-AFP}) in Isabelle/HOL to achieve the flexibility and extensibility of transition labels modeling.
-
+\section{The Isabelle/HOL Formalization of SFTs}
+
+We present the basic architecture of our SFT formalization in Isabelle/HOL, which consists of three distinct layers.
+
+The \textbf{abstract layer} defines transition labels as \textbf{sets}, states are stored in sets as well.
+This abstract formalization, which abstracts away implementation details, enables straightforward correctness proofs for SFT operations.
+
+The \textbf{implementation layer} refines the abstract transition labels to a locale where a boolean algebra is defined as type variable \texttt{'b} together with a collection of operations and assumptions.
+State sets are refined using Refine\_Monadic \cite{Refine_Monadic-AFP}, a framework in Isabelle/HOL for data refinement that automatically transforms general sets into efficient data structures such as hashmaps and red-black trees.
+
+The \textbf{boolean algebra layer} refines the transition label locale to a concrete boolean algebra implementation. In this paper, we present an interval-based refinement.
+
+This three-layer architecture separates the fixed components (abstract and implementation layers) of the SFT formalization from the variable components (transition labels).
+This separation enables us to extend the SFT formalization to accommodate new transition label theories without modifying the first two layer formalization.
+
+\subsection{The Abstract Layer of SFTs}
+\label{sec:abstract-layer}
+
+We present the SFT formalization in abstract layer in Fig.
+\ref{fig-def-FT}.
+While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
+%
+The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
+This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
 \begin{figure}[hbt!]
@@
 
-Figure \ref{fig-def-FT} presents our formalization of SFTs in Isabelle/HOL.
-While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
-%
-The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
-This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
-
 Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
@@
 
-\section{The Product Operation of SFTs}
-\label{sec:product-operation}
-
-In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
+\paragraph{The product operation formalization.} We formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
+The SFA is formalized as a record type \texttt{NFA in Figure \ref{fig-def-FAs}}.
 An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
-Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of SFAs with $\varepsilon$-transitions, we extend the framework in CertiStr with a formalization of symbolic SFAs with $\varepsilon$-transitions (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
+%
+Therefore, we need to formalize $\varepsilon$SFAs, which is shown in Fig.
+\ref{fig-def-FAs} as \texttt{eNFA}.
+The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
 In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
@@
 \end{figure}
-
-Figure \ref{fig-def-FAs} presents the formalization of both SFAs and $\varepsilon$SFAs using Isabelle/HOL record types \texttt{NFA} and \texttt{eNFA}, respectively.
-The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
 
@@
 
-\section{Algorithm Level Refinement}
+\subsection{Implementation Layer Refinement}
 \label{sec_alg_refinement}
-Having established the abstract definition of the SFT product operation in Section \ref{sec:product-operation}, we now present its algorithmic refinement.
+Having established the abstract definition of the SFT product operation in Section \ref{sec:abstract-layer}, we now present its algorithmic refinement.
 This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
 
-\subsection{Intervals}
-
-An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
-To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
-This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
-
-Throughout the following discussion, we use the term "interval" to refer to interval lists. Our formalization provides a collection of interval operations through the following interface:
-
-\begin{enumerate}
-  \item $\texttt{semIs}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semIs}~(i,j) = \{e \mid i \leq e \leq j\}$.
-  \item $\texttt{emptyIs}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semIs}~i = \emptyset$.
-  \item $\texttt{nemptyIs}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semIs}~i \neq \emptyset$.
-  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \cap \texttt{semIs}~i_2$.
-  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \setminus \texttt{semIs}~i_2$.
-\end{enumerate}
-
-To facilitate formal reasoning and optimize performance, we introduce a canonical form for interval.
-An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
-\begin{enumerate}
-  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
-  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
-\end{enumerate}
-
-We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
-This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
-
-\subsection{Algorithmic Implementation of the Product Operation}
-
-In this subsection, we present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.}.
-%
+\paragraph{Boolean Algebra Locale}
+
+Figure \ref{fig-bool-algebra-locale} presents the simplified definition of a boolean algebra locale.
+The type variable \texttt{'b} represents any boolean algebra, while the type variable \texttt{'a} denotes the element type of the sets that the boolean algebra represents.
+The locale defines a collection of operations and assumptions that characterize boolean algebra behavior, including operations for set semantics, emptiness checks, non-emptiness checks, intersection, difference, and element membership.
+The function "\texttt{sem}" provides the semantic interpretation of a boolean algebra as a set, with all other operations' assumptions defined in terms of this set semantics.
+
+The locale serves as an interface to boolean algebra implementations.
+Any concrete boolean algebra implementation must provide a concrete instantiation of this locale, which includes defining the type variable \texttt{'b} and providing implementations for all operations along with proofs of the required assumptions.
+
+\begin{figure}[hbt!]
+\begin{lstlisting}[mathescape=true]
+locale bool_algebra =
+ fixes sem :: " 'b $\Rightarrow$ 'a::ord set"
+ fixes empty:: "'b $\Rightarrow$ bool"
+ fixes nempty:: "'b $\Rightarrow$ bool"
+ fixes intersect:: "'b $\Rightarrow$ 'b $\Rightarrow$ 'b"
+ fixes diff:: "'b $\Rightarrow$ 'b $\Rightarrow$ 'b"
+ fixes elem:: "'a  $\Rightarrow$ 'b $\Rightarrow$ bool"
+ assumes
+  empty_sem: "empty s = (sem s = {})"
+                                    and
+  nempty_sem: "nempty_intervals s =
+                     ($\neg$ (empty s))" and
+  inter_sem: "sem (intersect s1 s2) =
+               (sem s1) $\cap$ (sem s2)" and
+  diff_sem: "sem (diff f1 f2 s1 s2) =
+               (sem s1) - (sem s2)" and
+  elem_sem: "elem a s $\equiv$  (a $\in$ sem s)"
+\end{lstlisting}
+\caption{The boolean algebra locale}
+\label{fig-bool-algebra-locale}
+\end{figure}
+
+\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.} based on Refined\_monadic framework.
+
 Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
@@
  FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
-  (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then
+  (if (nempty (intersect |$\alpha$| |$\alpha'$|)) then
     do {
       D1 |$\leftarrow$| 
-       (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) 
+       (if (F (M f) (intersect |$\alpha$| |$\alpha'$|)) 
           |$\neq$| None) then
          let |$\alpha_i$| = the (F (M f) 
-              (intersectIs |$\alpha$| |$\alpha'$|)) in
+              (intersect |$\alpha$| |$\alpha'$|)) in
          RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
@@
       D2 |$\leftarrow$| 
-       (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) 
+       (if fe (M f) (intersect |$\alpha$| |$\alpha'$|) 
         then 
@@
 
+\subsection{Interval Implementation}
+
+Finally, we present the implementation of the interval algebra.
+An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
+To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
+This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
+For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
+
+Throughout the following discussion, we use the term "interval" to refer to interval lists.
+Our formalization provides all operations in the boolean algebra locale.
+For instance, we implement \texttt{intersect} for intervals, which computes the intersection of two intervals $i_1$ and $i_2$, yielding an interval $i$ such that $\texttt{sem}~i = \texttt{sem}~i_1 \cap \texttt{sem}~i_2$
+
+To facilitate formal reasoning and optimize performance, in addition to the above locale operations, we introduce a canonical form for interval.
+An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
+\begin{enumerate}
+  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
+  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
+\end{enumerate}
+
+We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
+This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
+
 \section{Applications and Evaluation}
@@
 
-One Important application of our formalization is in the context of string sanitization, specifically for web applications.
-Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before being rendered in a web browser.
-Listing \ref{lst-html-encode} presents a C\# code snippet that implements an HTML encoding function, which is a sanitization operation.
-The function iterates through each character in the input string and encodes characters that are not safe for HTML rendering.
+An important application of our formalization lies in string sanitization for web applications.
+Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before rendering in web browsers.
+Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function—a fundamental sanitization operation.
+This function iterates through each character in the input string and encodes characters that are unsafe for HTML rendering.
 
@@
 
-  The modeling of the aformentioned replacement operation will replace one occurrence of the substring that matches the regular expression pattern.
-We may also want to model left-most replacement and replace all occurrences of the substring that matches the regular expression pattern.
-
-  One way to model left-most replacement is illustrated in the following.
+  The replacement operation model described above replaces a single occurrence of any substring matching the regular expression pattern.
+To support different replacement semantics, we may also need to model left-most replacement (which replaces the first occurrence) and replace-all operations (which substitute all occurrences of matching substrings).
+
+  The approach to modeling left-most replacement is illustrated in Figure \ref{fig:replace-left-most}.
+This construction involves two SFTs: one for non-matching characters and one for the matching pattern.
+The non-matching SFT is constructed by first computing the negation of the pattern (the complement of the regular expression's NFA) and converting it into an SFT by adding the identity function $\lambda x.x$.
+The matching SFT is constructed similarly to the previous example in Figure \ref{fig-rearranged-automata}, but without self-loops at the initial state, creating an SFT that processes the pattern and generates the replacement string.
+
+  The two SFTs are composed by adding an $\varepsilon$-transition from the accepting state of the non-matching SFT to the initial state of the matching SFT.
+This transition enables the left-most replacement operation to proceed to pattern matching after processing any non-matching prefix.
+
+  The approach to modeling replace-all operations is illustrated in Figure \ref{fig:replace-all}.
+This construction is similar to left-most replacement but includes an additional $\varepsilon$-transition from the accepting state of the matching SFT back to the initial state of the non-matching SFT.
+This bidirectional connection allows the replace-all operation to continue processing the input string and replace all occurrences of matching substrings.
 
@@
 
-\subsection{Experiments}
-
-We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
+\subsection{Experiments of String Replacement Operations}
+
+We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations.
 %\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
-While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
-
-CertiStrR implements two kinds of SMT-LIB's replacement operations:
-\texttt{str.replace s p r}: a string-based replacement where pattern \texttt{p} is a constant string.
-\texttt{str.replace\_re s p r}: a regular expression-based replacement where pattern \texttt{p} is a regular expression
+CertiStr \cite{cpp/KanLRS22} is a certified string solver that only support some basic string operations, such as concatenation, and regular constraints.
+
+%While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
+
+CertiStrR implements three different pattern matching introduced above: (1) singleton matching at any position, (2) left-most matching, and (3) replace-all matching.
+The pattern can be a regular expression or a constant string, which is a special case of regular expression.
 
@@
 The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
-%
-
-As discussed in the previous subsection, our replacement semantics differs from the SMT-LIB standard semantics.
-Run the SMT solver CVC5 \cite{cvc5} with the code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
+But the left-matching will return UNSAT.
+%
 
@@
       \midrule
-      \texttt{replace\_str} & 142/173/8 & 0.27 & 323\\
-      \texttt{replace\_re} & 84/4/10 & 0.36 & 98\\
+      \texttt{replace singleton} & 142/173/8 & 0.27 & 323\\
+      \texttt{replace left-match} & 84/4/10 & 0.36 & 98\\
+      \texttt{replace all} & 84/4/10 & 0.36 & 98\\
       \bottomrule
@@
 
-We discuss the effort of certified SFT development in this subsection.
-Table \ref{tab:abstract_impl} provides an overview.
-The abstract-level development means all formalizations in Section \ref{sec:formalization} and  \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs .
-The implementation-level development means all formalizations in Section \ref{sec_alg_refinement} including the refinement of the product operation.
-The last row corresponds to the effort of the interval formalization.
-The most difficult part is the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
-Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
+This subsection discusses the development effort required for our certified SFT formalization. Table \ref{tab:abstract_impl} provides an overview of this effort.
+All efforts listed in the table are additional to the existing CertiStr framework that we build upon.
+
+The abstract-level development encompasses all formalizations presented in Sections \ref{sec:formalization} and \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs.
+The implementation-level development covers all formalizations in Section \ref{sec_alg_refinement}, including the refinement of the product operation.
+The final row corresponds to the interval formalization effort.
+The most challenging component proved to be the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
+
+The interval formalization complexity increased dramatically in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22}, primarily due to extending intervals from single pairs to lists.
 

### 2025-07-31 Paper/main.tex
polish paper

@@
 This experimental setup reflects realistic usage scenarios where diverse character sets and input patterns are encountered.
+The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a one-minute time limit per test.
 
@@
 The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
-But the left-matching will return UNSAT.
+But the left-matching will return \texttt{UNSAT}.
 %
@@
 We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are divided into two categories based on the replacement operations: (1) string-based replacement (\texttt{str.replace}) and (2) regular expression-based replacement (\texttt{str.replace\_re}).
-Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB language, we preprocess certain operations.
-For instance, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
+The benchmarks are categorized into three groups based on their pattern matching semantics: (1) general singleton matching, (2) left-most matching, and (3) replace-all matching.
+Patterns can be either regular expressions or constant strings.
+
+Due to CertiStrR's current frontend limitations in supporting the complete SMT-LIB language specification, we preprocess certain operations.
+For example, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when both \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
 
@@
 
-The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a time limit of one minute per test. 
-The results show average execution times of 0.27 seconds for \texttt{str.replace\_str} and 0.36 seconds for \texttt{str.replace\_re} operations.
+The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a one-minute time limit per test.
+The results show average execution times of 0.27 seconds for \texttt{replace singleton}, 0.36 seconds for \texttt{left-match}, and 0.36 seconds for \texttt{replace all}.
 Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
-%
-An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests are finished in one minute) but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
-Specifically, when the string constraints do not satisfy the tree property defined in \cite{cpp/KanLRS22}, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
+
+An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests completed within one minute), but due to the inherent incompleteness of the string solver's algorithm.
+An inconclusive result can occur when the solver cannot decide whether a string variable can be assigned a unique value to make the constraints satisfiable.
 
@@
 \begin{align}
-x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber \\
-x &= x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
+x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber ~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
 \end{align}

### 2025-08-08 Paper/abstract.tex
main.tex split

--- 2025-07-29 Paper/main.tex
@@
-\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%This is a template for producing LIPIcs articles. 
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-\usepackage{stmaryrd}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
-\usetikzlibrary{intersections}
-\usetikzlibrary{automata}
-\usetikzlibrary{positioning}
-\usepackage{booktabs}
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-\usepackage{listings}
-\usepackage{xcolor}
-
-\lstdefinelanguage{SMTLIB}{
-  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
-  sensitive=true,
-  morecomment=[l];,
-  morestring=[b]"
-}
-\lstset{
-  language=SMTLIB,
-  basicstyle=\ttfamily\small,
-  keywordstyle=\color{blue}\bfseries,
-  commentstyle=\color{green!50!black},
-  numbers=left,
-  numberstyle=\tiny\color{gray},
-  stepnumber=1,
-  numbersep=5pt,
-  backgroundcolor=\color{gray!10},
-  frame=single,
-  breaklines=true,
-  captionpos=b,
-  showstringspaces=false
-}
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\lstdefinelanguage{Isabelle}{
-    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
-              where,fixes,assumes,shows,if,then,else,let,in,case,of,
-              record,datatype,type_synonym,fun,function,primrec},
-    sensitive=true,
-    morecomment=[s]{(*}{*)},
-    morestring=[b]",
-  }
-
-\lstset{
-  language=Isabelle,
-	basicstyle=\ttfamily,
-	keywordstyle=\color{blue}\bfseries,
-	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and},
-	escapeinside={|}{|},
-  sensitive=true
-}
-
-\begin{document}
-
-\special{papersize=8.5in,11in}
-\setlength{\pdfpageheight}{\paperheight}
-\setlength{\pdfpagewidth}{\paperwidth}
-
-\title{Certified Symbolic Finite Transducers: Formalization and Applications to String Manipulation}
-
 \begin{abstract}
@@
 \end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Finite Automata (FA) and Finite Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
-For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
-Additionally, FAs and FTs find significant industrial applications, such as in the verification of AWS access control policies \cite{DBLP:conf/fmcad/BackesBCDGLRTV18}.
-
-Even though there are various formalizations of FAs and FTs in interactive proof assistants such as Isabelle \cite{isabelle-homepage} and Coq \cite{coq-homepage}, these are predominantly based on classical definitions.
-However, these traditional approaches present certain limitations when applied to practical scenarios.
-One significant drawback is that transition labels are typically non-symbolic and finite.
-A conventional transition is represented as $q\xrightarrow{a}q'$, where $a$ is a character from a finite alphabet.
-This simplistic representation can lead to a phenomenon known as transition explosion.
-For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
-Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the FA and FT operations highly inefficient.
-Furthermore, in practical applications, it is often necessary to consider infinite alphabets, such as the set of all integers.
-
-Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
-%
-These symbolic models utilize transition labels defined by first-order predicates over boolean algebras, allowing for more expressive representations.
-For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \mod 2 = 0$), denoting all even numbers.
-This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
-
-However, the formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
-Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse predicate representations of boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
-
-Prior work of CertiStr \cite{cpp/KanLRS22} has successfully formalized SFAs within Isabelle/HOL, demonstrating both the efficiency and effectiveness of SFAs in practice.
-However, the formalization of SFTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
-SFTs constitute a significantly more expressive and powerful theoretical framework compared to  SFAs, as evidenced by their capability to model complex string transformations such as replacement operations.
-%
-Furthermore, numerous studies have demonstrated the broad applicability of SFTs across diverse domains.
-The work of \cite{VeanesHLMB12Transducer} showcases SFTs in security-critical applications, particularly for cross-site scripting (XSS) prevention.
-\cite{uss/HooimeijerLMSV11} extends this security focus by applying SFTs to web application sanitizer analysis.
-In system verification, \cite{osdi/YaseenABCL20} employs SFTs for runtime behavior monitoring, while \cite{pldi/HuD17} demonstrates their utility in automated program transformation through systematic inversion techniques.
-
-In this work, we present a comprehensive formalization of SFTs.
-To address the extensibility challenges inherent in supporting diverse transition label theories, we adopt a refinement-based approach.
-At the abstraction level, transition labels are formalized through the fundamental mathematical concept of \emph{sets}.
-This abstraction facilitates subsequent refinement to various representations of Boolean algebras, such as intervals and arithmetic predicates, while maintaining theoretical consistency.
-
-The key operation of our formalization is the product operation between an SFT and an input regular language.
-Specifically, given an SFA $\mathcal{A}$ representing a regular language and an SFT $\mathcal{T}$, we define the product operation $\mathcal{T} \times\mathcal{A}$ that characterizes the output language generated by $\mathcal{T}$ when processing inputs from the language recognized by $\mathcal{A}$.
-
-In the refinement level, we implemented transition labels using an interval-based representation to examplify the refinement process.
-The formalization of interval algebra provides efficient set-theoretic operations—including membership checking, intersection, and difference computations—facilitating the refinement of transition labels from abstract sets to concrete intervals.
-Furthermore, leveraging the data refinement framework \cite{DBLP:conf/itp/Lammich13} in Isabelle/HOL, we store states and transitions using sophisticated data structures such as hashmaps and red-black trees, ensuring efficient automata manipulation.
-
-To evaluate the effectiveness and efficiency of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
-We evaluated our implementation using a set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
-The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
-
-In summary, our formalization makes the following contributions:
-\begin{enumerate}
-\item We present the first formalization of symbolic finite transducers in Isabelle/HOL proof assistant.
-\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
-\item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
-\item We demonstrate the practical utility of our formalization through its application to string constraint solving, specifically in modeling replacement operations with verified correctness guarantees.
-\end{enumerate}
-
-The remainder of this paper is organized as follows:
-Section \ref{sec:formalization} introduces the formalization of symbolic finite transducers.
-Section \ref{sec:product-operation} presents the product operation at the abstract level.
-Section \ref{sec_alg_refinement} details the algorithmic refinement of the product operation.
-Section \ref{sec-app-str-solver} demonstrates the application to string constraint solving.
-Section \ref{sec:related-work} discusses related work.
-Section \ref{sec:conclusion} concludes with future directions.
-
-\section{Formalization of SFTs}
-\label{sec:formalization}
-
-We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL. 
-\subsection{The Mathematical Definition of SFTs}
-
-Let $\mathcal{U}$ be a multi-sorted carrier set or background universe, which is equipped with functions and relations over the elements.
-We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
-We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type.
-
-A lambda term is defined as $\lambda x.~t$ of type $\tau_1 \rightarrow \tau_2$.
-When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate.
-Let $\phi$ be a predicate.
-We write $a\in \llbracket\phi \rrbracket$ if $\phi~a=\top$.
-For non-predicate lambda terms, we view them as functions that generate output elements of type $\tau_2$ given input terms of type $\tau_1$.
-With these notations and the above definitions, we can define SFTs as follows.
-
-\begin{definition}[Symbolic Finite Transducer]
-\label{def-sft}
-   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
-   \begin{itemize}
-   \item $\mathcal{Q}$ is a finite set of states,
-   \item $\mathcal{I}\subseteq \mathcal{Q}$ is the set of initial states,
-   \item $\mathcal{F} \subseteq\mathcal{Q}$ is the set of accepting states,
-   \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
-   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
-   \end{itemize}
-
-For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called an input,  then the application $(f~a)$ is the output.
-
-\end{definition}
-
-SFTs accept an input word and generate an output word. This can be defined by \emph{runs} of SFTs.
-An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i,
-\phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
-$\sigma$ is an accepting run when $q_n$ is an accepting state.
-
-For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ if and only if $a_i \in \llbracket\phi_i \rrbracket$ for $0 \leq i \leq n - 1$.
-When $w$ is accepted, run $\sigma$ generates an output sequence $w'= f_0~a_0, \ldots, f_{n-1}~a_{n-1}$.
-We define:
-\begin{itemize}
-\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
-\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, $ $(f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
-\end{itemize}
-
-If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
-For a trace $\pi$, we denote its input as $\mathit{in}(\pi)$ and its output as $\mathit{out}(\pi)$.
-%
-Given an SFT $\mathcal{T}$ and a word $w$, we define the \emph{product} operation of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with input $w$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
-\end{split}
-\]
-
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an SFA $\mathcal{A}$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \\\text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
-\end{split}
-\]
-
-\begin{figure}[hbt!]
-  \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm and 2cm, on grid, auto]
-    % Make q0 and q1 horizontally wide apart
-    \node[state, initial] (q_0)   {$q_0$}; 
-    \node[state] (q_1) [right=4cm of q_0] {$q_1$}; 
-    \node[state, accepting] (q_2) [below=2.2cm of q_1] {$q_2$};
-
-    \path[->] 
-    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
-    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
-  \end{tikzpicture}
-    \caption{An example of SFT}
-    \label{fig-example-ft}
-\end{figure}
-
-    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
-The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
-The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation.
-For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.
-
-%The transition labels can be more complex.
-For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$.
-This symbolic label corresponds to the set of even natural numbers.
-
-\subsection{The Isabelle/HOL Formalization of SFTs}
-
-Firstly, we present the basic idea of our formalization of SFTs in Isabelle/HOL, which consists of three layers.
-
-\begin{figure}[hbt!]
-\centering
-\begin{tikzpicture}[
-  node distance=1cm,
-  every node/.style={font=\small},
-  layer/.style={rectangle, draw, rounded corners, fill=blue!10, minimum width=5.5cm, minimum height=1.2cm, align=center},
-  arrow/.style={->, thick}
-]
-
-% Abstract Layer
-\node[layer] (abstract) {%
-  \textbf{Abstract Layer}\\
-  FT with transition labels as \textbf{sets}
-};
-
-% Locale Layer
-\node[layer, below=of abstract] (locale) {%
-  \textbf{Implementation Layer}\\
-  FT with transition labels as a \\boolean algebra locale
-};
-
-% Implementation Layer
-\node[layer, below=of locale] (impl) {%
-  \textbf{Boolean Algebra Layer}\\
-  Concrete Boolean algebra implementations\\
-  (e.g., intervals, bitvectors, predicates)
-};
-
-% Draw arrows with refinement labels
-\draw[arrow] (abstract) -- node[right] {refinement} (locale);
-\draw[arrow] (locale) -- node[right] {refinement} (impl);
-
-\end{tikzpicture}
-\caption{Three-layered formalization of SFTs in Isabelle/HOL. 
-\textbf{Abstract Layer:} Transition labels are modeled as sets, representing the most general case.
-\textbf{Locale Layer:} Transition labels are abstracted as a type variable \texttt{'a} within a locale, supporting any Boolean algebra.
-\textbf{Implementation Layer:} Concrete Boolean algebra instances, such as intervals, bitvectors, or predicates, are provided.}
-\label{fig:layered-formalization}
-\end{figure}
-
-As we have discussed the diversity of transition labels in Section \ref{sec:introduction}, we now present an extensible formalization of SFTs that accommodates this variety.
-Our approach leverages the refinement framework (Refine\_Monadic \cite{Refine_Monadic-AFP}) in Isabelle/HOL to achieve the flexibility and extensibility of transition labels modeling.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a,$| |$'i,$| |$'b$|) |$\texttt{NFT}$| =
-	|$\mathcal{Q}_t$| :: "|$'q$| set"
-	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'i$|) LTTS"
-	|$\mathcal{I}_t$| :: "|$'q$| set"
-	|$\mathcal{F}_t$| :: "|$'q$| set"
-        |$\mathcal{M}_t$| :: "|$'i\Rightarrow ('a,~'b)$| Tlabel"
-type_synonym |$('q,~'a,~'i)$| LTTS = "|$('q \times ('a \text{ set option }\times 'i) \times 'q)$| set"
-type_synonym |$('a,~'b)$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
-	\end{lstlisting}
-\caption{The formalization of $SFTs$ in Isabelle/HOL}
-\label{fig-def-FT}
-\end{figure}
-
-Figure \ref{fig-def-FT} presents our formalization of SFTs in Isabelle/HOL.
-While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
-%
-The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
-This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
-
-Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
-Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
-A predicate's semantics as introduced before represents a set of elements that make the predicate true.
-But predicates have various different forms.
-For instance, the interval $[1-9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
-The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
-All these different forms of labels are abstracted as sets in our formalization.
-%
-The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements.
-This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
-
-The second element of type $'i$ serves as an index into the output function space. 
-%
-We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
-%
-The mapping $\mathcal{M}_t$ associates each index with a specific output function.
-These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
-This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria.
-Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
-
-%Now let us look at an example.
-In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
-This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}).
-%We can easily model this operation using our formalization of NFTs with the following steps:
-%\begin{itemize}
-%\item Step 1, construct a corresponding NFA from \texttt{pattern}).
-%\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
-%\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
-%\item Step 4, unlabel previous accepting states as non-accepting states. 
-%\end{itemize}
-
-%We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
-
-\section{The Product Operation of SFTs}
-\label{sec:product-operation}
-
-In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
-An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
-Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of SFAs with $\varepsilon$-transitions, we extend the framework in CertiStr with a formalization of symbolic SFAs with $\varepsilon$-transitions (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
-In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a$|) |$\texttt{NFA}$| =
-	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\mathcal{I}$| :: "|$'q$| set"
-	|$\mathcal{F}$| :: "|$'q$| set"
-
-record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
-	|$\mathcal{Q}_e$| :: "|$'q$| set"
-	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\Delta_e'$| :: "|$('q * 'q)$| set"
-	|$\mathcal{I}_e$| :: "|$'q$| set"
-	|$\mathcal{F}_e$| :: "|$'q$| set"
-
-type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
-	\end{lstlisting}
-\caption{The formalization of $\varepsilon$SFAs and SFAs}
-\label{fig-def-FAs}
-\end{figure}
-
-Figure \ref{fig-def-FAs} presents the formalization of both SFAs and $\varepsilon$SFAs using Isabelle/HOL record types \texttt{NFA} and \texttt{eNFA}, respectively.
-The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
-
-Having established these foundational definitions, we can now formalize the product operation.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| 
-   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$| 
-   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
-"productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
-  |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
-       |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
-        |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
-      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
-       |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
-       |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
-          |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
-  |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-       |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
-       |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
-      |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-        |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}$|
-        |$\land~\sigma_1\cap \sigma_2 \neq \emptyset~\land\exists x\in(\sigma_1\cap\sigma_2).$|
-        |$~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
-  |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
-  |$\mathcal{F}_e = \mathcal{F}_t~\mathcal{T} \times \mathcal{F}~\mathcal{A}$| |$\rrparenthesis$|"
-	\end{lstlisting}
-\caption{The formalization of product operation}
-\label{fig-def-FTProd}
-\end{figure}
-
-Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA. The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively. The 
-result of the product operation is an $\varepsilon$SFA.
-But we need to explain the role of parameter $\text{\texttt{F}}$.
-The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$.
-$\text{\texttt{F}}$ extends $f$ to apply $f$ to a set of elements.
-More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as follows:
-
-\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].
-
-The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions.
-In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels.
-Let us consider $\Delta_e$ first.
-
-\begin{enumerate}
-\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e.
-the input character is $\varepsilon$, and $f~\text{None} \neq \text{None}$.
-Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output "$((\mathcal{M}_t~\mathcal{T})~f)~\text{None}$".
-Remember that $f$ is just an index, $(\mathcal{M}_t~\mathcal{T})~f$ is the output function.
-
-\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
-The resulting output is $\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)$.
-\end{enumerate}
-
-The transitions in $\Delta_e'$ follow a similar pattern with analogous cases for $\varepsilon$ and non-$\varepsilon$ transitions.
-
-To establish the correctness specification of the product operation, we begin by formalizing the concept of SFT \emph{traces} as introduced in Definition \ref{def-sft}.
-In our formalization, traces are represented by the type $('a\;\text{option} \times 'b\;\text{option})\;\texttt{list}$, as shown in Figure \ref{fig-def-output}.
-%
-Given a trace $\pi$, we define two key projection functions:
-(1) \texttt{inputE}, which corresponds to ${in}(\pi)$ and extracts the input sequence
-(2) \texttt{outputE}, which corresponds to ${out}(\pi)$ and extracts the output sequence.
-
-The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-fun inputE :: 
-   "|$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$|"
-where
-  "inputE [] = []" |$\mid$|
-  "inputE ((Some a, |$\_$|) |$\#$| l) = 
-                      a |$\#$| (inputE l)" |$\mid$|
-  "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"
-
-fun outputE :: 
-    "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" 
-where
-  "outputE [] = []" |$\mid$|
-  "outputE ((_,Some a) # l) = 
-                    a # (outputE l)" |$\mid$|
-  "outputE ((_,None) # l) = (outputE l)"
-
-definition outputL :: 
-   "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
-                               ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. 
-    |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T}~\land$| 
-    |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land$||$ \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}"
-	\end{lstlisting}
-\caption{The formalization of traces in SFTs}
-\label{fig-def-output}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-lemma productT_correct:
-  fixes |$\mathcal{T}$| |$\mathcal{A}$| F
-  assumes 
-   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
-       |$\text{ F}~f~s = \text{None}"$|
-   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
-        |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
-        |$f~(\text{Some }e) = \text{ Some S}\})$|"
-      and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-  |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL 
-          |$\mathcal{T}~\mathcal{A}$|"
-
-	\end{lstlisting}
-\caption{The correctness lemma of the product operation}
-\label{fig-def-product-correct}
-\end{figure}
-
-Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation.
-The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
-The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.
-The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.
-
-\section{Algorithm Level Refinement}
-\label{sec_alg_refinement}
-Having established the abstract definition of the SFT product operation in Section \ref{sec:product-operation}, we now present its algorithmic refinement.
-This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
-
-\subsection{Intervals}
-
-An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
-To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
-This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
-
-Throughout the following discussion, we use the term "interval" to refer to interval lists. Our formalization provides a collection of interval operations through the following interface:
-
-\begin{enumerate}
-  \item $\texttt{semIs}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semIs}~(i,j) = \{e \mid i \leq e \leq j\}$.
-  \item $\texttt{emptyIs}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semIs}~i = \emptyset$.
-  \item $\texttt{nemptyIs}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semIs}~i \neq \emptyset$.
-  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \cap \texttt{semIs}~i_2$.
-  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \setminus \texttt{semIs}~i_2$.
-\end{enumerate}
-
-To facilitate formal reasoning and optimize performance, we introduce a canonical form for interval.
-An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
-\begin{enumerate}
-  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
-  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
-\end{enumerate}
-
-We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
-This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
-
-\subsection{Algorithmic Implementation of the Product Operation}
-
-In this subsection, we present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.}.
-%
-Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}.
-These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
-
-The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets.
-This function employs the \texttt{FOREACH} construct, a higher-order iteration operator analogous to OCaml's \texttt{Set.fold}.
-Specifically, given a set $S$, a function $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$, and an initial accumulator $I$ of type $'b$, the expression $\texttt{FOREACH}~S~f~I$ systematically applies $f$ to each element in $S$, accumulating results in a principled manner.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT_impl where
-"productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
-    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) 
-    (nfa_states |$\mathcal{A}$|);
-    (D1, D2) |$\leftarrow$| trans_comp_imp 
-     (nft_tranfun |$\mathcal{T}$|) F fe (nft_trans |$\mathcal{T}$|) 
-      (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
-    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|)
-      (nfa_initial |$\mathcal{A}$|);
-    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) 
-      (nfa_accepting |$\mathcal{A}$|);
-    RETURN (Q, D1, D2, I, F)
-  }"
-\end{lstlisting}
-\caption{The computation of SFT product}
-\label{fig-compute-nft-product}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition prods_imp where
-"prods_imp Q1 Q2 =
-   FOREACH {q. q |$\in$| Q1} (|$\lambda$| q Q. do {
-    S |$\leftarrow$| FOREACH {q. q |$\in$| Q2}
-      (|$\lambda$| q|$'$| Q|$'$|. RETURN ({(q,q|$'$|)} |$\cup$| Q|$'$|)) |$\emptyset$|;
-    RETURN (Q |$\cup$| S)
-   }) |$\emptyset$|"
-\end{lstlisting}
-\caption{The computation of Cartesian product of two state sets}
-\label{fig-def-prods_imp}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition trans_comp_imp where
-"trans_comp_imp M F fe T1 T2 Q =
-  FOREACH {t. t |$\in$| T1}
-   (|$\lambda$|(q,(|$\alpha$|,f),q|$'$|) (D1,D2). 
-    (if (|$\alpha$|=None) then 
-     (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe 
-                              T2 D1 D2)
-     else
-     (subtrans_comp M q (the |$\alpha$|) f 
-            q|$'$| F fe T2 D1 D2))) (|$\emptyset$|, |$\emptyset$|)"
-\end{lstlisting}
-\caption{The computation of \texttt{trans\_comp\_imp}}
-\label{fig-def-prods-imp}
-\end{figure}
-
-A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ (corresponding to $\Delta_e$) and $\texttt{D2}$ (corresponding to $\Delta_e'$) in Figure \ref{fig-compute-nft-product}.
-As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
-This function decomposes the synchronization process into two distinct cases, each handled by a specialized function:
-
-\begin{enumerate}
-  \item \texttt{subtrans\_comp\_$\varepsilon$}: Processes $\varepsilon$-transitions in the SFT, where transitions consume no input but may produce output
-  \item \texttt{subtrans\_comp}: Processes standard transitions in the SFT, where both input consumption and output generation may occur
-\end{enumerate}
-
-\begin{figure}[hbt!]
-\begin{lstlisting}[language=SMTLIB]
-definition subtrans_comp where
-"subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
- FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
-  (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then
-    do {
-      D1 |$\leftarrow$| 
-       (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) 
-          |$\neq$| None) then
-         let |$\alpha_i$| = the (F (M f) 
-              (intersectIs |$\alpha$| |$\alpha'$|)) in
-         RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
-                |$\cup$| D1
-        else RETURN D1);
-      D2 |$\leftarrow$| 
-       (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) 
-        then 
-         RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 
-        else RETURN D2);
-      RETURN (D1, D2) }
-   else (RETURN (D1, D2)))) (D1, D2)"
-    \end{lstlisting}
-    \caption{The computation of \texttt{subtrans\_comp}}
-    \label{fig-def-subtrans_comp}
-    \end{figure}
-
-    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
-For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
-For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
-
-    \begin{enumerate}
-      \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
-      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
-The checking is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
-    \end{enumerate}
-
-  \begin{figure}[hbt!]
-    \begin{lstlisting}
-  lemma productT_imp_correct:
-  assumes finite_TT: "|$\text{finite } (\Delta_t~\mathcal{T})$|"
-      and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
-      and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
-      and finite_TQ: "|$\text{finite } (\mathcal{Q}_t~\mathcal{T})$|"
-      and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
-      and finite_TI: "|$\text{finite } (\mathcal{I}_t~\mathcal{T})$|"
-      and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
-      and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
-  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe}\leq$|
-            |$\text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
-  \end{lstlisting}
-  \caption{The refinement relation between \texttt{productT\_imp} and \texttt{productT}}
-  \label{fig-def-productT_imp_correct}
-  \end{figure}
-
-  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
-This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.
-
-  The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$.
-More precisely,  the concrete implementation must produce an $\varepsilon$SFA that is structurally equivalent or isomorphic to the one produced by the abstract algorithm \texttt{productT}.
-
-  To establish this equivalence, we must prove that the $\varepsilon$SFAs produced by \texttt{productT\_imp} and \texttt{productT} are isomorphic in all essential components: the set of states, transition relations, $\varepsilon$-transition relations, initial and accepting state sets.
-
-The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
-The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 
-In our formalization, we refine the sets of storing states and transitions to red-black trees.
-
-\section{Applications and Evaluation}
-
-We evaluate our formalization and implementation of the SFT product operation in two key application domains: string sanitization and string replacement operations in string solving.
-These applications demonstrate the practical utility of our certified SFT product operation in real-world scenarios.
-\subsection{Sanitizer}
-
-One Important application of our formalization is in the context of string sanitization, specifically for web applications.
-Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before being rendered in a web browser.
-Listing \ref{lst-html-encode} presents a C\# code snippet that implements an HTML encoding function, which is a sanitization operation.
-The function iterates through each character in the input string and encodes characters that are not safe for HTML rendering.
-
-\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
-static string EncodeHtml(string t)
-{
-  if (t == null) { return null; }
-  if (t.Length == 0) {
-    return string.Empty;
-  }
-  StringBuilder builder = new
-    StringBuilder("", t.Length * 2);
-  foreach (char c in t) {
-    if ((((c > ''') && (c < '{')) |$\mid\mid$|
-         ((c > '@') && (c < '['))) |$\mid\mid$|
-         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
-         (((c == '.') || (c == ',')) |$\mid\mid$|
-         ((c == '-') || (c == '_'))))) {
-          builder.Append(c);
-      }
-      else {
-          builder.Append("&#" +
-          ((int) c).ToString() + ";");
-      }
-  }
-  return builder.ToString();
-}
-\end{lstlisting}
-
-Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
-The transducer accepts any input string and produces an output where safe characters are preserved using the identity function $\texttt{id}$, while unsafe characters are encoded as HTML entities.
-The set of safe characters is defined as $\texttt{safe\_chars} = [\texttt{' '}, \texttt{','}, \texttt{'.'}, \texttt{'-'}, \texttt{'\_'}] \cup [\texttt{'0'}-\texttt{'9'}] \cup [\texttt{'A'}-\texttt{'Z'}] \cup [\texttt{'a'}-\texttt{'z'}]$, which can be efficiently represented using the interval algebra we formalized.
-Unsafe characters are transformed using the encoding function $\texttt{encode}(c) = \texttt{``\&\#''} + \texttt{toString}(\texttt{ord}(c)) + \texttt{``\;''}$.
-When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
-  % States
-  \node[state, initial, accepting] (q0) {$q_0$};
-  \node[state] (q1) [right=3.5cm of q0] {$q_1$};
-  \node[state] (q2) [below=of q0] {$q_2$};
-
-  \path[->] 
-    % Transition for safe characters
-    (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
-    % Transition for unsafe characters  
-    (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
-    % Epsilon transitions back to q0
-    (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
-    (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
-\end{tikzpicture}
-\caption{SFT representation of the HTML encoding sanitizer function.}
-\label{fig:html-sanitizer-sft}
-\end{figure}
-
-We evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT with SFAs constructed from randomly generated Unicode user inputs.
-This experimental setup reflects realistic usage scenarios where diverse character sets and input patterns are encountered.
-
-Figure \ref{fig:benchmark-results} demonstrates the scalability characteristics of our SFT product implementation.
-The computation requires approximately 1 second for 100,000-character inputs and 10 seconds for 1,000,000-character inputs.
-These performance metrics confirm that our certified implementation maintains computational efficiency suitable for practical string processing applications at scale.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.45]{benchmark_loglog.pdf}
-\caption{Experimental results for the sanitizer product}
-\label{fig:benchmark-results}
-\end{figure}
-
-Beyond performance evaluation, our SFT formalization enables formal verification of string sanitizer correctness.
-For example, this verification framework models user input as an SFA, the sanitizer as an SFT, and potential attacks as an SFA accepting unsafe strings.
-A sanitizer is verified correct if the product of the sanitizer SFT and user input SFA produces no outputs that intersect with the attack SFA's language.
-This correctness property can be checked by computing the intersection of the product SFT's output language with the attack SFA, where emptiness of this intersection guarantees sanitizer safety.
-
-\subsection{Modeling the Replacement Operation}
-
-The string replacement operation, denoted as 
-$\texttt{replace}(\texttt{str}, \newline\texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
-\begin{itemize}
-  \item $\texttt{str}$: The input string to be transformed.
-  \item $\texttt{pattern}$: A regular expression defining the matching criteria.  
-  \item $\texttt{replacement}$: The string to be substituted for the matched substring.
-\end{itemize}
-
-The semantics of the replacement operation can be formally characterized by two distinct cases:
-\begin{enumerate}
-  \item When there exists at least one substring $s'$ in \texttt{str} such that $s'$ matches $\texttt{pattern}$, then $s'$ is replaced with \texttt{replacement}.
-  \item When no substring of \texttt{str} matches $\texttt{pattern}$, the operation returns \texttt{str} unchanged.
-\end{enumerate}
-
-The first case can be modeled by SFTs.
-We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}\newline(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
-We model this replacement operation by an SFT with the following 3 steps:
-
-\begin{enumerate}
-  \item First, construct an SFA that recognizes the regular expression pattern $\texttt{/[0-9]+/}$.
-Transform this SFA into an SFT by augmenting each transition with an output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
-  \item Second, construct an SFA that accepts the replacement string "\texttt{NUM}".
-Since a constant string can be viewed as a specialized regular expression, we can construct its SFA representation.
-Convert this SFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
-  \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
-These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and the identity function $\texttt{id}=\lambda x.~x$, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
-\end{enumerate}
-
-\noindent\emph{Step 1.}
-Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
-The left side shows the SFA that recognizes the regular expression $\texttt{/[0-9]+/}$, while the right side presents its transformation into an SFT.
-This transformation is achieved by augmenting each transition with the output function $\texttt{f} = \lambda x.~\texttt{None}$, which consistently produces the empty string, effectively "consuming" the matched digits without generating output.
-
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
-  % Left: SFA for [0-9]+ (vertical)
-  \node[state, initial] (q0)   {$q_0$}; 
-  \node[state, accepting, below=of q0] (q1) {$q_1$};
-
-   \path[->] 
-   (q0) edge [bend left=20] node[right] {\texttt{[0-9]}} (q1)
-   (q1) edge [loop below] node {\texttt{[0-9]}} (q1);
-
-  % Right: SFT for [0-9]+ (vertical)
-  \node[state, initial, right=4cm of q0] (q0') {$q_0'$}; 
-  \node[state, accepting, below=of q0'] (q1') {$q_1'$};
-
-   \path[->] 
-   (q0') edge [bend left=20] node[right] {\texttt{[0-9]},$\texttt{f}$} (q1')
-   (q1') edge [loop below] node {\texttt{[0-9]},$\texttt{f}$} (q1');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for \texttt{/[0-9]+/}}
-\label{fig-snfa-pattern}
-\end{figure}
-
-\noindent\emph{Step 2.}
-Figure \ref{fig-snfa-replacement} depicts the automata for the replacement string "\texttt{NUM}".
-The left side shows the SFA that accepts this constant string, while the right side presents its SFT transformation.
-The transformation employs an indexed output function \texttt{g} that emits characters of the replacement string sequentially:
-\[
-\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
-\begin{cases}
-1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
-2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
-3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
-\_ \mapsto \texttt{None}
-\end{cases}
-\]
-This function maps transition indices to their corresponding character outputs, using ASCII codes to represent the string "\texttt{NUM}" character by character.
-
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-  % SFA for "NUM" (top)
-  \node[state, initial above] (q0)   {$q_0$}; 
-  \node[state] (q1) [right=1cm of q0] {$q_1$}; 
-  \node[state] (q2) [below=1cm of q1] {$q_2$}; 
-  \node[state, accepting] (q3) [below=1cm of q0] {$q_3$};
-
-  \path[->] 
-  (q0) edge node {\texttt{N}} (q1)
-  (q1) edge node {\texttt{U}} (q2)
-  (q2) edge node {\texttt{M}} (q3);
-
-  % SFT for "NUM" (right)
-  \node[state, initial above] (q0') [right=1.5cm of q1] {$q_0'$}; 
-  \node[state] (q1') [right=of q0'] {$q_1'$}; 
-  \node[state] (q2') [below=1cm of q1'] {$q_2'$}; 
-  \node[state, accepting] (q3') [below=1cm of q0'] {$q_3'$};
-
-  \path[->] 
-  (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
-  (q1') edge node[left] {\texttt{None}, \texttt{(g 2)}} (q2')
-  (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for "\texttt{NUM}"}
-\label{fig-snfa-replacement}
-\end{figure}
-
-\noindent\emph{Step 3.}
-Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT (Fig.
-\ref{fig-snfa-pattern}) with the replacement-generating SFT (Fig.
-\ref{fig-snfa-replacement}).
-The composition process involves two key modifications:
-\begin{enumerate}
-  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
-  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{id}$", where $\Sigma$ represents the full alphabet.
-For the string solver, $\Sigma$ is the set of all unicode characters.
-These transitions enable the SFT to process arbitrary input prefixes and suffixes while preserving the matched substring for replacement
-\end{enumerate}
-
-\begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-    % Second automaton from the first figure
-    \node[state, initial] (q0') {$q_0'$}; 
-    \node[state] (q1') [right=of q0'] {$q_1'$};
-
-     \path[->] 
-     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q0') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q0');
-
-    % Second automaton from the second figure
-    \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
-    \node[state] (q1'') [right=of q0''] {$q_1''$}; 
-    \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
-    \node[state] (q2'') [right=of q3''] {$q_2''$};
-
-    \path[->] 
-    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
-    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
-    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q3'') edge [loop left] node {$\Sigma$, $\texttt{id}$} (q3'')
-    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
-  \end{tikzpicture}
-  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
-  \label{fig-rearranged-automata}
-  \end{figure}
-
-  Having constructed the SFT, we can now compute the forward image of the replacement operation.
-Consider the string constraint $s' = \texttt{replace}(s,~\texttt{/[0-9]+/},~\text{"}\texttt{NUM}\text{"})$, where we aim to characterize the possible values of $s'$.
-Let $\mathcal{T}$ denote the constructed SFT modeling the replacement operation, and let $\mathcal{A}$ be the SFA representing the domain of possible values for the input string $s$.
-The forward image of this replacement operation is given by the product $\mathcal{T} \times \mathcal{A}$, which precisely captures the set of all possible output strings that can be produced by applying the replacement operation to any input string accepted by $\mathcal{A}$.
-  Our string solver uses this forward image to do forward propagation for further string solving.
-
-  \emph{Comparison with SMT-LIB Semantics.} It is important to note that our formalization differs from the standard SMT-LIB semantics for the replacement operation.
-In SMT-LIB, the \texttt{str.replace} operation is defined to replace only the \emph{first} occurrence of a substring matching the given pattern.
-In contrast, our semantics allows for a more general interpretation where any matching substring may be replaced.
-But it still replaces only one occurrence of them.
-
-  The modeling of the aformentioned replacement operation will replace one occurrence of the substring that matches the regular expression pattern.
-We may also want to model left-most replacement and replace all occurrences of the substring that matches the regular expression pattern.
-
-  One way to model left-most replacement is illustrated in the following.
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-\end{tikzpicture}
-\caption{The modeling of left-most replacement operation.}
-\label{fig:replace-left-most}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-  % Epsilon transition from accepting state of second SFT to initial state of first SFT
-  \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
-
-\end{tikzpicture}
-\caption{The modeling of replace all operation.}
-\label{fig:replace-all}
-\end{figure}
-
-\subsection{Experiments}
-
-We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
-%\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
-While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
-
-CertiStrR implements two kinds of SMT-LIB's replacement operations:
-\texttt{str.replace s p r}: a string-based replacement where pattern \texttt{p} is a constant string.
-\texttt{str.replace\_re s p r}: a regular expression-based replacement where pattern \texttt{p} is a regular expression
-
-Listing \ref{lst-smtlib-code} demonstrates the regular expression variant through an example that replaces numeric sequences with the string "\texttt{NUM}".
-The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
-%
-
-As discussed in the previous subsection, our replacement semantics differs from the SMT-LIB standard semantics.
-Run the SMT solver CVC5 \cite{cvc5} with the code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
-
-\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}, float=htbp]
-  (set-logic QF_S)
-  (declare-fun a () String)
-  (declare-fun b () String)
-  (assert (= a "2024,2025"))
-  (assert (= b "2024,NUM"))
-  (assert (= b (str.replace_re a 
-      (re.+ (re.range "0" "9")) "NUM")))
-  (check-sat)
-  \end{lstlisting}
-
-  We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are divided into two categories based on the replacement operations: (1) string-based replacement (\texttt{str.replace}) and (2) regular expression-based replacement (\texttt{str.replace\_re}).
-Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB language, we preprocess certain operations.
-For instance, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
-
-% replace_re
-% Average real time: 0.37 seconds
-% Total number of real entries: 98
-% Average user time: 0.35 seconds
-% Total number of user entries: 98
-% Average sys time: 0.01 seconds
-% Total number of sys entries: 98
-% Total number of SAT entries: 61
-% Total number of UNSAT entries: 4
-% Total number of inconclusive entries: 30
-
-% replace_str   
-% Average real time: 0.27 seconds
-% Total number of real entries: 323
-% Average user time: 0.25 seconds
-% Total number of user entries: 323
-% Average sys time: 0.01 seconds
-% Total number of sys entries: 323
-% Total number of SAT entries: 315
-% Total number of UNSAT entries: 173
-% Total number of inconclusive entries: 8
-
-\begin{table}[h]
-  \centering
-  \small
-  \begin{tabular}{lcccc}
-      \toprule
-      & \textbf{SAT/UNSAT/Inc} & \textbf{Time (s)} & \textbf{Tests} \\
-      \midrule
-      \texttt{replace\_str} & 142/173/8 & 0.27 & 323\\
-      \texttt{replace\_re} & 84/4/10 & 0.36 & 98\\
-      \bottomrule
-  \end{tabular}
-  \caption{Experimental results}
-  \label{tab:string_operations}
-\end{table}
-
-The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a time limit of one minute per test. 
-The results show average execution times of 0.27 seconds for \texttt{str.replace\_str} and 0.36 seconds for \texttt{str.replace\_re} operations.
-Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
-%
-An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests are finished in one minute) but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
-Specifically, when the string constraints do not satisfy the tree property defined in \cite{cpp/KanLRS22}, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
-
-Our performance analysis revealed that while the SFT-based replacement operation modeling is efficient, the primary computational bottleneck stems from automata accumulation during  forward-propagation.
-Consider the following string constraints:
-\begin{align}
-x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber \\
-x &= x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
-\end{align}
-where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}\newline($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
-Each accumulation step requires computing the product of the current automaton with the previous result.
-Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state explosion.
-
-\subsection{Effort of Certified Development}
-
-We discuss the effort of certified SFT development in this subsection.
-Table \ref{tab:abstract_impl} provides an overview.
-The abstract-level development means all formalizations in Section \ref{sec:formalization} and  \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs .
-The implementation-level development means all formalizations in Section \ref{sec_alg_refinement} including the refinement of the product operation.
-The last row corresponds to the effort of the interval formalization.
-The most difficult part is the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
-Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
-
-\begin{table}[h]
-  \centering
-  \small
-  \begin{tabular}{lccc}
-      \toprule
-      & \textbf{Defs} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
-      \midrule
-      Abstract & 17 & 21 & 3274 \\
-      Implementation & 51 & 43 & 2700 \\
-      Interval & 15 & 29 & 1500 \\
-      \bottomrule
-  \end{tabular}
-  \caption{Overview of the effort of certified development}
-  \label{tab:abstract_impl}
-\end{table}
-
-\section{Related Work}
-\label{sec:related-work}
-
-\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
-This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
-Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
-
-\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
-These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
-Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.
-
-\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
-To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
-Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.
-
-\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
-While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
-As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
-We believe that our work will benefit the community by providing a formal foundation for string solvers development.
-
-\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains.
-For example, the work by Shi et al.
-\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL.
-Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.
-
-To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}.
-This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
-The experimental results confirm both the efficiency and effectiveness of our approach.
-While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.
-
-Future work includes extending our symbolic formalization in two key directions:
-%
-(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
-(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
-These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.
-
-\bibliography{literature}
-\end{document}

### 2025-08-08 Paper/application.tex
main.tex split

--- 2025-07-31 Paper/main.tex
@@
-\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%This is a template for producing LIPIcs articles. 
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-\usepackage{stmaryrd}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
-\usetikzlibrary{intersections}
-\usetikzlibrary{automata}
-\usetikzlibrary{positioning}
-\usepackage{booktabs}
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-\usepackage{listings}
-\usepackage{xcolor}
-
-\lstdefinelanguage{SMTLIB}{
-  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
-  sensitive=true,
-  morecomment=[l];,
-  morestring=[b]"
-}
-\lstset{
-  language=SMTLIB,
-  basicstyle=\ttfamily\small,
-  keywordstyle=\color{blue}\bfseries,
-  commentstyle=\color{green!50!black},
-  numbers=left,
-  numberstyle=\tiny\color{gray},
-  stepnumber=1,
-  numbersep=5pt,
-  backgroundcolor=\color{gray!10},
-  frame=single,
-  breaklines=true,
-  captionpos=b,
-  showstringspaces=false
-}
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\lstdefinelanguage{Isabelle}{
-    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
-              where,fixes,assumes,shows,if,then,else,let,in,case,of,
-              record,datatype,type_synonym,fun,function,primrec},
-    sensitive=true,
-    morecomment=[s]{(*}{*)},
-    morestring=[b]",
-  }
-
-\lstset{
-  language=Isabelle,
-	basicstyle=\ttfamily,
-	keywordstyle=\color{blue}\bfseries,
-	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and, locale},
-	escapeinside={|}{|},
-  sensitive=true
-}
-
-\begin{document}
-
-\special{papersize=8.5in,11in}
-\setlength{\pdfpageheight}{\paperheight}
-\setlength{\pdfpagewidth}{\paperwidth}
-
-\title{Certified Symbolic Finite Transducers: Formalization and Applications to String Manipulation}
-
-\begin{abstract}
-Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
-FTs have wide applications, including program analysis, string constraint solving, and security-critical sanitizer analysis.
-However, there is currently no scalable formalization of FTs.
-This is primarily because transition labels in FTs are not formalized symbolically, resulting in transition explosion and significant performance bottlenecks in practical applications.
-
-In this paper, we present a formalization of FTs in the symbolic setting using Isabelle/HOL, where transition labels can be instantiated with various Boolean algebras, such as intervals and arithmetic predicates.
-This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
-%
-To evaluate the effectiveness of our formalization, we apply the formalized SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for preventing XSS attacks, and (2) a string solver that models replacement operations with various semantics, including left-most and replace-all matching.
-Experimental results demonstrate that our approach achieves computational efficiency in constraint-solving scenarios.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:introduction}
-
-Finite Automata (FA) and Finite Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
-For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
-Additionally, FAs and FTs find significant industrial applications, such as in the verification of AWS access control policies \cite{DBLP:conf/fmcad/BackesBCDGLRTV18}.
-
-Even though there are various formalizations of FAs and FTs in interactive proof assistants such as Isabelle \cite{isabelle-homepage} and Coq \cite{coq-homepage}, these are predominantly based on classical definitions.
-However, these traditional approaches present certain limitations when applied to practical scenarios.
-One significant drawback is that transition labels are typically non-symbolic and finite.
-A conventional transition is represented as $q\xrightarrow{a}q'$, where $a$ is a character from a finite alphabet.
-This simplistic representation can lead to a phenomenon known as transition explosion.
-For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
-Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the FA and FT operations highly inefficient.
-Furthermore, in practical applications, it is often necessary to consider infinite alphabets, such as the set of all integers.
-
-Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
-%
-These symbolic models utilize transition labels defined by first-order predicates over boolean algebras, allowing for more expressive representations.
-For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \mod 2 = 0$), denoting all even numbers.
-This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
-
-However, the formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
-Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse predicate representations of boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
-
-Prior work of CertiStr \cite{cpp/KanLRS22} has successfully formalized SFAs within Isabelle/HOL, demonstrating both the efficiency and effectiveness of SFAs in practice.
-However, the formalization of SFTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
-SFTs constitute a significantly more expressive and powerful theoretical framework compared to  SFAs, as evidenced by their capability to model complex string transformations such as replacement operations.
-%
-Furthermore, numerous studies have demonstrated the broad applicability of SFTs across diverse domains.
-The work of \cite{VeanesHLMB12Transducer} showcases SFTs in security-critical applications, particularly for cross-site scripting (XSS) prevention.
-\cite{uss/HooimeijerLMSV11} extends this security focus by applying SFTs to web application sanitizer analysis.
-In system verification, \cite{osdi/YaseenABCL20} employs SFTs for runtime behavior monitoring, while \cite{pldi/HuD17} demonstrates their utility in automated program transformation through systematic inversion techniques.
-
-In this work, we present a comprehensive formalization of SFTs.
-To address the extensibility challenges inherent in supporting diverse transition label theories, we adopt a refinement-based approach.
-At the abstraction level, transition labels are formalized through the fundamental mathematical concept of \emph{sets}.
-This abstraction facilitates subsequent refinement to various representations of Boolean algebras, such as intervals and arithmetic predicates, while maintaining theoretical consistency.
-
-The key operation of our formalization is the product operation between an SFT and an input regular language.
-Specifically, given an SFA $\mathcal{A}$ representing a regular language and an SFT $\mathcal{T}$, we define the product operation $\mathcal{T} \times\mathcal{A}$ that characterizes the output language generated by $\mathcal{T}$ when processing inputs from the language recognized by $\mathcal{A}$.
-
-In the refinement level, we implemented transition labels using an interval-based representation to examplify the refinement process.
-The formalization of interval algebra provides efficient set-theoretic operations—including membership checking, intersection, and difference computations—facilitating the refinement of transition labels from abstract sets to concrete intervals.
-Furthermore, leveraging the data refinement framework \cite{DBLP:conf/itp/Lammich13} in Isabelle/HOL, we store states and transitions using sophisticated data structures such as hashmaps and red-black trees, ensuring efficient automata manipulation.
-
-To evaluate the effectiveness and efficiency of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
-We evaluated our implementation using a set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
-The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
-
-In summary, our formalization makes the following contributions:
-\begin{enumerate}
-\item We present the first formalization of symbolic finite transducers in Isabelle/HOL proof assistant.
-\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
-\item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
-\item We demonstrate the practical utility of our formalization through its application to string constraint solving, specifically in modeling replacement operations with verified correctness guarantees.
-\end{enumerate}
-
-The remainder of this paper is organized as follows:
-Section \ref{sec:formalization} introduces the formalization of symbolic finite transducers.
-Section \ref{sec:product-operation} presents the product operation at the abstract level.
-Section \ref{sec_alg_refinement} details the algorithmic refinement of the product operation.
-Section \ref{sec-app-str-solver} demonstrates the application to string constraint solving.
-Section \ref{sec:related-work} discusses related work.
-Section \ref{sec:conclusion} concludes with future directions.
-
-\section{Symbolic Finite Transducers}
-\label{sec:formalization}
-
-We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL.
-
-Let $\mathcal{U}$ be a multi-sorted carrier set or background universe, which is equipped with functions and relations over the elements.
-We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
-We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type.
-
-A lambda term is defined as $\lambda x.~t$ of type $\tau_1 \rightarrow \tau_2$.
-When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate.
-Let $\phi$ be a predicate.
-We write $a\in \llbracket\phi \rrbracket$ if $\phi~a=\top$.
-For non-predicate lambda terms, we view them as functions that generate output elements of type $\tau_2$ given input terms of type $\tau_1$.
-With these notations and the above definitions, we can define SFTs as follows.
-
-\begin{definition}[Symbolic Finite Transducer]
-\label{def-sft}
-   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
-   \begin{itemize}
-   \item $\mathcal{Q}$ is a finite set of states,
-   \item $\mathcal{I}\subseteq \mathcal{Q}$ is the set of initial states,
-   \item $\mathcal{F} \subseteq\mathcal{Q}$ is the set of accepting states,
-   \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
-   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
-   \end{itemize}
-
-For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called an input,  then the application $(f~a)$ is the output.
-
-\end{definition}
-
-SFTs accept an input word and generate an output word. This can be defined by \emph{runs} of SFTs.
-An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i,
-\phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
-$\sigma$ is an accepting run when $q_n$ is an accepting state.
-
-For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ if and only if $a_i \in \llbracket\phi_i \rrbracket$ for $0 \leq i \leq n - 1$.
-When $w$ is accepted, run $\sigma$ generates an output sequence $w'= f_0~a_0, \ldots, f_{n-1}~a_{n-1}$.
-We define:
-\begin{itemize}
-\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
-\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, $ $(f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
-\end{itemize}
-
-If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
-For a trace $\pi$, we denote its input as $\mathit{in}(\pi)$ and its output as $\mathit{out}(\pi)$.
-%
-Given an SFT $\mathcal{T}$ and a word $w$, we define the \emph{product} operation of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with input $w$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\ \mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
-\end{split}
-\]
-
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an SFA $\mathcal{A}$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \\\text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
-\end{split}
-\]
-
-\begin{figure}[hbt!]
-  \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm and 2cm, on grid, auto]
-    % Make q0 and q1 horizontally wide apart
-    \node[state, initial] (q_0)   {$q_0$}; 
-    \node[state] (q_1) [right=4cm of q_0] {$q_1$}; 
-    \node[state, accepting] (q_2) [below=2.2cm of q_1] {$q_2$};
-
-    \path[->] 
-    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
-    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
-  \end{tikzpicture}
-    \caption{An example of SFT}
-    \label{fig-example-ft}
-\end{figure}
-
-    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
-The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
-The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation.
-For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.
-
-%The transition labels can be more complex.
-For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$.
-This symbolic label corresponds to the set of even natural numbers.
-
-\section{The Isabelle/HOL Formalization of SFTs}
-
-We present the basic architecture of our SFT formalization in Isabelle/HOL, which consists of three distinct layers.
-
-The \textbf{abstract layer} defines transition labels as \textbf{sets}, states are stored in sets as well.
-This abstract formalization, which abstracts away implementation details, enables straightforward correctness proofs for SFT operations.
-
-The \textbf{implementation layer} refines the abstract transition labels to a locale where a boolean algebra is defined as type variable \texttt{'b} together with a collection of operations and assumptions.
-State sets are refined using Refine\_Monadic \cite{Refine_Monadic-AFP}, a framework in Isabelle/HOL for data refinement that automatically transforms general sets into efficient data structures such as hashmaps and red-black trees.
-
-The \textbf{boolean algebra layer} refines the transition label locale to a concrete boolean algebra implementation. In this paper, we present an interval-based refinement.
-
-This three-layer architecture separates the fixed components (abstract and implementation layers) of the SFT formalization from the variable components (transition labels).
-This separation enables us to extend the SFT formalization to accommodate new transition label theories without modifying the first two layer formalization.
-
-\subsection{The Abstract Layer of SFTs}
-\label{sec:abstract-layer}
-
-We present the SFT formalization in abstract layer in Fig.
-\ref{fig-def-FT}.
-While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
-%
-The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
-This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a,$| |$'i,$| |$'b$|) |$\texttt{NFT}$| =
-	|$\mathcal{Q}_t$| :: "|$'q$| set"
-	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'i$|) LTTS"
-	|$\mathcal{I}_t$| :: "|$'q$| set"
-	|$\mathcal{F}_t$| :: "|$'q$| set"
-        |$\mathcal{M}_t$| :: "|$'i\Rightarrow ('a,~'b)$| Tlabel"
-type_synonym |$('q,~'a,~'i)$| LTTS = "|$('q \times ('a \text{ set option }\times 'i) \times 'q)$| set"
-type_synonym |$('a,~'b)$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
-	\end{lstlisting}
-\caption{The formalization of $SFTs$ in Isabelle/HOL}
-\label{fig-def-FT}
-\end{figure}
-
-Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
-Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
-A predicate's semantics as introduced before represents a set of elements that make the predicate true.
-But predicates have various different forms.
-For instance, the interval $[1-9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
-The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
-All these different forms of labels are abstracted as sets in our formalization.
-%
-The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements.
-This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
-
-The second element of type $'i$ serves as an index into the output function space. 
-%
-We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
-%
-The mapping $\mathcal{M}_t$ associates each index with a specific output function.
-These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
-This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria.
-Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
-
-%Now let us look at an example.
-In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
-This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}).
-%We can easily model this operation using our formalization of NFTs with the following steps:
-%\begin{itemize}
-%\item Step 1, construct a corresponding NFA from \texttt{pattern}).
-%\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
-%\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
-%\item Step 4, unlabel previous accepting states as non-accepting states. 
-%\end{itemize}
-
-%We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
-
-\paragraph{The product operation formalization.} We formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
-The SFA is formalized as a record type \texttt{NFA in Figure \ref{fig-def-FAs}}.
-An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
-%
-Therefore, we need to formalize $\varepsilon$SFAs, which is shown in Fig.
-\ref{fig-def-FAs} as \texttt{eNFA}.
-The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
-In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a$|) |$\texttt{NFA}$| =
-	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\mathcal{I}$| :: "|$'q$| set"
-	|$\mathcal{F}$| :: "|$'q$| set"
-
-record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
-	|$\mathcal{Q}_e$| :: "|$'q$| set"
-	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\Delta_e'$| :: "|$('q * 'q)$| set"
-	|$\mathcal{I}_e$| :: "|$'q$| set"
-	|$\mathcal{F}_e$| :: "|$'q$| set"
-
-type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
-	\end{lstlisting}
-\caption{The formalization of $\varepsilon$SFAs and SFAs}
-\label{fig-def-FAs}
-\end{figure}
-
-Having established these foundational definitions, we can now formalize the product operation.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| 
-   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$| 
-   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
-"productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
-  |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
-       |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
-        |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
-      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
-       |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
-       |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
-          |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
-  |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-       |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
-       |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
-      |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-        |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}$|
-        |$\land~\sigma_1\cap \sigma_2 \neq \emptyset~\land\exists x\in(\sigma_1\cap\sigma_2).$|
-        |$~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
-  |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
-  |$\mathcal{F}_e = \mathcal{F}_t~\mathcal{T} \times \mathcal{F}~\mathcal{A}$| |$\rrparenthesis$|"
-	\end{lstlisting}
-\caption{The formalization of product operation}
-\label{fig-def-FTProd}
-\end{figure}
-
-Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA. The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively. The 
-result of the product operation is an $\varepsilon$SFA.
-But we need to explain the role of parameter $\text{\texttt{F}}$.
-The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$.
-$\text{\texttt{F}}$ extends $f$ to apply $f$ to a set of elements.
-More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as follows:
-
-\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].
-
-The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions.
-In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels.
-Let us consider $\Delta_e$ first.
-
-\begin{enumerate}
-\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e.
-the input character is $\varepsilon$, and $f~\text{None} \neq \text{None}$.
-Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output "$((\mathcal{M}_t~\mathcal{T})~f)~\text{None}$".
-Remember that $f$ is just an index, $(\mathcal{M}_t~\mathcal{T})~f$ is the output function.
-
-\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
-The resulting output is $\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)$.
-\end{enumerate}
-
-The transitions in $\Delta_e'$ follow a similar pattern with analogous cases for $\varepsilon$ and non-$\varepsilon$ transitions.
-
-To establish the correctness specification of the product operation, we begin by formalizing the concept of SFT \emph{traces} as introduced in Definition \ref{def-sft}.
-In our formalization, traces are represented by the type $('a\;\text{option} \times 'b\;\text{option})\;\texttt{list}$, as shown in Figure \ref{fig-def-output}.
-%
-Given a trace $\pi$, we define two key projection functions:
-(1) \texttt{inputE}, which corresponds to ${in}(\pi)$ and extracts the input sequence
-(2) \texttt{outputE}, which corresponds to ${out}(\pi)$ and extracts the output sequence.
-
-The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-fun inputE :: 
-   "|$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$|"
-where
-  "inputE [] = []" |$\mid$|
-  "inputE ((Some a, |$\_$|) |$\#$| l) = 
-                      a |$\#$| (inputE l)" |$\mid$|
-  "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"
-
-fun outputE :: 
-    "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" 
-where
-  "outputE [] = []" |$\mid$|
-  "outputE ((_,Some a) # l) = 
-                    a # (outputE l)" |$\mid$|
-  "outputE ((_,None) # l) = (outputE l)"
-
-definition outputL :: 
-   "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
-                               ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. 
-    |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T}~\land$| 
-    |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land$||$ \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}"
-	\end{lstlisting}
-\caption{The formalization of traces in SFTs}
-\label{fig-def-output}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-lemma productT_correct:
-  fixes |$\mathcal{T}$| |$\mathcal{A}$| F
-  assumes 
-   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
-       |$\text{ F}~f~s = \text{None}"$|
-   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
-        |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
-        |$f~(\text{Some }e) = \text{ Some S}\})$|"
-      and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-  |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL 
-          |$\mathcal{T}~\mathcal{A}$|"
-
-	\end{lstlisting}
-\caption{The correctness lemma of the product operation}
-\label{fig-def-product-correct}
-\end{figure}
-
-Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation.
-The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
-The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.
-The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.
-
-\subsection{Implementation Layer Refinement}
-\label{sec_alg_refinement}
-Having established the abstract definition of the SFT product operation in Section \ref{sec:abstract-layer}, we now present its algorithmic refinement.
-This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
-
-\paragraph{Boolean Algebra Locale}
-
-Figure \ref{fig-bool-algebra-locale} presents the simplified definition of a boolean algebra locale.
-The type variable \texttt{'b} represents any boolean algebra, while the type variable \texttt{'a} denotes the element type of the sets that the boolean algebra represents.
-The locale defines a collection of operations and assumptions that characterize boolean algebra behavior, including operations for set semantics, emptiness checks, non-emptiness checks, intersection, difference, and element membership.
-The function "\texttt{sem}" provides the semantic interpretation of a boolean algebra as a set, with all other operations' assumptions defined in terms of this set semantics.
-
-The locale serves as an interface to boolean algebra implementations.
-Any concrete boolean algebra implementation must provide a concrete instantiation of this locale, which includes defining the type variable \texttt{'b} and providing implementations for all operations along with proofs of the required assumptions.
-
-\begin{figure}[hbt!]
-\begin{lstlisting}[mathescape=true]
-locale bool_algebra =
- fixes sem :: " 'b $\Rightarrow$ 'a::ord set"
- fixes empty:: "'b $\Rightarrow$ bool"
- fixes nempty:: "'b $\Rightarrow$ bool"
- fixes intersect:: "'b $\Rightarrow$ 'b $\Rightarrow$ 'b"
- fixes diff:: "'b $\Rightarrow$ 'b $\Rightarrow$ 'b"
- fixes elem:: "'a  $\Rightarrow$ 'b $\Rightarrow$ bool"
- assumes
-  empty_sem: "empty s = (sem s = {})"
-                                    and
-  nempty_sem: "nempty_intervals s =
-                     ($\neg$ (empty s))" and
-  inter_sem: "sem (intersect s1 s2) =
-               (sem s1) $\cap$ (sem s2)" and
-  diff_sem: "sem (diff f1 f2 s1 s2) =
-               (sem s1) - (sem s2)" and
-  elem_sem: "elem a s $\equiv$  (a $\in$ sem s)"
-\end{lstlisting}
-\caption{The boolean algebra locale}
-\label{fig-bool-algebra-locale}
-\end{figure}
-
-\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.} based on Refined\_monadic framework.
-
-Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}.
-These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
-
-The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets.
-This function employs the \texttt{FOREACH} construct, a higher-order iteration operator analogous to OCaml's \texttt{Set.fold}.
-Specifically, given a set $S$, a function $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$, and an initial accumulator $I$ of type $'b$, the expression $\texttt{FOREACH}~S~f~I$ systematically applies $f$ to each element in $S$, accumulating results in a principled manner.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT_impl where
-"productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
-    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) 
-    (nfa_states |$\mathcal{A}$|);
-    (D1, D2) |$\leftarrow$| trans_comp_imp 
-     (nft_tranfun |$\mathcal{T}$|) F fe (nft_trans |$\mathcal{T}$|) 
-      (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
-    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|)
-      (nfa_initial |$\mathcal{A}$|);
-    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) 
-      (nfa_accepting |$\mathcal{A}$|);
-    RETURN (Q, D1, D2, I, F)
-  }"
-\end{lstlisting}
-\caption{The computation of SFT product}
-\label{fig-compute-nft-product}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition prods_imp where
-"prods_imp Q1 Q2 =
-   FOREACH {q. q |$\in$| Q1} (|$\lambda$| q Q. do {
-    S |$\leftarrow$| FOREACH {q. q |$\in$| Q2}
-      (|$\lambda$| q|$'$| Q|$'$|. RETURN ({(q,q|$'$|)} |$\cup$| Q|$'$|)) |$\emptyset$|;
-    RETURN (Q |$\cup$| S)
-   }) |$\emptyset$|"
-\end{lstlisting}
-\caption{The computation of Cartesian product of two state sets}
-\label{fig-def-prods_imp}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition trans_comp_imp where
-"trans_comp_imp M F fe T1 T2 Q =
-  FOREACH {t. t |$\in$| T1}
-   (|$\lambda$|(q,(|$\alpha$|,f),q|$'$|) (D1,D2). 
-    (if (|$\alpha$|=None) then 
-     (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe 
-                              T2 D1 D2)
-     else
-     (subtrans_comp M q (the |$\alpha$|) f 
-            q|$'$| F fe T2 D1 D2))) (|$\emptyset$|, |$\emptyset$|)"
-\end{lstlisting}
-\caption{The computation of \texttt{trans\_comp\_imp}}
-\label{fig-def-prods-imp}
-\end{figure}
-
-A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ (corresponding to $\Delta_e$) and $\texttt{D2}$ (corresponding to $\Delta_e'$) in Figure \ref{fig-compute-nft-product}.
-As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
-This function decomposes the synchronization process into two distinct cases, each handled by a specialized function:
-
-\begin{enumerate}
-  \item \texttt{subtrans\_comp\_$\varepsilon$}: Processes $\varepsilon$-transitions in the SFT, where transitions consume no input but may produce output
-  \item \texttt{subtrans\_comp}: Processes standard transitions in the SFT, where both input consumption and output generation may occur
-\end{enumerate}
-
-\begin{figure}[hbt!]
-\begin{lstlisting}[language=SMTLIB]
-definition subtrans_comp where
-"subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
- FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
-  (if (nempty (intersect |$\alpha$| |$\alpha'$|)) then
-    do {
-      D1 |$\leftarrow$| 
-       (if (F (M f) (intersect |$\alpha$| |$\alpha'$|)) 
-          |$\neq$| None) then
-         let |$\alpha_i$| = the (F (M f) 
-              (intersect |$\alpha$| |$\alpha'$|)) in
-         RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
-                |$\cup$| D1
-        else RETURN D1);
-      D2 |$\leftarrow$| 
-       (if fe (M f) (intersect |$\alpha$| |$\alpha'$|) 
-        then 
-         RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 
-        else RETURN D2);
-      RETURN (D1, D2) }
-   else (RETURN (D1, D2)))) (D1, D2)"
-    \end{lstlisting}
-    \caption{The computation of \texttt{subtrans\_comp}}
-    \label{fig-def-subtrans_comp}
-    \end{figure}
-
-    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
-For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
-For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
-
-    \begin{enumerate}
-      \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
-      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
-The checking is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
-    \end{enumerate}
-
-  \begin{figure}[hbt!]
-    \begin{lstlisting}
-  lemma productT_imp_correct:
-  assumes finite_TT: "|$\text{finite } (\Delta_t~\mathcal{T})$|"
-      and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
-      and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
-      and finite_TQ: "|$\text{finite } (\mathcal{Q}_t~\mathcal{T})$|"
-      and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
-      and finite_TI: "|$\text{finite } (\mathcal{I}_t~\mathcal{T})$|"
-      and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
-      and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
-  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe}\leq$|
-            |$\text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
-  \end{lstlisting}
-  \caption{The refinement relation between \texttt{productT\_imp} and \texttt{productT}}
-  \label{fig-def-productT_imp_correct}
-  \end{figure}
-
-  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
-This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.
-
-  The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$.
-More precisely,  the concrete implementation must produce an $\varepsilon$SFA that is structurally equivalent or isomorphic to the one produced by the abstract algorithm \texttt{productT}.
-
-  To establish this equivalence, we must prove that the $\varepsilon$SFAs produced by \texttt{productT\_imp} and \texttt{productT} are isomorphic in all essential components: the set of states, transition relations, $\varepsilon$-transition relations, initial and accepting state sets.
-
-The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
-The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 
-In our formalization, we refine the sets of storing states and transitions to red-black trees.
-
-\subsection{Interval Implementation}
-
-Finally, we present the implementation of the interval algebra.
-An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
-To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
-This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
-
-Throughout the following discussion, we use the term "interval" to refer to interval lists.
-Our formalization provides all operations in the boolean algebra locale.
-For instance, we implement \texttt{intersect} for intervals, which computes the intersection of two intervals $i_1$ and $i_2$, yielding an interval $i$ such that $\texttt{sem}~i = \texttt{sem}~i_1 \cap \texttt{sem}~i_2$
-
-To facilitate formal reasoning and optimize performance, in addition to the above locale operations, we introduce a canonical form for interval.
-An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
-\begin{enumerate}
-  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
-  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
-\end{enumerate}
-
-We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
-This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
-
 \section{Applications and Evaluation}
@@
 
-  \emph{Comparison with SMT-LIB Semantics.} It is important to note that our formalization differs from the standard SMT-LIB semantics for the replacement operation.
-In SMT-LIB, the \texttt{str.replace} operation is defined to replace only the \emph{first} occurrence of a substring matching the given pattern.
-In contrast, our semantics allows for a more general interpretation where any matching substring may be replaced.
-But it still replaces only one occurrence of them.
-
-  The replacement operation model described above replaces a single occurrence of any substring matching the regular expression pattern.
-To support different replacement semantics, we may also need to model left-most replacement (which replaces the first occurrence) and replace-all operations (which substitute all occurrences of matching substrings).
-
-  The approach to modeling left-most replacement is illustrated in Figure \ref{fig:replace-left-most}.
-This construction involves two SFTs: one for non-matching characters and one for the matching pattern.
-The non-matching SFT is constructed by first computing the negation of the pattern (the complement of the regular expression's NFA) and converting it into an SFT by adding the identity function $\lambda x.x$.
-The matching SFT is constructed similarly to the previous example in Figure \ref{fig-rearranged-automata}, but without self-loops at the initial state, creating an SFT that processes the pattern and generates the replacement string.
-
-  The two SFTs are composed by adding an $\varepsilon$-transition from the accepting state of the non-matching SFT to the initial state of the matching SFT.
-This transition enables the left-most replacement operation to proceed to pattern matching after processing any non-matching prefix.
+  The replacement operation model described above replaces a single occurrence of any substring matching the regular expression pattern. To support different replacement semantics, 
+  we model replace-all operations (which substitute all occurrences of matching substrings).
 
   The approach to modeling replace-all operations is illustrated in Figure \ref{fig:replace-all}.
-This construction is similar to left-most replacement but includes an additional $\varepsilon$-transition from the accepting state of the matching SFT back to the initial state of the non-matching SFT.
-This bidirectional connection allows the replace-all operation to continue processing the input string and replace all occurrences of matching substrings.
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-\end{tikzpicture}
-\caption{The modeling of left-most replacement operation.}
-\label{fig:replace-left-most}
-\end{figure}
 
@@
 
-CertiStrR implements three different pattern matching introduced above: (1) singleton matching at any position, (2) left-most matching, and (3) replace-all matching.
+CertiStrR implements two different pattern matching introduced above: (1) singleton matching at any position, (2) replace-all matching.
 The pattern can be a regular expression or a constant string, which is a special case of regular expression.
 
-Listing \ref{lst-smtlib-code} demonstrates the regular expression variant through an example that replaces numeric sequences with the string "\texttt{NUM}".
+% Listing \ref{lst-smtlib-code} demonstrates the regular expression variant through an example that replaces numeric sequences with the string "\texttt{NUM}".
 The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
-But the left-matching will return \texttt{UNSAT}.
+% %
 %
-
-\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}, float=htbp]
-  (set-logic QF_S)
-  (declare-fun a () String)
-  (declare-fun b () String)
-  (assert (= a "2024,2025"))
-  (assert (= b "2024,NUM"))
-  (assert (= b (str.replace_re a 
-      (re.+ (re.range "0" "9")) "NUM")))
-  (check-sat)
-  \end{lstlisting}
+%
+% \begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}, float=htbp]
+%   (set-logic QF_S)
+%   (declare-fun a () String)
+%   (declare-fun b () String)
+%   (assert (= a "2024,2025"))
+%   (assert (= b "2024,NUM"))
+%   (assert (= b (str.replace_re a 
+%       (re.+ (re.range "0" "9")) "NUM")))
+%   (check-sat)
+%   \end{lstlisting}
 
 We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are categorized into three groups based on their pattern matching semantics: (1) general singleton matching, (2) left-most matching, and (3) replace-all matching.
+The benchmarks are categorized into three groups based on their pattern matching semantics: (1) general singleton matching and (2) replace-all matching.
 Patterns can be either regular expressions or constant strings.
@@
       \texttt{replace singleton} & 142/173/8 & 0.27 & 323\\
-      \texttt{replace left-match} & 84/4/10 & 0.36 & 98\\
       \texttt{replace all} & 84/4/10 & 0.36 & 98\\
@@
 The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a one-minute time limit per test.
-The results show average execution times of 0.27 seconds for \texttt{replace singleton}, 0.36 seconds for \texttt{left-match}, and 0.36 seconds for \texttt{replace all}.
+The results show average execution times of 0.27 seconds for \texttt{replace singleton} and 0.36 seconds for \texttt{replace all}.
 Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
@@
 This subsection discusses the development effort required for our certified SFT formalization. Table \ref{tab:abstract_impl} provides an overview of this effort.
+Note that, to model replacement operation, we not only need the formalization of SFTs but also the formalization of negation of an SFA, $\epsilon$NFA.
+
 All efforts listed in the table are additional to the existing CertiStr framework that we build upon.
@@
 \end{table}
-
-\section{Related Work}
-\label{sec:related-work}
-
-\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
-This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
-Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
-
-\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
-These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
-Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.
-
-\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
-To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
-Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.
-
-\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
-While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
-As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
-We believe that our work will benefit the community by providing a formal foundation for string solvers development.
-
-\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains.
-For example, the work by Shi et al.
-\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL.
-Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.
-
-To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}.
-This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
-The experimental results confirm both the efficiency and effectiveness of our approach.
-While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.
-
-Future work includes extending our symbolic formalization in two key directions:
-%
-(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
-(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
-These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.
-
-\bibliography{literature}
-\end{document}

### 2025-08-08 Paper/introduction.tex
main.tex split

--- 2025-07-29 Paper/main.tex
@@
-\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%This is a template for producing LIPIcs articles. 
-%See lipics-v2021-authors-guidelines.pdf for further information.
-%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
-%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
-%for section-numbered lemmas etc., use "numberwithinsect"
-%for enabling cleveref support, use "cleveref"
-%for enabling autoref support, use "autoref"
-%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
-%for enabling thm-restate support, use "thm-restate"
-%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
-%for producing a PDF according the PDF/A standard, add "pdfa"
-
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-\usepackage{stmaryrd}
-\usepackage{tikz}
-\usetikzlibrary{shapes,arrows}
-\usetikzlibrary{intersections}
-\usetikzlibrary{automata}
-\usetikzlibrary{positioning}
-\usepackage{booktabs}
-%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
-\usepackage{listings}
-\usepackage{xcolor}
-
-\lstdefinelanguage{SMTLIB}{
-  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
-  sensitive=true,
-  morecomment=[l];,
-  morestring=[b]"
-}
-\lstset{
-  language=SMTLIB,
-  basicstyle=\ttfamily\small,
-  keywordstyle=\color{blue}\bfseries,
-  commentstyle=\color{green!50!black},
-  numbers=left,
-  numberstyle=\tiny\color{gray},
-  stepnumber=1,
-  numbersep=5pt,
-  backgroundcolor=\color{gray!10},
-  frame=single,
-  breaklines=true,
-  captionpos=b,
-  showstringspaces=false
-}
-
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\lstdefinelanguage{Isabelle}{
-    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
-              where,fixes,assumes,shows,if,then,else,let,in,case,of,
-              record,datatype,type_synonym,fun,function,primrec},
-    sensitive=true,
-    morecomment=[s]{(*}{*)},
-    morestring=[b]",
-  }
-
-\lstset{
-  language=Isabelle,
-	basicstyle=\ttfamily,
-	keywordstyle=\color{blue}\bfseries,
-	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and},
-	escapeinside={|}{|},
-  sensitive=true
-}
-
-\begin{document}
-
-\special{papersize=8.5in,11in}
-\setlength{\pdfpageheight}{\paperheight}
-\setlength{\pdfpagewidth}{\paperwidth}
-
-\title{Certified Symbolic Finite Transducers: Formalization and Applications to String Manipulation}
-
-\begin{abstract}
-Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
-FTs have wide applications, including program analysis, string constraint solving, and security-critical sanitizer analysis.
-However, there is currently no scalable formalization of FTs.
-This is primarily because transition labels in FTs are not formalized symbolically, resulting in transition explosion and significant performance bottlenecks in practical applications.
-
-In this paper, we present a formalization of FTs in the symbolic setting using Isabelle/HOL, where transition labels can be instantiated with various Boolean algebras, such as intervals and arithmetic predicates.
-This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
-%
-To evaluate the effectiveness of our formalization, we apply the formalized SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for preventing XSS attacks, and (2) a string solver that models replacement operations with various semantics, including left-most and replace-all matching.
-Experimental results demonstrate that our approach achieves computational efficiency in constraint-solving scenarios.
-\end{abstract}
-
-\maketitle
-
 \section{Introduction}
@@
 Section \ref{sec:conclusion} concludes with future directions.
-
-\section{Formalization of SFTs}
-\label{sec:formalization}
-
-We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL. 
-\subsection{The Mathematical Definition of SFTs}
-
-Let $\mathcal{U}$ be a multi-sorted carrier set or background universe, which is equipped with functions and relations over the elements.
-We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
-We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type.
-
-A lambda term is defined as $\lambda x.~t$ of type $\tau_1 \rightarrow \tau_2$.
-When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate.
-Let $\phi$ be a predicate.
-We write $a\in \llbracket\phi \rrbracket$ if $\phi~a=\top$.
-For non-predicate lambda terms, we view them as functions that generate output elements of type $\tau_2$ given input terms of type $\tau_1$.
-With these notations and the above definitions, we can define SFTs as follows.
-
-\begin{definition}[Symbolic Finite Transducer]
-\label{def-sft}
-   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
-   \begin{itemize}
-   \item $\mathcal{Q}$ is a finite set of states,
-   \item $\mathcal{I}\subseteq \mathcal{Q}$ is the set of initial states,
-   \item $\mathcal{F} \subseteq\mathcal{Q}$ is the set of accepting states,
-   \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
-   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
-   \end{itemize}
-
-For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called an input,  then the application $(f~a)$ is the output.
-
-\end{definition}
-
-SFTs accept an input word and generate an output word. This can be defined by \emph{runs} of SFTs.
-An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i,
-\phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
-$\sigma$ is an accepting run when $q_n$ is an accepting state.
-
-For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ if and only if $a_i \in \llbracket\phi_i \rrbracket$ for $0 \leq i \leq n - 1$.
-When $w$ is accepted, run $\sigma$ generates an output sequence $w'= f_0~a_0, \ldots, f_{n-1}~a_{n-1}$.
-We define:
-\begin{itemize}
-\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
-\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, $ $(f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
-\end{itemize}
-
-If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
-For a trace $\pi$, we denote its input as $\mathit{in}(\pi)$ and its output as $\mathit{out}(\pi)$.
-%
-Given an SFT $\mathcal{T}$ and a word $w$, we define the \emph{product} operation of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with input $w$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
-\end{split}
-\]
-
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an SFA $\mathcal{A}$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \\\text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
-\end{split}
-\]
-
-\begin{figure}[hbt!]
-  \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm and 2cm, on grid, auto]
-    % Make q0 and q1 horizontally wide apart
-    \node[state, initial] (q_0)   {$q_0$}; 
-    \node[state] (q_1) [right=4cm of q_0] {$q_1$}; 
-    \node[state, accepting] (q_2) [below=2.2cm of q_1] {$q_2$};
-
-    \path[->] 
-    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
-    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
-  \end{tikzpicture}
-    \caption{An example of SFT}
-    \label{fig-example-ft}
-\end{figure}
-
-    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
-The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
-The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation.
-For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.
-
-%The transition labels can be more complex.
-For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$.
-This symbolic label corresponds to the set of even natural numbers.
-
-\subsection{The Isabelle/HOL Formalization of SFTs}
-
-Firstly, we present the basic idea of our formalization of SFTs in Isabelle/HOL, which consists of three layers.
-
-\begin{figure}[hbt!]
-\centering
-\begin{tikzpicture}[
-  node distance=1cm,
-  every node/.style={font=\small},
-  layer/.style={rectangle, draw, rounded corners, fill=blue!10, minimum width=5.5cm, minimum height=1.2cm, align=center},
-  arrow/.style={->, thick}
-]
-
-% Abstract Layer
-\node[layer] (abstract) {%
-  \textbf{Abstract Layer}\\
-  FT with transition labels as \textbf{sets}
-};
-
-% Locale Layer
-\node[layer, below=of abstract] (locale) {%
-  \textbf{Implementation Layer}\\
-  FT with transition labels as a \\boolean algebra locale
-};
-
-% Implementation Layer
-\node[layer, below=of locale] (impl) {%
-  \textbf{Boolean Algebra Layer}\\
-  Concrete Boolean algebra implementations\\
-  (e.g., intervals, bitvectors, predicates)
-};
-
-% Draw arrows with refinement labels
-\draw[arrow] (abstract) -- node[right] {refinement} (locale);
-\draw[arrow] (locale) -- node[right] {refinement} (impl);
-
-\end{tikzpicture}
-\caption{Three-layered formalization of SFTs in Isabelle/HOL. 
-\textbf{Abstract Layer:} Transition labels are modeled as sets, representing the most general case.
-\textbf{Locale Layer:} Transition labels are abstracted as a type variable \texttt{'a} within a locale, supporting any Boolean algebra.
-\textbf{Implementation Layer:} Concrete Boolean algebra instances, such as intervals, bitvectors, or predicates, are provided.}
-\label{fig:layered-formalization}
-\end{figure}
-
-As we have discussed the diversity of transition labels in Section \ref{sec:introduction}, we now present an extensible formalization of SFTs that accommodates this variety.
-Our approach leverages the refinement framework (Refine\_Monadic \cite{Refine_Monadic-AFP}) in Isabelle/HOL to achieve the flexibility and extensibility of transition labels modeling.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a,$| |$'i,$| |$'b$|) |$\texttt{NFT}$| =
-	|$\mathcal{Q}_t$| :: "|$'q$| set"
-	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'i$|) LTTS"
-	|$\mathcal{I}_t$| :: "|$'q$| set"
-	|$\mathcal{F}_t$| :: "|$'q$| set"
-        |$\mathcal{M}_t$| :: "|$'i\Rightarrow ('a,~'b)$| Tlabel"
-type_synonym |$('q,~'a,~'i)$| LTTS = "|$('q \times ('a \text{ set option }\times 'i) \times 'q)$| set"
-type_synonym |$('a,~'b)$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
-	\end{lstlisting}
-\caption{The formalization of $SFTs$ in Isabelle/HOL}
-\label{fig-def-FT}
-\end{figure}
-
-Figure \ref{fig-def-FT} presents our formalization of SFTs in Isabelle/HOL.
-While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
-%
-The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
-This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
-
-Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
-Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
-A predicate's semantics as introduced before represents a set of elements that make the predicate true.
-But predicates have various different forms.
-For instance, the interval $[1-9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
-The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
-All these different forms of labels are abstracted as sets in our formalization.
-%
-The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements.
-This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
-
-The second element of type $'i$ serves as an index into the output function space. 
-%
-We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
-%
-The mapping $\mathcal{M}_t$ associates each index with a specific output function.
-These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
-This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria.
-Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
-
-%Now let us look at an example.
-In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
-This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}).
-%We can easily model this operation using our formalization of NFTs with the following steps:
-%\begin{itemize}
-%\item Step 1, construct a corresponding NFA from \texttt{pattern}).
-%\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
-%\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
-%\item Step 4, unlabel previous accepting states as non-accepting states. 
-%\end{itemize}
-
-%We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
-
-\section{The Product Operation of SFTs}
-\label{sec:product-operation}
-
-In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
-An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
-Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of SFAs with $\varepsilon$-transitions, we extend the framework in CertiStr with a formalization of symbolic SFAs with $\varepsilon$-transitions (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
-In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a$|) |$\texttt{NFA}$| =
-	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\mathcal{I}$| :: "|$'q$| set"
-	|$\mathcal{F}$| :: "|$'q$| set"
-
-record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
-	|$\mathcal{Q}_e$| :: "|$'q$| set"
-	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\Delta_e'$| :: "|$('q * 'q)$| set"
-	|$\mathcal{I}_e$| :: "|$'q$| set"
-	|$\mathcal{F}_e$| :: "|$'q$| set"
-
-type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
-	\end{lstlisting}
-\caption{The formalization of $\varepsilon$SFAs and SFAs}
-\label{fig-def-FAs}
-\end{figure}
-
-Figure \ref{fig-def-FAs} presents the formalization of both SFAs and $\varepsilon$SFAs using Isabelle/HOL record types \texttt{NFA} and \texttt{eNFA}, respectively.
-The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
-
-Having established these foundational definitions, we can now formalize the product operation.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| 
-   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$| 
-   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
-"productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
-  |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
-       |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
-        |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
-      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
-       |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
-       |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
-          |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
-  |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-       |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
-       |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
-      |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-        |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}$|
-        |$\land~\sigma_1\cap \sigma_2 \neq \emptyset~\land\exists x\in(\sigma_1\cap\sigma_2).$|
-        |$~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
-  |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
-  |$\mathcal{F}_e = \mathcal{F}_t~\mathcal{T} \times \mathcal{F}~\mathcal{A}$| |$\rrparenthesis$|"
-	\end{lstlisting}
-\caption{The formalization of product operation}
-\label{fig-def-FTProd}
-\end{figure}
-
-Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA. The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively. The 
-result of the product operation is an $\varepsilon$SFA.
-But we need to explain the role of parameter $\text{\texttt{F}}$.
-The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$.
-$\text{\texttt{F}}$ extends $f$ to apply $f$ to a set of elements.
-More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as follows:
-
-\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].
-
-The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions.
-In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels.
-Let us consider $\Delta_e$ first.
-
-\begin{enumerate}
-\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e.
-the input character is $\varepsilon$, and $f~\text{None} \neq \text{None}$.
-Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output "$((\mathcal{M}_t~\mathcal{T})~f)~\text{None}$".
-Remember that $f$ is just an index, $(\mathcal{M}_t~\mathcal{T})~f$ is the output function.
-
-\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
-The resulting output is $\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)$.
-\end{enumerate}
-
-The transitions in $\Delta_e'$ follow a similar pattern with analogous cases for $\varepsilon$ and non-$\varepsilon$ transitions.
-
-To establish the correctness specification of the product operation, we begin by formalizing the concept of SFT \emph{traces} as introduced in Definition \ref{def-sft}.
-In our formalization, traces are represented by the type $('a\;\text{option} \times 'b\;\text{option})\;\texttt{list}$, as shown in Figure \ref{fig-def-output}.
-%
-Given a trace $\pi$, we define two key projection functions:
-(1) \texttt{inputE}, which corresponds to ${in}(\pi)$ and extracts the input sequence
-(2) \texttt{outputE}, which corresponds to ${out}(\pi)$ and extracts the output sequence.
-
-The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-fun inputE :: 
-   "|$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$|"
-where
-  "inputE [] = []" |$\mid$|
-  "inputE ((Some a, |$\_$|) |$\#$| l) = 
-                      a |$\#$| (inputE l)" |$\mid$|
-  "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"
-
-fun outputE :: 
-    "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" 
-where
-  "outputE [] = []" |$\mid$|
-  "outputE ((_,Some a) # l) = 
-                    a # (outputE l)" |$\mid$|
-  "outputE ((_,None) # l) = (outputE l)"
-
-definition outputL :: 
-   "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
-                               ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. 
-    |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T}~\land$| 
-    |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land$||$ \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}"
-	\end{lstlisting}
-\caption{The formalization of traces in SFTs}
-\label{fig-def-output}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-lemma productT_correct:
-  fixes |$\mathcal{T}$| |$\mathcal{A}$| F
-  assumes 
-   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
-       |$\text{ F}~f~s = \text{None}"$|
-   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
-        |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
-        |$f~(\text{Some }e) = \text{ Some S}\})$|"
-      and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-  |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL 
-          |$\mathcal{T}~\mathcal{A}$|"
-
-	\end{lstlisting}
-\caption{The correctness lemma of the product operation}
-\label{fig-def-product-correct}
-\end{figure}
-
-Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation.
-The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
-The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.
-The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.
-
-\section{Algorithm Level Refinement}
-\label{sec_alg_refinement}
-Having established the abstract definition of the SFT product operation in Section \ref{sec:product-operation}, we now present its algorithmic refinement.
-This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
-
-\subsection{Intervals}
-
-An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
-To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
-This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
-
-Throughout the following discussion, we use the term "interval" to refer to interval lists. Our formalization provides a collection of interval operations through the following interface:
-
-\begin{enumerate}
-  \item $\texttt{semIs}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semIs}~(i,j) = \{e \mid i \leq e \leq j\}$.
-  \item $\texttt{emptyIs}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semIs}~i = \emptyset$.
-  \item $\texttt{nemptyIs}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semIs}~i \neq \emptyset$.
-  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \cap \texttt{semIs}~i_2$.
-  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \setminus \texttt{semIs}~i_2$.
-\end{enumerate}
-
-To facilitate formal reasoning and optimize performance, we introduce a canonical form for interval.
-An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
-\begin{enumerate}
-  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
-  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
-\end{enumerate}
-
-We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
-This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
-
-\subsection{Algorithmic Implementation of the Product Operation}
-
-In this subsection, we present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.}.
-%
-Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}.
-These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
-
-The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets.
-This function employs the \texttt{FOREACH} construct, a higher-order iteration operator analogous to OCaml's \texttt{Set.fold}.
-Specifically, given a set $S$, a function $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$, and an initial accumulator $I$ of type $'b$, the expression $\texttt{FOREACH}~S~f~I$ systematically applies $f$ to each element in $S$, accumulating results in a principled manner.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT_impl where
-"productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
-    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) 
-    (nfa_states |$\mathcal{A}$|);
-    (D1, D2) |$\leftarrow$| trans_comp_imp 
-     (nft_tranfun |$\mathcal{T}$|) F fe (nft_trans |$\mathcal{T}$|) 
-      (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
-    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|)
-      (nfa_initial |$\mathcal{A}$|);
-    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) 
-      (nfa_accepting |$\mathcal{A}$|);
-    RETURN (Q, D1, D2, I, F)
-  }"
-\end{lstlisting}
-\caption{The computation of SFT product}
-\label{fig-compute-nft-product}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition prods_imp where
-"prods_imp Q1 Q2 =
-   FOREACH {q. q |$\in$| Q1} (|$\lambda$| q Q. do {
-    S |$\leftarrow$| FOREACH {q. q |$\in$| Q2}
-      (|$\lambda$| q|$'$| Q|$'$|. RETURN ({(q,q|$'$|)} |$\cup$| Q|$'$|)) |$\emptyset$|;
-    RETURN (Q |$\cup$| S)
-   }) |$\emptyset$|"
-\end{lstlisting}
-\caption{The computation of Cartesian product of two state sets}
-\label{fig-def-prods_imp}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition trans_comp_imp where
-"trans_comp_imp M F fe T1 T2 Q =
-  FOREACH {t. t |$\in$| T1}
-   (|$\lambda$|(q,(|$\alpha$|,f),q|$'$|) (D1,D2). 
-    (if (|$\alpha$|=None) then 
-     (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe 
-                              T2 D1 D2)
-     else
-     (subtrans_comp M q (the |$\alpha$|) f 
-            q|$'$| F fe T2 D1 D2))) (|$\emptyset$|, |$\emptyset$|)"
-\end{lstlisting}
-\caption{The computation of \texttt{trans\_comp\_imp}}
-\label{fig-def-prods-imp}
-\end{figure}
-
-A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ (corresponding to $\Delta_e$) and $\texttt{D2}$ (corresponding to $\Delta_e'$) in Figure \ref{fig-compute-nft-product}.
-As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
-This function decomposes the synchronization process into two distinct cases, each handled by a specialized function:
-
-\begin{enumerate}
-  \item \texttt{subtrans\_comp\_$\varepsilon$}: Processes $\varepsilon$-transitions in the SFT, where transitions consume no input but may produce output
-  \item \texttt{subtrans\_comp}: Processes standard transitions in the SFT, where both input consumption and output generation may occur
-\end{enumerate}
-
-\begin{figure}[hbt!]
-\begin{lstlisting}[language=SMTLIB]
-definition subtrans_comp where
-"subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
- FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
-  (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then
-    do {
-      D1 |$\leftarrow$| 
-       (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) 
-          |$\neq$| None) then
-         let |$\alpha_i$| = the (F (M f) 
-              (intersectIs |$\alpha$| |$\alpha'$|)) in
-         RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
-                |$\cup$| D1
-        else RETURN D1);
-      D2 |$\leftarrow$| 
-       (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) 
-        then 
-         RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 
-        else RETURN D2);
-      RETURN (D1, D2) }
-   else (RETURN (D1, D2)))) (D1, D2)"
-    \end{lstlisting}
-    \caption{The computation of \texttt{subtrans\_comp}}
-    \label{fig-def-subtrans_comp}
-    \end{figure}
-
-    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
-For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
-For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
-
-    \begin{enumerate}
-      \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
-      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
-The checking is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
-    \end{enumerate}
-
-  \begin{figure}[hbt!]
-    \begin{lstlisting}
-  lemma productT_imp_correct:
-  assumes finite_TT: "|$\text{finite } (\Delta_t~\mathcal{T})$|"
-      and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
-      and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
-      and finite_TQ: "|$\text{finite } (\mathcal{Q}_t~\mathcal{T})$|"
-      and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
-      and finite_TI: "|$\text{finite } (\mathcal{I}_t~\mathcal{T})$|"
-      and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
-      and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
-  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe}\leq$|
-            |$\text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
-  \end{lstlisting}
-  \caption{The refinement relation between \texttt{productT\_imp} and \texttt{productT}}
-  \label{fig-def-productT_imp_correct}
-  \end{figure}
-
-  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
-This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.
-
-  The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$.
-More precisely,  the concrete implementation must produce an $\varepsilon$SFA that is structurally equivalent or isomorphic to the one produced by the abstract algorithm \texttt{productT}.
-
-  To establish this equivalence, we must prove that the $\varepsilon$SFAs produced by \texttt{productT\_imp} and \texttt{productT} are isomorphic in all essential components: the set of states, transition relations, $\varepsilon$-transition relations, initial and accepting state sets.
-
-The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
-The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 
-In our formalization, we refine the sets of storing states and transitions to red-black trees.
-
-\section{Applications and Evaluation}
-
-We evaluate our formalization and implementation of the SFT product operation in two key application domains: string sanitization and string replacement operations in string solving.
-These applications demonstrate the practical utility of our certified SFT product operation in real-world scenarios.
-\subsection{Sanitizer}
-
-One Important application of our formalization is in the context of string sanitization, specifically for web applications.
-Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before being rendered in a web browser.
-Listing \ref{lst-html-encode} presents a C\# code snippet that implements an HTML encoding function, which is a sanitization operation.
-The function iterates through each character in the input string and encodes characters that are not safe for HTML rendering.
-
-\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
-static string EncodeHtml(string t)
-{
-  if (t == null) { return null; }
-  if (t.Length == 0) {
-    return string.Empty;
-  }
-  StringBuilder builder = new
-    StringBuilder("", t.Length * 2);
-  foreach (char c in t) {
-    if ((((c > ''') && (c < '{')) |$\mid\mid$|
-         ((c > '@') && (c < '['))) |$\mid\mid$|
-         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
-         (((c == '.') || (c == ',')) |$\mid\mid$|
-         ((c == '-') || (c == '_'))))) {
-          builder.Append(c);
-      }
-      else {
-          builder.Append("&#" +
-          ((int) c).ToString() + ";");
-      }
-  }
-  return builder.ToString();
-}
-\end{lstlisting}
-
-Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
-The transducer accepts any input string and produces an output where safe characters are preserved using the identity function $\texttt{id}$, while unsafe characters are encoded as HTML entities.
-The set of safe characters is defined as $\texttt{safe\_chars} = [\texttt{' '}, \texttt{','}, \texttt{'.'}, \texttt{'-'}, \texttt{'\_'}] \cup [\texttt{'0'}-\texttt{'9'}] \cup [\texttt{'A'}-\texttt{'Z'}] \cup [\texttt{'a'}-\texttt{'z'}]$, which can be efficiently represented using the interval algebra we formalized.
-Unsafe characters are transformed using the encoding function $\texttt{encode}(c) = \texttt{``\&\#''} + \texttt{toString}(\texttt{ord}(c)) + \texttt{``\;''}$.
-When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
-  % States
-  \node[state, initial, accepting] (q0) {$q_0$};
-  \node[state] (q1) [right=3.5cm of q0] {$q_1$};
-  \node[state] (q2) [below=of q0] {$q_2$};
-
-  \path[->] 
-    % Transition for safe characters
-    (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
-    % Transition for unsafe characters  
-    (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
-    % Epsilon transitions back to q0
-    (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
-    (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
-\end{tikzpicture}
-\caption{SFT representation of the HTML encoding sanitizer function.}
-\label{fig:html-sanitizer-sft}
-\end{figure}
-
-We evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT with SFAs constructed from randomly generated Unicode user inputs.
-This experimental setup reflects realistic usage scenarios where diverse character sets and input patterns are encountered.
-
-Figure \ref{fig:benchmark-results} demonstrates the scalability characteristics of our SFT product implementation.
-The computation requires approximately 1 second for 100,000-character inputs and 10 seconds for 1,000,000-character inputs.
-These performance metrics confirm that our certified implementation maintains computational efficiency suitable for practical string processing applications at scale.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.45]{benchmark_loglog.pdf}
-\caption{Experimental results for the sanitizer product}
-\label{fig:benchmark-results}
-\end{figure}
-
-Beyond performance evaluation, our SFT formalization enables formal verification of string sanitizer correctness.
-For example, this verification framework models user input as an SFA, the sanitizer as an SFT, and potential attacks as an SFA accepting unsafe strings.
-A sanitizer is verified correct if the product of the sanitizer SFT and user input SFA produces no outputs that intersect with the attack SFA's language.
-This correctness property can be checked by computing the intersection of the product SFT's output language with the attack SFA, where emptiness of this intersection guarantees sanitizer safety.
-
-\subsection{Modeling the Replacement Operation}
-
-The string replacement operation, denoted as 
-$\texttt{replace}(\texttt{str}, \newline\texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
-\begin{itemize}
-  \item $\texttt{str}$: The input string to be transformed.
-  \item $\texttt{pattern}$: A regular expression defining the matching criteria.  
-  \item $\texttt{replacement}$: The string to be substituted for the matched substring.
-\end{itemize}
-
-The semantics of the replacement operation can be formally characterized by two distinct cases:
-\begin{enumerate}
-  \item When there exists at least one substring $s'$ in \texttt{str} such that $s'$ matches $\texttt{pattern}$, then $s'$ is replaced with \texttt{replacement}.
-  \item When no substring of \texttt{str} matches $\texttt{pattern}$, the operation returns \texttt{str} unchanged.
-\end{enumerate}
-
-The first case can be modeled by SFTs.
-We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}\newline(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
-We model this replacement operation by an SFT with the following 3 steps:
-
-\begin{enumerate}
-  \item First, construct an SFA that recognizes the regular expression pattern $\texttt{/[0-9]+/}$.
-Transform this SFA into an SFT by augmenting each transition with an output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
-  \item Second, construct an SFA that accepts the replacement string "\texttt{NUM}".
-Since a constant string can be viewed as a specialized regular expression, we can construct its SFA representation.
-Convert this SFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
-  \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
-These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and the identity function $\texttt{id}=\lambda x.~x$, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
-\end{enumerate}
-
-\noindent\emph{Step 1.}
-Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
-The left side shows the SFA that recognizes the regular expression $\texttt{/[0-9]+/}$, while the right side presents its transformation into an SFT.
-This transformation is achieved by augmenting each transition with the output function $\texttt{f} = \lambda x.~\texttt{None}$, which consistently produces the empty string, effectively "consuming" the matched digits without generating output.
-
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
-  % Left: SFA for [0-9]+ (vertical)
-  \node[state, initial] (q0)   {$q_0$}; 
-  \node[state, accepting, below=of q0] (q1) {$q_1$};
-
-   \path[->] 
-   (q0) edge [bend left=20] node[right] {\texttt{[0-9]}} (q1)
-   (q1) edge [loop below] node {\texttt{[0-9]}} (q1);
-
-  % Right: SFT for [0-9]+ (vertical)
-  \node[state, initial, right=4cm of q0] (q0') {$q_0'$}; 
-  \node[state, accepting, below=of q0'] (q1') {$q_1'$};
-
-   \path[->] 
-   (q0') edge [bend left=20] node[right] {\texttt{[0-9]},$\texttt{f}$} (q1')
-   (q1') edge [loop below] node {\texttt{[0-9]},$\texttt{f}$} (q1');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for \texttt{/[0-9]+/}}
-\label{fig-snfa-pattern}
-\end{figure}
-
-\noindent\emph{Step 2.}
-Figure \ref{fig-snfa-replacement} depicts the automata for the replacement string "\texttt{NUM}".
-The left side shows the SFA that accepts this constant string, while the right side presents its SFT transformation.
-The transformation employs an indexed output function \texttt{g} that emits characters of the replacement string sequentially:
-\[
-\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
-\begin{cases}
-1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
-2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
-3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
-\_ \mapsto \texttt{None}
-\end{cases}
-\]
-This function maps transition indices to their corresponding character outputs, using ASCII codes to represent the string "\texttt{NUM}" character by character.
-
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-  % SFA for "NUM" (top)
-  \node[state, initial above] (q0)   {$q_0$}; 
-  \node[state] (q1) [right=1cm of q0] {$q_1$}; 
-  \node[state] (q2) [below=1cm of q1] {$q_2$}; 
-  \node[state, accepting] (q3) [below=1cm of q0] {$q_3$};
-
-  \path[->] 
-  (q0) edge node {\texttt{N}} (q1)
-  (q1) edge node {\texttt{U}} (q2)
-  (q2) edge node {\texttt{M}} (q3);
-
-  % SFT for "NUM" (right)
-  \node[state, initial above] (q0') [right=1.5cm of q1] {$q_0'$}; 
-  \node[state] (q1') [right=of q0'] {$q_1'$}; 
-  \node[state] (q2') [below=1cm of q1'] {$q_2'$}; 
-  \node[state, accepting] (q3') [below=1cm of q0'] {$q_3'$};
-
-  \path[->] 
-  (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
-  (q1') edge node[left] {\texttt{None}, \texttt{(g 2)}} (q2')
-  (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for "\texttt{NUM}"}
-\label{fig-snfa-replacement}
-\end{figure}
-
-\noindent\emph{Step 3.}
-Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT (Fig.
-\ref{fig-snfa-pattern}) with the replacement-generating SFT (Fig.
-\ref{fig-snfa-replacement}).
-The composition process involves two key modifications:
-\begin{enumerate}
-  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
-  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{id}$", where $\Sigma$ represents the full alphabet.
-For the string solver, $\Sigma$ is the set of all unicode characters.
-These transitions enable the SFT to process arbitrary input prefixes and suffixes while preserving the matched substring for replacement
-\end{enumerate}
-
-\begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-    % Second automaton from the first figure
-    \node[state, initial] (q0') {$q_0'$}; 
-    \node[state] (q1') [right=of q0'] {$q_1'$};
-
-     \path[->] 
-     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q0') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q0');
-
-    % Second automaton from the second figure
-    \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
-    \node[state] (q1'') [right=of q0''] {$q_1''$}; 
-    \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
-    \node[state] (q2'') [right=of q3''] {$q_2''$};
-
-    \path[->] 
-    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
-    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
-    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q3'') edge [loop left] node {$\Sigma$, $\texttt{id}$} (q3'')
-    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
-  \end{tikzpicture}
-  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
-  \label{fig-rearranged-automata}
-  \end{figure}
-
-  Having constructed the SFT, we can now compute the forward image of the replacement operation.
-Consider the string constraint $s' = \texttt{replace}(s,~\texttt{/[0-9]+/},~\text{"}\texttt{NUM}\text{"})$, where we aim to characterize the possible values of $s'$.
-Let $\mathcal{T}$ denote the constructed SFT modeling the replacement operation, and let $\mathcal{A}$ be the SFA representing the domain of possible values for the input string $s$.
-The forward image of this replacement operation is given by the product $\mathcal{T} \times \mathcal{A}$, which precisely captures the set of all possible output strings that can be produced by applying the replacement operation to any input string accepted by $\mathcal{A}$.
-  Our string solver uses this forward image to do forward propagation for further string solving.
-
-  \emph{Comparison with SMT-LIB Semantics.} It is important to note that our formalization differs from the standard SMT-LIB semantics for the replacement operation.
-In SMT-LIB, the \texttt{str.replace} operation is defined to replace only the \emph{first} occurrence of a substring matching the given pattern.
-In contrast, our semantics allows for a more general interpretation where any matching substring may be replaced.
-But it still replaces only one occurrence of them.
-
-  The modeling of the aformentioned replacement operation will replace one occurrence of the substring that matches the regular expression pattern.
-We may also want to model left-most replacement and replace all occurrences of the substring that matches the regular expression pattern.
-
-  One way to model left-most replacement is illustrated in the following.
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-\end{tikzpicture}
-\caption{The modeling of left-most replacement operation.}
-\label{fig:replace-left-most}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-  % Epsilon transition from accepting state of second SFT to initial state of first SFT
-  \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
-
-\end{tikzpicture}
-\caption{The modeling of replace all operation.}
-\label{fig:replace-all}
-\end{figure}
-
-\subsection{Experiments}
-
-We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
-%\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
-While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
-
-CertiStrR implements two kinds of SMT-LIB's replacement operations:
-\texttt{str.replace s p r}: a string-based replacement where pattern \texttt{p} is a constant string.
-\texttt{str.replace\_re s p r}: a regular expression-based replacement where pattern \texttt{p} is a regular expression
-
-Listing \ref{lst-smtlib-code} demonstrates the regular expression variant through an example that replaces numeric sequences with the string "\texttt{NUM}".
-The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
-%
-
-As discussed in the previous subsection, our replacement semantics differs from the SMT-LIB standard semantics.
-Run the SMT solver CVC5 \cite{cvc5} with the code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
-
-\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}, float=htbp]
-  (set-logic QF_S)
-  (declare-fun a () String)
-  (declare-fun b () String)
-  (assert (= a "2024,2025"))
-  (assert (= b "2024,NUM"))
-  (assert (= b (str.replace_re a 
-      (re.+ (re.range "0" "9")) "NUM")))
-  (check-sat)
-  \end{lstlisting}
-
-  We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are divided into two categories based on the replacement operations: (1) string-based replacement (\texttt{str.replace}) and (2) regular expression-based replacement (\texttt{str.replace\_re}).
-Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB language, we preprocess certain operations.
-For instance, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
-
-% replace_re
-% Average real time: 0.37 seconds
-% Total number of real entries: 98
-% Average user time: 0.35 seconds
-% Total number of user entries: 98
-% Average sys time: 0.01 seconds
-% Total number of sys entries: 98
-% Total number of SAT entries: 61
-% Total number of UNSAT entries: 4
-% Total number of inconclusive entries: 30
-
-% replace_str   
-% Average real time: 0.27 seconds
-% Total number of real entries: 323
-% Average user time: 0.25 seconds
-% Total number of user entries: 323
-% Average sys time: 0.01 seconds
-% Total number of sys entries: 323
-% Total number of SAT entries: 315
-% Total number of UNSAT entries: 173
-% Total number of inconclusive entries: 8
-
-\begin{table}[h]
-  \centering
-  \small
-  \begin{tabular}{lcccc}
-      \toprule
-      & \textbf{SAT/UNSAT/Inc} & \textbf{Time (s)} & \textbf{Tests} \\
-      \midrule
-      \texttt{replace\_str} & 142/173/8 & 0.27 & 323\\
-      \texttt{replace\_re} & 84/4/10 & 0.36 & 98\\
-      \bottomrule
-  \end{tabular}
-  \caption{Experimental results}
-  \label{tab:string_operations}
-\end{table}
-
-The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a time limit of one minute per test. 
-The results show average execution times of 0.27 seconds for \texttt{str.replace\_str} and 0.36 seconds for \texttt{str.replace\_re} operations.
-Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
-%
-An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests are finished in one minute) but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
-Specifically, when the string constraints do not satisfy the tree property defined in \cite{cpp/KanLRS22}, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
-
-Our performance analysis revealed that while the SFT-based replacement operation modeling is efficient, the primary computational bottleneck stems from automata accumulation during  forward-propagation.
-Consider the following string constraints:
-\begin{align}
-x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber \\
-x &= x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
-\end{align}
-where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}\newline($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
-Each accumulation step requires computing the product of the current automaton with the previous result.
-Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state explosion.
-
-\subsection{Effort of Certified Development}
-
-We discuss the effort of certified SFT development in this subsection.
-Table \ref{tab:abstract_impl} provides an overview.
-The abstract-level development means all formalizations in Section \ref{sec:formalization} and  \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs .
-The implementation-level development means all formalizations in Section \ref{sec_alg_refinement} including the refinement of the product operation.
-The last row corresponds to the effort of the interval formalization.
-The most difficult part is the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
-Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
-
-\begin{table}[h]
-  \centering
-  \small
-  \begin{tabular}{lccc}
-      \toprule
-      & \textbf{Defs} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
-      \midrule
-      Abstract & 17 & 21 & 3274 \\
-      Implementation & 51 & 43 & 2700 \\
-      Interval & 15 & 29 & 1500 \\
-      \bottomrule
-  \end{tabular}
-  \caption{Overview of the effort of certified development}
-  \label{tab:abstract_impl}
-\end{table}
-
-\section{Related Work}
-\label{sec:related-work}
-
-\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
-This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
-Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
-
-\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
-These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
-Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.
-
-\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
-To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
-Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.
-
-\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
-While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
-As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
-We believe that our work will benefit the community by providing a formal foundation for string solvers development.
-
-\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains.
-For example, the work by Shi et al.
-\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL.
-Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.
-
-To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}.
-This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
-The experimental results confirm both the efficiency and effectiveness of our approach.
-While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.
-
-Future work includes extending our symbolic formalization in two key directions:
-%
-(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
-(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
-These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.
-
-\bibliography{literature}
-\end{document}

### 2025-08-08 Paper/main.tex
main.tex split

--- 2025-07-29 Paper/main.tex
@@
 
-\begin{abstract}
-Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
-FTs have wide applications, including program analysis, string constraint solving, and security-critical sanitizer analysis.
-However, there is currently no scalable formalization of FTs.
-This is primarily because transition labels in FTs are not formalized symbolically, resulting in transition explosion and significant performance bottlenecks in practical applications.
-
-In this paper, we present a formalization of FTs in the symbolic setting using Isabelle/HOL, where transition labels can be instantiated with various Boolean algebras, such as intervals and arithmetic predicates.
-This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
-%
-To evaluate the effectiveness of our formalization, we apply the formalized SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for preventing XSS attacks, and (2) a string solver that models replacement operations with various semantics, including left-most and replace-all matching.
-Experimental results demonstrate that our approach achieves computational efficiency in constraint-solving scenarios.
-\end{abstract}
-
+\input{abstract.tex}
 \maketitle
 
-\section{Introduction}
-\label{sec:introduction}
+\input{introduction.tex}
 
-Finite Automata (FA) and Finite Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
-For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
-Additionally, FAs and FTs find significant industrial applications, such as in the verification of AWS access control policies \cite{DBLP:conf/fmcad/BackesBCDGLRTV18}.
+\input{SFT.tex}
 
-Even though there are various formalizations of FAs and FTs in interactive proof assistants such as Isabelle \cite{isabelle-homepage} and Coq \cite{coq-homepage}, these are predominantly based on classical definitions.
-However, these traditional approaches present certain limitations when applied to practical scenarios.
-One significant drawback is that transition labels are typically non-symbolic and finite.
-A conventional transition is represented as $q\xrightarrow{a}q'$, where $a$ is a character from a finite alphabet.
-This simplistic representation can lead to a phenomenon known as transition explosion.
-For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
-Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the FA and FT operations highly inefficient.
-Furthermore, in practical applications, it is often necessary to consider infinite alphabets, such as the set of all integers.
+\input{formalization.tex}
 
-Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
-%
-These symbolic models utilize transition labels defined by first-order predicates over boolean algebras, allowing for more expressive representations.
-For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \mod 2 = 0$), denoting all even numbers.
-This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
+\input{application.tex}
 
-However, the formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
-Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse predicate representations of boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
+\input{relatedwork.tex}
 
-Prior work of CertiStr \cite{cpp/KanLRS22} has successfully formalized SFAs within Isabelle/HOL, demonstrating both the efficiency and effectiveness of SFAs in practice.
-However, the formalization of SFTs remains an open challenge, primarily due to their inherent complexity in two aspects: the formalization of transition labels and the specification of transition output functions.
-SFTs constitute a significantly more expressive and powerful theoretical framework compared to  SFAs, as evidenced by their capability to model complex string transformations such as replacement operations.
-%
-Furthermore, numerous studies have demonstrated the broad applicability of SFTs across diverse domains.
-The work of \cite{VeanesHLMB12Transducer} showcases SFTs in security-critical applications, particularly for cross-site scripting (XSS) prevention.
-\cite{uss/HooimeijerLMSV11} extends this security focus by applying SFTs to web application sanitizer analysis.
-In system verification, \cite{osdi/YaseenABCL20} employs SFTs for runtime behavior monitoring, while \cite{pldi/HuD17} demonstrates their utility in automated program transformation through systematic inversion techniques.
-
-In this work, we present a comprehensive formalization of SFTs.
-To address the extensibility challenges inherent in supporting diverse transition label theories, we adopt a refinement-based approach.
-At the abstraction level, transition labels are formalized through the fundamental mathematical concept of \emph{sets}.
-This abstraction facilitates subsequent refinement to various representations of Boolean algebras, such as intervals and arithmetic predicates, while maintaining theoretical consistency.
-
-The key operation of our formalization is the product operation between an SFT and an input regular language.
-Specifically, given an SFA $\mathcal{A}$ representing a regular language and an SFT $\mathcal{T}$, we define the product operation $\mathcal{T} \times\mathcal{A}$ that characterizes the output language generated by $\mathcal{T}$ when processing inputs from the language recognized by $\mathcal{A}$.
-
-In the refinement level, we implemented transition labels using an interval-based representation to examplify the refinement process.
-The formalization of interval algebra provides efficient set-theoretic operations—including membership checking, intersection, and difference computations—facilitating the refinement of transition labels from abstract sets to concrete intervals.
-Furthermore, leveraging the data refinement framework \cite{DBLP:conf/itp/Lammich13} in Isabelle/HOL, we store states and transitions using sophisticated data structures such as hashmaps and red-black trees, ensuring efficient automata manipulation.
-
-To evaluate the effectiveness and efficiency of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
-We evaluated our implementation using a set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
-The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
-
-In summary, our formalization makes the following contributions:
-\begin{enumerate}
-\item We present the first formalization of symbolic finite transducers in Isabelle/HOL proof assistant.
-\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
-\item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
-\item We demonstrate the practical utility of our formalization through its application to string constraint solving, specifically in modeling replacement operations with verified correctness guarantees.
-\end{enumerate}
-
-The remainder of this paper is organized as follows:
-Section \ref{sec:formalization} introduces the formalization of symbolic finite transducers.
-Section \ref{sec:product-operation} presents the product operation at the abstract level.
-Section \ref{sec_alg_refinement} details the algorithmic refinement of the product operation.
-Section \ref{sec-app-str-solver} demonstrates the application to string constraint solving.
-Section \ref{sec:related-work} discusses related work.
-Section \ref{sec:conclusion} concludes with future directions.
-
-\section{Formalization of SFTs}
-\label{sec:formalization}
-
-We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL. 
-\subsection{The Mathematical Definition of SFTs}
-
-Let $\mathcal{U}$ be a multi-sorted carrier set or background universe, which is equipped with functions and relations over the elements.
-We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
-We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type.
-
-A lambda term is defined as $\lambda x.~t$ of type $\tau_1 \rightarrow \tau_2$.
-When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate.
-Let $\phi$ be a predicate.
-We write $a\in \llbracket\phi \rrbracket$ if $\phi~a=\top$.
-For non-predicate lambda terms, we view them as functions that generate output elements of type $\tau_2$ given input terms of type $\tau_1$.
-With these notations and the above definitions, we can define SFTs as follows.
-
-\begin{definition}[Symbolic Finite Transducer]
-\label{def-sft}
-   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
-   \begin{itemize}
-   \item $\mathcal{Q}$ is a finite set of states,
-   \item $\mathcal{I}\subseteq \mathcal{Q}$ is the set of initial states,
-   \item $\mathcal{F} \subseteq\mathcal{Q}$ is the set of accepting states,
-   \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
-   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
-   \end{itemize}
-
-For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called an input,  then the application $(f~a)$ is the output.
-
-\end{definition}
-
-SFTs accept an input word and generate an output word. This can be defined by \emph{runs} of SFTs.
-An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i,
-\phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
-$\sigma$ is an accepting run when $q_n$ is an accepting state.
-
-For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ if and only if $a_i \in \llbracket\phi_i \rrbracket$ for $0 \leq i \leq n - 1$.
-When $w$ is accepted, run $\sigma$ generates an output sequence $w'= f_0~a_0, \ldots, f_{n-1}~a_{n-1}$.
-We define:
-\begin{itemize}
-\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
-\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, $ $(f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
-\end{itemize}
-
-If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
-For a trace $\pi$, we denote its input as $\mathit{in}(\pi)$ and its output as $\mathit{out}(\pi)$.
-%
-Given an SFT $\mathcal{T}$ and a word $w$, we define the \emph{product} operation of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with input $w$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
-\end{split}
-\]
-
-To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an SFA $\mathcal{A}$.
-More precisely,
-\[
-\begin{split}
-\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \\\text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
-\end{split}
-\]
-
-\begin{figure}[hbt!]
-  \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm and 2cm, on grid, auto]
-    % Make q0 and q1 horizontally wide apart
-    \node[state, initial] (q_0)   {$q_0$}; 
-    \node[state] (q_1) [right=4cm of q_0] {$q_1$}; 
-    \node[state, accepting] (q_2) [below=2.2cm of q_1] {$q_2$};
-
-    \path[->] 
-    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
-    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
-  \end{tikzpicture}
-    \caption{An example of SFT}
-    \label{fig-example-ft}
-\end{figure}
-
-    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
-The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
-The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation.
-For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.
-
-%The transition labels can be more complex.
-For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$.
-This symbolic label corresponds to the set of even natural numbers.
-
-\subsection{The Isabelle/HOL Formalization of SFTs}
-
-Firstly, we present the basic idea of our formalization of SFTs in Isabelle/HOL, which consists of three layers.
-
-\begin{figure}[hbt!]
-\centering
-\begin{tikzpicture}[
-  node distance=1cm,
-  every node/.style={font=\small},
-  layer/.style={rectangle, draw, rounded corners, fill=blue!10, minimum width=5.5cm, minimum height=1.2cm, align=center},
-  arrow/.style={->, thick}
-]
-
-% Abstract Layer
-\node[layer] (abstract) {%
-  \textbf{Abstract Layer}\\
-  FT with transition labels as \textbf{sets}
-};
-
-% Locale Layer
-\node[layer, below=of abstract] (locale) {%
-  \textbf{Implementation Layer}\\
-  FT with transition labels as a \\boolean algebra locale
-};
-
-% Implementation Layer
-\node[layer, below=of locale] (impl) {%
-  \textbf{Boolean Algebra Layer}\\
-  Concrete Boolean algebra implementations\\
-  (e.g., intervals, bitvectors, predicates)
-};
-
-% Draw arrows with refinement labels
-\draw[arrow] (abstract) -- node[right] {refinement} (locale);
-\draw[arrow] (locale) -- node[right] {refinement} (impl);
-
-\end{tikzpicture}
-\caption{Three-layered formalization of SFTs in Isabelle/HOL. 
-\textbf{Abstract Layer:} Transition labels are modeled as sets, representing the most general case.
-\textbf{Locale Layer:} Transition labels are abstracted as a type variable \texttt{'a} within a locale, supporting any Boolean algebra.
-\textbf{Implementation Layer:} Concrete Boolean algebra instances, such as intervals, bitvectors, or predicates, are provided.}
-\label{fig:layered-formalization}
-\end{figure}
-
-As we have discussed the diversity of transition labels in Section \ref{sec:introduction}, we now present an extensible formalization of SFTs that accommodates this variety.
-Our approach leverages the refinement framework (Refine\_Monadic \cite{Refine_Monadic-AFP}) in Isabelle/HOL to achieve the flexibility and extensibility of transition labels modeling.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a,$| |$'i,$| |$'b$|) |$\texttt{NFT}$| =
-	|$\mathcal{Q}_t$| :: "|$'q$| set"
-	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'i$|) LTTS"
-	|$\mathcal{I}_t$| :: "|$'q$| set"
-	|$\mathcal{F}_t$| :: "|$'q$| set"
-        |$\mathcal{M}_t$| :: "|$'i\Rightarrow ('a,~'b)$| Tlabel"
-type_synonym |$('q,~'a,~'i)$| LTTS = "|$('q \times ('a \text{ set option }\times 'i) \times 'q)$| set"
-type_synonym |$('a,~'b)$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
-	\end{lstlisting}
-\caption{The formalization of $SFTs$ in Isabelle/HOL}
-\label{fig-def-FT}
-\end{figure}
-
-Figure \ref{fig-def-FT} presents our formalization of SFTs in Isabelle/HOL.
-While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
-%
-The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$.
-This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
-
-Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
-Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
-A predicate's semantics as introduced before represents a set of elements that make the predicate true.
-But predicates have various different forms.
-For instance, the interval $[1-9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$.
-The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
-All these different forms of labels are abstracted as sets in our formalization.
-%
-The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements.
-This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
-
-The second element of type $'i$ serves as an index into the output function space. 
-%
-We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
-%
-The mapping $\mathcal{M}_t$ associates each index with a specific output function.
-These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
-This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria.
-Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
-
-%Now let us look at an example.
-In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$.
-This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}).
-%We can easily model this operation using our formalization of NFTs with the following steps:
-%\begin{itemize}
-%\item Step 1, construct a corresponding NFA from \texttt{pattern}).
-%\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
-%\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
-%\item Step 4, unlabel previous accepting states as non-accepting states. 
-%\end{itemize}
-
-%We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.
-
-\section{The Product Operation of SFTs}
-\label{sec:product-operation}
-
-In this section, we formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
-An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
-Since CertiStr \cite{cpp/KanLRS22} does not include a formalization of SFAs with $\varepsilon$-transitions, we extend the framework in CertiStr with a formalization of symbolic SFAs with $\varepsilon$-transitions (denoted as $\varepsilon$SFA) and provide a verified conversion to standard SFAs.
-In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-record (|$'q,$||$~'a$|) |$\texttt{NFA}$| =
-	|$\mathcal{Q}$| :: "|$'q$| set"
-	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\mathcal{I}$| :: "|$'q$| set"
-	|$\mathcal{F}$| :: "|$'q$| set"
-
-record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
-	|$\mathcal{Q}_e$| :: "|$'q$| set"
-	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
-	|$\Delta_e'$| :: "|$('q * 'q)$| set"
-	|$\mathcal{I}_e$| :: "|$'q$| set"
-	|$\mathcal{F}_e$| :: "|$'q$| set"
-
-type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
-	\end{lstlisting}
-\caption{The formalization of $\varepsilon$SFAs and SFAs}
-\label{fig-def-FAs}
-\end{figure}
-
-Figure \ref{fig-def-FAs} presents the formalization of both SFAs and $\varepsilon$SFAs using Isabelle/HOL record types \texttt{NFA} and \texttt{eNFA}, respectively.
-The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
-
-Having established these foundational definitions, we can now formalize the product operation.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| 
-   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$| 
-   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
-"productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
-  |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
-       |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
-        |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
-      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
-       |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
-       |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
-          |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
-  |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
-       |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
-       |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
-      |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
-        |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}$|
-        |$\land~\sigma_1\cap \sigma_2 \neq \emptyset~\land\exists x\in(\sigma_1\cap\sigma_2).$|
-        |$~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
-  |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
-  |$\mathcal{F}_e = \mathcal{F}_t~\mathcal{T} \times \mathcal{F}~\mathcal{A}$| |$\rrparenthesis$|"
-	\end{lstlisting}
-\caption{The formalization of product operation}
-\label{fig-def-FTProd}
-\end{figure}
-
-Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA. The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively. The 
-result of the product operation is an $\varepsilon$SFA.
-But we need to explain the role of parameter $\text{\texttt{F}}$.
-The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$.
-$\text{\texttt{F}}$ extends $f$ to apply $f$ to a set of elements.
-More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as follows:
-
-\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].
-
-The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions.
-In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels.
-Let us consider $\Delta_e$ first.
-
-\begin{enumerate}
-\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e.
-the input character is $\varepsilon$, and $f~\text{None} \neq \text{None}$.
-Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output "$((\mathcal{M}_t~\mathcal{T})~f)~\text{None}$".
-Remember that $f$ is just an index, $(\mathcal{M}_t~\mathcal{T})~f$ is the output function.
-
-\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$, where $\sigma_2$ represents the input label of the corresponding SFA transition.
-The resulting output is $\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)$.
-\end{enumerate}
-
-The transitions in $\Delta_e'$ follow a similar pattern with analogous cases for $\varepsilon$ and non-$\varepsilon$ transitions.
-
-To establish the correctness specification of the product operation, we begin by formalizing the concept of SFT \emph{traces} as introduced in Definition \ref{def-sft}.
-In our formalization, traces are represented by the type $('a\;\text{option} \times 'b\;\text{option})\;\texttt{list}$, as shown in Figure \ref{fig-def-output}.
-%
-Given a trace $\pi$, we define two key projection functions:
-(1) \texttt{inputE}, which corresponds to ${in}(\pi)$ and extracts the input sequence
-(2) \texttt{outputE}, which corresponds to ${out}(\pi)$ and extracts the output sequence.
-
-The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-fun inputE :: 
-   "|$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$|"
-where
-  "inputE [] = []" |$\mid$|
-  "inputE ((Some a, |$\_$|) |$\#$| l) = 
-                      a |$\#$| (inputE l)" |$\mid$|
-  "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"
-
-fun outputE :: 
-    "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" 
-where
-  "outputE [] = []" |$\mid$|
-  "outputE ((_,Some a) # l) = 
-                    a # (outputE l)" |$\mid$|
-  "outputE ((_,None) # l) = (outputE l)"
-
-definition outputL :: 
-   "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
-                               ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
-  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. 
-    |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T}~\land$| 
-    |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land$||$ \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}"
-	\end{lstlisting}
-\caption{The formalization of traces in SFTs}
-\label{fig-def-output}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-lemma productT_correct:
-  fixes |$\mathcal{T}$| |$\mathcal{A}$| F
-  assumes 
-   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
-       |$\text{ F}~f~s = \text{None}"$|
-   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
-        |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
-        |$f~(\text{Some }e) = \text{ Some S}\})$|"
-      and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
-  |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL 
-          |$\mathcal{T}~\mathcal{A}$|"
-
-	\end{lstlisting}
-\caption{The correctness lemma of the product operation}
-\label{fig-def-product-correct}
-\end{figure}
-
-Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation.
-The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
-The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.
-The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.
-
-\section{Algorithm Level Refinement}
-\label{sec_alg_refinement}
-Having established the abstract definition of the SFT product operation in Section \ref{sec:product-operation}, we now present its algorithmic refinement.
-This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
-
-\subsection{Intervals}
-
-An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.
-To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$.
-This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting.
-For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.
-
-Throughout the following discussion, we use the term "interval" to refer to interval lists. Our formalization provides a collection of interval operations through the following interface:
-
-\begin{enumerate}
-  \item $\texttt{semIs}~i$: Defines the semantic interpretation of an interval as a set. For an interval $(i,j)$, $\texttt{semIs}~(i,j) = \{e \mid i \leq e \leq j\}$.
-  \item $\texttt{emptyIs}~i$: Tests whether an interval represents an empty set, i.e., whether $\texttt{semIs}~i = \emptyset$.
-  \item $\texttt{nemptyIs}~i$: Tests whether an interval represents a non-empty set, i.e., whether $\texttt{semIs}~i \neq \emptyset$.
-  \item $\texttt{intersectIs}~i_1~i_2$: Computes the intersection of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \cap \texttt{semIs}~i_2$.
-  \item $\texttt{diffIs}~i_1~i_2$: Computes the set difference of two intervals, yielding an interval $i$ such that $\texttt{semIs}~i = \texttt{semIs}~i_1 \setminus \texttt{semIs}~i_2$.
-\end{enumerate}
-
-To facilitate formal reasoning and optimize performance, we introduce a canonical form for interval.
-An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
-\begin{enumerate}
-  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
-  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
-\end{enumerate}
-
-We prove that all interval operations preserve canonical form when applied to canonically-formed inputs.
-This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
-
-\subsection{Algorithmic Implementation of the Product Operation}
-
-In this subsection, we present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.}.
-%
-Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}.
-These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
-
-The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets.
-This function employs the \texttt{FOREACH} construct, a higher-order iteration operator analogous to OCaml's \texttt{Set.fold}.
-Specifically, given a set $S$, a function $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$, and an initial accumulator $I$ of type $'b$, the expression $\texttt{FOREACH}~S~f~I$ systematically applies $f$ to each element in $S$, accumulating results in a principled manner.
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition productT_impl where
-"productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
-    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) 
-    (nfa_states |$\mathcal{A}$|);
-    (D1, D2) |$\leftarrow$| trans_comp_imp 
-     (nft_tranfun |$\mathcal{T}$|) F fe (nft_trans |$\mathcal{T}$|) 
-      (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
-    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|)
-      (nfa_initial |$\mathcal{A}$|);
-    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) 
-      (nfa_accepting |$\mathcal{A}$|);
-    RETURN (Q, D1, D2, I, F)
-  }"
-\end{lstlisting}
-\caption{The computation of SFT product}
-\label{fig-compute-nft-product}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition prods_imp where
-"prods_imp Q1 Q2 =
-   FOREACH {q. q |$\in$| Q1} (|$\lambda$| q Q. do {
-    S |$\leftarrow$| FOREACH {q. q |$\in$| Q2}
-      (|$\lambda$| q|$'$| Q|$'$|. RETURN ({(q,q|$'$|)} |$\cup$| Q|$'$|)) |$\emptyset$|;
-    RETURN (Q |$\cup$| S)
-   }) |$\emptyset$|"
-\end{lstlisting}
-\caption{The computation of Cartesian product of two state sets}
-\label{fig-def-prods_imp}
-\end{figure}
-
-\begin{figure}[hbt!]
-	\begin{lstlisting}
-definition trans_comp_imp where
-"trans_comp_imp M F fe T1 T2 Q =
-  FOREACH {t. t |$\in$| T1}
-   (|$\lambda$|(q,(|$\alpha$|,f),q|$'$|) (D1,D2). 
-    (if (|$\alpha$|=None) then 
-     (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe 
-                              T2 D1 D2)
-     else
-     (subtrans_comp M q (the |$\alpha$|) f 
-            q|$'$| F fe T2 D1 D2))) (|$\emptyset$|, |$\emptyset$|)"
-\end{lstlisting}
-\caption{The computation of \texttt{trans\_comp\_imp}}
-\label{fig-def-prods-imp}
-\end{figure}
-
-A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ (corresponding to $\Delta_e$) and $\texttt{D2}$ (corresponding to $\Delta_e'$) in Figure \ref{fig-compute-nft-product}.
-As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA.
-This function decomposes the synchronization process into two distinct cases, each handled by a specialized function:
-
-\begin{enumerate}
-  \item \texttt{subtrans\_comp\_$\varepsilon$}: Processes $\varepsilon$-transitions in the SFT, where transitions consume no input but may produce output
-  \item \texttt{subtrans\_comp}: Processes standard transitions in the SFT, where both input consumption and output generation may occur
-\end{enumerate}
-
-\begin{figure}[hbt!]
-\begin{lstlisting}[language=SMTLIB]
-definition subtrans_comp where
-"subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
- FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
-  (if (nemptyIs (intersectIs |$\alpha$| |$\alpha'$|)) then
-    do {
-      D1 |$\leftarrow$| 
-       (if (F (M f) (intersectIs |$\alpha$| |$\alpha'$|)) 
-          |$\neq$| None) then
-         let |$\alpha_i$| = the (F (M f) 
-              (intersectIs |$\alpha$| |$\alpha'$|)) in
-         RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
-                |$\cup$| D1
-        else RETURN D1);
-      D2 |$\leftarrow$| 
-       (if fe (M f) (intersectIs |$\alpha$| |$\alpha'$|) 
-        then 
-         RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 
-        else RETURN D2);
-      RETURN (D1, D2) }
-   else (RETURN (D1, D2)))) (D1, D2)"
-    \end{lstlisting}
-    \caption{The computation of \texttt{subtrans\_comp}}
-    \label{fig-def-subtrans_comp}
-    \end{figure}
-
-    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
-For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.
-For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):
-
-    \begin{enumerate}
-      \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
-      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}.
-The checking is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
-    \end{enumerate}
-
-  \begin{figure}[hbt!]
-    \begin{lstlisting}
-  lemma productT_imp_correct:
-  assumes finite_TT: "|$\text{finite } (\Delta_t~\mathcal{T})$|"
-      and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
-      and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
-      and finite_TQ: "|$\text{finite } (\mathcal{Q}_t~\mathcal{T})$|"
-      and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
-      and finite_TI: "|$\text{finite } (\mathcal{I}_t~\mathcal{T})$|"
-      and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
-      and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
-  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe}\leq$|
-            |$\text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
-  \end{lstlisting}
-  \caption{The refinement relation between \texttt{productT\_imp} and \texttt{productT}}
-  \label{fig-def-productT_imp_correct}
-  \end{figure}
-
-  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
-This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.
-
-  The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$.
-More precisely,  the concrete implementation must produce an $\varepsilon$SFA that is structurally equivalent or isomorphic to the one produced by the abstract algorithm \texttt{productT}.
-
-  To establish this equivalence, we must prove that the $\varepsilon$SFAs produced by \texttt{productT\_imp} and \texttt{productT} are isomorphic in all essential components: the set of states, transition relations, $\varepsilon$-transition relations, initial and accepting state sets.
-
-The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
-The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 
-In our formalization, we refine the sets of storing states and transitions to red-black trees.
-
-\section{Applications and Evaluation}
-
-We evaluate our formalization and implementation of the SFT product operation in two key application domains: string sanitization and string replacement operations in string solving.
-These applications demonstrate the practical utility of our certified SFT product operation in real-world scenarios.
-\subsection{Sanitizer}
-
-One Important application of our formalization is in the context of string sanitization, specifically for web applications.
-Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before being rendered in a web browser.
-Listing \ref{lst-html-encode} presents a C\# code snippet that implements an HTML encoding function, which is a sanitization operation.
-The function iterates through each character in the input string and encodes characters that are not safe for HTML rendering.
-
-\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
-static string EncodeHtml(string t)
-{
-  if (t == null) { return null; }
-  if (t.Length == 0) {
-    return string.Empty;
-  }
-  StringBuilder builder = new
-    StringBuilder("", t.Length * 2);
-  foreach (char c in t) {
-    if ((((c > ''') && (c < '{')) |$\mid\mid$|
-         ((c > '@') && (c < '['))) |$\mid\mid$|
-         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
-         (((c == '.') || (c == ',')) |$\mid\mid$|
-         ((c == '-') || (c == '_'))))) {
-          builder.Append(c);
-      }
-      else {
-          builder.Append("&#" +
-          ((int) c).ToString() + ";");
-      }
-  }
-  return builder.ToString();
-}
-\end{lstlisting}
-
-Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
-The transducer accepts any input string and produces an output where safe characters are preserved using the identity function $\texttt{id}$, while unsafe characters are encoded as HTML entities.
-The set of safe characters is defined as $\texttt{safe\_chars} = [\texttt{' '}, \texttt{','}, \texttt{'.'}, \texttt{'-'}, \texttt{'\_'}] \cup [\texttt{'0'}-\texttt{'9'}] \cup [\texttt{'A'}-\texttt{'Z'}] \cup [\texttt{'a'}-\texttt{'z'}]$, which can be efficiently represented using the interval algebra we formalized.
-Unsafe characters are transformed using the encoding function $\texttt{encode}(c) = \texttt{``\&\#''} + \texttt{toString}(\texttt{ord}(c)) + \texttt{``\;''}$.
-When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
-  % States
-  \node[state, initial, accepting] (q0) {$q_0$};
-  \node[state] (q1) [right=3.5cm of q0] {$q_1$};
-  \node[state] (q2) [below=of q0] {$q_2$};
-
-  \path[->] 
-    % Transition for safe characters
-    (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
-    % Transition for unsafe characters  
-    (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
-    % Epsilon transitions back to q0
-    (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
-    (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
-\end{tikzpicture}
-\caption{SFT representation of the HTML encoding sanitizer function.}
-\label{fig:html-sanitizer-sft}
-\end{figure}
-
-We evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT with SFAs constructed from randomly generated Unicode user inputs.
-This experimental setup reflects realistic usage scenarios where diverse character sets and input patterns are encountered.
-
-Figure \ref{fig:benchmark-results} demonstrates the scalability characteristics of our SFT product implementation.
-The computation requires approximately 1 second for 100,000-character inputs and 10 seconds for 1,000,000-character inputs.
-These performance metrics confirm that our certified implementation maintains computational efficiency suitable for practical string processing applications at scale.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.45]{benchmark_loglog.pdf}
-\caption{Experimental results for the sanitizer product}
-\label{fig:benchmark-results}
-\end{figure}
-
-Beyond performance evaluation, our SFT formalization enables formal verification of string sanitizer correctness.
-For example, this verification framework models user input as an SFA, the sanitizer as an SFT, and potential attacks as an SFA accepting unsafe strings.
-A sanitizer is verified correct if the product of the sanitizer SFT and user input SFA produces no outputs that intersect with the attack SFA's language.
-This correctness property can be checked by computing the intersection of the product SFT's output language with the attack SFA, where emptiness of this intersection guarantees sanitizer safety.
-
-\subsection{Modeling the Replacement Operation}
-
-The string replacement operation, denoted as 
-$\texttt{replace}(\texttt{str}, \newline\texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
-\begin{itemize}
-  \item $\texttt{str}$: The input string to be transformed.
-  \item $\texttt{pattern}$: A regular expression defining the matching criteria.  
-  \item $\texttt{replacement}$: The string to be substituted for the matched substring.
-\end{itemize}
-
-The semantics of the replacement operation can be formally characterized by two distinct cases:
-\begin{enumerate}
-  \item When there exists at least one substring $s'$ in \texttt{str} such that $s'$ matches $\texttt{pattern}$, then $s'$ is replaced with \texttt{replacement}.
-  \item When no substring of \texttt{str} matches $\texttt{pattern}$, the operation returns \texttt{str} unchanged.
-\end{enumerate}
-
-The first case can be modeled by SFTs.
-We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}\newline(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
-We model this replacement operation by an SFT with the following 3 steps:
-
-\begin{enumerate}
-  \item First, construct an SFA that recognizes the regular expression pattern $\texttt{/[0-9]+/}$.
-Transform this SFA into an SFT by augmenting each transition with an output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
-  \item Second, construct an SFA that accepts the replacement string "\texttt{NUM}".
-Since a constant string can be viewed as a specialized regular expression, we can construct its SFA representation.
-Convert this SFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
-  \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
-These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and the identity function $\texttt{id}=\lambda x.~x$, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
-\end{enumerate}
-
-\noindent\emph{Step 1.}
-Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
-The left side shows the SFA that recognizes the regular expression $\texttt{/[0-9]+/}$, while the right side presents its transformation into an SFT.
-This transformation is achieved by augmenting each transition with the output function $\texttt{f} = \lambda x.~\texttt{None}$, which consistently produces the empty string, effectively "consuming" the matched digits without generating output.
-
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
-  % Left: SFA for [0-9]+ (vertical)
-  \node[state, initial] (q0)   {$q_0$}; 
-  \node[state, accepting, below=of q0] (q1) {$q_1$};
-
-   \path[->] 
-   (q0) edge [bend left=20] node[right] {\texttt{[0-9]}} (q1)
-   (q1) edge [loop below] node {\texttt{[0-9]}} (q1);
-
-  % Right: SFT for [0-9]+ (vertical)
-  \node[state, initial, right=4cm of q0] (q0') {$q_0'$}; 
-  \node[state, accepting, below=of q0'] (q1') {$q_1'$};
-
-   \path[->] 
-   (q0') edge [bend left=20] node[right] {\texttt{[0-9]},$\texttt{f}$} (q1')
-   (q1') edge [loop below] node {\texttt{[0-9]},$\texttt{f}$} (q1');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for \texttt{/[0-9]+/}}
-\label{fig-snfa-pattern}
-\end{figure}
-
-\noindent\emph{Step 2.}
-Figure \ref{fig-snfa-replacement} depicts the automata for the replacement string "\texttt{NUM}".
-The left side shows the SFA that accepts this constant string, while the right side presents its SFT transformation.
-The transformation employs an indexed output function \texttt{g} that emits characters of the replacement string sequentially:
-\[
-\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
-\begin{cases}
-1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
-2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
-3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
-\_ \mapsto \texttt{None}
-\end{cases}
-\]
-This function maps transition indices to their corresponding character outputs, using ASCII codes to represent the string "\texttt{NUM}" character by character.
-
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-  % SFA for "NUM" (top)
-  \node[state, initial above] (q0)   {$q_0$}; 
-  \node[state] (q1) [right=1cm of q0] {$q_1$}; 
-  \node[state] (q2) [below=1cm of q1] {$q_2$}; 
-  \node[state, accepting] (q3) [below=1cm of q0] {$q_3$};
-
-  \path[->] 
-  (q0) edge node {\texttt{N}} (q1)
-  (q1) edge node {\texttt{U}} (q2)
-  (q2) edge node {\texttt{M}} (q3);
-
-  % SFT for "NUM" (right)
-  \node[state, initial above] (q0') [right=1.5cm of q1] {$q_0'$}; 
-  \node[state] (q1') [right=of q0'] {$q_1'$}; 
-  \node[state] (q2') [below=1cm of q1'] {$q_2'$}; 
-  \node[state, accepting] (q3') [below=1cm of q0'] {$q_3'$};
-
-  \path[->] 
-  (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
-  (q1') edge node[left] {\texttt{None}, \texttt{(g 2)}} (q2')
-  (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for "\texttt{NUM}"}
-\label{fig-snfa-replacement}
-\end{figure}
-
-\noindent\emph{Step 3.}
-Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT (Fig.
-\ref{fig-snfa-pattern}) with the replacement-generating SFT (Fig.
-\ref{fig-snfa-replacement}).
-The composition process involves two key modifications:
-\begin{enumerate}
-  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
-  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{id}$", where $\Sigma$ represents the full alphabet.
-For the string solver, $\Sigma$ is the set of all unicode characters.
-These transitions enable the SFT to process arbitrary input prefixes and suffixes while preserving the matched substring for replacement
-\end{enumerate}
-
-\begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-    % Second automaton from the first figure
-    \node[state, initial] (q0') {$q_0'$}; 
-    \node[state] (q1') [right=of q0'] {$q_1'$};
-
-     \path[->] 
-     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q0') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q0');
-
-    % Second automaton from the second figure
-    \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
-    \node[state] (q1'') [right=of q0''] {$q_1''$}; 
-    \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
-    \node[state] (q2'') [right=of q3''] {$q_2''$};
-
-    \path[->] 
-    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
-    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
-    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q3'') edge [loop left] node {$\Sigma$, $\texttt{id}$} (q3'')
-    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
-  \end{tikzpicture}
-  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
-  \label{fig-rearranged-automata}
-  \end{figure}
-
-  Having constructed the SFT, we can now compute the forward image of the replacement operation.
-Consider the string constraint $s' = \texttt{replace}(s,~\texttt{/[0-9]+/},~\text{"}\texttt{NUM}\text{"})$, where we aim to characterize the possible values of $s'$.
-Let $\mathcal{T}$ denote the constructed SFT modeling the replacement operation, and let $\mathcal{A}$ be the SFA representing the domain of possible values for the input string $s$.
-The forward image of this replacement operation is given by the product $\mathcal{T} \times \mathcal{A}$, which precisely captures the set of all possible output strings that can be produced by applying the replacement operation to any input string accepted by $\mathcal{A}$.
-  Our string solver uses this forward image to do forward propagation for further string solving.
-
-  \emph{Comparison with SMT-LIB Semantics.} It is important to note that our formalization differs from the standard SMT-LIB semantics for the replacement operation.
-In SMT-LIB, the \texttt{str.replace} operation is defined to replace only the \emph{first} occurrence of a substring matching the given pattern.
-In contrast, our semantics allows for a more general interpretation where any matching substring may be replaced.
-But it still replaces only one occurrence of them.
-
-  The modeling of the aformentioned replacement operation will replace one occurrence of the substring that matches the regular expression pattern.
-We may also want to model left-most replacement and replace all occurrences of the substring that matches the regular expression pattern.
-
-  One way to model left-most replacement is illustrated in the following.
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-\end{tikzpicture}
-\caption{The modeling of left-most replacement operation.}
-\label{fig:replace-left-most}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
-
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
-
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
-
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
-
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
-
-  % Epsilon transition from accepting state of second SFT to initial state of first SFT
-  \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
-
-\end{tikzpicture}
-\caption{The modeling of replace all operation.}
-\label{fig:replace-all}
-\end{figure}
-
-\subsection{Experiments}
-
-We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations
-%\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
-While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
-
-CertiStrR implements two kinds of SMT-LIB's replacement operations:
-\texttt{str.replace s p r}: a string-based replacement where pattern \texttt{p} is a constant string.
-\texttt{str.replace\_re s p r}: a regular expression-based replacement where pattern \texttt{p} is a regular expression
-
-Listing \ref{lst-smtlib-code} demonstrates the regular expression variant through an example that replaces numeric sequences with the string "\texttt{NUM}".
-The constraints are satisfiable because the input string \texttt{a} = "\texttt{2024,2025}" contains a substring "\texttt{2025}" that matches the regular expression \texttt{re.+ (re.range "0" "9")} (equivalent to \texttt{/[0-9]+/}), and replacing this match with "\texttt{NUM}" yields the expected output string \texttt{b} = "\texttt{2024,NUM}".
-%
-
-As discussed in the previous subsection, our replacement semantics differs from the SMT-LIB standard semantics.
-Run the SMT solver CVC5 \cite{cvc5} with the code in Listing \ref{lst-smtlib-code}, the result is UNSAT because CVC5 only matches the first occurrence of the pattern.
-
-\begin{lstlisting}[language=SMTLIB, caption={Example SMT-LIB Code}, label={lst-smtlib-code}, float=htbp]
-  (set-logic QF_S)
-  (declare-fun a () String)
-  (declare-fun b () String)
-  (assert (= a "2024,2025"))
-  (assert (= b "2024,NUM"))
-  (assert (= b (str.replace_re a 
-      (re.+ (re.range "0" "9")) "NUM")))
-  (check-sat)
-  \end{lstlisting}
-
-  We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are divided into two categories based on the replacement operations: (1) string-based replacement (\texttt{str.replace}) and (2) regular expression-based replacement (\texttt{str.replace\_re}).
-Due to CertiStrR's current front-end limitations in supporting the full SMT-LIB language, we preprocess certain operations.
-For instance, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
-
-% replace_re
-% Average real time: 0.37 seconds
-% Total number of real entries: 98
-% Average user time: 0.35 seconds
-% Total number of user entries: 98
-% Average sys time: 0.01 seconds
-% Total number of sys entries: 98
-% Total number of SAT entries: 61
-% Total number of UNSAT entries: 4
-% Total number of inconclusive entries: 30
-
-% replace_str   
-% Average real time: 0.27 seconds
-% Total number of real entries: 323
-% Average user time: 0.25 seconds
-% Total number of user entries: 323
-% Average sys time: 0.01 seconds
-% Total number of sys entries: 323
-% Total number of SAT entries: 315
-% Total number of UNSAT entries: 173
-% Total number of inconclusive entries: 8
-
-\begin{table}[h]
-  \centering
-  \small
-  \begin{tabular}{lcccc}
-      \toprule
-      & \textbf{SAT/UNSAT/Inc} & \textbf{Time (s)} & \textbf{Tests} \\
-      \midrule
-      \texttt{replace\_str} & 142/173/8 & 0.27 & 323\\
-      \texttt{replace\_re} & 84/4/10 & 0.36 & 98\\
-      \bottomrule
-  \end{tabular}
-  \caption{Experimental results}
-  \label{tab:string_operations}
-\end{table}
-
-The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a time limit of one minute per test. 
-The results show average execution times of 0.27 seconds for \texttt{str.replace\_str} and 0.36 seconds for \texttt{str.replace\_re} operations.
-Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
-%
-An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests are finished in one minute) but rather due to inherent limitations of the forward-propagation algorithm inherited from CertiStr \cite{cpp/KanLRS22}.
-Specifically, when the string constraints do not satisfy the tree property defined in \cite{cpp/KanLRS22}, the forward-propagation algorithm may be unable to reach a definitive conclusion, even when the variable domains remain non-empty after propagation.
-
-Our performance analysis revealed that while the SFT-based replacement operation modeling is efficient, the primary computational bottleneck stems from automata accumulation during  forward-propagation.
-Consider the following string constraints:
-\begin{align}
-x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber \\
-x &= x_4;~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
-\end{align}
-where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}\newline($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
-Each accumulation step requires computing the product of the current automaton with the previous result.
-Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state explosion.
-
-\subsection{Effort of Certified Development}
-
-We discuss the effort of certified SFT development in this subsection.
-Table \ref{tab:abstract_impl} provides an overview.
-The abstract-level development means all formalizations in Section \ref{sec:formalization} and  \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs .
-The implementation-level development means all formalizations in Section \ref{sec_alg_refinement} including the refinement of the product operation.
-The last row corresponds to the effort of the interval formalization.
-The most difficult part is the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
-Interval formalization complexity dramatic increase in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22} due to the extension of intervals to a list.
-
-\begin{table}[h]
-  \centering
-  \small
-  \begin{tabular}{lccc}
-      \toprule
-      & \textbf{Defs} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
-      \midrule
-      Abstract & 17 & 21 & 3274 \\
-      Implementation & 51 & 43 & 2700 \\
-      Interval & 15 & 29 & 1500 \\
-      \bottomrule
-  \end{tabular}
-  \caption{Overview of the effort of certified development}
-  \label{tab:abstract_impl}
-\end{table}
-
-\section{Related Work}
-\label{sec:related-work}
-
-\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets.
-This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
-Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
-
-\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
-These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
-Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.
-
-\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
-To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
-Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.
-
-\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
-While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
-As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
-We believe that our work will benefit the community by providing a formal foundation for string solvers development.
-
-\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains.
-For example, the work by Shi et al.
-\cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL.
-Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.
-
-To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}.
-This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
-The experimental results confirm both the efficiency and effectiveness of our approach.
-While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.
-
-Future work includes extending our symbolic formalization in two key directions:
-%
-(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
-(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
-These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.
+\input{conclusion.tex}
 

### 2025-08-08 Paper/relatedwork.tex
main.tex split

\section{Related Work}
\label{sec:related-work}

\emph{Symbolic Automata and Transducers.} Symbolic Automata and Transducers \cite{cav/DAntoniV17,VeanesHLMB12Transducer, popl/DAntoniV14, entcs/DAntoniKW18, sofsem/TammV18} represent a significant advancement in automata theory, offering improved efficiency in operations and enhanced expressiveness through algebraic theories that support infinite alphabets. This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures. Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.

\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications. These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}. Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.

\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification. To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant. Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.

\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade. While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
We believe that our work will benefit the community by providing a formal foundation for string solvers development.

\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains. For example, the work by Shi et al. \cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.

### 2025-08-15 Paper/application.tex
add applications

@@
 Beyond performance evaluation, our SFT formalization enables formal verification of string sanitizer correctness.
-For example, this verification framework models user input as an SFA, the sanitizer as an SFT, and potential attacks as an SFA accepting unsafe strings.
-A sanitizer is verified correct if the product of the sanitizer SFT and user input SFA produces no outputs that intersect with the attack SFA's language.
-This correctness property can be checked by computing the intersection of the product SFT's output language with the attack SFA, where emptiness of this intersection guarantees sanitizer safety.
+A santizer is used to transform user inputs into safe outputs, ensuring that the output does not contain any unsafe characters.
+However, santizers just like programs, which are eror-prone.
+Manual reviews of sanitizers are often insufficient to guarantee correctness.
+
+Based on our formalization, we can verify sanitizers by modeling the sanitizer as an SFT and the user input as an SFA.
+We can also model potential attacks as SFAs that accept unsafe strings.
+Formally speaking, Let $\mathcal{T}$ be an SFT, $\mathcal{I}$ is the possible user inputs modeled by an SFA, and $\mathcal{A}$ is the attack SFA that accepts unsafe strings.
+The santizer correctness checking against the attack is to check whether the intersection of the product $\mathcal{T} \times \mathcal{I}$ and the attack SFA $\mathcal{A}$ is empty, i.e., $\mathcal{T} \times \mathcal{I} \cap \mathcal{A} = \emptyset$.
+
+Moreover, santizers are usually used in combination with other sanitizers to ensure comprehensive safety.
+For example, the HTML encoding sanitizer can be combined with a string trimming operation to remove leading and trailing whitespace characters.
+For this kind of sanitizer composition, it can be modeled by our SFT product as well.
+Let $\mathcal{T}_1$ and $\mathcal{T}_2$ correspond to two sanitizers, and $\mathcal{T}_1$ is applied before $\mathcal{T}_2$.
+What we need to check is $\mathcal{T}_2 \times (\mathcal{T}_1 \times \mathcal{I}) \cap \mathcal{A} = \emptyset$.
+
+\begin{table}[h]
+  \centering
+  \small
+  \setlength{\tabcolsep}{4pt} % Reduced column gap
+  \begin{tabular}{p{2.2cm}ll} % Increased width for automatic line break
+    \toprule
+    \textbf{Attack Model} & \textbf{Verfication Result} & \textbf{Size} \\
+    \midrule
+    attack\_html           & GA HtmlEscape: \textbf{Unsafe} & 1/6\\
+                           & Trim + GA HtmlEscape: \textbf{Safe} & 4/12 \\
+                           & escapeString: \textbf{Unsafe} & 1/11 \\
+                           & Trim + escapeString: \textbf{Safe} & 4/17 \\
+                           & Trim + OWASP HTMLEncode: \textbf{Safe} & 4/18 \\
+    \midrule
+    attack\_javascript     & GA Htmlescape + GA PreEscape + & 3/20 \\
+                           &  gaSnippetesc: \textbf{safe} & \\
+    \midrule
+    attack\_json           & Trim + GA JSONESc : \textbf{Safe} & 4/16\\
+    \midrule
+    attack\_xml            & Trim + GA XMLEsc: \textbf{safe} & 4/16\\
+    \midrule
+    attack\_ccs            & Trim + GA CleanseCSS : \textbf{Safe} & 25/46\\
+    \bottomrule
+  \end{tabular}
+  \caption{Sanitizer verfication result}
+  \label{tab:sanitizer_attack}
+\end{table}
+
+Table \ref{tab:sanitizer_attack} summarizes the verification results for various sanitizers against different attack models.
+The first column lists the attack models, the second column indicates whether the sanitizer is classified as safe or unsafe with respect to the attack, and the third column reports the size of the transducers used in the verification, given as [number of states/number of transitions].
+For composed SFTs, the size reflects the total number of states and transitions across all components.
+
+We evaluate a range of attack models for HTML, JavaScript, JSON, XML, and CSS.
+Listing \ref{lst-css-attack} shows the CSS attack model, which is used to verify the corresponding sanitizer.
+The attack model includes comments, declaration breakers, and known dangerous tokens such as \textbf{expression(}, \textbf{@import}, \textbf{behavior:}, \textbf{-moz-binding}, and \textbf{url(} (unless the URL is separately validated).
+Sanitizers include GA HtmlEscape, GA JSONEscape, GA XMLEscape, escapeString, and GA CleanseCSS, where "GA" means from Google.
+Sanitizers are often composed with preprocessing steps such as Trim, which removes leading and trailing whitespace characters.
+As shown in the table, GAHtmlEscape and escapeString are unsafe with respect to the HTML attack model without preprocessing, but become safe when combined with Trim.
+The \textbf{Size} column demonstrates the expressive power and succinctness of SFTs in modeling real-world transducers.
+
+\begin{lstlisting}[language={}, caption={CSS attack model for sanitizer verification.}, label={lst-css-attack}, float=htbp]
+|$\Sigma$|* ( / \* [\s\S]*? \*/
+  |$\mid$| ; |$\mid$| \{ |$\mid$| \}
+  |$\mid$| [eE][xX][pP][rR][eE][sS][sS][iI]
+    [oO][nN]\s*\(
+  |$\mid$| @\s*[iI][mM][pP][oO][rR][tT]
+  |$\mid$| [bB][eE][hH][aA][vV][iI][oO][rR]\s*:
+  |$\mid$| -\s*[mM][oO][zZ]-\s*[bB][iI][nN]
+    [dD][iI][nN][gG]
+  |$\mid$| (?!allowUrl) [uU][rR][lL]\s*\(
+ ) |$\Sigma$|*
+\end{lstlisting}
 

### 2025-08-17 Paper/SFT.tex

\section{Symbolic Finite Transducers}
\label{sec:formalization}

We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL.

Let $\mathcal{U}$ be a multi-sorted carrier set or background universe, which is equipped with functions and relations over the elements. We use $\tau$ as a sort and $\mathcal{U}^\tau$ denotes the sub-universe of elements of type $\tau$.
We have a special type $\mathbb{B}$ with $\mathcal{U}^\mathbb{B} = \{ \top, \bot\}$, which corresponds to the boolean type.

A lambda term is defined as $\lambda x.~t$ of type $\tau_1 \rightarrow \tau_2$.
When $\tau_2$ is $\mathbb{B}$, this lambda term is a predicate. Let $\phi$ be a predicate. We write $a\in \llbracket\phi \rrbracket$ if $\phi~a=\top$. For non-predicate lambda terms, we view them as functions that generate output elements of type $\tau_2$ given input terms of type $\tau_1$.
With these notations and the above definitions, we can define SFTs as follows.

\begin{definition}[Symbolic Finite Transducer]
\label{def-sft}
   A Symbolic Finite Transducer over $\tau_1\rightarrow \tau_2$ is a quadruple $\mathcal{T} = (\mathcal{Q}, \Delta, \mathcal{I}, \mathcal{F})$, where 
   \begin{itemize}
   \item $\mathcal{Q}$ is a finite set of states,
   \item $\mathcal{I}\subseteq \mathcal{Q}$ is the set of initial states,
   \item $\mathcal{F} \subseteq\mathcal{Q}$ is the set of accepting states,
   \item $\Delta$ is the set of transition relations. Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
   \end{itemize}

For each transition $q\xrightarrow{\phi, f} q'$, if there exists an element $a\in \llbracket \phi \rrbracket$, where $a$ is called an input,  then the application $(f~a)$ is the output.

\end{definition}

SFTs accept an input word and generate an output word. This can be defined by \emph{runs} of SFTs.
An SFT run $\sigma$ is a sequence $(q_0, \phi_0, f_0, q_1),(q_1, \phi_1, f_1, q_2),\ldots, (q_{n-1}, \phi_{n-1}, f_{n-1}, q_n)$ such that $q_0\in \mathcal{I}$ and $(q_i,
\phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
$\sigma$ is an accepting run when $q_n$ is an accepting state.

For a word $w = a_0,\ldots, a_{n-1}$, it is accepted by $\sigma$ if and only if $a_i \in \llbracket\phi_i \rrbracket$ for $0 \leq i \leq n - 1$. When $w$ is accepted, run $\sigma$ generates an output sequence $w'= f_0~a_0, \ldots, f_{n-1}~a_{n-1}$. We define:
\begin{itemize}
\item  $(a_0,(f_0~a_0)), \ldots, (a_{n-1},(f_{n-1}~a_{n-1}))$ a \emph{trace},
\item $a_0,\ldots,a_{n-1}$ the \emph{input} of the trace and $(f_0~a_0), \ldots, $ $(f_{n-1}~a_{n-1})$ the \emph{output} of the trace.
\end{itemize}

If $a_0,\ldots,a_{n-1}$ is accepted by an accepting run in $\mathcal{T}$, we say that the trace is an accepting trace of $\mathcal{T}$.
For a trace $\pi$, we denote its input as $\mathit{in}(\pi)$ and its output as $\mathit{out}(\pi)$.
%
Given an SFT $\mathcal{T}$ and a word $w$, we define the \emph{product} operation of $\mathcal{T}$ and $w$ (denoted as $\mathcal{T}\times\{w\}$) as the set of outputs generated by $\mathcal{T}$ with input $w$. More precisely,
\[
\begin{split}
\mathcal{T}\times\{w\} = \{w'\mid \exists \pi.~\pi \text{ is an accepting trace of } \mathcal{T} \land \\ \mathit{in}(\pi) = w \land \mathit{out}(\pi) = w'\}.
\end{split}
\]

To make the operation \emph{product} more general, we extend the operation to an SFT and a set of input words represented by a regular language, which can be denoted by an SFA $\mathcal{A}$. More precisely,
\[
\begin{split}
\mathcal{T}\times \mathcal{A} = \{w'\mid \exists w.~w\in \mathcal{L}(\mathcal{A})\land w' \in \mathcal{T}\times\{w\}\}, \\\text{ where }\mathcal{L}(\mathcal{A})\text{ denotes the language of }\mathcal{A}.
\end{split}
\]

\begin{figure}[hbt!]
  \centering
  \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm and 2cm, on grid, auto]
    % Make q0 and q1 horizontally wide apart
    \node[state, initial] (q_0)   {$q_0$}; 
    \node[state] (q_1) [right=4cm of q_0] {$q_1$}; 
    \node[state, accepting] (q_2) [below=2.2cm of q_1] {$q_2$};

    \path[->] 
    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
  \end{tikzpicture}
    \caption{An example of SFT}
    \label{fig-example-ft}
\end{figure}

    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$. The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$. The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation. For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.

%The transition labels can be more complex. For instance, it can be a predicate over integers as: $\lambda x.~ x >0 \land x \text{ mod } 2 = 0$. This symbolic label corresponds to the set of even natural numbers.

### 2025-08-17 Paper/conclusion.tex

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL. Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.

To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}. This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}. The experimental results confirm both the efficiency and effectiveness of our approach. While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.

Future work includes extending our symbolic formalization in two key directions:
%
(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.

### 2025-08-17 Paper/formalization.tex

\section{The Isabelle/HOL Formalization of SFTs}

We present the basic architecture of our SFT formalization in Isabelle/HOL, which consists of three distinct layers.

The \textbf{abstract layer} defines transition labels as \textbf{sets}, states are stored in sets as well.  This abstract formalization, which abstracts away implementation details, enables straightforward correctness proofs for SFT operations.

The \textbf{implementation layer} refines the abstract transition labels to a locale where a boolean algebra is defined as type variable \texttt{'b} together with a collection of operations and assumptions. State sets are refined using Refine\_Monadic \cite{Refine_Monadic-AFP}, a framework in Isabelle/HOL for data refinement that automatically transforms general sets into efficient data structures such as hashmaps and red-black trees.

The \textbf{boolean algebra layer} refines the transition label locale to a concrete boolean algebra implementation. In this paper, we present an interval-based refinement.

This three-layer architecture separates the fixed components (abstract and implementation layers) of the SFT formalization from the variable components (transition labels). This separation enables us to extend the SFT formalization to accommodate new transition label theories without modifying the first two layer formalization.

\subsection{The Abstract Layer of SFTs}
\label{sec:abstract-layer}

We present the SFT formalization in abstract layer in Fig. \ref{fig-def-FT}. While the elements $\mathcal{Q}_t$, $\mathcal{I}_t$, and $\mathcal{F}_t$ directly correspond to their counterparts ($\mathcal{Q}$, $\mathcal{I}$, and $\mathcal{F}$) in Definition \ref{def-sft}, the transition relations are not exactly the same.
%
The transition relations are formalized through \texttt{LTTS} (Labeled Transducer Transition System), where each transition is represented as a triple $'q \times ('a \texttt{ set option }\times 'i) \times 'q$. This representation reflects several key design decisions aimed at enhancing the abstraction and flexibility of our SFT formalization.
\begin{figure}[hbt!]
	\begin{lstlisting}
record (|$'q,$||$~'a,$| |$'i,$| |$'b$|) |$\texttt{NFT}$| =
	|$\mathcal{Q}_t$| :: "|$'q$| set"
	|$\Delta_t$| :: "(|$'q,$||$~'a,$| |$'i$|) LTTS"
	|$\mathcal{I}_t$| :: "|$'q$| set"
	|$\mathcal{F}_t$| :: "|$'q$| set"
        |$\mathcal{M}_t$| :: "|$'i\Rightarrow ('a,~'b)$| Tlabel"
type_synonym |$('q,~'a,~'i)$| LTTS = "|$('q \times ('a \text{ set option }\times 'i) \times 'q)$| set"
type_synonym |$('a,~'b)$| Tlabel = "|$'a \text{ option}~\Rightarrow~'b$ set option"
	\end{lstlisting}
\caption{The formalization of $SFTs$ in Isabelle/HOL}
\label{fig-def-FT}
\end{figure}

Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate. A predicate's semantics as introduced before represents a set of elements that make the predicate true. But predicates have various different forms. For instance, the interval $[1-9]$ represents the set $\{e \mid 1 \leq e \leq 9\}$. The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position. All these different forms of labels are abstracted as sets in our formalization.
%
The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements. This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.

The second element of type $'i$ serves as an index into the output function space. 
%
We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
%
The mapping $\mathcal{M}_t$ associates each index with a specific output function. These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element. This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria. Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.

%Now let us look at an example. In some modern programming languages, there are support for replacement operations : $\texttt{replace}(\texttt{str}, \texttt{pattern}, \texttt{replacement})$. This operation replaces the first occurrence of the substring  in \texttt{str} that matches \texttt{pattern} (which usually a regular expression) with the string \texttt{replacement}). 
%We can easily model this operation using our formalization of NFTs with the following steps:
%\begin{itemize}
%\item Step 1, construct a corresponding NFA from \texttt{pattern}).
%\item Step 2, convert each transition label for transducers by adding a function that maps each input character to empty, that is, generate nothing.
%\item Step 3, Add a new accepting state, and redirect existing accepting states to the new accepting state and transition label as $(\texttt{None}, \lambda x. \texttt{replacement})$.
%\item Step 4, unlabel previous accepting states as non-accepting states. 
%\end{itemize}

%We can see that the definition of transition labels in NFTs makes it is very easy to model the replacement operation.

\paragraph{The product operation formalization.} We formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}. The SFA is formalized as a record type \texttt{NFA in Figure \ref{fig-def-FAs}}.
An additional consideration in this operation is the presence of $\varepsilon$-transitions in our SFT formalization, which implies that the resulting automata may also contain $\varepsilon$-transitions.
%
Therefore, we need to formalize $\varepsilon$SFAs, which is shown in Fig. \ref{fig-def-FAs} as \texttt{eNFA}. The $\varepsilon$SFA formalization extends the standard SFA structure by introducing $\Delta_e'$, which captures $\varepsilon$-transitions as pairs of states, while maintaining the same labeled transition relation $\Delta$ as in standard SFAs.
In this paper, we present only the definitions of  $\varepsilon$SFAs and SFAs, while the complete formalization, including correctness proofs, is available in our Isabelle development.

\begin{figure}[hbt!]
	\begin{lstlisting}
record (|$'q,$||$~'a$|) |$\texttt{NFA}$| =
	|$\mathcal{Q}$| :: "|$'q$| set"
	|$\Delta$| :: "(|$'q,$||$~'a$|) LTS"
	|$\mathcal{I}$| :: "|$'q$| set"
	|$\mathcal{F}$| :: "|$'q$| set"

record (|$'q,$||$~'a$|) |$\texttt{eNFA}$| =
	|$\mathcal{Q}_e$| :: "|$'q$| set"
	|$\Delta_e$| :: "(|$'q,$||$~'a$|) LTS"
	|$\Delta_e'$| :: "|$('q * 'q)$| set"
	|$\mathcal{I}_e$| :: "|$'q$| set"
	|$\mathcal{F}_e$| :: "|$'q$| set"

type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
	\end{lstlisting}
\caption{The formalization of $\varepsilon$SFAs and SFAs}
\label{fig-def-FAs}
\end{figure}

Having established these foundational definitions, we can now formalize the product operation.

\begin{figure}[hbt!]
	\begin{lstlisting}
definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$| 
   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$| 
   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
"productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
  |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
       |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
        |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
       |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
       |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
          |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
  |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
       |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
       |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
      |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
        |$ (p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land (p', \sigma_2, q')\in\Delta~\mathcal{A}$|
        |$\land~\sigma_1\cap \sigma_2 \neq \emptyset~\land\exists x\in(\sigma_1\cap\sigma_2).$|
        |$~((\mathcal{M}_t~\mathcal{T})~f)~(\text{Some } x) = \text{None} \}$|,
  |$\mathcal{I}_e = \mathcal{I}_t~\mathcal{T} \times \mathcal{I}~\mathcal{A}$|
  |$\mathcal{F}_e = \mathcal{F}_t~\mathcal{T} \times \mathcal{F}~\mathcal{A}$| |$\rrparenthesis$|"
	\end{lstlisting}
\caption{The formalization of product operation}
\label{fig-def-FTProd}
\end{figure}

Figure \ref{fig-def-FTProd} depicts the abstract level formalization for the product of an SFT and an SFA. The parameters $\mathcal{T}$ and $\mathcal{A}$ are an SFT and an SFA, respectively. The 
result of the product operation is an $\varepsilon$SFA.
But we need to explain the role of parameter $\text{\texttt{F}}$. The output function $f$ for each transition in $\Delta_t$ is of type "$'a\;\text{option} \Rightarrow 'b\;\text{set option}$", which applies to a \emph{single} element of type $'a$ or $\varepsilon$. $\text{\texttt{F}}$ extends $f$ to apply $f$ to a set of elements. More precisely, let $f$ be an output function, the semantics of $\text{\texttt{F}}$ is defined as follows:

\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].

The transition relations $\Delta_e$ and $\Delta_e'$ are determinted by considering two distinct cases based on the nature of transitions in the SFT: $\varepsilon$-transitions and non-$\varepsilon$-transitions. In both cases, the transition labels in the resulting $\varepsilon$SFA are derived from the composition of the SFT's output function and the SFA's input labels. Let us consider $\Delta_e$ first.

\begin{enumerate}
\item When $(p, (\text{None}, f), q)$ is a transition in $\Delta_t$, i.e. the input character is $\varepsilon$, and $f~\text{None} \neq \text{None}$. Consequently, the SFA $\mathcal{A}$ remains in its current state, and the product transition produces the output "$((\mathcal{M}_t~\mathcal{T})~f)~\text{None}$". Remember that $f$ is just an index, $(\mathcal{M}_t~\mathcal{T})~f$ is the output function.

\item When $(p, (\text{Some}~\sigma_1, f), q)$ is a transition in $\Delta_t$, synchronization is possible only with SFA transitions that share characters with $\sigma_1$, i.e., $\sigma_1\cap\sigma_2\neq \emptyset$, where $\sigma_2$ represents the input label of the corresponding SFA transition. The resulting output is $\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)$.
\end{enumerate}

The transitions in $\Delta_e'$ follow a similar pattern with analogous cases for $\varepsilon$ and non-$\varepsilon$ transitions.

To establish the correctness specification of the product operation, we begin by formalizing the concept of SFT \emph{traces} as introduced in Definition \ref{def-sft}. In our formalization, traces are represented by the type $('a\;\text{option} \times 'b\;\text{option})\;\texttt{list}$, as shown in Figure \ref{fig-def-output}.
%
Given a trace $\pi$, we define two key projection functions:
(1) \texttt{inputE}, which corresponds to ${in}(\pi)$ and extracts the input sequence
(2) \texttt{outputE}, which corresponds to ${out}(\pi)$ and extracts the output sequence.

The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$. The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.

\begin{figure}[hbt!]
	\begin{lstlisting}
fun inputE :: 
   "|$('a\;\text{option}\;\times\;'b\;\text{option})\;\text{list}\;\Rightarrow\;'a\;\text{list}$|"
where
  "inputE [] = []" |$\mid$|
  "inputE ((Some a, |$\_$|) |$\#$| l) = 
                      a |$\#$| (inputE l)" |$\mid$|
  "inputE ((None, |$\_$|) |$\#$| l) = (inputE l)"

fun outputE :: 
    "|$('a \text{ option} \times 'b \text{ option}) \text{ list} \Rightarrow 'b \text{ list}$|" 
where
  "outputE [] = []" |$\mid$|
  "outputE ((_,Some a) # l) = 
                    a # (outputE l)" |$\mid$|
  "outputE ((_,None) # l) = (outputE l)"

definition outputL :: 
   "|$('q, 'a, 'i, 'b)\text{  NFT} \Rightarrow 
                               ('q, 'a) \text{ NFA} \Rightarrow 'b\text{ list set}$|" where
  "outputL |$\mathcal{T}$| |$\mathcal{A}$| = {outputE |$\pi~\mid~\pi~q~q'$|. 
    |$q\in \mathcal{I}_t~\mathcal{T} \land q' \in \mathcal{F}_t~\mathcal{T}~\land$| 
    |$ \text{LTTS\_reachable }\mathcal{T}~q~\pi~q'\land$||$ \text{inputE }\pi\in \mathcal{L}~\mathcal{A}$|}"
	\end{lstlisting}
\caption{The formalization of traces in SFTs}
\label{fig-def-output}
\end{figure}

\begin{figure}[hbt!]
	\begin{lstlisting}
lemma productT_correct:
  fixes |$\mathcal{T}$| |$\mathcal{A}$| F
  assumes 
   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
       |$\text{ F}~f~s = \text{None}"$|
   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
        |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
        |$f~(\text{Some }e) = \text{ Some S}\})$|"
      and wfTA: "NFT_wf |$\mathcal{T}~\land \text{ NFA } \mathcal{A}$|"
  |$\textcolor{blue}{\text{shows}}$| "|$\mathcal{L}_e$| (productT |$\mathcal{T}~\mathcal{A}$| F) = outputL 
          |$\mathcal{T}~\mathcal{A}$|"

	\end{lstlisting}
\caption{The correctness lemma of the product operation}
\label{fig-def-product-correct}
\end{figure}

Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation. The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}. 
The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.
The conclusion, marked by \textcolor{blue}{\texttt{shows}}, demonstrates that the language of the constructed $\varepsilon$SFA ($\mathcal{L}_e$ denotes the language of $\varepsilon$SFA) from $\mathcal{T} \times \mathcal{A}$ coincides with the mathematical semantics defined by \texttt{outputL}, thereby establishing semantic preservation of the product construction.

\subsection{Implementation Layer Refinement}
\label{sec_alg_refinement}
Having established the abstract definition of the SFT product operation in Section \ref{sec:abstract-layer}, we now present its algorithmic refinement. This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.

\paragraph{Boolean Algebra Locale}

Figure \ref{fig-bool-algebra-locale} presents the simplified definition of a boolean algebra locale. The type variable \texttt{'b} represents any boolean algebra, while the type variable \texttt{'a} denotes the element type of the sets that the boolean algebra represents. The locale defines a collection of operations and assumptions that characterize boolean algebra behavior, including operations for set semantics, emptiness checks, non-emptiness checks, intersection, difference, and element membership. The function "\texttt{sem}" provides the semantic interpretation of a boolean algebra as a set, with all other operations' assumptions defined in terms of this set semantics.

The locale serves as an interface to boolean algebra implementations. Any concrete boolean algebra implementation must provide a concrete instantiation of this locale, which includes defining the type variable \texttt{'b} and providing implementations for all operations along with proofs of the required assumptions.

\begin{figure}[hbt!]
\begin{lstlisting}[mathescape=true]
locale bool_algebra =
 fixes sem :: " 'b $\Rightarrow$ 'a::ord set"
 fixes empty:: "'b $\Rightarrow$ bool"
 fixes nempty:: "'b $\Rightarrow$ bool"
 fixes intersect:: "'b $\Rightarrow$ 'b $\Rightarrow$ 'b"
 fixes diff:: "'b $\Rightarrow$ 'b $\Rightarrow$ 'b"
 fixes elem:: "'a  $\Rightarrow$ 'b $\Rightarrow$ bool"
 assumes
  empty_sem: "empty s = (sem s = {})"
                                    and
  nempty_sem: "nempty_intervals s =
                     ($\neg$ (empty s))" and
  inter_sem: "sem (intersect s1 s2) =
               (sem s1) $\cap$ (sem s2)" and
  diff_sem: "sem (diff f1 f2 s1 s2) =
               (sem s1) - (sem s2)" and
  elem_sem: "elem a s $\equiv$  (a $\in$ sem s)"
\end{lstlisting}
\caption{The boolean algebra locale}
\label{fig-bool-algebra-locale}
\end{figure}

\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.} based on Refined\_monadic framework.

Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework. Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}. These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.

The operation \texttt{prods\_imp}, shown in Figure \ref{fig-def-prods_imp}, computes the Cartesian product of two state sets. This function employs the \texttt{FOREACH} construct, a higher-order iteration operator analogous to OCaml's \texttt{Set.fold}. Specifically, given a set $S$, a function $f$ of type $'a \Rightarrow 'b \Rightarrow 'b$, and an initial accumulator $I$ of type $'b$, the expression $\texttt{FOREACH}~S~f~I$ systematically applies $f$ to each element in $S$, accumulating results in a principled manner.

\begin{figure}[hbt!]
	\begin{lstlisting}
definition productT_impl where
"productT_impl |$\mathcal{T}$| |$\mathcal{A}$| F fe = do {
    Q |$\leftarrow$| prods_imp (nft_states |$\mathcal{T}$|) 
    (nfa_states |$\mathcal{A}$|);
    (D1, D2) |$\leftarrow$| trans_comp_imp 
     (nft_tranfun |$\mathcal{T}$|) F fe (nft_trans |$\mathcal{T}$|) 
      (nfa_trans |$\mathcal{A}$|) (nfa_states |$\mathcal{A}$|);
    I |$\leftarrow$| prods_imp (nft_initial |$\mathcal{T}$|)
      (nfa_initial |$\mathcal{A}$|);
    F |$\leftarrow$| prods_imp (nft_accepting |$\mathcal{T}$|) 
      (nfa_accepting |$\mathcal{A}$|);
    RETURN (Q, D1, D2, I, F)
  }"
\end{lstlisting}
\caption{The computation of SFT product}
\label{fig-compute-nft-product}
\end{figure}

\begin{figure}[hbt!]
	\begin{lstlisting}
definition prods_imp where
"prods_imp Q1 Q2 =
   FOREACH {q. q |$\in$| Q1} (|$\lambda$| q Q. do {
    S |$\leftarrow$| FOREACH {q. q |$\in$| Q2}
      (|$\lambda$| q|$'$| Q|$'$|. RETURN ({(q,q|$'$|)} |$\cup$| Q|$'$|)) |$\emptyset$|;
    RETURN (Q |$\cup$| S)
   }) |$\emptyset$|"
\end{lstlisting}
\caption{The computation of Cartesian product of two state sets}
\label{fig-def-prods_imp}
\end{figure}

\begin{figure}[hbt!]
	\begin{lstlisting}
definition trans_comp_imp where
"trans_comp_imp M F fe T1 T2 Q =
  FOREACH {t. t |$\in$| T1}
   (|$\lambda$|(q,(|$\alpha$|,f),q|$'$|) (D1,D2). 
    (if (|$\alpha$|=None) then 
     (subtrans_comp_|$\varepsilon$| M q f q|$'$| F fe 
                              T2 D1 D2)
     else
     (subtrans_comp M q (the |$\alpha$|) f 
            q|$'$| F fe T2 D1 D2))) (|$\emptyset$|, |$\emptyset$|)"
\end{lstlisting}
\caption{The computation of \texttt{trans\_comp\_imp}}
\label{fig-def-prods-imp}
\end{figure}

A central algorithmic challenge in our implementation lies in the computation of transition sets $\texttt{D1}$ (corresponding to $\Delta_e$) and $\texttt{D2}$ (corresponding to $\Delta_e'$) in Figure \ref{fig-compute-nft-product}. As shown in Figure \ref{fig-def-prods-imp}, we implement this computation through the function \texttt{trans\_comp\_imp}, which computes the synchronization of transitions between the SFT and SFA. This function decomposes the synchronization process into two distinct cases, each handled by a specialized function:

\begin{enumerate}
  \item \texttt{subtrans\_comp\_$\varepsilon$}: Processes $\varepsilon$-transitions in the SFT, where transitions consume no input but may produce output
  \item \texttt{subtrans\_comp}: Processes standard transitions in the SFT, where both input consumption and output generation may occur
\end{enumerate}

\begin{figure}[hbt!]
\begin{lstlisting}[language=SMTLIB]
definition subtrans_comp where
"subtrans_comp M q |$\alpha$| f q|$'$| F fe T D1 D2 =
 FOREACH {t.t|$\in$|T} (|$\lambda$| (q1,|$\alpha'$|,q1|$'$|) (D1,D2).
  (if (nempty (intersect |$\alpha$| |$\alpha'$|)) then
    do {
      D1 |$\leftarrow$| 
       (if (F (M f) (intersect |$\alpha$| |$\alpha'$|)) 
          |$\neq$| None) then
         let |$\alpha_i$| = the (F (M f) 
              (intersect |$\alpha$| |$\alpha'$|)) in
         RETURN {((q,q1), |$\alpha_i$|, (q|$'$|,q1|$'$|))} 
                |$\cup$| D1
        else RETURN D1);
      D2 |$\leftarrow$| 
       (if fe (M f) (intersect |$\alpha$| |$\alpha'$|) 
        then 
         RETURN {((q,q1), (q|$'$|,q1|$'$|))} |$\cup$| D2 
        else RETURN D2);
      RETURN (D1, D2) }
   else (RETURN (D1, D2)))) (D1, D2)"
    \end{lstlisting}
    \caption{The computation of \texttt{subtrans\_comp}}
    \label{fig-def-subtrans_comp}
    \end{figure}

    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles). For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}. For each transition $(q_1, \alpha', q_1')\in \texttt{T}$, the function performs two key operations when the intersection of input labels is non-empty ($\alpha \cap \alpha' \neq \emptyset$, verified using \texttt{nemptyIs}):

    \begin{enumerate}
      \item Computes non-$\varepsilon$-transitions ($\texttt{D1}$): When the output function applied to the intersection $\alpha \cap \alpha'$ yields a non-empty set, a new transition is added to $\texttt{D1}$ with the computed output label.
      \item Generates $\varepsilon$-transitions ($\texttt{D2}$): When there exists at least one input in the intersection $\alpha \cap \alpha'$ that produces an empty string (verified by checking if $\texttt{M}~f$ maps any element to \texttt{None}. The checking is implemented by \texttt{fe}), a corresponding $\varepsilon$-transition is added to $\texttt{D2}$.
    \end{enumerate}

  \begin{figure}[hbt!]
    \begin{lstlisting}
  lemma productT_imp_correct:
  assumes finite_TT: "|$\text{finite } (\Delta_t~\mathcal{T})$|"
      and finite_TA: "|$\text{finite } (\Delta~\mathcal{A})$|"
      and finite_Q: "|$\text{finite } (\mathcal{Q}~\mathcal{A})$|"
      and finite_TQ: "|$\text{finite } (\mathcal{Q}_t~\mathcal{T})$|"
      and finite_I: "|$\text{finite } (\mathcal{I}~\mathcal{A})$|"
      and finite_TI: "|$\text{finite } (\mathcal{I}_t~\mathcal{T})$|"
      and finite_F: "|$\text{finite } (\mathcal{F}~\mathcal{A})$|"
      and finite_TF: "|$\text{finite } (\mathcal{F}_t~\mathcal{T})$|"
  shows "|$\text{productT\_imp } \mathcal{T}~\mathcal{A}~F~\text{fe}\leq$|
            |$\text{SPEC } (\lambda A.~A = \text{productT } \mathcal{T}~\mathcal{A}~F)$|"
  \end{lstlisting}
  \caption{The refinement relation between \texttt{productT\_imp} and \texttt{productT}}
  \label{fig-def-productT_imp_correct}
  \end{figure}

  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}). This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.

  The refinement is expressed through the relation $\mathbf{C} \leq \texttt{SPEC}~\mathbf{A}$, which asserts that the concrete implementation $\mathbf{C}$ is an element of the abstract specification $\mathbf{A}$. More precisely,  the concrete implementation must produce an $\varepsilon$SFA that is structurally equivalent or isomorphic to the one produced by the abstract algorithm \texttt{productT}.

  To establish this equivalence, we must prove that the $\varepsilon$SFAs produced by \texttt{productT\_imp} and \texttt{productT} are isomorphic in all essential components: the set of states, transition relations, $\varepsilon$-transition relations, initial and accepting state sets.

The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 
In our formalization, we refine the sets of storing states and transitions to red-black trees.

\subsection{Interval Implementation}

Finally, we present the implementation of the interval algebra.
An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$. To achieve greater expressiveness, our formalization extends this notion to interval lists of the form $[(i_1, j_1), \ldots, (i_n, j_n)]$, which denote the set $\bigcup_{1\leq k\leq n}\{e \mid i_k \leq e \leq j_k\}$. This generalization offers two key advantages: it enables more compact representation of transitions in SFAs and SFTs through merging, and it allows for efficient handling of interval operations without unnecessary splitting. For example, the set difference between intervals $(1, 5)$ and $(3, 4)$ can be directly represented as the interval list $[(1, 2), (5, 5)]$, which is also an interval.

Throughout the following discussion, we use the term "interval" to refer to interval lists. Our formalization provides all operations in the boolean algebra locale. For instance, we implement \texttt{intersect} for intervals, which computes the intersection of two intervals $i_1$ and $i_2$, yielding an interval $i$ such that $\texttt{sem}~i = \texttt{sem}~i_1 \cap \texttt{sem}~i_2$

To facilitate formal reasoning and optimize performance, in addition to the above locale operations, we introduce a canonical form for interval. An interval $[(i_1, j_1), $ $\ldots, (i_n, j_n)]$ is in canonical form if it satisfies two key properties:
\begin{enumerate}
  \item Each $(i_k, j_k)$ is well-formed: $i_k \leq j_k$ for all $k \in \{1,\ldots,n\}$
  \item Intervals are ordered and non-overlapping: $j_k < i_{k+1}$ for all $k \in \{1,\ldots,n-1\}$
\end{enumerate}

We prove that all interval operations preserve canonical form when applied to canonically-formed inputs. This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.

### 2025-08-18 CertiStr/test/sat/sat.sh
add one more replace-all

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 for test in ./*.smt2; do
-    result=$(CertiStr $test)
+    result=$(CertiStr --left-most $test)
     if ( [ $result != 'SAT' ])
     then
         echo "Test ($test) Fail:"

### 2025-08-18 CertiStr/test/unsat/unsat.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 for test in ./*.smt2; do
-    result=$(CertiStr $test --left-most)
+    result=$(CertiStr --left-most $test)
     if ( [ $result != 'UNSAT' ])
     then
         echo "Test ($test) Fail:"

### 2025-08-20 Paper/application.tex
polish application
finish experiments

@@
 The \textbf{Size} column demonstrates the expressive power and succinctness of SFTs in modeling real-world transducers.
+We did not list the execution time of verification because all of them are quite efficent with less than 0.02 second to finish the verificaiton.
 
@@
 
-The string replacement operation, denoted as
-$\texttt{replace}(\texttt{str}, \newline\texttt{pattern}, \texttt{replacement})$, is a fundamental string transformation that takes three parameters:
+The string replacement, denoted as
+$\texttt{replace}(\texttt{str}, \texttt{pattern}, \newline\texttt{replacement}, \texttt{flag})$, is a fundamental string transformation accepting four parameters:
 \begin{itemize}
@@
   \item $\texttt{pattern}$: A regular expression defining the matching criteria.
-  \item $\texttt{replacement}$: The string to be substituted for the matched substring.
+  \item $\texttt{replacement}$: The string that replaces matched substrings.
+  \item $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single occurrence replacement and $g$ denotes replace-all operation.
 \end{itemize}
 
-The semantics of the replacement operation can be formally characterized by two distinct cases:
-\begin{enumerate}
-  \item When there exists at least one substring $s'$ in \texttt{str} such that $s'$ matches $\texttt{pattern}$, then $s'$ is replaced with \texttt{replacement}.
-  \item When no substring of \texttt{str} matches $\texttt{pattern}$, the operation returns \texttt{str} unchanged.
-\end{enumerate}
+The operation's semantics is to replace the first occurrence (or all occurrences) of any substring $s'$ in \texttt{str} that matches $\texttt{pattern}$ with \texttt{replacement}.
 
-The first case can be modeled by SFTs.
-We illustrate this case by an example.
-Given a replacement operation $\texttt{replace}\newline(s, \texttt{/[0-9]+/}, "\texttt{NUM}")$, which means replacing a occurrence of the substring that matches the regular expression $\texttt{/[0-9]+/}$ with the string "\texttt{NUM}".
-We model this replacement operation by an SFT with the following 3 steps:
+To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}", \texttt{flag})$. We construct the SFT representation through three steps:
 
 \begin{enumerate}
-  \item First, construct an SFA that recognizes the regular expression pattern $\texttt{/[0-9]+/}$.
-Transform this SFA into an SFT by augmenting each transition with an output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
-  \item Second, construct an SFA that accepts the replacement string "\texttt{NUM}".
-Since a constant string can be viewed as a specialized regular expression, we can construct its SFA representation.
-Convert this SFA into an SFT by adding $\varepsilon$-transitions  and appropriate output functions that emit the characters of "\texttt{NUM}" in sequence.
-  \item Finally, compose the two SFTs through concatenation and augment the resulting transducer with self-loop transitions at the initial and final states.
-These additional transitions, labeled with $\Sigma$ (the set of all characters in the alphabet) and the identity function $\texttt{id}=\lambda x.~x$, enable the SFT to process arbitrary prefixes and suffixes of the input string while preserving the replacement behavior on matched substrings.
+  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
+  \item Construct an SFA accepting the replacement string "\texttt{NUM}".
+Since constant strings are specialized regular expressions, we build their SFA representation and convert it to an SFT by adding $\varepsilon$-transitions with output functions that emit the replacement characters sequentially.
+  \item Compose the two SFTs by connecting the accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT via $\varepsilon$-transitions.
 \end{enumerate}
@@
 \ref{fig-snfa-replacement}).
-The composition process involves two key modifications:
-\begin{enumerate}
-  \item Connect the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT
-  \item Augment the resulting transducer with self-loop transitions at both ends, labeled with "$\Sigma,~\texttt{id}$", where $\Sigma$ represents the full alphabet.
-For the string solver, $\Sigma$ is the set of all unicode characters.
-These transitions enable the SFT to process arbitrary input prefixes and suffixes while preserving the matched substring for replacement
-\end{enumerate}
+The composition process involves connecting the two SFTs by adding $\varepsilon$-transitions (labeled with "\texttt{None}, \texttt{f}") from each accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT.
 
@@
      (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q0') edge [loop below] node {$\Sigma$, $\texttt{id}$} (q0');
+     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1');
 
@@
     (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q3'') edge [loop left] node {$\Sigma$, $\texttt{id}$} (q3'')
     (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
@@
 
+  We can extend the above construction to model both singleton and replace-all operations.
+In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}.
+The SFT $\mathcal{T}_2$ combines pattern matching and replacement using the construction just described.
+The SFT $\mathcal{T}_1$ matches strings that do not contain any substrings matching the pattern.
+Two $\varepsilon$-transitions connect these components: one from $\mathcal{T}_1$'s accepting state to $\mathcal{T}_2$'s initial state, and another from $\mathcal{T}_2$'s accepting state back to $\mathcal{T}_1$'s initial state.
+This cyclic structure enables the replace-all operation by repeatedly applying pattern replacement until no more matches exist in the input string.
+
   Having constructed the SFT, we can now compute the forward image of the replacement operation.
@@
   Our string solver uses this forward image to do forward propagation for further string solving.
-
-  The replacement operation model described above replaces a single occurrence of any substring matching the regular expression pattern. To support different replacement semantics, 
-  we model replace-all operations (which substitute all occurrences of matching substrings).
-
-  The approach to modeling replace-all operations is illustrated in Figure \ref{fig:replace-all}.
 
@@
 
-CertiStrR implements two different pattern matching introduced above: (1) singleton matching at any position, (2) replace-all matching.
+CertiStrR implements two different pattern matching introduced above: (1) singleton matching at first occurrence, (2) replace-all matching.
 The pattern can be a regular expression or a constant string, which is a special case of regular expression.
@@
 We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are categorized into three groups based on their pattern matching semantics: (1) general singleton matching and (2) replace-all matching.
-Patterns can be either regular expressions or constant strings.
+The benchmarks are categorized into two groups: (1) singleton matching and (2) replace-all matching.
 
@@
 
+%\begin{table}[h]
+%  \centering
+%  \small
+%  \begin{tabular}{lcccc}
+%      \toprule
+%      & \textbf{SAT/UNSAT/Inc} & \textbf{Time (s)} & \textbf{Tests} %\\
+%      \midrule
+%      \texttt{replace singleton} & 142/173/8 & 0.27 & 323\\
+%      \texttt{replace all} & 84/4/10 & 0.36 & 98\\
+%      \bottomrule
+%  \end{tabular}
+%  \caption{Experimental results}
+%  \label{tab:string_operations}
+%\end{table}
+
 \begin{table}[h]
@@
 \small
-  \begin{tabular}{lcccc}
+\setlength{\tabcolsep}{5pt}
+\renewcommand{\arraystretch}{1.1}
+\begin{tabular}{lcccccc}
   \toprule
-      & \textbf{SAT/UNSAT/Inc} & \textbf{Time (s)} & \textbf{Tests} \\
+  \textbf{Subgroup} & \textbf{Tests} & \textbf{Time (s)} & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inc} \\
   \midrule
-      \texttt{replace singleton} & 142/173/8 & 0.27 & 323\\
-      \texttt{replace all} & 84/4/10 & 0.36 & 98\\
+  \multicolumn{6}{c}{\textbf{Replace Singleton}} \\
+  Replace RE  & 94  & 0.79 & 77  & 16  & 1 \\
+  Replace Str & 323 & 1.94 & 103 & 211 & 9 \\
+  \midrule
+  \multicolumn{6}{c}{\textbf{Replace All}} \\
+  PCP       & 1000 & 0.14 & 751 & 249 & 0 \\
+  RNA SAT   & 500  & 2.81 & 500 & 0   & 0 \\
+  RNA UNSAT & 500  & 3.14 & 0   & 500 & 0 \\
+  Rest      & 255  & 0.79 & 40  & 215 & 0 \\
   \bottomrule
@@
 The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a one-minute time limit per test.
-The results show average execution times of 0.27 seconds for \texttt{replace singleton} and 0.36 seconds for \texttt{replace all}.
+The results show average execution times of 0.97 seconds for "\texttt{Replace RE}" (patterns are regular expressions) and 1.94 seconds for "\texttt{Replace Str}" (patterns are strings).
 Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
 
-An "Inconclusive" result indicates that the solver cannot determine satisfiability, not due to timeout (all tests completed within one minute), but due to the inherent incompleteness of the string solver's algorithm.
-An inconclusive result can occur when the solver cannot decide whether a string variable can be assigned a unique value to make the constraints satisfiable.
+An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one minute).
+This occurs when the solver cannot determine whether string variables can be assigned values that satisfy all constraints.
 
-Our performance analysis revealed that while the SFT-based replacement operation modeling is efficient, the primary computational bottleneck stems from automata accumulation during  forward-propagation.
-Consider the following string constraints:
-\begin{align}
-x &= x_1\texttt{++}x_2;~x = \texttt{replace}(x_3, p, r); \nonumber ~x = \texttt{replace\_re}(x_5, p_1, r_1); \nonumber
-\end{align}
-where \texttt{++} denotes string concatenation.
-The variable $x$ appears multiple times on the left-hand side of the equations, causing the forward-propagation algorithm to accumulate automata representations for all constraints: $x_1\texttt{++}x_2$, \texttt{replace}\newline($x_3, p, r$), $x_4$, and \texttt{replace\_re}($x_5, p_1, r_1$).
-Each accumulation step requires computing the product of the current automaton with the previous result.
-Given that the product operation has a worst-case complexity of $O(n^2)$, where $n$ represents the automaton size, this repeated accumulation can lead to state explosion.
+The most significant results appear in the "\texttt{Replace All}" subgroup, where CertiStrR efficiently handles diverse replace-all operations: 0.14 seconds for PCP (Post Correspondence Problem) tests, 2.81 seconds for RNA SAT, and 3.14 seconds for RNA UNSAT.
+The "\texttt{Rest}" category, encompassing various other string constraints, averages 0.79 seconds.
+
+Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string solvers, including CVC4 \cite{cvc4} and Z3 \cite{Z3-str3}, yet CertiStrR solves them efficiently.
+Among existing solvers, only Ostrich \cite{pacmpl/ChenFHHHKLRW22} can handle these benchmarks.
+
+\paragraph{Transducers vs. Replacement Operations}.
+%TODO
 

### 2025-08-20 CertiStr/test/benchmark/parse_results.py

@@ -1,0 +1,174 @@
+#!/usr/bin/env python3
+"""
+Script to parse test.result files and generate statistics for benchmark categories.
+"""
+
+import os
+import re
+import statistics
+from collections import defaultdict
+
+def parse_test_result_file(filepath):
+    """Parse a test.result file and extract test data."""
+    tests = []
+    current_test = {}
+    
+    with open(filepath, 'r') as f:
+        lines = f.readlines()
+    
+    i = 0
+    while i < len(lines):
+        line = lines[i].strip()
+        
+        # Look for test start pattern
+        if line.startswith('Test (') and line.endswith('):'):
+            # Extract test filename
+            test_match = re.search(r'Test \((.+)\):', line)
+            if test_match:
+                if current_test:  # Save previous test if exists
+                    tests.append(current_test)
+                current_test = {'filename': test_match.group(1)}
+        
+        # Look for timing information
+        elif line.startswith('real'):
+            time_match = re.search(r'real\s+(\d+)m([\d.]+)s', line)
+            if time_match:
+                minutes = int(time_match.group(1))
+                seconds = float(time_match.group(2))
+                total_seconds = minutes * 60 + seconds
+                current_test['time'] = total_seconds
+        
+        # Look for result status
+        elif line in ['SAT', 'UNSAT', 'inconclusive']:
+            current_test['result'] = line
+            
+        i += 1
+    
+    # Don't forget the last test
+    if current_test:
+        tests.append(current_test)
+    
+    return tests
+
+def calculate_folder_stats(folder_path):
+    """Calculate statistics for a benchmark folder."""
+    result_file = os.path.join(folder_path, 'test.result')
+    
+    if not os.path.exists(result_file):
+        print(f"Warning: {result_file} does not exist")
+        return None
+    
+    tests = parse_test_result_file(result_file)
+    
+    if not tests:
+        print(f"Warning: No tests found in {result_file}")
+        return None
+    
+    # Calculate statistics
+    times = [test.get('time', 0) for test in tests if 'time' in test]
+    results = [test.get('result', 'Unknown') for test in tests if 'result' in test]
+    
+    stats = {
[…]

### 2025-08-20 CertiStr/test/benchmark/replace_all/test.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    echo "Test ($test):"
+    { time result=$(dune exec CertiStr -- $test); } 2>&1
+    echo "$result"
+done

### 2025-08-20 CertiStr/test/benchmark/replace_re/test.sh

@@ -2,6 +2,6 @@
 
 for test in ./*.smt2; do
     echo "Test ($test):"
-    { time result=$(dune exec CertiStr -- $test); } 2>&1
+    { time result=$(dune exec CertiStr -- --left-most $test); } 2>&1
     echo "$result"
 done

### 2025-08-21 Paper/application.tex
finish application and evaluation part

@@
 
-\paragraph{Transducers vs. Replacement Operations}.
-%TODO
+\paragraph{Transducers vs. Replacement Operations.}  
+Although both transducers and replacement operations can express string transformations for security-related verification, transducers are strictly more expressive because they can capture dependencies between the input pattern and the replacement string.
+In contrast, all replacement operations in SMT-LIB treat the pattern and the replacement string as independent.
+For example, in the SFT shown in Figure~\ref{fig:html-sanitizer-sft}, a transition may be defined over an interval such as $['a'-'z']$, and the output function computes its output directly based on the specific input character.
+Replacement operations, however, cannot express such dependencies.
+
+Consider the transition $['a'-'z']/id$, where $id$ is the identity function.
+To model this transition with replacement operations, the interval must be explicitly expanded into the individual characters $\texttt{'a'}$ through $\texttt{'z'}$, requiring 26 nested \texttt{replace-all} operations—one for each character.
+Thus, modeling $['a'-'z']/id$ with replacement operations yields an expression of the form:
+\[
+\texttt{replace}(\texttt{replace}(\ldots,~\texttt{"b"},~\texttt{"b"},~g),~\texttt{"a"},~\texttt{"a"},~g).
+\]
+
+When the interval contains a large set of characters—as is common in web applications that support the full Unicode range of $0x10FFFF$ code points—this nested structure becomes impractical, leading to significant inefficiency and complexity in the replacement-based approach.
 
@@
 
-This subsection discusses the development effort required for our certified SFT formalization. Table \ref{tab:abstract_impl} provides an overview of this effort.
-Note that, to model replacement operation, we not only need the formalization of SFTs but also the formalization of negation of an SFA, $\epsilon$NFA.
+This subsection discusses the development effort required for our certified SFT formalization.
+While SFTs constitute the central contribution of this paper, they are not sufficient on their own to support the applications we target.
+A more comprehensive symbolic framework for automata libraries is necessary.
+Beyond the existing automata library supported by CertiStr \cite{cpp/KanLRS22}, which already includes operations such as SFA product and SFA concatenation, we needed to develop additional \emph{certified} SFA operations from scratch.
+These include $\varepsilon$SFAs, translation from $\varepsilon$SFAs to SFAs, translation from SFAs to Symbolic Deterministic Finite Automata (SDFA), and negation of SFAs.
+Table~\ref{tab:abstract_impl} summarizes the certified development effort that was carried out specifically for this work.
 
-All efforts listed in the table are additional to the existing CertiStr framework that we build upon.
+The abstract-level development encompasses all formalizations presented in Section~\ref{sec:abstract-layer}, including SFTs and the additional SFA operations used in this paper's application.
+The implementation-level development covers all formalizations described in Section~\ref{sec_alg_refinement}, including the refinement of the product operation and other SFA operations used in this paper's application.
+The final row of Table~\ref{tab:abstract_impl} corresponds to the effort devoted to interval formalization.
+The most challenging component proved to be the correctness proof of the product operation at the abstract level (Figure~\ref{fig-def-product-correct}).
 
-The abstract-level development encompasses all formalizations presented in Sections \ref{sec:formalization} and \ref{sec:product-operation}, including SFTs and $\varepsilon$SFAs.
-The implementation-level development covers all formalizations in Section \ref{sec_alg_refinement}, including the refinement of the product operation.
-The final row corresponds to the interval formalization effort.
-The most challenging component proved to be the correctness proof of the product operation at the abstract level (Figure \ref{fig-def-product-correct}).
-
-The interval formalization complexity increased dramatically in CertiStrR compared to the interval formalization in CertiStr \cite{cpp/KanLRS22}, primarily due to extending intervals from single pairs to lists.
+Compared to CertiStr \cite{cpp/KanLRS22}, the interval formalization in CertiStrR is substantially more complex, primarily because intervals were extended from single pairs to lists.
 

### 2025-08-26 Paper/application.tex
Added new section
motivation seciton

@@
 \section{Applications and Evaluation}
-
-We evaluate our formalization and implementation of the SFT product operation in two key application domains: string sanitization and string replacement operations in string solving.
-These applications demonstrate the practical utility of our certified SFT product operation in real-world scenarios.
+We evaluate our formalization and implementation of the SFT product operation across two key application domains: string sanitization and string replacement in string solving.
+These case studies highlight the practical value of our certified SFT product operation in real-world scenarios.
 \subsection{Sanitizer}
@@
 \caption{SFT representation of the HTML encoding sanitizer function.}
-\label{fig:html-sanitizer-sft}
+\label{fig:html-sanitizer-sft1}
 \end{figure}
@@
 
-Beyond performance evaluation, our SFT formalization enables formal verification of string sanitizer correctness.
-A santizer is used to transform user inputs into safe outputs, ensuring that the output does not contain any unsafe characters.
-However, santizers just like programs, which are eror-prone.
-Manual reviews of sanitizers are often insufficient to guarantee correctness.
+Beyond performance evaluation, our SFT formalization also supports the formal verification of string sanitizer correctness.
+A sanitizer transforms user inputs into safe outputs, ensuring that the result contains no unsafe characters.
+However, sanitizers—like any other programs—are error-prone.
+Manual reviews or testing alone are often insufficient to guarantee their correctness.
 
@@
 Formally speaking, Let $\mathcal{T}$ be an SFT, $\mathcal{I}$ is the possible user inputs modeled by an SFA, and $\mathcal{A}$ is the attack SFA that accepts unsafe strings.
-The santizer correctness checking against the attack is to check whether the intersection of the product $\mathcal{T} \times \mathcal{I}$ and the attack SFA $\mathcal{A}$ is empty, i.e., $\mathcal{T} \times \mathcal{I} \cap \mathcal{A} = \emptyset$.
+The santizer correctness checking against the attack is to check whether the intersection of the product $\mathcal{T} \times \mathcal{I}$ and the attack SFA $\mathcal{A}$ is empty, i.e., $(\mathcal{T} \times \mathcal{I}) \cap \mathcal{A} = \emptyset$.
 
@@
 Let $\mathcal{T}_1$ and $\mathcal{T}_2$ correspond to two sanitizers, and $\mathcal{T}_1$ is applied before $\mathcal{T}_2$.
-What we need to check is $\mathcal{T}_2 \times (\mathcal{T}_1 \times \mathcal{I}) \cap \mathcal{A} = \emptyset$.
+What we need to check is $(\mathcal{T}_2 \times (\mathcal{T}_1 \times \mathcal{I})) \cap \mathcal{A} = \emptyset$.
 
@@
 The first column lists the attack models, the second column indicates whether the sanitizer is classified as safe or unsafe with respect to the attack, and the third column reports the size of the transducers used in the verification, given as [number of states/number of transitions].
-For composed SFTs, the size reflects the total number of states and transitions across all components.
+For composed SFTs, the size reflects the total number of states and transitions across all SFTs.
 
@@
 \begin{enumerate}
-  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function $f = \lambda x.~\texttt{None}$ that produces the empty string.
+  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$f = \lambda x.~\texttt{None}$" that produces the empty string.
   \item Construct an SFA accepting the replacement string "\texttt{NUM}".
@@
   \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
-  \label{fig-rearranged-automata}
+  \label{fig-rearranged-automata1}
   \end{figure}
@@
 \caption{The modeling of replace all operation.}
-\label{fig:replace-all}
+\label{fig:replace-all1}
 \end{figure}
@@
 
-We evaluate CertiStrR using benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
+We evaluate CertiStrR using benchmarks from SMT-LIB 2025 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
 The benchmarks are categorized into two groups: (1) singleton matching and (2) replace-all matching.
@@
 An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one minute).
-This occurs when the solver cannot determine whether string variables can be assigned values that satisfy all constraints.
+This occurs when the solver cannot determine whether string variables need to be assigned different values that satisfy all constraints.
 
@@
 
-Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string solvers, including CVC4 \cite{cvc4} and Z3 \cite{Z3-str3}, yet CertiStrR solves them efficiently.
-Among existing solvers, only Ostrich \cite{pacmpl/ChenFHHHKLRW22} can handle these benchmarks.
+Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string
+solvers, including CVC4 \cite{cvc4} and Z3 \cite{Z3-str3}, yet CertiStrR solves
+them efficiently. Among existing solvers, only OSTRICH \cite{pacmpl/ChenFHHHKLRW22} can handle these benchmarks.
 

### 2025-08-26 Paper/formalization.tex
finish application and evaluation part
motivation seciton

@@
 \section{The Isabelle/HOL Formalization of SFTs}
+\label{sec:formalization}
 
-We present the basic architecture of our SFT formalization in Isabelle/HOL, which consists of three distinct layers.
+In this section, we present our formalization of SFTs. We begin by outlining the basic architecture of our Isabelle/HOL formalization, which is organized into three distinct layers.
 
@@
 
-The \textbf{implementation layer} refines the abstract transition labels to a locale where a boolean algebra is defined as type variable \texttt{'b} together with a collection of operations and assumptions.
+The \textbf{implementation layer} refines the abstract transition labels to a locale where a boolean algebra is defined as type variable $'\texttt{b}$ together with a collection of operations and assumptions.
 State sets are refined using Refine\_Monadic \cite{Refine_Monadic-AFP}, a framework in Isabelle/HOL for data refinement that automatically transforms general sets into efficient data structures such as hashmaps and red-black trees.

### 2025-08-27 Paper/abstract.tex
New title and abstract

@@
 \begin{abstract}
-Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by enabling the transformation of input strings into output strings, thereby providing a more expressive framework for operations that encompass both recognition and transformation.
-FTs have wide applications, including program analysis, string constraint solving, and security-critical sanitizer analysis.
-However, there is currently no scalable formalization of FTs.
+    Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by
+    enabling the transformation of input strings into output strings. In many
+    practical applications --- including program analysis, string constraint
+    solving, and analysis of security-critical sanitizers --- Symbolic FTs 
+    (SFTs) and Symbolic FAs (SFAs) are used instead of the explicitly
+    represented models. To circumvent to the notorious state-space explosion 
+    problem caused by the extremely large alphabet size (e.g. UTF-16), SFTs
+    and SFAs allow the representation of the alphabet as an effective boolean
+    algebra including finite unions of intervals, as well as 
+    SMT-Algebras.
+    %can be instantiated with various Boolean algebras, 
+    The
+    security-critical nature of many of these applications demands trustworthy
+    implementations of such systems. To this end, we present the first 
+    formalization of SFTs and their most important algorithms in
+    Isabelle/HOL. 
+    %This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
+%
+    To evaluate the effectiveness of our formalization, we apply the formalized 
+    SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for 
+    preventing XSS attacks, and (2) string solving, which increasingly
+    employs intricate string replacement operations.
+    %a string solver that models replacement operations with various semantics, including left-most and replace-all matching. 
+    Our experimental results demonstrate that our methods are competitive with
+    the existing unverified implementations.
+
+    %that our approach achieves computational efficiency in constraint-solving scenarios.
+
+    %, thereby providing a more expressive framework for operations that encompass both recognition and transformation. 
+%FTs have wide applications, including program analysis, string constraint 
+    %solving, and security-critical sanitizer analysis. 
+%However, there is currently no scalable formalization of FTs.
 This is primarily because transition labels in FTs are not formalized symbolically, resulting in transition explosion and significant performance bottlenecks in practical applications.
 
+    \OMIT{
 In this paper, we present a formalization of FTs in the symbolic setting using Isabelle/HOL, where transition labels can be instantiated with various Boolean algebras, such as intervals and arithmetic predicates.
@@
 Experimental results demonstrate that our approach achieves computational efficiency in constraint-solving scenarios.
+    }
 \end{abstract}

### 2025-08-30 Paper/introduction.tex
Continuing with intro

@@
 
-Finite Automata (FA) and Finite Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming languages and software engineering.
-For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
+Finite Automata (FA) and Finite Transducers (FT) are fundamental constructs in the theory of formal languages, with extensive applications in programming 
+languages and software engineering. Despite this, many practical applications
+that employ FAs and FTs --- for example, string analysis for web applications
+%character sanitization in web
+%applications 
+(which easily prone to XSS vulnerabilities) and string constraint solving
+--- require extremely large alphabets (e.g. UTF-16), which are beyond the
+capabilities of classical automata algorithms
+\cite{DV21,cav/DAntoniV17,uss/HooimeijerLMSV11,Berkeley-JavaScript}.
+More precisely, explicit automata/transducer representations would yield 
+an extremely large number of transitions (potentially, one for 
+each letter).
+%For example, recent advancements in string solvers, as demonstrated by \cite{pacmpl/ChenFHHHKLRW22}, have illustrated the relationship between regular expressions in modern programming languages and various forms of FAs and FTs.
 Additionally, FAs and FTs find significant industrial applications, such as in the verification of AWS access control policies \cite{DBLP:conf/fmcad/BackesBCDGLRTV18}.
 
-Even though there are various formalizations of FAs and FTs in interactive proof assistants such as Isabelle \cite{isabelle-homepage} and Coq \cite{coq-homepage}, these are predominantly based on classical definitions.
-However, these traditional approaches present certain limitations when applied to practical scenarios.
-One significant drawback is that transition labels are typically non-symbolic and finite.
+%One significant drawback from classical automata approach is that transition 
+%labels are typically non-symbolic and finite.
 A conventional transition is represented as $q\xrightarrow{a}q'$, where $a$ is a character from a finite alphabet.
 This simplistic representation can lead to a phenomenon known as transition explosion.
-For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
+%For example, if the alphabet $\Sigma$ encompasses the entire Unicode range, which is common in modern programming languages, it consists of $\texttt{0x110000}$ distinct characters.
 Defining a transition from state $q$ to $q'$ that accepts any character in $\Sigma$ would require splitting into $\texttt{0x110000}$ individual transitions, rendering the FA and FT operations highly inefficient.
@@
 
-Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT) \cite{cav/DAntoniV17, VeanesHLMB12Transducer} represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
+%Even though there are various formalizations of FAs and FTs in interactive proof assistants such as Isabelle \cite{isabelle-homepage} and Coq \cite{coq-homepage}, these are predominantly based on classical definitions.
+However, these traditional approaches present certain limitations when applied to practical scenarios.
+
+Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT)
+\cite{cav/DAntoniV17, VeanesHLMB12Transducer,DV21} were proposed around 15 years ago
+to deal with the alphabet-explosion problem that often arises in practical
+applications. 
+%represent advanced extensions of classic FAs and  FTs, enhancing their applicability in practical scenarios.
 %
-These symbolic models utilize transition labels defined by first-order predicates over boolean algebras, allowing for more expressive representations.
-For example, a transition label might be specified as an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from $'\texttt{a}'$ to $'\texttt{z}'$, or as an arithmetic condition ($x \mod 2 = 0$), denoting all even numbers.
-This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
+These symbolic extensions permit transition labels defined by an effective 
+boolean algebra, allowing for a succinct representation of a set of transitions
+that can be conveniently manipulated. 
+%more expressive representations. 
+For example, in the interval algebra, a transition label might be specified as 
+an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from
+$'\texttt{a}'$ to $'\texttt{z}'$. As another example, in the SMT-Algebra
+(particularly, of the theory of Linear Integer Arithmetic), %or 
+one could associate an arithmetic condition ($x \mod 2 = 0$) to a transition, 
+denoting all even numbers. Efficient algorithms for manipulating (e.g. computing
+closures) and analyzing (e.g. checking emptiness) SFA and SFT are by now
+rather mature (e.g see the CACM article \cite{DV21}).
+%This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
 

### 2025-08-30 Paper/introduction.tex
More on intro

@@
 
+Owing to the security-critical nature of many applications of string analysis
+(e.g. analysis of web applications against XSS vulnerabilities), there is an 
+increasing need of trustworthy implementations in such application domains. 
+Unfortunately, string analysis implementations are often quite buggy. For
+example, even state-of-the-art SMT solvers over the theory of strings --- 
+which were used at AWS, esp. for analyzing Role-Based Access Control (RBAC) 
+policies --- were found to be buggy through such techniques as fuzzing
+\cite{DBLP:conf/cav/BlotskyMBZKG18,BM20,Mansur20}. This has motivated
+initial formalization efforts for the theory of strings in Isabelle/HOL, e.g., a verified solver
+\cite{cpp/KanLRS22}
+for handling a small fragment of SMT-LIB 2.6 unicode theory of strings
+consisting only of equality, concatenation and regular constraints, and a
+formalization of the semantics of this theory \cite{verified-verifying}.
+
+Despite these, existing verified implementations for string analysis do not
+handle the following indispensable string operations for analysis of web
+applications (e.g. see
+\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}):
+\emph{string transductions} and \emph{replaceall} operators. In particular, 
+string transductions are used frequently in web applications (e.g. for
+sanitization, as well as implicitly applied transductions, for instance
+through \texttt{innerHTML}). In particular, SMT-LIB string constraints that 
+arise from such applications require the \emph{replaceall} operator (part of
+SMT-LIB), which is currently not handled by any existing verified
+implementation. This paper tackles the formalization of SFA/SFT and
+their associated algorithms, as well as applications to analysis of sanitizers
+and verified solver for string constraints. \textcolor{red}{Updated up to here
+so far}
+
+\OMIT{
+SFA and SFT (e.g. 
+analysis of web applications, e.g., against XSS vulnerabilities),
+
 However, the formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants.
 Two fundamental considerations arise: first, the representation of transition labels must be sufficiently expressive to accommodate diverse predicate representations of boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
+
+}
 

### 2025-08-31 Paper/introduction.tex
More or less done with intro

@@
 rather mature (e.g see the CACM article \cite{DV21}).
+SFA and SFT have been applied in numerous domains including analysis of web
+applications against XSS vulnerabilities \cite{VeanesHLMB12Transducer,uss/HooimeijerLMSV11},
+runtime behavior monitoring \cite{osdi/YaseenABCL20}, automated program
+transformation \cite{pldi/HuD17}, and finally SMT solving for the unicode theory
+of strings (e.g. see \cite{pacmpl/ChenFHHHKLRW22,CHL+19}).
+%Furthermore, numerous studies have demonstrated the broad applicability of SFTs across diverse domains.
+The work of \cite{VeanesHLMB12Transducer} showcases SFTs in security-critical applications, particularly for cross-site scripting (XSS) prevention.
+\cite{uss/HooimeijerLMSV11} extends this security focus by applying SFTs to web application sanitizer analysis.
+In system verification, \cite{osdi/YaseenABCL20} employs SFTs for runtime behavior monitoring, while \cite{pldi/HuD17} demonstrates their utility in automated program transformation through systematic inversion techniques.
 %This symbolic approach not only provides a more succinct representation but also supports infinite alphabets, thereby extending the expressive capabilities of FAs and FTs.
@@
 Unfortunately, string analysis implementations are often quite buggy. For
-example, even state-of-the-art SMT solvers over the theory of strings --- 
-which were used at AWS, esp. for analyzing Role-Based Access Control (RBAC) 
-policies --- were found to be buggy through such techniques as fuzzing
-\cite{DBLP:conf/cav/BlotskyMBZKG18,BM20,Mansur20}. This has motivated
-initial formalization efforts for the theory of strings in Isabelle/HOL, e.g., a verified solver
-\cite{cpp/KanLRS22}
+example, even state-of-the-art SMT solvers over the theory of strings 
+were found to be buggy through such techniques as fuzzing
+\cite{DBLP:conf/cav/BlotskyMBZKG18,BM20,Mansur20}. 
+String solvers are among others employed at AWS for analyzing Role-Based 
+Access Control (RBAC) policies, which are security-critical.
+These have motivated initial formalization efforts for the theory of strings in 
+Isabelle/HOL, e.g., a verified solver \cite{cpp/KanLRS22}
 for handling a small fragment of SMT-LIB 2.6 unicode theory of strings
@@
 their associated algorithms, as well as applications to analysis of sanitizers
-and verified solver for string constraints. \textcolor{red}{Updated up to here
-so far}
+and verified solver for string constraints. 
+%\textcolor{red}{Updated up to here
+%so far}
 
@@
 
+\OMIT{
 Prior work of CertiStr \cite{cpp/KanLRS22} has successfully formalized SFAs within Isabelle/HOL, demonstrating both the efficiency and effectiveness of SFAs in practice.
@@
 In system verification, \cite{osdi/YaseenABCL20} employs SFTs for runtime behavior monitoring, while \cite{pldi/HuD17} demonstrates their utility in automated program transformation through systematic inversion techniques.
+}
 
+\paragraph{Contributions.} 
 In this work, we present a comprehensive formalization of SFTs. 
-To address the extensibility challenges inherent in supporting diverse transition label theories, we adopt a refinement-based approach.
+%However, the formalization of transition labels in SFAs and SFTs presents significant challenges within interactive proof assistants. 
+Two fundamental considerations arise in such a formalization: first, the 
+representation of transition labels must be sufficiently expressive to accommodate diverse predicate representations of boolean algebras; second, the formalization framework must be designed with extensibility in mind to facilitate the incorporation of new boolean algebras while minimizing redundant proof efforts.
+
+To address these challenges inherent, %in supporting diverse transition label theories, 
+we adopt a refinement-based approach.
 At the abstraction level, transition labels are formalized through the fundamental mathematical concept of \emph{sets}.
@@
 
-To evaluate the effectiveness and efficiency of our formalization, we developed a string solver compliant with SMT-LIB language \cite{smtlib}, focusing particularly on replacement operations.
-We evaluated our implementation using a set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}.
+To evaluate the effectiveness and efficiency of our formalization, we 
+provide two applications: (1) analysis of the sanitizer \texttt{HTMLdecode},
+which is used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
+compliant certified string solver handling equality, concatenation, regular
+constraints, and more important the string replaceall operator, which has been
+argued to be indispensable for analysis of XSS 
+\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}).
 The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
+
+\OMIT{
+which increasingly
+extended
+
+the SMT-LIB \cite{smtlib} compliant certified string solver \cite{cpp/KanLRS22}
+with replacement operations. We evaluated our implementation using a set of benchmarks from the SMT-LIB repository \cite{smtlib_benchmarks}. 
+}
 
@@
 \item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
-\item We demonstrate the practical utility of our formalization through its application to string constraint solving, specifically in modeling replacement operations with verified correctness guarantees.
+\item We demonstrate the practical utility of our formalization through its
+    application to sanitizer analysis and a certified implementation of
+    a string constraint solver that handles equality, concatenation, regular
+        constraints, and the string replacement operator.
 \end{enumerate}
 
+\paragraph{Organization.}
 The remainder of this paper is organized as follows:
+Section \ref{sec:motivation} provides several concrete motivating examples.
 Section \ref{sec:formalization} introduces the formalization of symbolic finite transducers.

### 2025-09-01 CertiStr/test/benchmark/pcp-3-4-hard/run_ostrich_benchmark.sh
experimental results

@@ -1,0 +1,60 @@
+#!/bin/bash
+
+# Script to run ostrich on all .smt2 files and collect timing data
+# Usage: ./run_ostrich_benchmark.sh
+
+echo "Running ostrich benchmark on all .smt2 files..."
+echo "File,Time(seconds),Result" > ostrich_results.csv
+
+total_time=0
+count=0
+successful_runs=0
+
+for file in unsolved_pcp_instance_*.smt2; do
+    if [ -f "$file" ]; then
+        echo "Processing: $file"
+        
+        # Measure execution time using time command
+        start_time=$(date +%s.%N)
+        
+        # Run ostrich and capture both stdout and return code
+        if timeout 60 ostrich "$file" > /dev/null 2>&1; then
+            end_time=$(date +%s.%N)
+            execution_time=$(echo "$end_time - $start_time" | bc)
+            result="completed"
+            total_time=$(echo "$total_time + $execution_time" | bc)
+            successful_runs=$((successful_runs + 1))
+        else
+            end_time=$(date +%s.%N)
+            execution_time=$(echo "$end_time - $start_time" | bc)
+            result="timeout/error"
+        fi
+        
+        # Log to CSV
+        echo "$file,$execution_time,$result" >> ostrich_results.csv
+        
+        count=$((count + 1))
+        
+        # Progress indicator
+        if [ $((count % 50)) -eq 0 ]; then
+            echo "Processed $count files..."
+        fi
+    fi
+done
+
+# Calculate and display results
+if [ $successful_runs -gt 0 ]; then
+    average_time=$(echo "scale=6; $total_time / $successful_runs" | bc)
+else
+    average_time=0
+fi
+
+echo ""
+echo "=== Benchmark Results ==="
+echo "Total files processed: $count"
+echo "Successful runs: $successful_runs"
+echo "Failed/timeout runs: $((count - successful_runs))"
+echo "Total execution time (successful): ${total_time}s"
+echo "Average execution time: ${average_time}s"
+echo ""
+echo "Detailed results saved to: ostrich_results.csv"

### 2025-09-01 CertiStr/test/benchmark/pcp-3-4-hard/run_ostrich_corrected.sh

@@ -1,0 +1,69 @@
+#!/bin/bash
+
+# Corrected script to run ostrich on all .smt2 files and collect timing data
+# Usage: ./run_ostrich_corrected.sh
+
+echo "Running ostrich benchmark on all .smt2 files..."
+echo "File,Time(seconds),Result,Output" > ostrich_corrected_results.csv
+
+total_time=0
+count=0
+successful_runs=0
+
+for file in unsolved_pcp_instance_*.smt2; do
+    if [ -f "$file" ]; then
+        echo "Processing: $file"
+        
+        # Measure execution time using time command
+        start_time=$(date +%s.%N)
+        
+        # Run ostrich and capture both stdout and return code
+        output=$(timeout 120 ostrich "$file" 2>&1)
+        exit_code=$?
+        
+        end_time=$(date +%s.%N)
+        execution_time=$(echo "$end_time - $start_time" | bc)
+        
+        if [ $exit_code -eq 0 ]; then
+            result="completed"
+            # Extract result from output (sat/unsat/unknown)
+            solver_result=$(echo "$output" | head -1)
+            total_time=$(echo "$total_time + $execution_time" | bc)
+            successful_runs=$((successful_runs + 1))
+        elif [ $exit_code -eq 124 ]; then
+            result="timeout"
+            solver_result="timeout"
+        else
+            result="error"
+            solver_result="error"
+        fi
+        
+        # Log to CSV (escape commas in output)
+        clean_output=$(echo "$solver_result" | tr ',' ';')
+        echo "$file,$execution_time,$result,$clean_output" >> ostrich_corrected_results.csv
+        
+        count=$((count + 1))
+        
+        # Progress indicator
+        if [ $((count % 50)) -eq 0 ]; then
+            echo "Processed $count files..."
+        fi
+    fi
+done
+
+# Calculate and display results
+if [ $successful_runs -gt 0 ]; then
+    average_time=$(echo "scale=6; $total_time / $successful_runs" | bc)
+else
+    average_time=0
+fi
+
+echo ""
+echo "=== Corrected Benchmark Results ==="
+echo "Total files processed: $count"
+echo "Successful runs: $successful_runs"
+echo "Failed/timeout runs: $((count - successful_runs))"
+echo "Total execution time (successful): ${total_time}s"
+echo "Average execution time: ${average_time}s"
+echo ""
+echo "Detailed results saved to: ostrich_corrected_results.csv"

### 2025-09-01 CertiStr/test/benchmark/replace_all/test167/test.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    echo "Test ($test):"
+    { time result=$(gtimeout 1800 ostrich -portfolio=strings $test); } 2>&1 #dune exec CertiStr -- --left-most
+    if [ $? -eq 124 ]; then
+        echo "TIMEOUT (1 hour limit exceeded)"
+    else
+        echo "$result"
+    fi
+done

### 2025-09-01 CertiStr/test/benchmark/replace_re/analyze_test_results.py

@@ -1,0 +1,99 @@
+#!/usr/bin/env python3
+"""
+Analyze test results from test2.result file.
+Count timeouts and calculate average execution time for non-timeout tests.
+"""
+
+import re
+import sys
+import subprocess
+
+def parse_test_results(file_path):
+    """
+    Parse test results file and extract timing information.
+    Handles binary files by using strings command first.
+    
+    Returns:
+        tuple: (timeout_count, non_timeout_times, total_tests)
+    """
+    timeout_count = 0
+    non_timeout_times = []
+    total_tests = 0
+    
+    # Use strings command to extract readable text from potentially binary file
+    try:
+        result = subprocess.run(['strings', file_path], capture_output=True, text=True)
+        if result.returncode != 0:
+            # Fall back to regular file reading
+            with open(file_path, 'r', encoding='utf-8', errors='ignore') as f:
+                content = f.read()
+        else:
+            content = result.stdout
+    except:
+        # Final fallback to regular file reading
+        with open(file_path, 'r', encoding='utf-8', errors='ignore') as f:
+            content = f.read()
+    
+    # Split by test cases - look for "Test (" pattern
+    test_blocks = re.split(r'Test \([^)]+\):', content)[1:]  # Skip first empty element
+    
+    for block in test_blocks:
+        total_tests += 1
+        
+        # Check if this test timed out
+        if 'TIMEOUT' in block:
+            timeout_count += 1
+        else:
+            # Extract real time - look for "real\t<time>" pattern
+            time_match = re.search(r'real\s+(\d+)m([\d.]+)s', block)
+            if time_match:
+                minutes = int(time_match.group(1))
+                seconds = float(time_match.group(2))
+                total_seconds = minutes * 60 + seconds
+                non_timeout_times.append(total_seconds)
+    
+    return timeout_count, non_timeout_times, total_tests
+
+def main():
+    file_path = 'test2.result'
+    
+    try:
+        timeout_count, non_timeout_times, total_tests = parse_test_results(file_path)
+        
+        print("=== Test Results Analysis ===")
+        print(f"Total tests: {total_tests}")
+        print(f"Timeout tests: {timeout_count}")
+        print(f"Non-timeout tests: {len(non_timeout_times)}")
+        
+        if non_timeout_times:
+            avg_time = sum(non_timeout_times) / len(non_timeout_times)
+            min_time = min(non_timeout_times)
+            max_time = max(non_timeout_times)
+            
[…]

### 2025-09-01 CertiStr/test/benchmark/replace_re/test89/test.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+for test in ./*.smt2; do
+    echo "Test ($test):"
+    { time result=$(timeout 600 cvc5 $test); } 2>&1 #dune exec CertiStr -- --left-most
+    if [ $? -eq 124 ]; then
+        echo "TIMEOUT (1 hour limit exceeded)"
+    else
+        echo "$result"
+    fi
+done

### 2025-09-01 CertiStr/test/benchmark/replace_str/analyze_results.py

@@ -1,0 +1,116 @@
+#!/usr/bin/env python3
+
+import re
+import sys
+
+def parse_time(time_str):
+    """Parse time string like '0m0.008s' or '1m22.497s' to seconds"""
+    match = re.match(r'(\d+)m(\d+\.\d+)s', time_str)
+    if match:
+        minutes = int(match.group(1))
+        seconds = float(match.group(2))
+        return minutes * 60 + seconds
+    return None
+
+def analyze_results(filename):
+    """Analyze test results from the given file"""
+    timeout_count = 0
+    total_tests = 0
+    execution_times = []
+    timeouts = []
+    
+    with open(filename, 'r') as f:
+        content = f.read()
+    
+    # Split by test entries
+    test_blocks = re.split(r'Test \([^)]+\):', content)[1:]  # Skip first empty split
+    
+    for block in test_blocks:
+        total_tests += 1
+        lines = block.strip().split('\n')
+        
+        # Check for timeout
+        timeout_found = False
+        real_time = None
+        
+        for line in lines:
+            if 'TIMEOUT' in line:
+                timeout_found = True
+                timeouts.append(line.strip())
+                break
+            elif line.startswith('real'):
+                # Extract time from real time line
+                time_match = re.search(r'real\s+(.+)', line)
+                if time_match:
+                    time_str = time_match.group(1)
+                    real_time = parse_time(time_str)
+        
+        if timeout_found:
+            timeout_count += 1
+        elif real_time is not None:
+            execution_times.append(real_time)
+    
+    # Calculate statistics
+    if execution_times:
+        avg_time = sum(execution_times) / len(execution_times)
+        min_time = min(execution_times)
+        max_time = max(execution_times)
+        median_time = sorted(execution_times)[len(execution_times) // 2]
+    else:
+        avg_time = min_time = max_time = median_time = 0
+    
+    # Print results
+    print("=" * 50)
+    print("EXPERIMENTAL RESULTS ANALYSIS")
+    print("=" * 50)
+    print(f"Total tests: {total_tests}")
+    print(f"Timeouts: {timeout_count}")
+    print(f"Successful tests: {len(execution_times)}")
+    print()
+    
+    if timeouts:
+        print("Timeout details:")
+        for i, timeout in enumerate(timeouts, 1):
+            print(f"  {i}. {timeout}")
+        print()
+    
+    if execution_times:
[…]

### 2025-09-01 CertiStr/test/benchmark/replace_str/analyze_test3_results.py

@@ -1,0 +1,96 @@
+#!/usr/bin/env python3
+"""
+Analyze test3.result to count timeouts and compute average execution time.
+
+This script:
+1. Counts the number of timeout tests
+2. Excludes timeout tests and computes average execution time for the rest
+"""
+
+import re
+import sys
+
+def analyze_test_results(filename):
+    """
+    Analyze test results from the given file.
+    
+    Returns:
+        tuple: (timeout_count, average_time, total_tests, valid_tests)
+    """
+    timeout_count = 0
+    execution_times = []
+    total_tests = 0
+    
+    try:
+        with open(filename, 'r', encoding='utf-8', errors='ignore') as f:
+            content = f.read()
+    except FileNotFoundError:
+        print(f"Error: File '{filename}' not found.")
+        return None, None, None, None
+    except Exception as e:
+        print(f"Error reading file: {e}")
+        return None, None, None, None
+    
+    # Split content into test blocks
+    # Each test starts with "Test (./filename.smt2):"
+    test_blocks = re.split(r'Test \([^)]+\.smt2\):', content)[1:]  # Skip first empty element
+    
+    total_tests = len(test_blocks)
+    
+    for block in test_blocks:
+        # Check if this test timed out
+        if 'TIMEOUT' in block:
+            timeout_count += 1
+        else:
+            # Look for execution time at the end of the block
+            # The pattern seems to be a standalone number followed by "ms" at the end
+            lines = block.strip().split('\n')
+            for line in reversed(lines):
+                line = line.strip()
+                # Look for standalone time in milliseconds
+                time_match = re.match(r'^(\d+)ms$', line)
+                if time_match:
+                    execution_time = int(time_match.group(1))
+                    execution_times.append(execution_time)
+                    break
+    
+    valid_tests = len(execution_times)
+    
+    if execution_times:
+        average_time = sum(execution_times) / len(execution_times)
+    else:
+        average_time = 0
+    
+    return timeout_count, average_time, total_tests, valid_tests
+
+def main():
+    filename = 'test3.result'
+    
+    print("Analyzing test3.result...")
+    print("=" * 50)
+    
+    timeout_count, average_time, total_tests, valid_tests = analyze_test_results(filename)
+    
+    if timeout_count is None:
+        sys.exit(1)
+    
+    print(f"Total tests found: {total_tests}")
+    print(f"Timeout tests: {timeout_count}")
[…]

### 2025-09-01 CertiStr/test/benchmark/replace_str/test.sh

@@ -2,6 +2,10 @@
 
 for test in ./*.smt2; do
     echo "Test ($test):"
-    { time result=$(dune exec CertiStr -- --left-most $test); } 2>&1
-    echo "$result"
+    { time result=$(gtimeout 1800 ostrich -portfolio=strings $test); } 2>&1 #dune exec CertiStr -- --left-most $test
+    if [ $? -eq 124 ]; then
+        echo "TIMEOUT (1 hour limit exceeded)"
+    else
+        echo "$result"
+    fi
 done

### 2025-09-01 CertiStr/test/benchmark/rna-sat/calculate_average_time.py

@@ -1,0 +1,40 @@
+#!/usr/bin/env python3
+import re
+
+def parse_time(time_str):
+    """Parse time string like '0m14.643s' to seconds"""
+    match = re.match(r'(\d+)m([\d.]+)s', time_str)
+    if match:
+        minutes = int(match.group(1))
+        seconds = float(match.group(2))
+        return minutes * 60 + seconds
+    return 0
+
+def calculate_average_real_time(filename):
+    """Calculate average real execution time from test result file"""
+    real_times = []
+    
+    with open(filename, 'r') as f:
+        for line in f:
+            line = line.strip()
+            if line.startswith('real'):
+                # Extract time part after 'real\t'
+                time_part = line.split('\t')[1]
+                seconds = parse_time(time_part)
+                if seconds > 0:
+                    real_times.append(seconds)
+    
+    if not real_times:
+        print("No real time entries found")
+        return
+    
+    average_time = sum(real_times) / len(real_times)
+    
+    print(f"Total test cases: {len(real_times)}")
+    print(f"Average real time: {average_time:.3f} seconds")
+    print(f"Average real time: {int(average_time // 60)}m{average_time % 60:.3f}s")
+    print(f"Min time: {min(real_times):.3f}s")
+    print(f"Max time: {max(real_times):.3f}s")
+
+if __name__ == "__main__":
+    calculate_average_real_time("test1.result")

### 2025-09-01 Paper/application.tex
add applications

@@
 
-\subsection{Modeling the Replacement Operation}
+%\subsection{Modeling the Replacement Operation}
 
-The string replacement, denoted as
-$\texttt{replace}(\texttt{str}, \texttt{pattern}, \newline\texttt{replacement}, \texttt{flag})$, is a fundamental string transformation accepting four parameters:
-\begin{itemize}
-  \item $\texttt{str}$: The input string to be transformed.
-  \item $\texttt{pattern}$: A regular expression defining the matching criteria.
-  \item $\texttt{replacement}$: The string that replaces matched substrings.
-  \item $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single occurrence replacement and $g$ denotes replace-all operation.
-\end{itemize}
+% The string replacement, denoted as
+% $\texttt{replace}(\texttt{str}, \texttt{pattern}, \newline\texttt{replacement}, \texttt{flag})$, is a fundamental string transformation accepting four parameters:
+% \begin{itemize}
+%   \item $\texttt{str}$: The input string to be transformed.
+%   \item $\texttt{pattern}$: A regular expression defining the matching criteria.
+%   \item $\texttt{replacement}$: The string that replaces matched substrings.
+%   \item $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single occurrence replacement and $g$ denotes replace-all operation.
+% \end{itemize}
 
-The operation's semantics is to replace the first occurrence (or all occurrences) of any substring $s'$ in \texttt{str} that matches $\texttt{pattern}$ with \texttt{replacement}.
+% The operation's semantics is to replace the first occurrence (or all occurrences) of any substring $s'$ in \texttt{str} that matches $\texttt{pattern}$ with \texttt{replacement}.
 
-To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}", \texttt{flag})$. We construct the SFT representation through three steps:
+% To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}", \texttt{flag})$. We construct the SFT representation through three steps:
 
-\begin{enumerate}
-  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$f = \lambda x.~\texttt{None}$" that produces the empty string.
-  \item Construct an SFA accepting the replacement string "\texttt{NUM}".
+% \begin{enumerate}
+%   \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$f = \lambda x.~\texttt{None}$" that produces the empty string.
+%   \item Construct an SFA accepting the replacement string "\texttt{NUM}".
 Since constant strings are specialized regular expressions, we build their SFA representation and convert it to an SFT by adding $\varepsilon$-transitions with output functions that emit the replacement characters sequentially.
-  \item Compose the two SFTs by connecting the accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT via $\varepsilon$-transitions.
-\end{enumerate}
+%   \item Compose the two SFTs by connecting the accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT via $\varepsilon$-transitions.
+% \end{enumerate}
 
-\noindent\emph{Step 1.}
-Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
+% \noindent\emph{Step 1.}
+% Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
 The left side shows the SFA that recognizes the regular expression $\texttt{/[0-9]+/}$, while the right side presents its transformation into an SFT.
@@
 
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
-  % Left: SFA for [0-9]+ (vertical)
-  \node[state, initial] (q0)   {$q_0$}; 
-  \node[state, accepting, below=of q0] (q1) {$q_1$};
+% \begin{figure}[h] \centering
+% \begin{tikzpicture}[shorten >=1pt, node distance=1.8cm, auto]
+%   % Left: SFA for [0-9]+ (vertical)
+%   \node[state, initial] (q0)   {$q_0$}; 
+%   \node[state, accepting, below=of q0] (q1) {$q_1$};
 
-   \path[->] 
-   (q0) edge [bend left=20] node[right] {\texttt{[0-9]}} (q1)
-   (q1) edge [loop below] node {\texttt{[0-9]}} (q1);
+%    \path[->] 
+%    (q0) edge [bend left=20] node[right] {\texttt{[0-9]}} (q1)
+%    (q1) edge [loop below] node {\texttt{[0-9]}} (q1);
 
-  % Right: SFT for [0-9]+ (vertical)
-  \node[state, initial, right=4cm of q0] (q0') {$q_0'$}; 
-  \node[state, accepting, below=of q0'] (q1') {$q_1'$};
+%   % Right: SFT for [0-9]+ (vertical)
+%   \node[state, initial, right=4cm of q0] (q0') {$q_0'$}; 
+%   \node[state, accepting, below=of q0'] (q1') {$q_1'$};
 
-   \path[->] 
-   (q0') edge [bend left=20] node[right] {\texttt{[0-9]},$\texttt{f}$} (q1')
-   (q1') edge [loop below] node {\texttt{[0-9]},$\texttt{f}$} (q1');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for \texttt{/[0-9]+/}}
-\label{fig-snfa-pattern}
-\end{figure}
+%    \path[->] 
+%    (q0') edge [bend left=20] node[right] {\texttt{[0-9]},$\texttt{f}$} (q1')
+%    (q1') edge [loop below] node {\texttt{[0-9]},$\texttt{f}$} (q1');
+% \end{tikzpicture}
+% \caption{Corresponding SFA and SFT for \texttt{/[0-9]+/}}
+% \label{fig-snfa-pattern}
+% \end{figure}
 
-\noindent\emph{Step 2.}
-Figure \ref{fig-snfa-replacement} depicts the automata for the replacement string "\texttt{NUM}".
+% \noindent\emph{Step 2.}
+% Figure \ref{fig-snfa-replacement} depicts the automata for the replacement string "\texttt{NUM}".
 The left side shows the SFA that accepts this constant string, while the right side presents its SFT transformation.
 The transformation employs an indexed output function \texttt{g} that emits characters of the replacement string sequentially:
-\[
-\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
-\begin{cases}
-1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
-2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
-3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
-\_ \mapsto \texttt{None}
-\end{cases}
-\]
-This function maps transition indices to their corresponding character outputs, using ASCII codes to represent the string "\texttt{NUM}" character by character.
+% \[
+% \texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
+% \begin{cases}
+% 1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
+% 2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
+% 3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
+% \_ \mapsto \texttt{None}
+% \end{cases}
+% \]
+% This function maps transition indices to their corresponding character outputs, using ASCII codes to represent the string "\texttt{NUM}" character by character.
 
-\begin{figure}[h] \centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-  % SFA for "NUM" (top)
-  \node[state, initial above] (q0)   {$q_0$}; 
-  \node[state] (q1) [right=1cm of q0] {$q_1$}; 
-  \node[state] (q2) [below=1cm of q1] {$q_2$}; 
-  \node[state, accepting] (q3) [below=1cm of q0] {$q_3$};
+% \begin{figure}[h] \centering
+% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+%   % SFA for "NUM" (top)
+%   \node[state, initial above] (q0)   {$q_0$}; 
+%   \node[state] (q1) [right=1cm of q0] {$q_1$}; 
+%   \node[state] (q2) [below=1cm of q1] {$q_2$}; 
+%   \node[state, accepting] (q3) [below=1cm of q0] {$q_3$};
 
-  \path[->] 
-  (q0) edge node {\texttt{N}} (q1)
-  (q1) edge node {\texttt{U}} (q2)
-  (q2) edge node {\texttt{M}} (q3);
+%   \path[->] 
+%   (q0) edge node {\texttt{N}} (q1)
+%   (q1) edge node {\texttt{U}} (q2)
+%   (q2) edge node {\texttt{M}} (q3);
 
-  % SFT for "NUM" (right)
-  \node[state, initial above] (q0') [right=1.5cm of q1] {$q_0'$}; 
-  \node[state] (q1') [right=of q0'] {$q_1'$}; 
-  \node[state] (q2') [below=1cm of q1'] {$q_2'$}; 
-  \node[state, accepting] (q3') [below=1cm of q0'] {$q_3'$};
+%   % SFT for "NUM" (right)
+%   \node[state, initial above] (q0') [right=1.5cm of q1] {$q_0'$}; 
+%   \node[state] (q1') [right=of q0'] {$q_1'$}; 
+%   \node[state] (q2') [below=1cm of q1'] {$q_2'$}; 
+%   \node[state, accepting] (q3') [below=1cm of q0'] {$q_3'$};
 
-  \path[->] 
-  (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
-  (q1') edge node[left] {\texttt{None}, \texttt{(g 2)}} (q2')
-  (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
-\end{tikzpicture}
-\caption{Corresponding SFA and SFT for "\texttt{NUM}"}
-\label{fig-snfa-replacement}
-\end{figure}
+%   \path[->] 
+%   (q0') edge node {\texttt{None}, \texttt{(g 1)}} (q1')
+%   (q1') edge node[left] {\texttt{None}, \texttt{(g 2)}} (q2')
+%   (q2') edge node[below] {\texttt{None}, \texttt{(g 3)}} (q3');
+% \end{tikzpicture}
+% \caption{Corresponding SFA and SFT for "\texttt{NUM}"}
+% \label{fig-snfa-replacement}
+% \end{figure}
 
-\noindent\emph{Step 3.}
-Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT (Fig.
+% \noindent\emph{Step 3.}
+% Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT (Fig.
 \ref{fig-snfa-pattern}) with the replacement-generating SFT (Fig.
@@
 
-\begin{figure}[h] \centering
-  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
-    % Second automaton from the first figure
-    \node[state, initial] (q0') {$q_0'$}; 
-    \node[state] (q1') [right=of q0'] {$q_1'$};
+% \begin{figure}[h] \centering
+%   \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+%     % Second automaton from the first figure
+%     \node[state, initial] (q0') {$q_0'$}; 
+%     \node[state] (q1') [right=of q0'] {$q_1'$};
 
-     \path[->] 
-     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1');
+%      \path[->] 
+%      (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
+%      (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1');
 
-    % Second automaton from the second figure
-    \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
-    \node[state] (q1'') [right=of q0''] {$q_1''$}; 
-    \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
-    \node[state] (q2'') [right=of q3''] {$q_2''$};
+%     % Second automaton from the second figure
+%     \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
+%     \node[state] (q1'') [right=of q0''] {$q_1''$}; 
+%     \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
+%     \node[state] (q2'') [right=of q3''] {$q_2''$};
 
-    \path[->] 
-    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
-    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
-    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
-  \end{tikzpicture}
-  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
-  \label{fig-rearranged-automata1}
-  \end{figure}
+%     \path[->] 
+%     (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
+%     (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
+%     (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
+%     (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
+%   \end{tikzpicture}
+%   \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
+%   \label{fig-rearranged-automata1}
+%   \end{figure}
 
-  We can extend the above construction to model both singleton and replace-all operations.
+%   We can extend the above construction to model both singleton and replace-all operations.
 In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}.
@@
 
-  Having constructed the SFT, we can now compute the forward image of the replacement operation.
+%   Having constructed the SFT, we can now compute the forward image of the replacement operation.
 Consider the string constraint $s' = \texttt{replace}(s,~\texttt{/[0-9]+/},~\text{"}\texttt{NUM}\text{"})$, where we aim to characterize the possible values of $s'$.
@@
 The forward image of this replacement operation is given by the product $\mathcal{T} \times \mathcal{A}$, which precisely captures the set of all possible output strings that can be produced by applying the replacement operation to any input string accepted by $\mathcal{A}$.
-  Our string solver uses this forward image to do forward propagation for further string solving.
+%   Our string solver uses this forward image to do forward propagation for further string solving.
 
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
-  % First SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
-  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
+% \begin{figure}[htbp]
+% \centering
+% \begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
+%   % First SFT rectangle
+%   \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
+%   \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
 
-  % States inside first SFT
-  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
-  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
-  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
+%   % States inside first SFT
+%   \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
+%   \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
+%   \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
 
-  % Second SFT rectangle
-  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
-  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
+%   % Second SFT rectangle
+%   \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
+%   \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
 
-  % States inside second SFT
-  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
-  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
-  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
+%   % States inside second SFT
+%   \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
+%   \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
+%   \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
 
-  % Epsilon transition from accepting state of first SFT to initial state of second SFT
-  \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
+%   % Epsilon transition from accepting state of first SFT to initial state of second SFT
+%   \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
 
-  % Epsilon transition from accepting state of second SFT to initial state of first SFT
-  \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
+%   % Epsilon transition from accepting state of second SFT to initial state of first SFT
+%   \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
 
-\end{tikzpicture}
-\caption{The modeling of replace all operation.}
-\label{fig:replace-all1}
-\end{figure}
+% \end{tikzpicture}
+% \caption{The modeling of replace all operation.}
+% \label{fig:replace-all1}
+% \end{figure}
 
@@
 
-We have implemented CertiStrR, an extension of  CertiStr \cite{cpp/KanLRS22}, to support string replacement operations.
-%\footnote{The implementation is available at \url{https://github.com/ShlKan/certified-str-solver}}. 
-CertiStr \cite{cpp/KanLRS22} is a certified string solver that only support some basic string operations, such as concatenation, and regular constraints.
+We evaluate our transducer formalization and implementation by extending CertiStr \cite{cpp/KanLRS22} to support string replacement operations, which is a certified string solver that originally supports only basic string operations, such as concatenation and regular constraints.
+By integrating our SFT-based modeling of replacement operations, we develop CertiStrR, an extended string solver that supports two types of string replacement: (1) replacing the first occurrence of a substring matching a given pattern, and (2) replacing all occurrences of substrings matching a given pattern.
+The patterns can be either constant strings or regular expressions.
 
 %While the core solving algorithm maintains the certification guarantees of CertiStrR, the frontend components are implemented using established non-certified  OCaml libraries: dolmen \cite{dolmen} for SMT-LIB parsing and ocaml-re-nfa \cite{ocaml-re-nfa} for regular expression to NFA conversion.
-
-CertiStrR implements two different pattern matching introduced above: (1) singleton matching at first occurrence, (2) replace-all matching.
-The pattern can be a regular expression or a constant string, which is a special case of regular expression.
 
@@
 
-We evaluate CertiStrR using benchmarks from SMT-LIB 2025 \cite{smtlib_benchmarks}, focusing on the QF\_S and QF\_SLIA logic fragments.
-The benchmarks are categorized into two groups: (1) singleton matching and (2) replace-all matching.
-
-Due to CertiStrR's current frontend limitations in supporting the complete SMT-LIB language specification, we preprocess certain operations.
-For example, conjunctive assertions of the form \texttt{(assert (and c1 c2))} are decomposed into separate assertions \texttt{(assert c1)} and \texttt{(assert c2)} when both \texttt{c1} and \texttt{c2} are string constraints supported by CertiStrR.
+We evaluate {CertiStrR} using benchmarks from SMT-LIB 2025~\cite{smtlib_benchmarks}, focusing on the \textsf{QF\_S} and \textsf{QF\_SLIA} logic fragments.
+The benchmarks are divided into two categories: (1) singleton matching and (2) replace-all matching.
+In addition to replacement, they also include string concatenation and regular constraints.
 
@@
 
-\begin{table}[h]
+\begin{table*}[h]
 \centering
 \small
-\setlength{\tabcolsep}{5pt}
+\setlength{\tabcolsep}{4pt}
 \renewcommand{\arraystretch}{1.1}
-\begin{tabular}{lcccccc}
+\begin{tabular}{llccccccc}
   \toprule
-  \textbf{Subgroup} & \textbf{Tests} & \textbf{Time (s)} & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inc} \\
+  \textbf{Subgroup} & \textbf{Test} & \textbf{Count} & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inconclusive} & \textbf{CertiStrR (s)} & \textbf{Ostrich (s/timeout)} & \textbf{CVC5 (s/timeout)} \\
   \midrule
-  \multicolumn{6}{c}{\textbf{Replace Singleton}} \\
-  Replace RE  & 94  & 0.79 & 77  & 16  & 1 \\
-  Replace Str & 323 & 1.94 & 103 & 211 & 9 \\
-  \midrule
-  \multicolumn{6}{c}{\textbf{Replace All}} \\
-  PCP       & 1000 & 0.14 & 751 & 249 & 0 \\
-  RNA SAT   & 500  & 2.81 & 500 & 0   & 0 \\
-  RNA UNSAT & 500  & 3.14 & 0   & 500 & 0 \\
-  Rest      & 255  & 0.79 & 40  & 215 & 0 \\
+  \multirow{2}{*}{Replace Singleton} & Replace RE  & 94  & 77  & 16  & 1 & 0.79 & 166.27/27 & 2.67/20 \\
+   & Replace Str & 323 & 103 & 211 & 9 & 1.94 & 7.10/20 & 0.81/1 \\
+  \multirow{4}{*}{Replace All} & PCP & 1000 & 751 & 249 & 0 & 0.14 & 3.412/0 & unknown \\
+   & RNA SAT   & 500  & 500 & 0   & 0 & 2.81 & 55.61/0 & unknown \\
+   & RNA UNSAT & 500  & 0   & 500 & 0 & 3.14 & 64.185/0 & unknown \\
+   & Rest      & 255  & 40  & 215 & 0 & 0.79 & 23.67/17 & 8.379/1 \\
   \bottomrule
@@
 \label{tab:string_operations}
-\end{table}
+\end{table*}
 
-The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a one-minute time limit per test.
-The results show average execution times of 0.97 seconds for "\texttt{Replace RE}" (patterns are regular expressions) and 1.94 seconds for "\texttt{Replace Str}" (patterns are strings).
-Test outcomes were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
-
+The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a \textbf{one-hour} time limit per test.
+The results were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
 An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one minute).
 This occurs when the solver cannot determine whether string variables need to be assigned different values that satisfy all constraints.
+
+We also compare with two state-of-the-art string solvers: Ostrich \cite{pacmpl/ChenFHHHKLRW22} and CVC5 \cite{cvc5}. From the experimental results, we observe that string replacement
+
+Table~\ref{tab:string_operations} summarizes the results, including the number of tests, counts of SAT, UNSAT, and Inconclusive results, as well as the average execution time for each subgroup of benchmarks.
 

### 2025-09-02 Paper/application.tex
refactor isabelle code

@@
 The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a \textbf{one-hour} time limit per test.
-The results were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
+Table~\ref{tab:string_operations} summarizes the results.
+The results from CertiStrR were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
 An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one minute).
@@
 
-We also compare with two state-of-the-art string solvers: Ostrich \cite{pacmpl/ChenFHHHKLRW22} and CVC5 \cite{cvc5}. From the experimental results, we observe that string replacement
+We also compare our approach with two state-of-the-art string solvers: Ostrich \cite{pacmpl/ChenFHHHKLRW22} and CVC5 \cite{cvc5}.
+The experimental results demonstrate that string replacement operations, particularly replace-all operations, pose significant challenges for existing state-of-the-art solvers.
+The "unknown" label in the table indicates instances where CVC5 cannot determine satisfiability.
+CertiStrR outperforms both Ostrich and CVC5 across most benchmark subgroups, with the notable exception of the "Replace Str" category where CVC5 achieves better performance (though it experiences one timeout).
+CertiStrR demonstrates particularly strong performance in the "Replace All" category.
 
-Table~\ref{tab:string_operations} summarizes the results, including the number of tests, counts of SAT, UNSAT, and Inconclusive results, as well as the average execution time for each subgroup of benchmarks.
-
-The most significant results appear in the "\texttt{Replace All}" subgroup, where CertiStrR efficiently handles diverse replace-all operations: 0.14 seconds for PCP (Post Correspondence Problem) tests, 2.81 seconds for RNA SAT, and 3.14 seconds for RNA UNSAT.
-The "\texttt{Rest}" category, encompassing various other string constraints, averages 0.79 seconds.
+CertiStrR employs a straightforward forward-propagation algorithm \cite{cpp/KanLRS22} combined with our SFT-based modeling of replacement operations.
+This combination not only resolves numerous test cases that state-of-the-art solvers fail to handle (either timing out or returning unknown results), but also achieves superior performance across most benchmark subgroups.
+These results demonstrate the effectiveness and efficiency of our certified SFT formalization and implementation.
 

### 2025-09-02 Paper/application.tex
polish application

@@
 Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before rendering in web browsers.
-Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function—a fundamental sanitization operation.
+
+% Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function—a fundamental sanitization operation.
 This function iterates through each character in the input string and encodes characters that are unsafe for HTML rendering.
 
-\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
-static string EncodeHtml(string t)
-{
-  if (t == null) { return null; }
-  if (t.Length == 0) {
-    return string.Empty;
-  }
-  StringBuilder builder = new
-    StringBuilder("", t.Length * 2);
-  foreach (char c in t) {
-    if ((((c > ''') && (c < '{')) |$\mid\mid$|
-         ((c > '@') && (c < '['))) |$\mid\mid$|
-         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
-         (((c == '.') || (c == ',')) |$\mid\mid$|
-         ((c == '-') || (c == '_'))))) {
-          builder.Append(c);
-      }
-      else {
-          builder.Append("&#" +
-          ((int) c).ToString() + ";");
-      }
-  }
-  return builder.ToString();
-}
-\end{lstlisting}
+% \begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
+% static string EncodeHtml(string t)
+% {
+%   if (t == null) { return null; }
+%   if (t.Length == 0) {
+%     return string.Empty;
+%   }
+%   StringBuilder builder = new
+%     StringBuilder("", t.Length * 2);
+%   foreach (char c in t) {
+%     if ((((c > ''') && (c < '{')) |$\mid\mid$|
+%          ((c > '@') && (c < '['))) |$\mid\mid$|
+%          (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
+%          (((c == '.') || (c == ',')) |$\mid\mid$|
+%          ((c == '-') || (c == '_'))))) {
+%           builder.Append(c);
+%       }
+%       else {
+%           builder.Append("&#" +
+%           ((int) c).ToString() + ";");
+%       }
+%   }
+%   return builder.ToString();
+% }
+% \end{lstlisting}
 
-Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
+% Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
 The transducer accepts any input string and produces an output where safe characters are preserved using the identity function $\texttt{id}$, while unsafe characters are encoded as HTML entities.
@@
 When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
-\begin{figure}[htbp]
-\centering
-\begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
-  % States
-  \node[state, initial, accepting] (q0) {$q_0$};
-  \node[state] (q1) [right=3.5cm of q0] {$q_1$};
-  \node[state] (q2) [below=of q0] {$q_2$};
+% \begin{figure}[htbp]
+% \centering
+% \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
+%   % States
+%   \node[state, initial, accepting] (q0) {$q_0$};
+%   \node[state] (q1) [right=3.5cm of q0] {$q_1$};
+%   \node[state] (q2) [below=of q0] {$q_2$};
 
-  \path[->] 
-    % Transition for safe characters
-    (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
-    % Transition for unsafe characters  
-    (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
-    % Epsilon transitions back to q0
-    (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
-    (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
-\end{tikzpicture}
-\caption{SFT representation of the HTML encoding sanitizer function.}
-\label{fig:html-sanitizer-sft1}
-\end{figure}
+%   \path[->] 
+%     % Transition for safe characters
+%     (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
+%     % Transition for unsafe characters  
+%     (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
+%     % Epsilon transitions back to q0
+%     (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
+%     (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
+% \end{tikzpicture}
+% \caption{SFT representation of the HTML encoding sanitizer function.}
+% \label{fig:html-sanitizer-sft1}
+% \end{figure}
 
-We evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT with SFAs constructed from randomly generated Unicode user inputs.
-This experimental setup reflects realistic usage scenarios where diverse character sets and input patterns are encountered.
-The experimental evaluation was conducted on a laptop with an Apple M4 processor and 24 GB of memory, with a one-minute time limit per test.
-
-Figure \ref{fig:benchmark-results} demonstrates the scalability characteristics of our SFT product implementation.
-The computation requires approximately 1 second for 100,000-character inputs and 10 seconds for 1,000,000-character inputs.
-These performance metrics confirm that our certified implementation maintains computational efficiency suitable for practical string processing applications at scale.
+We first evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT shown in Figure \ref{fig:html-sanitizer-sft} (see Section \ref{sec:motivation} for motivation) with SFAs constructed from randomly generated Unicode sequences.
+This product represents the sanitization result produced by the SFT.
+The experimental evaluation was conducted on a laptop equipped with an Apple M4 processor and 24 GB of memory, with a one-minute time limit imposed for each test.
 
@@
 \end{figure}
+
+Figure \ref{fig:benchmark-results} demonstrates the scalability characteristics of our SFT product implementation.
+The computation requires approximately 1 second for 100,000-character inputs and 10 seconds for 1,000,000-character inputs.
+These performance metrics confirm that our certified implementation maintains computational efficiency suitable for practical string processing applications at scale.
 
@@
 The "unknown" label in the table indicates instances where CVC5 cannot determine satisfiability. 
-CertiStrR outperforms both Ostrich and CVC5 across most benchmark subgroups, with the notable exception of the "Replace Str" category where CVC5 achieves better performance (though it experiences one timeout).
+Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string
+solvers, including CVC5 \cite{cvc5} and Z3Str \cite{Z3-str3}, yet CertiStrR solves
+them efficiently. Among existing solvers, only OSTRICH \cite{pacmpl/ChenFHHHKLRW22} can handle these benchmarks.
+CertiStrR outperforms both Ostrich and CVC5 across most benchmark subgroups, with the only exception of the "Replace Str" category where CVC5 achieves better performance (though it experiences one timeout).
 CertiStrR demonstrates particularly strong performance in the "Replace All" category.
@@
 These results demonstrate the effectiveness and efficiency of our certified SFT formalization and implementation.
-
-Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string
-solvers, including CVC4 \cite{cvc4} and Z3 \cite{Z3-str3}, yet CertiStrR solves
-them efficiently. Among existing solvers, only OSTRICH \cite{pacmpl/ChenFHHHKLRW22} can handle these benchmarks.
 

### 2025-09-03 Paper/formalization.tex
polish the formalization parts

@@
 
-The \textbf{abstract layer} defines transition labels as \textbf{sets}, states are stored in sets as well.
-This abstract formalization, which abstracts away implementation details, enables straightforward correctness proofs for SFT operations.
+The \textbf{abstract layer} represents transition labels as \textbf{sets}, with states likewise stored in sets.
+This abstraction, which deliberately omits implementation details, closely aligns with the SFT definitions presented in Section~\ref{sec:sft}.
 
@@
 
-The second element of type $'i$ serves as an index into the output function space. 
+The second element of type $'i$ in LTTS serves as an index into the output function space. 
 %
@@
 The definition \texttt{outputL} generalizes \texttt{outputE} to characterize the set of all possible outputs that an SFT $\mathcal{T}$ can generate when processing inputs from the language accepted by the SFA $\mathcal{A}$.
-The reachability of a trace $\pi$ between states $q$ and $q'$ in an SFT $\mathcal{T}$ is verified by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
+The reachability of a trace $\pi$ between states $q$ and $q'$ is checked by the predicate $\texttt{LTTS\_reachable}~\mathcal{T}~q~\pi~q'$.
 
@@
 \label{sec_alg_refinement}
-Having established the abstract definition of the SFT product operation in Section \ref{sec:abstract-layer}, we now present its algorithmic refinement.
-This section introduces an efficient implementation of the product construction and refines the abstract representation of transition labels to a concrete interval algebra, enabling practical computation.
+Having introduced the abstract definition of the SFT product operation in Section \ref{sec:abstract-layer}, we now turn to its algorithmic refinement.
+In this section, we present an efficient implementation of the product construction and refine the representation of transition labels into a Boolean algebra locale, which serves as the interface structure.
 
@@
 \begin{figure}[hbt!]
-\begin{lstlisting}[language=SMTLIB]
+\begin{lstlisting}
 definition subtrans_comp where
@@
 
-Finally, we present the implementation of the interval algebra.
+Finally, we present the implementation of the interval algebra which implements all operations defined in boolean algebra locale.
 An interval is defined as a pair $(i, j)$ (represented as $[i-j]$ as well in the paper) representing the set $\{e \mid i \leq e \leq j\}$.

### 2025-09-04 Paper/application.tex
polish application

@@
 \section{Applications and Evaluation}
+\label{sec:application}
 We evaluate our formalization and implementation of the SFT product operation across two key application domains: string sanitization and string replacement in string solving.

### 2025-09-04 Paper/introduction.tex
polish application

@@
 The remainder of this paper is organized as follows:
-Section \ref{sec:motivation} provides several concrete motivating examples.
-Section \ref{sec:formalization} introduces the formalization of symbolic finite transducers.
-Section \ref{sec:product-operation} presents the product operation at the abstract level.
-Section \ref{sec_alg_refinement} details the algorithmic refinement of the product operation.
-Section \ref{sec-app-str-solver} demonstrates the application to string constraint solving.
+Section \ref{sec:motivation} provides concrete motivating examples.
+Section \ref{sec:sft} introduces the definition of symbolic finite transducers and their product operations.
+Section \ref{sec:formalization} presents our formalization of SFTs in Isabelle/HOL.
+Section \ref{sec:application} demonstrates applications to sanitizer analysis and string constraint solving.
 Section \ref{sec:related-work} discusses related work.
-Section \ref{sec:conclusion} concludes with future directions.
+Section \ref{sec:conclusion} concludes and outlines future directions.

### 2025-09-04 Paper/application.tex
change benchmark name from Rest to WebApp

@@
    & RNA UNSAT & 500  & 0   & 500 & 0 & 3.14 & 64.185/0 & unknown \\
-   & Rest      & 255  & 40  & 215 & 0 & 0.79 & 23.67/17 & 8.379/1 \\
+   & WebApp      & 255  & 40  & 215 & 0 & 0.79 & 23.67/17 & 8.379/1 \\
   \bottomrule

### 2025-09-08 Paper/application.tex
repolish applications

@@
 We evaluate our formalization and implementation of the SFT product operation across two key application domains: string sanitization and string replacement in string solving.
-These case studies highlight the practical value of our certified SFT product operation in real-world scenarios.
+These case studies highlight the practical value of our certified SFT formalization in real-world scenarios.
 \subsection{Sanitizer}
 
-An important application of our formalization lies in string sanitization for web applications.
 Sanitizers are crucial for preventing security vulnerabilities such as cross-site scripting (XSS) attacks by ensuring that user inputs are properly encoded before rendering in web browsers.
 
-% Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function—a fundamental sanitization operation.
+Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function for Anti-XSS.
 This function iterates through each character in the input string and encodes characters that are unsafe for HTML rendering.
 
-% \begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
-% static string EncodeHtml(string t)
-% {
-%   if (t == null) { return null; }
-%   if (t.Length == 0) {
-%     return string.Empty;
-%   }
-%   StringBuilder builder = new
-%     StringBuilder("", t.Length * 2);
-%   foreach (char c in t) {
-%     if ((((c > ''') && (c < '{')) |$\mid\mid$|
-%          ((c > '@') && (c < '['))) |$\mid\mid$|
-%          (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
-%          (((c == '.') || (c == ',')) |$\mid\mid$|
-%          ((c == '-') || (c == '_'))))) {
-%           builder.Append(c);
-%       }
-%       else {
-%           builder.Append("&#" +
-%           ((int) c).ToString() + ";");
-%       }
-%   }
-%   return builder.ToString();
-% }
-% \end{lstlisting}
+\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
+static string EncodeHtml(string t)
+{
+  if (t == null) { return null; }
+  if (t.Length == 0) {
+    return string.Empty;
+  }
+  StringBuilder builder = new
+    StringBuilder("", t.Length * 2);
+  foreach (char c in t) {
+    if ((((c > ''') && (c < '{')) |$\mid\mid$|
+         ((c > '@') && (c < '['))) |$\mid\mid$|
+         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
+         (((c == '.') || (c == ',')) |$\mid\mid$|
+         ((c == '-') || (c == '_'))))) {
+          builder.Append(c);
+      }
+      else {
+          builder.Append("&#" +
+          ((int) c).ToString() + ";");
+      }
+  }
+  return builder.ToString();
+}
+\end{lstlisting}
 
-% Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
-The transducer accepts any input string and produces an output where safe characters are preserved using the identity function $\texttt{id}$, while unsafe characters are encoded as HTML entities.
+Figure \ref{fig:html-sanitizer-sft} illustrates the SFT representation of the HTML encoding sanitizer function.
  The set of safe characters is defined as $\texttt{safe\_chars} = [\texttt{' '}, \texttt{','}, \texttt{'.'}, \texttt{'-'}, \texttt{'\_'}] \cup [\texttt{'0'}-\texttt{'9'}] \cup [\texttt{'A'}-\texttt{'Z'}] \cup [\texttt{'a'}-\texttt{'z'}]$, which can be efficiently represented using the interval algebra we formalized.
@@
 When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
-% \begin{figure}[htbp]
-% \centering
-% \begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
-%   % States
-%   \node[state, initial, accepting] (q0) {$q_0$};
-%   \node[state] (q1) [right=3.5cm of q0] {$q_1$};
-%   \node[state] (q2) [below=of q0] {$q_2$};
+\texttt{id}, \texttt{encode}, and $\lambda \_.\varepsilon$ are the identity, encode of unsafe characters, and empty word output functions, respectively.
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[shorten >=1pt, node distance=2.5cm, on grid, auto]
+  % States
+  \node[state, initial, accepting] (q0) {$q_0$};
+  \node[state] (q1) [right=3.5cm of q0] {$q_1$};
+  \node[state] (q2) [below=of q0] {$q_2$};
 
-%   \path[->] 
-%     % Transition for safe characters
-%     (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
-%     % Transition for unsafe characters  
-%     (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
-%     % Epsilon transitions back to q0
-%     (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda\varepsilon$} (q0)
-%     (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda\varepsilon$} (q0);
-% \end{tikzpicture}
-% \caption{SFT representation of the HTML encoding sanitizer function.}
-% \label{fig:html-sanitizer-sft1}
-% \end{figure}
+  \path[->] 
+    % Transition for safe characters
+    (q0) edge[bend left=20] node[above,midway=1] {$\texttt{safe\_chars}, \texttt{id}$} (q1)
+    % Transition for unsafe characters  
+    (q0) edge[bend right=20] node[left,midway] {$\neg\texttt{safe\_chars}, \texttt{encode}$} (q2)
+    % Epsilon transitions back to q0
+    (q1) edge[bend left=30] node[below,midway] {$\varepsilon, \lambda \_.\varepsilon$} (q0)
+    (q2) edge[bend right=20] node[right,midway] {$\varepsilon, \lambda \_.\varepsilon$} (q0);
+\end{tikzpicture}
+\caption{SFT representation of the HTML encoding sanitizer function.}
+\label{fig:html-sanitizer-sft}
+\end{figure}
 
-We first evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT shown in Figure \ref{fig:html-sanitizer-sft} (see Section \ref{sec:motivation} for motivation) with SFAs constructed from randomly generated Unicode sequences.
+We first evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT shown in Figure \ref{fig:html-sanitizer-sft} with SFAs constructed from randomly generated Unicode sequences.
 This product represents the sanitization result produced by the SFT.
@@
     \midrule
-    attack\_javascript     & GA Htmlescape + GA PreEscape + & 3/20 \\
+    attack\_javascript     & GA HtmlEscape + GA PreEscape + & 3/20 \\
                            &  gaSnippetesc: \textbf{safe} & \\
@@
 The attack model includes comments, declaration breakers, and known dangerous tokens such as \textbf{expression(}, \textbf{@import}, \textbf{behavior:}, \textbf{-moz-binding}, and \textbf{url(} (unless the URL is separately validated).
-Sanitizers include GA HtmlEscape, GA JSONEscape, GA XMLEscape, escapeString, and GA CleanseCSS, where "GA" means from Google.
+Sanitizers include GA HtmlEscape, GA JSONEscape, GA XMLEscape, escapeString, and CleanseCSS, where "GA" means from Google Analytics.
 Sanitizers are often composed with preprocessing steps such as Trim, which removes leading and trailing whitespace characters.
@@
 
-% The string replacement, denoted as
-% $\texttt{replace}(\texttt{str}, \texttt{pattern}, \newline\texttt{replacement}, \texttt{flag})$, is a fundamental string transformation accepting four parameters:
-% \begin{itemize}
-%   \item $\texttt{str}$: The input string to be transformed.
-%   \item $\texttt{pattern}$: A regular expression defining the matching criteria.
-%   \item $\texttt{replacement}$: The string that replaces matched substrings.
-%   \item $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single occurrence replacement and $g$ denotes replace-all operation.
-% \end{itemize}
+\subsection{Applications to String Replacement}
+
+The string replacement, denoted as
+$\texttt{replace}(\texttt{str}, \texttt{pattern}, \newline\texttt{replacement}, \texttt{flag})$, is a fundamental string transformation accepting four parameters:
+(1) $\texttt{str}$: The input string to be transformed.
+(2) $\texttt{pattern}$: A regular expression defining the matching criteria.
+(3) $\texttt{replacement}$: The string that replaces matched substrings.
+(4) $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single first occurrence replacement and $g$ denotes all occurrences replace operation.
 
@@
 
-% To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}", \texttt{flag})$. We construct the SFT representation through three steps:
+To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}", \texttt{flag})$. We construct the corresponding SFT by the following three steps.
 
-% \begin{enumerate}
-%   \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$f = \lambda x.~\texttt{None}$" that produces the empty string.
-%   \item Construct an SFA accepting the replacement string "\texttt{NUM}".
+\begin{enumerate}
+  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$f = \lambda x.~\texttt{None}$" that produces the empty string.
+  \item Construct an SFA accepting the replacement string "\texttt{NUM}".
 Since constant strings are specialized regular expressions, we build their SFA representation and convert it to an SFT by adding $\varepsilon$-transitions with output functions that emit the replacement characters sequentially.
-%   \item Compose the two SFTs by connecting the accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT via $\varepsilon$-transitions.
-% \end{enumerate}
+  \item Compose the two SFTs by connecting the accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT via $\varepsilon$-transitions.
+\end{enumerate}
 
-% \noindent\emph{Step 1.}
-% Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
+%\noindent\emph{Step 1.}
+%Figure \ref{fig-snfa-pattern} illustrates the construction of the pattern-matching component.
 The left side shows the SFA that recognizes the regular expression $\texttt{/[0-9]+/}$, while the right side presents its transformation into an SFT.
 This transformation is achieved by augmenting each transition with the output function $\texttt{f} = \lambda x.~\texttt{None}$, which consistently produces the empty string, effectively "consuming" the matched digits without generating output.
+
+Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT generated from \texttt{/[0-9]+/}  with the replacement-generating SFT generated from "\texttt{NUM}", where the function \texttt{f} is defined as "$f = \lambda x.~\texttt{None}$" and the function \texttt{g} is defined as:
+\[
+\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
+\begin{cases}
+1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
+2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
+3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
+\_ \mapsto \texttt{None}
+\end{cases}
+\]
+
+\begin{figure}[h] \centering
+  \begin{tikzpicture}[shorten >=1pt, node distance=2cm, auto]
+    % Second automaton from the first figure
+    \node[state, initial] (q0') {$q_0'$}; 
+    \node[state] (q1') [right=of q0'] {$q_1'$};
+
+     \path[->] 
+     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
+     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1');
+
+    % Second automaton from the second figure
+    \node[state] (q0'') [below=1cm of q1'] {$q_0''$}; 
+    \node[state] (q1'') [right=of q0''] {$q_1''$}; 
+    \node[state, accepting] (q3'') [below=1cm of q0''] {$q_3''$}; 
+    \node[state] (q2'') [right=of q3''] {$q_2''$};
+
+    \path[->] 
+    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
+    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
+    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
+    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
+  \end{tikzpicture}
+  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
+  \label{fig-rearranged-automata}
+  \end{figure}
+
+  We can extend the above construction to model both singleton and replace-all operations.
+In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}.
+The SFT $\mathcal{T}_2$ combines pattern matching and replacement using the construction just described.
+The SFT $\mathcal{T}_1$ matches strings that do not contain any substrings matching the pattern.
+Two $\varepsilon$-transitions connect these components: one from $\mathcal{T}_1$'s accepting state to $\mathcal{T}_2$'s initial state, and another from $\mathcal{T}_2$'s accepting state back to $\mathcal{T}_1$'s initial state.
+This cyclic structure enables the replace-all operation by repeatedly applying pattern replacement until no more matches exist in the input string.
+
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[node distance=2cm, state/.style={circle, draw, minimum size=0.8cm}]
+  % First SFT rectangle
+  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm] (sft1-box) {};
+  \node[above=0.1cm of sft1-box.north] {\textbf{SFT} $\mathcal{T}_1$ (Non-matching)};
+
+  % States inside first SFT
+  \node[state, initial] (p0) at ([xshift=-0.8cm]sft1-box.center) {$p_0$};
+  \node[state, accepting] (p1) at ([xshift=1.8cm]sft1-box.center) {$p_1$};
+  \path[->] (p0) edge[above] node {\shortstack{$\neg\texttt{pattern},$ \\ $\lambda x.x$}} (p1);
+
+  % Second SFT rectangle
+  \node[draw, rectangle, minimum width=5.2cm, minimum height=2.1cm, below=1.5cm of sft1-box] (sft2-box) {};
+  \node[below=0.1cm of sft2-box.south] {\textbf{SFT} $\mathcal{T}_2$ (Matching)};
+
+  % States inside second SFT
+  \node[state, initial] (s0) at ([xshift=-0.8cm]sft2-box.center) {$s_0$};
+  \node[state, accepting] (s1) at ([xshift=2cm]sft2-box.center) {$s_1$};
+  \path[->] (s0) edge[above] node {\shortstack{$\texttt{pattern},$ \\ $\texttt{replacement}$}} (s1);
+
+  % Epsilon transition from accepting state of first SFT to initial state of second SFT
+  \path[->] (p1) edge[bend right=30] node[left] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (s0);
+
+  % Epsilon transition from accepting state of second SFT to initial state of first SFT
+  \path[->] (s1) edge[bend right=30] node[right] {\shortstack{$\varepsilon$, $\lambda\varepsilon$}} (p0);
+
+\end{tikzpicture}
+\caption{The modeling of replace all operation.}
+\label{fig:replace-all}
+\end{figure}
 
@@
 
-\subsection{Experiments of String Replacement Operations}
+We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that originally supports only basic string operations such as concatenation and regular constraints.
+CertiStr employs a forward-propagation approach to solve string constraints, which requires computing the forward image of string operations.
+Our SFT product operation provides an exact forward image computation for replacement operations, enabling seamless integration of replacement functionality into CertiStr's solving framework.
 
-We evaluate our transducer formalization and implementation by extending CertiStr \cite{cpp/KanLRS22} to support string replacement operations, which is a certified string solver that originally supports only basic string operations, such as concatenation and regular constraints.
-By integrating our SFT-based modeling of replacement operations, we develop CertiStrR, an extended string solver that supports two types of string replacement: (1) replacing the first occurrence of a substring matching a given pattern, and (2) replacing all occurrences of substrings matching a given pattern.
+By incorporating our SFT-based modeling of replacement, we develop CertiStrR—an extended string solver that supports two types of string replacement: (1) replacing the first occurrence of a substring matching a given pattern, and (2) replacing all occurrences of substrings matching a given pattern.
 The patterns can be either constant strings or regular expressions.
@@
 The benchmarks are divided into two categories: (1) singleton matching and (2) replace-all matching.
-In addition to replacement, they also include string concatenation and regular constraints.
+In addition to replacement, they also include string concatenation and regular constraints in the tests.
 
@@
 CertiStrR demonstrates particularly strong performance in the "Replace All" category.
-
-CertiStrR employs a straightforward forward-propagation algorithm \cite{cpp/KanLRS22} combined with our SFT-based modeling of replacement operations.
-This combination not only resolves numerous test cases that state-of-the-art solvers fail to handle (either timing out or returning unknown results), but also achieves superior performance across most benchmark subgroups.
 These results demonstrate the effectiveness and efficiency of our certified SFT formalization and implementation.

### 2025-09-10 Paper/introduction.tex
Something wrong with the example

@@
 compliant certified string solver handling equality, concatenation, regular
-constraints, and more important the string replaceall operator, which has been
+constraints, and more importantly the string replaceall operator, which has been
 argued to be indispensable for analysis of XSS 

### 2025-09-10 Paper/abstract.tex
Continuing with intro
fix motivation

@@
     represented models. To circumvent to the notorious state-space explosion 
-    problem caused by the extremely large alphabet size (e.g. UTF-16), SFTs
+    problem caused by an extremely large alphabet size (e.g. UTF-16), SFTs
     and SFAs allow the representation of the alphabet as an effective boolean
@@
     To evaluate the effectiveness of our formalization, we apply the formalized 
-    SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for 
+    SFTs to two applications: (1) sanitizers for web applications used for 
     preventing XSS attacks, and (2) string solving, which increasingly

### 2025-09-10 Paper/introduction.tex
fix motivation

--- 2025-09-04 Paper/introduction.tex
@@
 To evaluate the effectiveness and efficiency of our formalization, we 
-provide two applications: (1) analysis of the sanitizer \texttt{HTMLdecode},
-which is used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
+provide two applications: (1) analysis of the web application sanitizers for html, css, json, and javascript
+which are used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
 compliant certified string solver handling equality, concatenation, regular
-constraints, and more important the string replaceall operator, which has been
+constraints, and more importantly the string replaceall operator, which has been
 argued to be indispensable for analysis of XSS 

### 2025-09-10 Paper/introduction.tex
Fixed typos

@@
 sanitization, as well as implicitly applied transductions, for instance
-through \texttt{innerHTML}). In particular, SMT-LIB string constraints that 
+through \texttt{innerHTML}) and are well-known to be critical for analyzing XSS
+vulnerabilities for web applications
+\cite{Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
 arise from such applications require the \emph{replaceall} operator (part of
@@
 
-To address these challenges inherent, %in supporting diverse transition label theories, 
+To address these challenges, %in supporting diverse transition label theories, 
 we adopt a refinement-based approach.
@@
 argued to be indispensable for analysis of XSS 
-\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}).
+\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}.
 The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.

### 2025-09-10 Paper/introduction.tex
Typos, more discussion added

@@
 applications (e.g. see
-\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}):
+\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,systematic-transduction,uss/HooimeijerLMSV11}):
 \emph{string transductions} and \emph{replaceall} operators. In particular, 
@@
 vulnerabilities for web applications
-\cite{Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
+\cite{Berkeley-JavaScript,systematic-transduction,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
 arise from such applications require the \emph{replaceall} operator (part of
@@
 argued to be indispensable for analysis of XSS 
-\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}.
+\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,systematic-transduction,uss/HooimeijerLMSV11}.
 The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.

### 2025-09-11 Paper/introduction.tex
contribution

@@
 \begin{enumerate}
-\item We present the first formalization of symbolic finite transducers in Isabelle/HOL proof assistant.
-\item We leverage the refinement framework to create an extensible SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
-\item We develop an interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
-\item We demonstrate the practical utility of our formalization through its
-    application to sanitizer analysis and a certified implementation of
-    a string constraint solver that handles equality, concatenation, regular
-        constraints, and the string replacement operator.
+\item We present the first formalization of symbolic finite transducers (SFTs) in the Isabelle/HOL. 
+We prove the correctness of key operations between SFTs and symbolic SFAs, such as the product operation between SFTs and SFAs.
+\item We leverage the refinement framework to create an extensible and efficient SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
+The correctness of the refined efficient implementation is proved .
+\item We develop a certified interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
+\item We demonstrate the practical utility of our formalization by applying it to sanitizer analysis and integrating our formalized SFT into the verified forward-propagation string solver CertiStr \cite{cpp/KanLRS22}.
+This integration extends CertiStr from supporting only simple concatenation and regular constraints to handling more sophisticated string replacement operations.
 \end{enumerate}

### 2025-09-11 Paper/application.tex
polish introduction

@@
  The set of safe characters is defined as $\texttt{safe\_chars} = [\texttt{' '}, \texttt{','}, \texttt{'.'}, \texttt{'-'}, \texttt{'\_'}] \cup [\texttt{'0'}-\texttt{'9'}] \cup [\texttt{'A'}-\texttt{'Z'}] \cup [\texttt{'a'}-\texttt{'z'}]$, which can be efficiently represented using the interval algebra we formalized.
-Unsafe characters are transformed using the encoding function $\texttt{encode}(c) = \texttt{``\&\#''} + \texttt{toString}(\texttt{ord}(c)) + \texttt{``\;''}$.
+Unsafe characters are transformed using the encoding function $\texttt{encode}(c) = \texttt{``\&\#''} + \texttt{toString}(\texttt{ord}(c)) + \texttt{``;''}$.
 When the alphabet is set to Unicode $[\mathtt{0x0000}, \mathtt{0x10FFFF}]$, the complement $\neg\texttt{safe\_chars}$ represents all Unicode characters outside the safe set, which can also be efficiently represented using our formalized interval operations.
@@
 
-Beyond performance evaluation, our SFT formalization also supports the formal verification of string sanitizer correctness.
+Beyond performance evaluation, our SFT formalization also supports the formal verification of string sanitizer's safety property.
 A sanitizer transforms user inputs into safe outputs, ensuring that the result contains no unsafe characters.
@@
 The attack model includes comments, declaration breakers, and known dangerous tokens such as \textbf{expression(}, \textbf{@import}, \textbf{behavior:}, \textbf{-moz-binding}, and \textbf{url(} (unless the URL is separately validated).
+These are unsafe and undesired strings in CSS contexts.
 Sanitizers include GA HtmlEscape, GA JSONEscape, GA XMLEscape, escapeString, and CleanseCSS, where "GA" means from Google Analytics.

### 2025-09-11 Paper/formalization.tex
polish introduction

@@
 
-\paragraph{The product operation formalization.} We formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:formalization}.
+\paragraph{The product operation formalization.} We formalize the product operation between an SFT $\mathcal{T}$ and an SFA $\mathcal{A}$, denoted as $\mathcal{T} \times \mathcal{A}$ in Section \ref{sec:sft}.
 The SFA is formalized as a record type \texttt{NFA in Figure \ref{fig-def-FAs}}.
@@
                                     and
-  nempty_sem: "nempty_intervals s =
-                     ($\neg$ (empty s))" and
+  nempty_sem: "nempty s = $\neg$ (empty s)" 
+                                    and
   inter_sem: "sem (intersect s1 s2) =
@@
 
-The implementation of \texttt{productT\_imp} follows \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
+The implementation of \texttt{productT\_imp} takes advantages of \emph{Refine\_Monadic} framework's interfaces for sets to store states and transition relations. 
 The \emph{Refine\_Monadic} framework provides a way to automatically refine these interfaces to more efficient data structure, such as red-black trees or hashmaps. 

### 2025-09-11 Paper/introduction.tex
polish introduction

@@
 For example, in the interval algebra, a transition label might be specified as 
-an interval ($'\texttt{a}'-'\texttt{z}'$), encompassing all characters from
-$'\texttt{a}'$ to $'\texttt{z}'$. As another example, in the SMT-Algebra
+an interval [$\texttt{a}-\texttt{z}$], encompassing all characters from
+$\texttt{a}$ to $\texttt{z}$. As another example, in the SMT-Algebra
 (particularly, of the theory of Linear Integer Arithmetic), %or 
@@
 \begin{enumerate}
-\item We present the first formalization of symbolic finite transducers (SFTs) in the Isabelle/HOL. 
+\item We present the first formalization of symbolic finite transducers (SFTs) in Isabelle/HOL.
 We prove the correctness of key operations between SFTs and symbolic SFAs, such as the product operation between SFTs and SFAs.
-\item We leverage the refinement framework to create an extensible and efficient SFT formalization that is capable of accommodating various boolean algebras, ensuring adaptability to future transition label developments.
-The correctness of the refined efficient implementation is proved .
-\item We develop a certified interval algebra with efficient set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
+\item We leverage the refinement framework to develop an extensible and efficient SFT formalization capable of supporting diverse Boolean algebras, ensuring adaptability to future transition label representations.
+The correctness of the refined, efficient implementation is proved via structural isomorphism between refinement layers.
+\item We develop a certified interval algebra with verified set-theoretic operations, enabling refinement of transition labels from abstract sets to concrete intervals.
 \item We demonstrate the practical utility of our formalization by applying it to sanitizer analysis and integrating our formalized SFT into the verified forward-propagation string solver CertiStr \cite{cpp/KanLRS22}.

### 2025-09-11 Paper/application.tex
polish application, conclusion, and related work

@@
 
-  We can extend the above construction to model both singleton and replace-all operations.
+  We can extend the above construction to model both replace-first and replace-all operations.
 In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}.
@@
 We evaluate {CertiStrR} using benchmarks from SMT-LIB 2025~\cite{smtlib_benchmarks}, focusing on the \textsf{QF\_S} and \textsf{QF\_SLIA} logic fragments.
-The benchmarks are divided into two categories: (1) singleton matching and (2) replace-all matching.
+The benchmarks are divided into two categories: (1) replace-first matching and (2) replace-all matching.
 In addition to replacement, they also include string concatenation and regular constraints in the tests.
@@
   \midrule
-  \multirow{2}{*}{Replace Singleton} & Replace RE  & 94  & 77  & 16  & 1 & 0.79 & 166.27/27 & 2.67/20 \\
+  \multirow{2}{*}{Replace-First} & Replace RE  & 94  & 77  & 16  & 1 & 0.79 & 166.27/27 & 2.67/20 \\
    & Replace Str & 323 & 103 & 211 & 9 & 1.94 & 7.10/20 & 0.81/1 \\
-  \multirow{4}{*}{Replace All} & PCP & 1000 & 751 & 249 & 0 & 0.14 & 3.412/0 & unknown \\
+  \multirow{4}{*}{Replace-All} & PCP & 1000 & 751 & 249 & 0 & 0.14 & 3.412/0 & unknown \\
    & RNA SAT   & 500  & 500 & 0   & 0 & 2.81 & 55.61/0 & unknown \\
@@
 In contrast, all replacement operations in SMT-LIB treat the pattern and the replacement string as independent.
-For example, in the SFT shown in Figure~\ref{fig:html-sanitizer-sft}, a transition may be defined over an interval such as $['a'-'z']$, and the output function computes its output directly based on the specific input character.
+For example, in the SFT shown in Figure~\ref{fig:html-sanitizer-sft}, a transition may be defined over an interval such as $[\texttt{a-z}]$, and the output function computes its output directly based on the input character.
 Replacement operations, however, cannot express such dependencies.
+Because the replacement string is not a function.
 
-Consider the transition $['a'-'z']/id$, where $id$ is the identity function.
+Consider the transition $[\texttt{a-z}]/id$, where $id$ is the identity function.
 To model this transition with replacement operations, the interval must be explicitly expanded into the individual characters $\texttt{'a'}$ through $\texttt{'z'}$, requiring 26 nested \texttt{replace-all} operations—one for each character.
-Thus, modeling $['a'-'z']/id$ with replacement operations yields an expression of the form:
+Thus, modeling $[\texttt{a-z}]/id$ with replacement operations yields an expression of the form:
 \[
@@
 
-When the interval contains a large set of characters—as is common in web applications that support the full Unicode range of $0x10FFFF$ code points—this nested structure becomes impractical, leading to significant inefficiency and complexity in the replacement-based approach.
+When the interval contains a large set of characters—as is common in web applications that support the full Unicode range of 
+\texttt{0x10FFFF} code points—this nested structure becomes impractical, leading to significant inefficiency and complexity in the replacement-based approach.
 
@@
       \toprule
-      & \textbf{Defs} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
+      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
       \midrule

### 2025-09-11 Paper/conclusion.tex
polish application, conclusion, and related work

@@
 
-In this paper, we have presented the first formalization of symbolic finite transducers in the proof assistant Isabelle/HOL.
-Our formalization provides flexible interfaces that facilitate diverse applications through two key features: support for $\varepsilon$-transitions in both inputs and outputs, and extensibility to various boolean algebras via the refinement framework.
+In this paper, we presented the first formalization of symbolic finite transducers (SFTs) in Isabelle/HOL.
+Our framework offers flexible interfaces that enable diverse applications, featuring support for $\varepsilon$-transitions on both inputs and outputs, as well as extensibility to various Boolean algebras through a refinement framework.
 
-To demonstrate the practical utility of our formalization, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}.
-This implementation adds support for string replacement operations and has been evaluated on benchmarks from SMT-LIB 2024 \cite{smtlib_benchmarks}.
-The experimental results confirm both the efficiency and effectiveness of our approach.
-While we focused on string solving applications, our extensible framework for transition labels is broadly applicable to other domains, including program verification.
-
-Future work includes extending our symbolic formalization in two key directions:
-%
-(1) Enriching the boolean algebra framework to support more complex theories, such as arithmetic constraints.
-(2) Incorporating prioritized transitions into SFTs, which would enable precise modeling of SMT-LIB's standard semantics for replacement operations, particularly the first-match behavior.
-These extensions will further enhance the expressiveness and practical applicability of our formalization while maintaining its formal verification guarantees.
+To demonstrate the practical utility of our formalization, we applied SFTs to model string sanitization functions commonly used in web applications and showed how to verify the correctness of these sanitizers.
+Furthermore, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}, which adds support for string replacement operations.
+We evaluated CertiStrR on benchmarks from SMT-LIB 2025 \cite{smtlib_benchmarks}, and the experimental results confirm both the efficiency and effectiveness of our approach.

### 2025-09-12 Paper/SFT.tex
polish the formalization parts
Working on related work, shortening etc.

@@
 \section{Symbolic Finite Transducers}
-\label{sec:formalization}
+\label{sec:sft}
 
-We begin by presenting a mathematical definition of SFTs \cite{VeanesHLMB12Transducer}, abstracting from the specific implementation details of Isabelle/HOL.
+We recall the definition of SFTs \cite{VeanesHLMB12Transducer}, 
+abstracting from specific implementation details in Isabelle/HOL.
 
@@
 \phi_i, f_i, q_{i+1}), 0 \leq i \leq n-1$ is a transition in $\Delta$.
-$\sigma$ is an accepting run when $q_n$ is an accepting state.
+$\sigma$ is \emph{accepting} when $q_n \in \mathcal{F}$.
 

### 2025-09-12 Paper/formalization.tex
Working on related work, shortening etc.

@@
 The predicate $\lambda b.~ b[7] = 1$, where $b$ is a bit vector of length 8, denotes the set of bit vectors that have a 1 in the 7th position.
-All these different forms of labels are abstracted as sets in our formalization.
+All different forms of labels are abstracted as sets in our formalization.
 %
@@
 %
-We use indices instead of functions themselves to enable the reuse of the same output functions for different transitions.
+We use indices instead of functions %themselves 
+to enable the reuse of the same output functions for different transitions.
 %
-The mapping $\mathcal{M}_t$ associates each index with a specific output function.
-These output functions, formalized by \texttt{Tlabel}, map a single input element to a set of possible output elements rather than to a single element.
-This design enables non-deterministic output behavior, where the transducer may select any element from the output set randomly or according to specified criteria.
-Additionally, output functions can produce the empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
+The mapping $\mathcal{M}_t$ associates each index with a specific output
+function. These output functions, formalized by \texttt{Tlabel}, map a single
+input element to a set of possible output elements rather than to a single
+element. This design enables nondeterministic output behavior, where the
+transducer may select any element from the output set arbitrarily or according
+to specified criteria. Further, output functions can produce empty string $\varepsilon$ by returning $\texttt{None}$, providing further flexibility in transition behavior.
 

### 2025-09-12 Paper/introduction.tex
Working on related work, shortening etc.

@@
 Owing to the security-critical nature of many applications of string analysis
-(e.g. analysis of web applications against XSS vulnerabilities), there is an 
+(e.g. analysis of web applications against XSS vulnerabilities
+\cite{Berkeley-JavaScript} and
+at AWS for analyzing Role-Based 
+Access Control (RBAC) policies \cite{neha,DBLP:conf/fmcad/BackesBCDGLRTV18}), 
+there is an 
 increasing need of trustworthy implementations in such application domains. 
@@
 \cite{DBLP:conf/cav/BlotskyMBZKG18,BM20,Mansur20}. 
-String solvers are among others employed at AWS for analyzing Role-Based 
-Access Control (RBAC) policies, which are security-critical.
-These have motivated initial formalization efforts for the theory of strings in 
+%Recent applications of string solvers in security critical applications
+%(e.g. at AWS for analyzing Role-Based 
+%Access Control (RBAC) policies \cite{neha,DBLP:conf/fmcad/BackesBCDGLRTV18})
+This
+has motivated initial formalization efforts for the theory of strings in 
 Isabelle/HOL, e.g., a verified solver \cite{cpp/KanLRS22}
@@
 \item We present the first formalization of symbolic finite transducers (SFTs) in Isabelle/HOL.
-We prove the correctness of key operations between SFTs and symbolic SFAs, such as the product operation between SFTs and SFAs.
+We prove the correctness of key operations of SFTs and symbolic SFAs, such as the product operation between SFTs and SFAs.
 \item We leverage the refinement framework to develop an extensible and efficient SFT formalization capable of supporting diverse Boolean algebras, ensuring adaptability to future transition label representations.

### 2025-09-12 Paper/relatedwork.tex
polish application, conclusion, and related work
Working on related work, shortening etc.

@@
 This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
-Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
+%Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
 
-\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and expressive power of symbolic automata and transducers have led to their widespread adoption in numerous real-world applications.
-These applications span diverse domains, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, and automated program inversion for software transformation \cite{pldi/HuD17}.
-Each application leverages the symbolic approach's ability to handle complex patterns and infinite alphabets efficiently.
+\emph{Applications of Symbolic Transducers.} The efficiency, scalability, and
+expressive power of symbolic automata and transducers have led to their
+widespread adoption in numerous applications, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, automated program inversion for software
+transformation \cite{pldi/HuD17}, and string solving
+\cite{pacmpl/ChenFHHHKLRW22,CHL+19}. Each application leverages the symbolic 
+approach's ability to handle complex
+patterns and large alphabets efficiently.
 
 \emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
-To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
+%To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
 Our work advances this frontier by extending the formal treatment to both SFTs and $\varepsilon$SFAs.
 
-\emph{String Solving.} A significant application of our certified transducer framework is in string constraint solving, a field that has seen intensive research development over the past decade.
-While our work provides formal verification guarantees, there exists a rich ecosystem of non-certified string solvers, each with distinct capabilities: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
-As more and more bugs have beein uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
-We believe that our work will benefit the community by providing a formal foundation for string solvers development.
+\emph{String Solving.} A significant application of our certified transducer
+framework is in string constraint solving, a field that has seen intensive
+research development over the past decade. While our work provides formal
+verification guarantees, there exists a rich ecosystem of non-certified string
+solvers, which has in the last few years led to a standardized SMT-LIB 2.6/2.7 
+unicode string theory and annual competition. SMT-LIB compatible string 
+solvers include Z3 \cite{z3}, Z3-alpha \cite{z3alpha}, Z3-Noodler 
+\cite{noodler,noodler-int,noodler-len, noodler-tool}, Z3str3RE \cite{z3str3re},
+cvc5 \cite{cvc5}, Z3str4 \cite{z3str4}, Trau \cite{trau0,trau-tool,trau2}, and 
+OSTRICH \cite{ostrich}. Recent SMT competition over string theory has revealed
+superiority of automata-based methods (based on restrictions of symbolic
+automata). For this reason,
+%: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
+%As more and more 
+%Numerous bugs have been uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20} 
+our work will benefit the community by providing a formal 
+foundation for string solvers development.
 
@@
 \cite{DBLP:conf/cav/ShiFLTWY20}, who developed a certified SMT solver for quantifier-free bit-vector theory, demonstrating the broader applicability of interactive theorem proving in certified SMT solver development.
+The work by Mansur et al.
+\cite{verified-verifying} define a formal semantics for SMT-LIB string
+to check compatibility between solvers and SMTLIB standard.

### 2025-09-12 Paper/application.tex
Added missing references. Improved related work and conc. Now 12 pages

@@
 
-We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that originally supports only basic string operations such as concatenation and regular constraints.
+We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that supports only basic string operations such as concatenation and regular constraints.
 CertiStr employs a forward-propagation approach to solve string constraints, which requires computing the forward image of string operations.
@@
 
-By incorporating our SFT-based modeling of replacement, we develop CertiStrR—an extended string solver that supports two types of string replacement: (1) replacing the first occurrence of a substring matching a given pattern, and (2) replacing all occurrences of substrings matching a given pattern.
-The patterns can be either constant strings or regular expressions.
+By incorporating our SFT-based modeling of replacement, we develop CertiStrR—an
+extended string solver that supports two types of string replacement: (1)
+replacing the first occurrence of a substring matching a given pattern, and (2)
+replacing all occurrences of substrings matching a given pattern, i.e., either 
+a constant string or a regular expression.
 
@@
 \paragraph{Transducers vs. Replacement Operations.}  
-Although both transducers and replacement operations can express string transformations for security-related verification, transducers are strictly more expressive because they can capture dependencies between the input pattern and the replacement string.
-In contrast, all replacement operations in SMT-LIB treat the pattern and the replacement string as independent.
-For example, in the SFT shown in Figure~\ref{fig:html-sanitizer-sft}, a transition may be defined over an interval such as $[\texttt{a-z}]$, and the output function computes its output directly based on the input character.
-Replacement operations, however, cannot express such dependencies.
-Because the replacement string is not a function.
+Although both transducers and replacement operations can express string
+transformations for security-related verification, transducers are strictly more
+expressive because they can capture dependencies between the input pattern and
+the replacement string. In contrast, all replacement operations in SMT-LIB treat
+the pattern and the replacement string as independent. For example, in the SFT
+in Figure~\ref{fig:html-sanitizer-sft}, a transition may be defined over
+an interval such as $[\texttt{a-z}]$, and the output function computes its
+output directly based on the input character. Replacement operations, however,
+cannot express such dependencies since the replacement string is not a function.
 
@@
 
-This subsection discusses the development effort required for our certified SFT formalization.
+%This subsection discusses the development effort required for our certified SFT formalization. 
 While SFTs constitute the central contribution of this paper, they are not sufficient on their own to support the applications we target.

### 2025-09-12 Paper/conclusion.tex
Added missing references. Improved related work and conc. Now 12 pages

@@
 
-In this paper, we presented the first formalization of symbolic finite transducers (SFTs) in Isabelle/HOL.
-Our framework offers flexible interfaces that enable diverse applications, featuring support for $\varepsilon$-transitions on both inputs and outputs, as well as extensibility to various Boolean algebras through a refinement framework.
-
-To demonstrate the practical utility of our formalization, we applied SFTs to model string sanitization functions commonly used in web applications and showed how to verify the correctness of these sanitizers.
+We have presented the first formalization of symbolic finite transducers (SFTs)
+and their most crucial algorithms in Isabelle/HOL. Our framework offers 
+flexible interfaces that enable diverse applications, featuring support for 
+$\varepsilon$-transitions on both inputs and outputs, as well as extensibility to various Boolean algebras through a refinement framework.
+To demonstrate the practical utility of our formalization, we applied formalized
+SFTs for analyzing sanitization for web applications, as well as certified 
+string solving for complex constraints involving equality, concatenation,
+regular constraints, and replaceall, for which we confirm the efficiency
+and effectiveness of our approach on SMT-LIB 2025 benchmarks
+\cite{smtlib_benchmarks}.
+%We envision that our work will benefit the string solving community by 
+%providing a formal foundation for string solvers development.
+%from SMT-LIB 2025 \cite{smtlib_benchmarks}, and the experimental results confirm both the efficiency and effectiveness of our approach.
+%applied SFTs to model string sanitization functions commonly used in web applications and showed how to verify the correctness of these sanitizers.
 Furthermore, we developed CertiStrR, an extension of CertiStr \cite{cpp/KanLRS22}, which adds support for string replacement operations.
-We evaluated CertiStrR on benchmarks from SMT-LIB 2025 \cite{smtlib_benchmarks}, and the experimental results confirm both the efficiency and effectiveness of our approach.

### 2025-09-12 Paper/relatedwork.tex
Added missing references. Improved related work and conc. Now 12 pages

@@
 unicode string theory and annual competition. SMT-LIB compatible string 
-solvers include Z3 \cite{z3}, Z3-alpha \cite{z3alpha}, Z3-Noodler 
+solvers include Z3 \cite{Z3}, Z3-alpha \cite{z3alpha}, Z3-Noodler 
 \cite{noodler,noodler-int,noodler-len, noodler-tool}, Z3str3RE \cite{z3str3re},
-cvc5 \cite{cvc5}, Z3str4 \cite{z3str4}, Trau \cite{trau0,trau-tool,trau2}, and 
-OSTRICH \cite{ostrich}. Recent SMT competition over string theory has revealed
+cvc5 \cite{cvc5}, Z3str4 \cite{z3str4}, Trau \cite{Z3-trau,trau-tool}, and 
+OSTRICH \cite{ostrich,CHL+19,pacmpl/ChenFHHHKLRW22}. 
+Recent SMT competition over string theory has revealed
 superiority of automata-based methods (based on restrictions of symbolic
-automata). For this reason,
+automata). 
+We envision that our work will benefit the string solving community by 
+providing a formal foundation for string solvers development.
 %: Kaluza \cite{Berkeley-JavaScript} specializes in JavaScript analysis, CVC5 \cite{cvc5}, Z3-str3 \cite{Z3-str3} builds on the Z3 framework, S3P \cite{DBLP:conf/ccs/TrinhCJ14}, Ostrich \cite{pacmpl/ChenFHHHKLRW22}, and SLOTH \cite{HJLRV18}.
@@
 %Numerous bugs have been uncovered in existing string solvers \cite{DBLP:conf/cav/BlotskyMBZKG18} and SMT solvers \cite{Mansur20}
-our work will benefit the community by providing a formal 
-foundation for string solvers development.
 
-\emph{Certified SMT Solvers.} Beyond string theories, certification efforts in SMT solving have extended to other domains.
+\emph{Certified SMT Solvers.} Certification efforts 
+for SMT (other than string theories) also exist.
 For example, the work by Shi et al.

### 2025-09-12 Paper/SFT.tex
polish

@@
     \path[->] 
-    (q_0) edge node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_1) edge node[right] {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2)
-    (q_1) edge[loop above] node {[$'\texttt{a}'-'\texttt{z}'$],$f_1$} (q_1)
-    (q_2) edge[loop below] node {[$'\texttt{A}'-'\texttt{Z}'$],$f_2$} (q_2);
+    (q_0) edge node {[$\texttt{a}-\texttt{z}$],$f_1$} (q_1)
+    (q_1) edge node[right] {[$\texttt{A}-\texttt{Z}$],$f_2$} (q_2)
+    (q_1) edge[loop above] node {[$\texttt{a}-\texttt{z}$],$f_1$} (q_1)
+    (q_2) edge[loop below] node {[$\texttt{A}-\texttt{Z}$],$f_2$} (q_2);
   \end{tikzpicture}
@@
 
-    Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/['\texttt{a}'-'\texttt{z}']+['\texttt{A}'-'\texttt{Z}']+/$.
+Figure \ref{fig-example-ft} illustrates an SFT that accepts words matching the regular expression $/[\texttt{a}-\texttt{z}]+[\texttt{A}-\texttt{Z}]+/$.
 The transition labels utilize intervals of the form $[i\text{-}j]$, which are interpreted as predicates $\lambda x.~i \leq x \leq j$.
 The output functions $f_1$ and $f_2$ perform case transformations: $f_1 = \lambda x.~\texttt{toUpper}(x)$ converts lowercase letters to uppercase, while $f_2 = \lambda x.~\texttt{toLower}(x)$ performs the inverse operation.
-For example, given the input string \texttt{"bigSMALL"}, this SFT produces the output \texttt{"BIGsmall"}.
+For example, given the input string \texttt{bigSMALL}, this SFT produces the output \texttt{BIGsmall}.
 

### 2025-09-12 Paper/application.tex
polish

@@
 (1) $\texttt{str}$: The input string to be transformed.
-(2) $\texttt{pattern}$: A regular expression defining the matching criteria.
+(2) $\texttt{pattern}$: A regular expression defining the matching pattern.
 (3) $\texttt{replacement}$: The string that replaces matched substrings.
-(4) $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single first occurrence replacement and $g$ denotes all occurrences replace operation.
+(4) $\texttt{flag}=s/g$: The replacement mode flag, where $s$ denotes single first occurrence replacement and $g$ denotes all occurrences replacement operation.
 
@@
 
-To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/}, "\texttt{NUM}", \texttt{flag})$. We construct the corresponding SFT by the following three steps.
+To illustrate how SFTs model this operation, consider $\texttt{replace}(s, \texttt{/[0-9]+/},~$\texttt{NUM}$,~\texttt{flag})$. We construct the corresponding SFT by the following three steps.
 
 \begin{enumerate}
-  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$f = \lambda x.~\texttt{None}$" that produces the empty string.
-  \item Construct an SFA accepting the replacement string "\texttt{NUM}".
-Since constant strings are specialized regular expressions, we build their SFA representation and convert it to an SFT by adding $\varepsilon$-transitions with output functions that emit the replacement characters sequentially.
+  \item Construct an SFA recognizing the pattern $\texttt{/[0-9]+/}$ and transform it into an SFT by augmenting each transition with output function "$\lambda x.~\texttt{None}$" that produces the empty string.
+  \item Construct an SFA accepting the replacement string \texttt{NUM} and convert it to an SFT by
+  replacing each transtion in the SFA $(q, c, q')$ with
+  $(q, (\texttt{None},~\lambda \_.~c ), q')$.
   \item Compose the two SFTs by connecting the accepting state of the pattern-matching SFT to the initial state of the replacement-generating SFT via $\varepsilon$-transitions.
@@
 
-Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT generated from \texttt{/[0-9]+/}  with the replacement-generating SFT generated from "\texttt{NUM}", where the function \texttt{f} is defined as "$f = \lambda x.~\texttt{None}$" and the function \texttt{g} is defined as:
-\[
-\texttt{g} = \lambda~i~x.~\texttt{match}~i~\texttt{with}~
-\begin{cases}
-1 \mapsto [(78, 78)] & \text{(ASCII for 'N')} \\
-2 \mapsto [(85, 85)] & \text{(ASCII for 'U')} \\
-3 \mapsto [(77, 77)] & \text{(ASCII for 'M')} \\
-\_ \mapsto \texttt{None}
-\end{cases}
-\]
+Figure \ref{fig-rearranged-automata} presents the complete SFT construction, obtained by composing the pattern-matching SFT generated from \texttt{/[0-9]+/}  with the replacement-generating SFT generated from \texttt{NUM}.
 
@@
      \path[->] 
-     (q0') edge [bend left] node {\texttt{[0-9]}, $\texttt{f}$} (q1')
-     (q1') edge [loop right] node {\texttt{[0-9]}, $\texttt{f}$} (q1');
+     (q0') edge [bend left] node {\texttt{[0-9]}, $\lambda~x.~\texttt{None}$} (q1')
+     (q1') edge [loop right] node {\texttt{[0-9]}, $\lambda~x.~\texttt{None}$} (q1');
 
@@
     \path[->] 
-    (q0'') edge node {\texttt{None}, \texttt{(g 1)}} (q1'')
-    (q1'') edge [bend left=5] node[left] {\texttt{None}, \texttt{(g 2)}} (q2'')
-    (q2'') edge node {\texttt{None}, \texttt{(g 3)}} (q3'')
-    (q1') edge [bend right=30] node {\texttt{None}, \texttt{f}} (q0'');
+    (q0'') edge node {\texttt{None}, $\lambda\_.\texttt{N}$} (q1'')
+    (q1'') edge [bend left=5] node[left] {\texttt{None}, $\lambda\_.\texttt{U}$} (q2'')
+    (q2'') edge node {\texttt{None}, $\lambda\_.\texttt{M}$} (q3'')
+    (q1') edge [bend right=30] node {\texttt{None}, $\lambda x.~\texttt{None}$} (q0'');
   \end{tikzpicture}
-  \caption{The SFT for the replacement operation \\ $\texttt{replace}(s, \texttt{/[0-9]+/}, \text{``NUM''})$}
+  \caption{The SFT for the replacement operation $\texttt{replace}(s,\texttt{/[0-9]+/}, \texttt{NUM})$}
   \label{fig-rearranged-automata}
@@
 
-We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that supports only basic string operations such as concatenation and regular constraints.
+We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that previously supported only basic string operations such as concatenation and regular constraints.
+The original CertiStr was incapable of verifying any XSS attack due to the lack of support for transducers and replacement operations.
 CertiStr employs a forward-propagation approach to solve string constraints, which requires computing the forward image of string operations.
@@
 Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string
-solvers, including CVC5 \cite{cvc5} and Z3Str \cite{Z3-str3}, yet CertiStrR solves
+solvers, including CVC5 \cite{cvc5}, yet CertiStrR solves
 them efficiently. Among existing solvers, only OSTRICH \cite{pacmpl/ChenFHHHKLRW22} can handle these benchmarks.
@@
 Consider the transition $[\texttt{a-z}]/id$, where $id$ is the identity function.
-To model this transition with replacement operations, the interval must be explicitly expanded into the individual characters $\texttt{'a'}$ through $\texttt{'z'}$, requiring 26 nested \texttt{replace-all} operations—one for each character.
+To model this transition with replacement operations, the interval must be explicitly expanded into the individual characters $\texttt{a}$ through $\texttt{z}$, requiring 26 nested \texttt{replace-all} operations—one for each character.
 Thus, modeling $[\texttt{a-z}]/id$ with replacement operations yields an expression of the form:
 \[
-\texttt{replace}(\texttt{replace}(\ldots,~\texttt{"b"},~\texttt{"b"},~g),~\texttt{"a"},~\texttt{"a"},~g).
+\texttt{replace}(\texttt{replace}(\ldots,~\texttt{b},~\texttt{b},~g),~\texttt{a},~\texttt{a},~g).
 \]

### 2025-09-12 Paper/formalization.tex
polish

@@
 	\end{lstlisting}
-\caption{The formalization of $SFTs$ in Isabelle/HOL}
+\caption{The formalization of SFTs in Isabelle/HOL}
 \label{fig-def-FT}
@@
 Figure \ref{fig-bool-algebra-locale} presents the simplified definition of a boolean algebra locale.
-The type variable \texttt{'b} represents any boolean algebra, while the type variable \texttt{'a} denotes the element type of the sets that the boolean algebra represents.
+The type variable \texttt{$'$b} represents any boolean algebra, while the type variable \texttt{$'$a} denotes the element type of the sets that the boolean algebra represents.
 The locale defines a collection of operations and assumptions that characterize boolean algebra behavior, including operations for set semantics, emptiness checks, non-emptiness checks, intersection, difference, and element membership.
@@
 The locale serves as an interface to boolean algebra implementations.
-Any concrete boolean algebra implementation must provide a concrete instantiation of this locale, which includes defining the type variable \texttt{'b} and providing implementations for all operations along with proofs of the required assumptions.
+Any concrete boolean algebra implementation must provide a concrete instantiation of this locale, which includes defining the type variable $'$\texttt{b} and providing implementations for all operations along with proofs of the required assumptions.
 
@@
                (sem s1) $\cap$ (sem s2)" and
-  diff_sem: "sem (diff f1 f2 s1 s2) =
+  diff_sem: "sem (diff s1 s2) =
                (sem s1) - (sem s2)" and
@@
 
-Figure \ref{fig-compute-nft-product} illustrates the core algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations: \texttt {nft\_tranfun}, \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, and \texttt{nfa\_accepting}.
-These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
+Figure \ref{fig-compute-nft-product} illustrates the algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
+Note that in the implementation, there are some refined operations:  \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nfa\_accepting},
+and \texttt{nft\_tranfun}. These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
 

### 2025-09-12 Paper/introduction.tex
polish

@@
 provide two applications: (1) analysis of the web application sanitizers for html, css, json, and javascript
-which are used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
-compliant certified string solver handling equality, concatenation, regular
+which are used for preventing XSS attacks, and (2) a certified string solver handling equality, concatenation, regular
 constraints, and more importantly the string replaceall operator, which has been
@@
 \begin{enumerate}
-\item We present the first formalization of symbolic finite transducers (SFTs) in Isabelle/HOL.
+\item We present the first formalization of symbolic finite transducers  in Isabelle/HOL.
 We prove the correctness of key operations of SFTs and symbolic SFAs, such as the product operation between SFTs and SFAs.

### 2025-09-12 Paper/application.tex
polish application introduction and motivation

@@
 
-We first evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer SFT shown in Figure \ref{fig:html-sanitizer-sft} with SFAs constructed from randomly generated Unicode sequences.
-This product represents the sanitization result produced by the SFT.
+We first evaluate the performance of our SFT product operation by computing the product of the HTML encoding sanitizer shown in Figure \ref{fig:html-sanitizer-sft} with SFAs constructed from randomly generated Unicode sequences.
+This product represents the sanitization result by the SFT for the input sequence.
 The experimental evaluation was conducted on a laptop equipped with an Apple M4 processor and 24 GB of memory, with a one-minute time limit imposed for each test.
@@
 
-  We can extend the above construction to model both replace-first and replace-all operations.
+  We can extend the above construction to model both replace-first and replace-all.
 In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}.
@@
 
-We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that previously supported only basic string operations such as concatenation and regular constraints.
+We evaluate our SFT-based replacement operation by extending CertiStr \cite{cpp/KanLRS22}, a certified string solver that previously supported only basic string operations: concatenation and regular constraints.
 The original CertiStr was incapable of verifying any XSS attack due to the lack of support for transducers and replacement operations.
@@
 We evaluate {CertiStrR} using benchmarks from SMT-LIB 2025~\cite{smtlib_benchmarks}, focusing on the \textsf{QF\_S} and \textsf{QF\_SLIA} logic fragments.
-The benchmarks are divided into two categories: (1) replace-first matching and (2) replace-all matching.
-In addition to replacement, they also include string concatenation and regular constraints in the tests.
+The benchmarks are divided into two categories: (1) replace-first and (2) replace-all.
+In addition to replacement, the tests also include string concatenation and regular constraints.
+Due to the limitation of CertiStrR's front-end, some constraints were rewritten or removed from the tests.
 
@@
    & Replace Str & 323 & 103 & 211 & 9 & 1.94 & 7.10/20 & 0.81/1 \\
+     \midrule
   \multirow{4}{*}{Replace-All} & PCP & 1000 & 751 & 249 & 0 & 0.14 & 3.412/0 & unknown \\
@@
 The results from CertiStrR were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
-An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one minute).
+An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one hour).
 This occurs when the solver cannot determine whether string variables need to be assigned different values that satisfy all constraints.
@@
 The experimental results demonstrate that string replacement operations, particularly replace-all operations, pose significant challenges for existing state-of-the-art solvers.
-The "unknown" label in the table indicates instances where CVC5 cannot determine satisfiability. 
+The "unknown" label in the table indicates the tests where CVC5 cannot determine satisfiability. 
 Notably, PCP and RNA benchmarks cannot be solved by most state-of-the-art string
@@
 CertiStrR outperforms both Ostrich and CVC5 across most benchmark subgroups, with the only exception of the "Replace Str" category where CVC5 achieves better performance (though it experiences one timeout).
-CertiStrR demonstrates particularly strong performance in the "Replace All" category.
-These results demonstrate the effectiveness and efficiency of our certified SFT formalization and implementation.
+CertiStrR demonstrates particularly strong performance in  "Replace All" category.
+These results demonstrate the effectiveness and efficiency of our SFT formalization and implementation.
 
@@
 
-Consider the transition $[\texttt{a-z}]/id$, where $id$ is the identity function.
-To model this transition with replacement operations, the interval must be explicitly expanded into the individual characters $\texttt{a}$ through $\texttt{z}$, requiring 26 nested \texttt{replace-all} operations—one for each character.
-Thus, modeling $[\texttt{a-z}]/id$ with replacement operations yields an expression of the form:
-\[
-\texttt{replace}(\texttt{replace}(\ldots,~\texttt{b},~\texttt{b},~g),~\texttt{a},~\texttt{a},~g).
-\]
+Consider a string transformation that maps each character in $[\texttt{a-z}]$ to its successor, with \texttt{z} wrapping around to \texttt{a}.
+Modeling this transformation using only replacement operations requires spliting the interval $[\texttt{a-z}]$ into individual characters, resulting in 26 separate \texttt{replace-all} operations—one for each letter.
+For example:
+\begin{verbatim}
+  replace(replace(..., b, c, g), a, b, g)
+\end{verbatim}
 
@@
 The most challenging component proved to be the correctness proof of the product operation at the abstract level (Figure~\ref{fig-def-product-correct}).
-
-Compared to CertiStr \cite{cpp/KanLRS22}, the interval formalization in CertiStrR is substantially more complex, primarily because intervals were extended from single pairs to lists.
+%
+%Compared to CertiStr \cite{cpp/KanLRS22}, the interval formalization in CertiStrR is substantially more complex, primarily because intervals were extended from single pairs to lists.
 

### 2025-09-12 Paper/introduction.tex
polish application introduction and motivation

@@
 provide two applications: (1) analysis of the web application sanitizers for html, css, json, and javascript
-which are used for preventing XSS attacks, and (2) a certified string solver handling equality, concatenation, regular
+which are used for preventing XSS attacks, and (2) string solving handling equality, concatenation, regular
 constraints, and more importantly the string replaceall operator, which has been
@@
 We prove the correctness of key operations of SFTs and symbolic SFAs, such as the product operation between SFTs and SFAs.
-\item We leverage the refinement framework to develop an extensible and efficient SFT formalization capable of supporting diverse Boolean algebras, ensuring adaptability to future transition label representations.
+\item We leverage the refinement framework to develop an extensible and efficient SFT implementation capable of supporting diverse Boolean algebras, ensuring adaptability to future transition label representations.
 The correctness of the refined, efficient implementation is proved via structural isomorphism between refinement layers.

### 2025-09-13 Paper/application.tex
polish application and intro

@@
 As shown in the table, GAHtmlEscape and escapeString are unsafe with respect to the HTML attack model without preprocessing, but become safe when combined with Trim.
-The \textbf{Size} column demonstrates the expressive power and succinctness of SFTs in modeling real-world transducers.
+The \textbf{Size} column demonstrates the succinctness of SFTs in modeling real-world transducers.
 We did not list the execution time of verification because all of them are quite efficent with less than 0.02 second to finish the verificaiton.
@@
 
-  We can extend the above construction to model both replace-first and replace-all.
+  We can extend the above construction to model both replace-first and replace-all for XSS attack verification.
 In this paper, we only illustrate the construction of replace-all in Figure \ref{fig:replace-all}.
@@
 This cyclic structure enables the replace-all operation by repeatedly applying pattern replacement until no more matches exist in the input string.
+Replace-first modeling has no transtion from $\mathcal{T}_2$'s accepting state back to $\mathcal{T}_1$'s initial state.
 
@@
 
-Consider a string transformation that maps each character in $[\texttt{a-z}]$ to its successor, with \texttt{z} wrapping around to \texttt{a}.
+Consider a string transformation that replace each character in $[\texttt{a-z}]$ with its ASCII code.
 Modeling this transformation using only replacement operations requires spliting the interval $[\texttt{a-z}]$ into individual characters, resulting in 26 separate \texttt{replace-all} operations—one for each letter.
@@
 \begin{verbatim}
-  replace(replace(..., b, c, g), a, b, g)
+  replace(replace(..., b, 0x62, g), a, 0x61, g)
 \end{verbatim}

### 2025-09-13 Paper/introduction.tex
polish application and intro

@@
 \cite{Berkeley-JavaScript,systematic-transduction,Kern,uss/HooimeijerLMSV11,DBLP:conf/popl/LinB16}. In particular, SMT-LIB string constraints that 
-arise from such applications require the \emph{replaceall} operator (part of
-SMT-LIB), which is currently not handled by any existing verified
+arise from such applications require the \emph{replaceall} operator, which is currently not handled by any existing verified
 implementation. This paper tackles the formalization of SFA/SFT and

### 2025-09-13 Paper/abstract.tex
abstract and intro

@@
     represented models. To circumvent to the notorious state-space explosion 
-    problem caused by an extremely large alphabet size (e.g. UTF-16), SFTs
+    problem caused by an extremely large alphabet size (e.g. Unicode), SFTs
     and SFAs allow the representation of the alphabet as an effective boolean

### 2025-11-24 Paper/main.tex
finish experiments
Added new section
motivation seciton
Removed motivating
New title and abstract
experimental results
add todo_list for camera ready

@@
-\documentclass[sigplan,10pt,anonymous,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %This is a template for producing LIPIcs articles. 
@@
 \usepackage{xcolor}
+\usepackage{multirow}
+
+\input{comm}
 
@@
 \setlength{\pdfpagewidth}{\paperwidth}
+\author{Shuanglong Kan}
+\affiliation{
+  \institution{Barkhausen Institute}
+  \country{Germany}
+}
 
-\title{Certified Symbolic Finite Transducers: Formalization and Applications to String Manipulation}
+\author{Anthony W. Lin}
+\affiliation{
+  \institution{TU Kaiserslautern \& MPI-SWS}
+  \country{Germany}
+}
+
+\title{Certified Symbolic Finite Transducers: Formalization and Applications to
+String Analysis}
 
@@
 \input{introduction.tex}
+
+\input{motivation.tex}
 

### 2025-11-25 Paper/SFT.tex
fix reviews

@@
    \item $\Delta$ is the set of transition relations.
-Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$.
-   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$.
-   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$. $f$ is called an \emph{output function}.
+Each element in $\Delta$ is of the form $(q, \phi, f, q')$ or written as $q\xrightarrow{\phi, f} q'$, where $q$ and $q'$ are states in $\mathcal{Q}$,
+   $\phi$ is a predicate of type $\tau_1\rightarrow \mathbb{B}$,
+   $f$ is a lambda term of type $\tau_1\rightarrow \tau_2$, and $f$ is called an \emph{output function}.
    \end{itemize}

### 2025-11-25 Paper/formalization.tex
fix reviews

@@
 
-Firstly, $'a \text{ set option }$ is the input type of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$. 
+Firstly, $'a \text{ set option }$ is the input type (\texttt{option} type of set) of the transition, it accepts a set of elements of type $'a$ or $\texttt{None}$ corresponding to empty string $\varepsilon$.
 Accepting a set of $'a$ elements aims to express the same but more abstract semantics of the input labels in Definition \ref{def-sft}, in which an input label is a predicate.
@@
 %
-The value $\texttt{None}$ represents the empty string $\varepsilon$ in our formalization, indicating a transition that consumes no input but may still produce output elements.
-This design choice facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
+The value $\texttt{None}$ (empty string) facilitates the modeling of real-world applications in SFTs, as we will demonstrate in our application to string solvers.
 
@@
 function. These output functions, formalized by \texttt{Tlabel}, map a single
-input element to a set of possible output elements rather than to a single
-element. This design enables nondeterministic output behavior, where the
+input element to a set of possible output elements to model nondeterministic output behavior, where the
 transducer may select any element from the output set arbitrarily or according
@@
 
-type_synonym |$('q, 'a)$| LTS = "|$'q \times 'a \text{ set }\times 'q$|"    
+type_synonym |$('q, 'a)$| LTS =
+          "|$'q~\times~'a \text{ set }\times ~'q$| set"
 	\end{lstlisting}
@@
   |$\mathcal{Q}e = \mathcal{Q}_t~\mathcal{T} \times \mathcal{Q} ~\mathcal{A}$|,
-  |$\Delta_e = \{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
+  |$\Delta_e =$| (* case of |$\epsilon$|-transitions *)
+      |$\{((p,p'),~\texttt{the }(((\mathcal{M}_t~\mathcal{T})~f)\texttt{ None}),~(q, p'))\mid $|
        |$p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A}~\land$| |$(p, (\texttt{None}, f), q)\in \Delta_t~\mathcal{T}$|
         |$ \land~\exists S.~(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{Some } S\} ~\cup$|
-      |$\{((p,p'),~\texttt{the }(F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
+      (* case of non-|$\epsilon$|-transitions *)
+      |$\{((p,p'),~\texttt{the }(\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2)),(q, q'))\mid $|
        |$p, p', q, \sigma_1, \sigma_2, q', f.~(p, (\text{Some }\sigma_1, f), q)\in \Delta_t~\mathcal{T} \land$|
        |$ (p', \sigma_2, q')\in\Delta~\mathcal{A}\land \sigma_1\cap \sigma_2 \neq \emptyset~\land$|
-          |$ \exists S.~F~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
-  |$\Delta_e' = \{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
+          |$ \exists S.~\texttt{F}~((\mathcal{M}_t~\mathcal{T})~f)~(\sigma_1\cap\sigma_2) = \text{Some } S\}$|,
+  |$\Delta_e' =$| (* case of |$\epsilon$|-transitions *)
+      |$\{((p,p'),~(q, p'))\mid p, p', q, f.~p'\in \mathcal{Q}~\mathcal{A} \land$|
        |$(p, (None, f), q)\in \Delta_t~\mathcal{T}~\land $|
        |$(\mathcal{M}_t~\mathcal{T})~f~\texttt{None} = \text{None}\} ~\cup$|
+      (* case of non-|$\epsilon$|-transitions *)
       |$\{((p,p'),~(q, q'))\mid p, p', q, \sigma_1, \sigma_2, q', f.~$|
@@
 
-\[\texttt{F}~f~A=\texttt{Some } (\bigcup_{a\in A} (\text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset))\].
+\[\texttt{F}~f~A=\texttt{Some } \bigcup_{a\in A} \text{if }f~a= \text{Some }S \texttt{ then } S \texttt{ else } \emptyset\].
 
@@
   assumes
-   F_ok1: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
-       |$\text{ F}~f~s = \text{None}"$|
-   and F_ok2: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| = 
+   F_None: "|$\forall~f~s.~(\forall e \in s.~f~(\text{Some } e) = \text{None})\longleftrightarrow$|
+       |$\text{ F}~f~s = \text{None}"$| and
+   F_Some: "|$\forall~f~s.~ \texttt{F}~f~s\neq \texttt{None} \longrightarrow$| F |$f~s$| =
         |$\text{Some }(\cup~\{\texttt{S} \mid e~\texttt{S}.~e \in s~\land~$|
@@
 Figure \ref{fig-def-product-correct} presents Lemma \texttt{productT\_correct}, which establishes the correctness of the product operation.
-The lemma's assumptions, \texttt{F\_ok1} and \texttt{F\_ok2}, specify the essential properties of function \texttt{F}.
+The assumptions, \texttt{F\_None} and \texttt{F\_Some}, specify the essential properties of function \texttt{F}.
 The assumptions $\texttt{NFT\_wf}~\mathcal{T}$ and $\texttt{NFA}~\mathcal{A}$ ensure that the SFT $\mathcal{T}$ and the SFA $\mathcal{A}$ are well-formed.

### 2025-11-25 Paper/introduction.tex
abstract and intro
fix reviews

@@
 (which easily prone to XSS vulnerabilities) and string constraint solving
---- require extremely large alphabets (e.g. UTF-16), which are beyond the
+--- require extremely large alphabets (e.g. Unicode), which are beyond the
 capabilities of classical automata algorithms
@@
 Symbolic Finite Automata (SFA) and Symbolic Finite Transducers (SFT)
-\cite{cav/DAntoniV17, VeanesHLMB12Transducer,DV21} were proposed around 15 years ago
+\cite{cav/DAntoniV17, VeanesHLMB12Transducer,DV21} were proposed around 2010
 to deal with the alphabet-explosion problem that often arises in practical
@@
 arise from such applications require the \emph{replaceall} operator, which is currently not handled by any existing verified
-implementation. This paper tackles the formalization of SFA/SFT and
+implementation, as far as we know. This paper tackles the formalization of SFA/SFT and
 their associated algorithms, as well as applications to analysis of sanitizers

### 2025-11-26 Paper/application.tex
finish fix reviews

@@
 Our SFT product operation provides an exact forward image computation for replacement operations, enabling seamless integration of replacement functionality into CertiStr's solving framework.
+The OCaml code of the SFT implementation is extracted from our Isabelle formalization using Isabelle's extraction mechanism, ensuring that the core solving algorithm of CertiStr remains certified.
 
@@
 Table~\ref{tab:string_operations} summarizes the results.
-The results from CertiStrR were classified into three categories: SAT (satisfiable), UNSAT (unsatisfiable), and Inconclusive.
+The results from CertiStrR were classified into three categories: SAT (satisfiable), indicating that there exist assignments of string variables satisfying the constraints; UNSAT (unsatisfiable), indicating that no such assignments exist; and Inconclusive.
 An "Inconclusive" result indicates that the solver cannot determine satisfiability due to the inherent incompleteness of the string solver's algorithm, not timeout (all tests completed within one hour).
@@
 These results demonstrate the effectiveness and efficiency of our SFT formalization and implementation.
+
+Our approach demonstrates superior efficiency compared to other string solving approaches for several key reasons.
+First, our SFTs are implemented with compact symbolic labels that enable efficient representation of infinite alphabets, and we employ red-black trees for storing states, ensuring logarithmic-time lookup and insertion operations.
+Second, our forward propagation strategy proves more efficient than alternative approaches such as term rewriting or backpropagation employed by other solvers.
+While forward propagation supports only a limited set of string constraints—for instance, it is only complete for string constraints satisfy tree properties \cite{cpp/KanLRS22}—this trade-off between completeness and efficiency enables our solver to handle large-scale benchmarks that remain intractable for more general but computationally expensive approaches.
 

### 2025-11-26 Paper/formalization.tex
finish fix reviews

@@
 	\begin{lstlisting}
-definition productT :: |$"('q,'a,'i,'b)"$| NFT |$\Rightarrow$|
-   |$('q,'a)$| NFA |$\Rightarrow$||$(('a,'b)$| Tlabel |$\Rightarrow 'a \text{ set}$|
-   |$\Rightarrow 'b\text{ set option})\Rightarrow$| |$('q\times 'q, ~'b)$| eNFA where
+definition productT :: |$"('{q},~'a,~'i,~'b)"$| NFT
+  |$\Rightarrow$| |$('q,~'a)$| NFA |$\Rightarrow$||$(('a,~'b)$| Tlabel |$\Rightarrow~'a \text{ set}$|
+  |$\Rightarrow~'b\text{ set option})\Rightarrow$| |$('q~\times~'q, ~'b)$| eNFA where
 "productT |$\mathcal{T}$| |$\mathcal{A}$| F = |$\llparenthesis$|
@@
 This invariant serves two purposes: it simplifies formal proofs by eliminating the need to reason about malformed or overlapping intervals, and it enables more efficient implementations of interval operations by reducing the number of cases to consider.
+
+The formalization of SFA/SFT presents distinct challenges compared to classical FA/FT, primarily stemming from the symbolic nature of transition labels.
+Unlike classical automata where labels are concrete alphabet symbols, SFAs and SFTs use Boolean algebras to represent potentially infinite label sets.
+This introduces two key complications.
+First, our formalization must accommodate different Boolean algebra implementations (intervals, bit vectors, predicates, etc.) while maintaining uniform semantics, necessitating the locale-based abstraction layer that separates the fixed automata structure from variable label theories.
+Second, operations on transition labels—such as intersection and membership testing—require Boolean algebra computations rather than simple syntactic comparisons.
+This demands explicit reasoning about semantic preservation: we must prove that operations on Boolean algebra representations (e.g., interval intersection) correctly preserve the set semantics defined by the \texttt{sem} function.
+These complications, absent in classical automata where label operations are trivial, required careful architectural design and substantial additional proof effort to establish correctness.

### 2025-12-01 Paper/application.tex
camera-ready

@@
 
-Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function for Anti-XSS.
+Figure \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function for Anti-XSS.
 This function iterates through each character in the input string and encodes characters that are unsafe for HTML rendering.
 
-\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
+\begin{figure}[hbt!]
+\begin{lstlisting}[language={[Sharp]C}]
 static string EncodeHtml(string t)
@@
 \end{lstlisting}
+\caption{C\# Code for AntiXSS.EncodeHtml version 2.0.}
+\label{lst-html-encode}
+\end{figure}
 
@@
   \small
+  \caption{Sanitizer verfication result}
+  \label{tab:sanitizer_attack}
   \setlength{\tabcolsep}{4pt} % Reduced column gap
@@
   \end{tabular}
-  \caption{Sanitizer verfication result}
-  \label{tab:sanitizer_attack}
 \end{table}
@@
 
-\begin{lstlisting}[language={}, caption={CSS attack model for sanitizer verification.}, label={lst-css-attack}, float=htbp]
+\begin{lstlisting}[language={}, caption={CSS attack model for sanitizer verification}, captionpos=b, label={lst-css-attack}, float=htbp]
 |$\Sigma$|* ( / \* [\s\S]*? \*/
@@
 \small
+\caption{Experimental results}
+\label{tab:string_operations}
 \setlength{\tabcolsep}{4pt}
@@
 \end{tabular}
-\caption{Experimental results}
-\label{tab:string_operations}
 \end{table*}
@@
   \small
+  \caption{Overview of the effort of certified development}
+  \label{tab:abstract_impl}
   \begin{tabular}{lccc}
@@
   \end{tabular}
-  \caption{Overview of the effort of certified development}
-  \label{tab:abstract_impl}
 \end{table}

### 2025-12-01 Paper/formalization.tex
camera-ready

@@
 
-\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.} based on Refined\_monadic framework.
+\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA based on Refined\_monadic framework.
 
@@
 
-  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
+  The correctness of the product operation is established by a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}).
 This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.

### 2025-12-01 Paper/main.tex
fix reviews
camera-ready

@@
-\documentclass[sigplan,10pt]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,screen]{acmart}
 %This is a template for producing LIPIcs articles. 
@@
 \usepackage{multirow}
+\usepackage{balance}
+\usepackage{microtype}
 
@@
 }
-\lstset{
-  language=SMTLIB,
-  basicstyle=\ttfamily\small,
-  keywordstyle=\color{blue}\bfseries,
-  commentstyle=\color{green!50!black},
-  numbers=left,
-  numberstyle=\tiny\color{gray},
-  stepnumber=1,
-  numbersep=5pt,
-  backgroundcolor=\color{gray!10},
-  frame=single,
-  breaklines=true,
-  captionpos=b,
-  showstringspaces=false
-}
 
-\bibliographystyle{plainurl}% the mandatory bibstyle
-
-%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \lstdefinelanguage{Isabelle}{
@@
 \lstset{
-  language=Isabelle,
   basicstyle=\ttfamily,
   keywordstyle=\color{blue}\bfseries,
-	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and, locale},
+  morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and,locale,set-logic,declare-fun,assert,check-sat,get-model},
   escapeinside={|}{|},
-  sensitive=true
+  sensitive=true,
+  commentstyle=\color{green!50!black},
+  morecomment=[l];,
+  morestring=[b]",
+  breaklines=true,
+  showstringspaces=false
 }
 
+\bibliographystyle{ACM-Reference-Format}% the mandatory bibstyle
+
+%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
+
+\title[Certified Symbolic Finite Transducers]{Certified Symbolic Finite Transducers: Formalization and Applications to
+String Analysis}
 
@@
 \setlength{\pdfpagewidth}{\paperwidth}
+
 \author{Shuanglong Kan}
-\affiliation{
-  \institution{Barkhausen Institute}
+\orcid{0000-0002-4132-2626}
+\affiliation{%
+  \department{ Verified System Design Automation Group}
+  \institution{Barkhausen Institut}
+  \city{Dresden}
   \country{Germany}
 }
+\email{shuanglong.kan@barkhauseninstitut.org}
 
 \author{Anthony W. Lin}
+\orcid{0000-0003-4715-5096}
 \affiliation{
-  \institution{TU Kaiserslautern \& MPI-SWS}
+  \department{Department of Computer Science}
+  \institution{University of Kaiserslautern-Landau \& MPI-SWS}
+  \city{Kaiserslautern}
   \country{Germany}
 }
-
-\title{Certified Symbolic Finite Transducers: Formalization and Applications to
-String Analysis}
+\email{awlin@mpi-sws.org}
 
 \input{abstract.tex}
+
+\begin{CCSXML}
+<ccs2012>
+   <concept>
+       <concept_id>10003752.10003790.10003794</concept_id>
+       <concept_desc>Theory of computation~Automated reasoning</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+   <concept>
+       <concept_id>10003752.10003766.10003773.10003774</concept_id>
+       <concept_desc>Theory of computation~Transducers</concept_desc>
+       <concept_significance>500</concept_significance>
+       </concept>
+ </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Automated reasoning}
+\ccsdesc[500]{Theory of computation~Transducers}
+
+\keywords{symbolic transducers, Isabelle/HOL, string theory, SMT solvers} %TODO mandatory; please add comma-separated list of keywords
 \maketitle
@@
 
+\balance
 \bibliography{literature}

### 2025-12-01 Paper/relatedwork.tex
Added citation for certified proof checkers
finish fix reviews
Added some additional work in the citation and discussion

@@
 This symbolic framework has been progressively extended to accommodate more complex structures: symbolic tree transducers \cite{ershov/VeanesB11} handle hierarchical data structures, while symbolic pushdown automata \cite{cav/DAntoniA14} manage nested word structures.
+Recent work has also explored regex decision procedures using symbolic techniques \cite{VarataluVZECAV25}.
 %Recent developments in 2024 have further expanded the scope of symbolic techniques to include B\"{u}chi automata and omega-regular languages \cite{pacmpl/VeanesBEZ25}, enabling verification of infinite-state systems and analysis of non-terminating computations.
@@
 expressive power of symbolic automata and transducers have led to their
-widespread adoption in numerous applications, including: constraint solving for program analysis \cite{lpar/VeanesBM10}, security-critical sanitizer analysis for web applications \cite{uss/HooimeijerLMSV11}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, automated program inversion for software
+widespread adoption in numerous applications, including: constraint solving for
+program analysis \cite{lpar/VeanesBM10,CHL+19}, security-critical sanitizer
+analysis for web applications \cite{uss/HooimeijerLMSV11,pacmpl/ChenFHHHKLRW22,HJLRV18}, runtime verification of system behaviors \cite{osdi/YaseenABCL20}, automated 
+program inversion for software 
 transformation \cite{pldi/HuD17}, and string solving
-\cite{pacmpl/ChenFHHHKLRW22,CHL+19}. Each application leverages the symbolic 
-approach's ability to handle complex
-patterns and large alphabets efficiently.
+\cite{pacmpl/ChenFHHHKLRW22,CHL+19,HJLRV18}. Each application leverages the 
+symbolic approach's ability to handle complex patterns and large alphabets 
+efficiently.
 
-\emph{Formalization of Symbolic Automata and Transducers.} While classical automata theories have been extensively formalized in interactive theorem provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic variants of automata and transducers remain largely unexplored in formal verification.
+\emph{Formalization of Symbolic Automata and Transducers.} While classical
+automata theories have been extensively formalized in interactive theorem
+provers \cite{Tuerk-NFA, Lammich2014TheCA, Peter14, cpp/DoczkalKS13}, with some
+work on transducer formalization \cite{afp/LochmannFSTS21}, the symbolic
+variants of automata and transducers remain largely unexplored in formal
+verification. Initial works on formalizing symbolic automata and symbolic
+regular expressions are available (e.g. see \cite{cpp/KanLRS22,ZMVE25,ZVE24}),
+but (to the best of our knowledge) our work is the first that formalizes
+symbolic transducers.
 %To our knowledge, CertiStr \cite{cpp/KanLRS22} represents the only existing work on symbolic automata formalization in a proof assistant.
@@
 to check compatibility between solvers and SMTLIB standard.
+Finally, there are also recent works on certified proof checkers for automated 
+reasoning and in particular SMT (cf. \cite{cacm-certified}).

### 2025-12-03 Paper/abstract.tex
Shortened abstract

@@
     SMT-Algebras.
-    %can be instantiated with various Boolean algebras, 
     The
@@
     Isabelle/HOL. 
-    %This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
-%
     To evaluate the effectiveness of our formalization, we apply the formalized 
@@
     employs intricate string replacement operations.
-    %a string solver that models replacement operations with various semantics, including left-most and replace-all matching. 
     Our experimental results demonstrate that our methods are competitive with
     the existing unverified implementations.
-
-    %that our approach achieves computational efficiency in constraint-solving scenarios.
-
-    %, thereby providing a more expressive framework for operations that encompass both recognition and transformation. 
-%FTs have wide applications, including program analysis, string constraint 
-    %solving, and security-critical sanitizer analysis. 
-%However, there is currently no scalable formalization of FTs.
-This is primarily because transition labels in FTs are not formalized symbolically, resulting in transition explosion and significant performance bottlenecks in practical applications.
-
-    \OMIT{
-In this paper, we present a formalization of FTs in the symbolic setting using Isabelle/HOL, where transition labels can be instantiated with various Boolean algebras, such as intervals and arithmetic predicates.
-This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
-%
-To evaluate the effectiveness of our formalization, we apply the formalized SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for preventing XSS attacks, and (2) a string solver that models replacement operations with various semantics, including left-most and replace-all matching.
-Experimental results demonstrate that our approach achieves computational efficiency in constraint-solving scenarios.
-    }
 \end{abstract}

### 2025-12-03 Paper/formalization.tex
fix errors

@@
 Figure \ref{fig-compute-nft-product} illustrates the algorithm \texttt{productT\_impl}, which is implemented using the \emph{Refine\_Monadic} framework.
-Note that in the implementation, there are some refined operations:  \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nfa\_accepting},
-and \texttt{nft\_tranfun}. These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
+Note that in the implementation, there are some refined operations:  \texttt{nfa\_states}, \texttt{nfa\_trans}, \texttt{nfa\_initial}, \texttt{nft\_tranfun},
+and \texttt{nfa\_accepting}. These are corresponding to the states, transitions, initial states, accepting states, and output function mapping of the SFT.
 
@@
 
-    We now present the implementation of \texttt{subtrans\_comp} in detail, as shown in Figure \ref{fig-def-subtrans_comp} (the implementation of \texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
+    We now present the implementation of \texttt{subtrans\_comp} in detail shown in Figure \ref{fig-def-subtrans_comp} (\texttt{subtrans\_comp\_$\varepsilon$} follows analogous principles).
 For a transtion $(q, (\text{Some}~\alpha, f), q')$ in the SFT, this function traverses all transitions in the SFA, represented by the set \texttt{T}.

### 2025-12-03 Paper/main.tex
Changed institutes
Removed dept
Added ack
minor update
fix errors

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% The following is specific to CPP '26 and the paper
+%%% 'Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis'
+%%% by Shuanglong Kan and Anthony W. Lin.
+%%%
+\acmDOI{10.1145/3779031.3779094}
+\acmYear{2026}
+\copyrightyear{2026}
+\acmISBN{979-8-4007-2341-4/2026/01}
+\acmConference[CPP '26]{Proceedings of the 15th ACM SIGPLAN International Conference on Certified Programs and Proofs}{January 12--13, 2026}{Rennes, France}
+\acmBooktitle{Proceedings of the 15th ACM SIGPLAN International Conference on Certified Programs and Proofs (CPP '26), January 12--13, 2026, Rennes, France}
+\received{2025-09-12}
+\received[accepted]{2025-11-13}
 
@@
 \affiliation{%
-  \department{ Verified System Design Automation Group}
-  \institution{Barkhausen Institut}
+  \institution{Barkhausen Institute}
   \city{Dresden}
@@
 \affiliation{
-  \department{Department of Computer Science}
-  \institution{University of Kaiserslautern-Landau \& MPI-SWS}
+  \institution{Max-Planck Institute for Software Systems}
+  \institution{TU Kaiserslautern}
   \city{Kaiserslautern}
@@
 
+\paragraph{Acknowledgment.} We thank Hongjian Jiang and Philipp Rümmer for the
+discussion. The authors are supported by European Research Council (LASD, 
+grant number 101089343).
+
 \balance
